### 2021-07-09 R/climbing.Rmd
Initial commit

---
title: "Climbing Sim"
output: rmdformats::robobook 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      fig.width = 6,
                      fig.height = 7)
```

```{r}
library(tidyverse)
theme_set(theme_light())
```

# Qualification

Note: 20 athletes participate in qualification. 8 advance to final round.

Let's try 10000 sims for now.

```{r}
set.seed(1)
nsim <- 10000
qual_sim <- list()
for (i in 1:nsim) {
  qual_sim[[i]] <-
    bind_cols(
      player = 1:20,
      e1 = sample(1:20, replace = FALSE),
      e2 = sample(1:20, replace = FALSE),
      e3 = sample(1:20, replace = FALSE)
    ) %>%
    mutate(sim = i)
}

qual <- bind_rows(qual_sim) # tidyverse's do.call-rbind version

qual <- qual %>% 
  mutate(qual_score = e1 * e2 * e3) %>% 
  group_by(sim) %>% 
  arrange(qual_score, .by_group = TRUE) %>% 
  mutate(qual_rank = row_number()) 
```

## If you win the first event, what is the probability that you advance?

```{r}
qual %>% 
  filter(e1 == 1 & qual_rank <= 8) %>% 
  ungroup() %>% 
  summarize(nrow(.)/nsim)
```

9948/10000 = 0.9948

## If you win ANY event, what is the probability you advance?

```{r}
qual %>% 
  filter((e1 == 1 | e2 == 1 | e3 == 1) & qual_rank <= 8) %>% 
  nrow()
```

28420/30000 = 0.9473333

## If you finish last in the first event, what is the probability of advancing?

```{r}
qual %>% 
  filter(e1 == 20 & qual_rank <= 8) %>% 
  nrow()
```

1833/10000 = 0.1833

## If you finish last in ANY event, what is the probability you advance?

```{r}
qual %>% 
  filter((e1 == 20 | e2 == 20 | e3 == 20) & qual_rank <= 8) %>% 
  nrow()
```

5376/30000 = 0.1792

## What is the score that guarantees that you advance / do not advance?

## What is the average score of someone who advances?

```{r}
qual %>% 
  filter(qual_rank <= 8) %>% 
  ungroup() %>% 
  group_by(qual_rank) %>% 
  summarize(avg_adv_score = mean(qual_score))
```

```{r}
qual %>% 
  filter(qual_rank <= 8) %>% 
  ungroup() %>% 
  summarize(avg_adv_score_overall = mean(qual_score))
```

## What is/are the distribution on the the first/second/etc. place qualifier?

```{r}
qual %>% 
  ggplot(aes(x = factor(qual_rank), y = qual_score)) +
  geom_dotplot(binaxis = "y", binwidth = 1) + 
  coord_flip()
```

```{r}
library(ggridges)
qual %>% 
  ggplot(aes(x = qual_score, y = factor(qual_rank))) +
  geom_density_ridges()
```

# Final Round

```{r}
set.seed(1)
nsim <- 10000
final_sim <- list()
for (i in 1:nsim) {
  final_sim[[i]] <-
    bind_cols(
      player = 1:8,
      e1 = sample(1:8, replace = FALSE),
      e2 = sample(1:8, replace = FALSE),
      e3 = sample(1:8, replace = FALSE)
    ) %>%
    mutate(sim = i)
}

final <- bind_rows(final_sim)

final <- final %>% 
  mutate(final_score = e1 * e2 * e3) %>% 
  group_by(sim) %>% 
  arrange(final_score, .by_group = TRUE) %>% 
  mutate(final_rank = row_number()) 
```

## If you win any event, what is the probability that you win a medal?

```{r}
final %>% 
  filter((e1 == 1 | e2 == 1 | e3 == 1) & final_rank <= 3) %>% 
  nrow()
```

21985/30000 = 0.7328333

## If you win the first event, what is the probability you win a medal?

```{r}
final %>% 
  filter(e1 == 1 & final_rank <= 3) %>% 
  nrow()
```

8508/10000 = 0.8508

## What is the distribution of the winners score if ranks are random?

```{r}
final %>% 
  filter(final_rank == 1) %>% 
  ggplot(aes(final_score)) + 
  geom_histogram(bins = 21, color = "white")
```

### 2021-07-09 manuscript/manuscript.Rnw
Initial commit

\documentclass{article}
\usepackage{natbib}      % cross referencing bibliography entries
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{graphicx}    % importing graphics into figures
\usepackage{multirow}    % tables with multiple rows
\usepackage{slashbox}    % tables with slash
\usepackage{rotating}    % for vertical words in table
\usepackage{color}       % for textcolor.. (temporary use before submission)
\usepackage{bm}
\usepackage[section]{placeins} % ensure floats do not go into the next section.

\begin{document}
\section{Introduction}
The 2020 Olympics (held in 2021) will feature the debut of five new sports (https://olympics.com/ioc/news/ioc-approves-five-new-sports-for-olympic-games-tokyo-2020): surfing, skateboarding, karate, baseball/softball (not really new, but absent since 2008) and sport climbing.  Sport climbing is particularly interesting as the International Olympic Committee (IOC) awarded only two sets of medals for the sport, one for men's and one for women's.

What makes this particularly interesting is that, sport climbing has several distinct disciplines: Speed climbing, bouldering, and lead climbing.  Speed climbing takes place on a standardized course and competitors try to reach the top of the course as fast as possible.  In the Olympics, speed climbing is being contested in a head-to-head format with ranks determined by how far a competitor advances in the bracket.   In bouldering, contestants have a fixed amount of time to complete as many courses as they can.  Winners are determined based on who completes the most courses and ties are broken based on who had the fewest attempts.  Ties are further broken by the competitor acheived the most ``zone holds", which are holds approximately half way through each course.  Finally, in lead climbing a competitor gets one point for each hold that they reach, so whoever reaches the highest point on the wall is the winner.  In lead climbing, each competitor only gets one attempt and when they fall their attempt is over.  %(https://www.cnet.com/news/climbing-at-the-tokyo-olympics-everything-you-need-to-know/)

These three different events demand different sets of skills and, often, athletes specialize in a single event.  However, since only one set of Olympic medals was awarded to climbing, rather than choosing only one of these disciplines to include in the Olympics, all three events were chosen to be included as a sort of climbing triathlon.

In order to declare a winner, a rather unique scoring system has been put in to place, with the full details to come, that rewards high finishes in each of the individual events and relatively ignores very poor finishes.  This article examines this unique scoring system.

The rest of this article is organized as follows:

FILL IN

\section{Scoring Method}
In the Olympic sport climbing, there are 20 competitors at the start (in both men's and women's).  All 20 competitors compete in each of the three events in the qualification round, and their performances in each event are ranked from 1 to 20.  A competitors final score is then computed as the product of their ranks in the three events and the lower product is better.  Specifically,

$$
Score_i = R^S_i \times R^B_i \times R^L_i
$$,

where $R^S_i$, $R^B_i$, and $R^L_i$ are the ranks of the $i$-th competitor in speed, bouldering, and lead, respectively.

The top 8 (I think it's 8) finishers in the qualification round advance to the finals where the 8 remaining climbers again compete in all three events, they are again ranked from 1 to 8, and their final score is the product of these three ranks in the final.  Whoever has the lowest product of ranks in the final wins the gold medal.

This type of scoring system heavily rewards high finishes and relatively ignores poor finishes.  For instance, if climber A finished 1st, 20th, and 20th and climber B finished 10th, 10th, and 10th, climber B would have a score of 1000 whereas climber A would have a much better score of 400, despie finishing last in 2 out of 3 of the events.

\section{Other scoring methods}

\section{Rank Product}
https://www.sciencedirect.com/science/article/pii/S0014579304009354

Youth Olympics Qualification video: %https://www.youtube.com/watch?v=3w8kCI5GCUQ&ab_channel=Olympics

%https://www.ifsc-climbing.org/images/media-centre/2017/2017_IFSC_Plenary_Assembly_Quebec_City_Olympic_Format.pdf

%https://www.climbing.com/competition/olympics/a-guide-to-the-olympic-climbing-format/

Soccer points: %https://journals.sagepub.com/doi/10.1177/1527002507301116

Cross country scoring: %https://link.springer.com/content/pdf/10.1007/s11127-007-9193-6.pdf

https://link.springer.com/article/10.1007/s11127-017-0494-0

Decathlon scoring: %https://research.ou.nl/en/publications/decathlon-towards-a-balanced-and-sustainable-performance-assessme

Rock Climbing scoring.

\section{Results}

\section{Conclusion}

\bibliography{rockclimbing}
\bibliographystyle{chicago}

\end{document}

### 2021-07-09 manuscript/manuscript.tex
Initial commit

--- 2021-07-09 manuscript/manuscript.Rnw
@@
-\documentclass{article}
+\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage{natbib}      % cross referencing bibliography entries
@@
 \usepackage[section]{placeins} % ensure floats do not go into the next section.
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 Speed climbing takes place on a standardized course and competitors try to reach the top of the course as fast as possible.
-In the Olympics, speed climbing is being contested in a head-to-head format with ranks determined by how far a competitor advances in the bracket.
 In bouldering, contestants have a fixed amount of time to complete as many courses as they can.
@@
 In lead climbing, each competitor only gets one attempt and when they fall their attempt is over.
-%(https://www.cnet.com/news/climbing-at-the-tokyo-olympics-everything-you-need-to-know/)
 
@@
 In the Olympic sport climbing, there are 20 competitors at the start (in both men's and women's).
-All 20 competitors compete in each of the three events in the qualification round, and their performances in each event are ranked from 1 to 20.
-A competitors final score is then computed as the product of their ranks in the three events and the lower product is better.
+All 20 competitors compete in each of the three events, and their performances in each event are ranked.
+A competitors final score is then computed as the product of their ranks in the three events.
 Specifically,
@@
 
-where $R^S_i$, $R^B_i$, and $R^L_i$ are the ranks of the $i$-th competitor in speed, bouldering, and lead, respectively.
-
-The top 8 (I think it's 8) finishers in the qualification round advance to the finals where the 8 remaining climbers again compete in all three events, they are again ranked from 1 to 8, and their final score is the product of these three ranks in the final.
-Whoever has the lowest product of ranks in the final wins the gold medal.
-
-This type of scoring system heavily rewards high finishes and relatively ignores poor finishes.
-For instance, if climber A finished 1st, 20th, and 20th and climber B finished 10th, 10th, and 10th, climber B would have a score of 1000 whereas climber A would have a much better score of 400, despie finishing last in 2 out of 3 of the events.
-
-\section{Other scoring methods}
-
-\section{Rank Product}
-https://www.sciencedirect.com/science/article/pii/S0014579304009354
+where $R^S_i$, $R^B_i$, and $R^L_i$
 

### 2021-07-09 data/youth_olympics_2018/scraping.R

@@ -1,0 +1,67 @@
+library(tidyverse)
+library(pdftools)
+txt <- pdf_text("https://www.dropbox.com/s/qoeoe4jwndoltpu/climbing_youth_olympics_2018.pdf?dl=1")
+# each index is a page
+
+# women qualification
+# page 15
+women_qual <- txt[15] %>% 
+  read_lines() %>% 
+  tibble(info = .) %>% 
+  mutate(info = str_squish(info),
+         event = "wq") %>%
+  filter(str_detect(info, "^[0-9]{1,2}\\s")) %>% 
+  mutate(info = str_replace(info, " Zaneta", "-Zaneta"),
+         info = str_replace(info, " Lara", "-Lara"),
+         info = str_remove(info, " Q")) %>% 
+  separate(info, into = c("rank", "bib", "last", "first", "from", "speed", "bould", "lead", "total"), sep = " ") %>% 
+  mutate(across(c(rank, speed:total), as.numeric))
+
+# write_csv(women_qual, "women_qual.csv")
+
+
+# women final
+# page 19
+women_final <- txt[19] %>% 
+  read_lines() %>% 
+  tibble(info = .) %>% 
+  mutate(info = str_squish(info),
+         event = "wf") %>% 
+  filter(str_detect(info, "^[1-6]{1}\\s")) %>% 
+  separate(info, into = c("rank", "bib", "last", "first", "from", "speed", "bould", "lead", "total"), sep = " ") %>% 
+  mutate(across(c(rank, speed:total), as.numeric))
+
+# write_csv(women_final, "women_final.csv")
+
+
+# men qualification
+# page 25
+men_qual <- txt[25] %>% 
+  read_lines() %>% 
+  tibble(info = .) %>% 
+  mutate(info = str_squish(info),
+         event = "wq") %>%
+  filter(str_detect(info, "^[0-9]{1,2}\\s")) %>% 
+  mutate(info = str_replace(info, "Chong Kiat ", "Chong-Kiat-"),
+         info = str_replace(info, "Lukas Joris ", "Lukas-Joris-"),
+         info = str_replace(info, "LINACISORO MOLINA Mikel ", "LINACISORO-MOLINA Mikel-"),
+         info = str_replace(info, "David ", "David-"),
+         info = str_remove(info, " Q")) %>% 
+  separate(info, into = c("rank", "bib", "last", "first", "from", "speed", "bould", "lead", "total"), sep = " ") %>% 
[…]

### 2021-07-09 data/youth_olympics_2018/scraping.R
Update scraping script

@@ -1,6 +1,6 @@
 library(tidyverse)
 library(pdftools)
-txt <- pdf_text("https://www.dropbox.com/s/qoeoe4jwndoltpu/climbing_youth_olympics_2018.pdf?dl=1")
+txt <- pdf_text("climbing_youth_olympics_2018.pdf")
 # each index is a page
 
 # women qualification

### 2021-07-09 data/youth_olympics_2018/scraping.R
Update scraping script

@@ -1,6 +1,6 @@
 library(tidyverse)
 library(pdftools)
-txt <- pdf_text("climbing_youth_olympics_2018.pdf")
+txt <- pdf_text("data/youth_olympics_2018/climbing_youth_olympics_2018.pdf")
 # each index is a page
 
 # women qualification

### 2021-07-09 data/youth_olympics_2018/scraping.R
Update scraping script

@@ -40,7 +40,7 @@ men_qual <- txt[25] %>%
   read_lines() %>% 
   tibble(info = .) %>% 
   mutate(info = str_squish(info),
-         event = "wq") %>%
+         event = "mq") %>%
   filter(str_detect(info, "^[0-9]{1,2}\\s")) %>% 
   mutate(info = str_replace(info, "Chong Kiat ", "Chong-Kiat-"),
          info = str_replace(info, "Lukas Joris ", "Lukas-Joris-"),

### 2021-07-10 R/displace_medalist.R
simulated probabilities for medalist displacement

@@ -1,0 +1,62 @@
+# Hannah Butler
+# 07-09-2021
+# Olympic Sport Climbing Scoring
+
+## Libraries
+library(tidyverse)
+## set constants
+
+## define functions
+assign_ranks <- function(c = 8) {
+  # randomly assign ranks in 3 events & compute total
+  # ranking in an event is considered to be independent from ranking in another event
+  output <- data.frame(climber = sample(LETTERS, c, replace = FALSE)
+                       , speed = sample(1:c, c, replace = FALSE)
+                       , bould = sample(1:c, c, replace = FALSE)
+                       , lead = sample(1:c, c, replace = FALSE)
+                       ) %>%
+    mutate(total = speed*bould*lead) %>%
+    arrange(total)
+  
+  return(output)
+}
+
+drop_climber <- function(results, r = 4) {
+  # drop a competitor and recompute rankings and totals
+  ud_results <- results[-r, ] %>%
+    mutate(speed = rank(speed)
+           , bould = rank(bould)
+           , lead = rank(lead)
+           , total = speed*bould*lead
+           ) %>%
+    arrange(total)
+  return(ud_results)
+}
+
+## begin code
+
+set.seed(80085)
+first <- vector(mode = "numeric", length = 1000)
+second <- vector(mode = "numeric", length = 1000)
+third <- vector(mode = "numeric", length = 1000)
+
+displace_climber <- function(nsim = 1000, drop_rank = 4, m_place = 1) {
+  # simulate 1000 rankings
+  rank_perms <- lapply(1:nsim, function(x) assign_ranks())
+  # drop ith competitor and recalculate ranks
+  dropout_ranks <- lapply(rank_perms, function(x) drop_climber(x, r = drop_rank))
+  
+  # compare mth place & calculate % not matched
+  pct_displ <- sapply(1:nsim, 
+                      function(x) {
+                        rank_perms[[x]]$climber[m_place] != dropout_ranks[[x]]$climber[m_place]
+                        }
+                      ) %>%
+    sum()/nsim
+  
+  return(pct_displ)
+}
+
+first <- replicate(100, displace_climber(m_place = 1))
+second <- replicate(100, displace_climber(m_place = 2))
+third <- replicate(100, displace_climber(m_place = 3))

### 2021-07-10 R/displace_medalist.R
added formatting + plotting

@@ -1,12 +1,14 @@
 # Hannah Butler
 # 07-09-2021
 # Olympic Sport Climbing Scoring
-
-## Libraries
+##### Description #####
+# estimate the probability of a medalist being displaced if another competitor drops out
+##### Libraries #####
 library(tidyverse)
+library(ggplot2)
 ## set constants
-
-## define functions
+load("R/medal_displacement_probabilities.rdata")
+##### Define Functions #####
 assign_ranks <- function(c = 8) {
   # randomly assign ranks in 3 events & compute total
   # ranking in an event is considered to be independent from ranking in another event
@@ -33,12 +35,8 @@ drop_climber <- function(results, r = 4) {
   return(ud_results)
 }
 
-## begin code
-
+##### begin code #####
 set.seed(80085)
-first <- vector(mode = "numeric", length = 1000)
-second <- vector(mode = "numeric", length = 1000)
-third <- vector(mode = "numeric", length = 1000)
 
 displace_climber <- function(nsim = 1000, drop_rank = 4, m_place = 1) {
   # simulate 1000 rankings
@@ -57,6 +55,39 @@ displace_climber <- function(nsim = 1000, drop_rank = 4, m_place = 1) {
   return(pct_displ)
 }
 
-first <- replicate(100, displace_climber(m_place = 1))
-second <- replicate(100, displace_climber(m_place = 2))
-third <- replicate(100, displace_climber(m_place = 3))
+first <- lapply(4:8, function(x) replicate(100, displace_climber(drop_rank = x, m_place = 1)))
+second <- lapply(4:8, function(x) replicate(100, displace_climber(drop_rank = x, m_place = 2)))
+third <- lapply(4:8, function(x) replicate(100, displace_climber(drop_rank = x, m_place = 3)))
+
+# save(first, second, third, file = "R/prob_displace_medalists.rdata")
+####################################
+# format data
+gold_df <- lapply(1:5
+                  , function(x) {
+                    data.frame(medal = "gold", dropped_rank = 3+x, displ_prob = first[[x]])
+                    }
+                  ) %>%
+  do.call(rbind, .)
+  
+silv_df <- lapply(1:5
+                  , function(x) {
+                    data.frame(medal = "silver", dropped_rank = 3+x, displ_prob = second[[x]])
+                    }
+                  ) %>%
+  do.call(rbind, .)
+
+bron_df <- lapply(1:5
+                  , function(x) {
[…]

### 2021-07-10 R/displace_medalist.R
view densities

@@ -6,8 +6,17 @@
 ##### Libraries #####
 library(tidyverse)
 library(ggplot2)
-## set constants
+##### set constants #####
+
+##### VISUALIZE #####
+# just look at this part
 load("R/medal_displacement_probabilities.rdata")
+
+displacement_probs %>%
+  ggplot() +
+  geom_density(aes(x = displ_prob, fill =  medal), alpha = 0.3) +
+  facet_wrap(~dropped_rank)
+
 ##### Define Functions #####
 assign_ranks <- function(c = 8) {
   # randomly assign ranks in 3 events & compute total

### 2021-07-10 R/displace_medalist.R
added tie-breaker

@@ -3,6 +3,11 @@
 # Olympic Sport Climbing Scoring
 ##### Description #####
 # estimate the probability of a medalist being displaced if another competitor drops out
+
+##### Notes ######
+# write code to deal with ties 
+# tiebreaker: determined by number of events with highest rank (super rudimentary, update)
+# in case of continued tie it doesn't matter, given the random nature of the ranks
 ##### Libraries #####
 library(tidyverse)
 library(ggplot2)
@@ -27,7 +32,15 @@ assign_ranks <- function(c = 8) {
                        , lead = sample(1:c, c, replace = FALSE)
                        ) %>%
     mutate(total = speed*bould*lead) %>%
-    arrange(total)
+    arrange(total) %>% # some questionable tie-breaking code
+    mutate(medal_rank_speed = ifelse(total %in% total[1:3] , rank(speed[1:3]), 0)
+           , medal_rank_bould = ifelse(total %in% total[1:3] , rank(bould[1:3]), 0)
+           , medal_rank_lead = ifelse(total %in% total[1:3] , rank(lead[1:3]), 0)
+           , tiebreaker = medal_rank_speed + medal_rank_bould + medal_rank_lead
+           ) %>%
+    arrange(total
+            , tiebreaker
+            )
   
   return(output)
 }
@@ -40,12 +53,17 @@ drop_climber <- function(results, r = 4) {
            , lead = rank(lead)
            , total = speed*bould*lead
            ) %>%
-    arrange(total)
+    arrange(total)%>%
+    mutate(medal_rank_speed = ifelse(total %in% total[1:3], rank(speed[1:3]), 0)
+           , medal_rank_bould = ifelse(total %in% total[1:3], rank(bould[1:3]), 0)
+           , medal_rank_lead = ifelse(total %in% total[1:3], rank(lead[1:3]), 0)
+           , tiebreaker = medal_rank_speed + medal_rank_bould + medal_rank_lead
+           ) %>%
+    arrange(total
+            , tiebreaker
+            )
   return(ud_results)
 }
-
-##### begin code #####
-set.seed(80085)
 
 displace_climber <- function(nsim = 1000, drop_rank = 4, m_place = 1) {
   # simulate 1000 rankings
@@ -57,18 +75,19 @@ displace_climber <- function(nsim = 1000, drop_rank = 4, m_place = 1) {
   pct_displ <- sapply(1:nsim, 
                       function(x) {
[…]

### 2021-07-13 manuscript/manuscript.Rnw
Manuscript update

@@
 
-Cross country scoring: %https://link.springer.com/content/pdf/10.1007/s11127-007-9193-6.pdf
+Cross country scoring: %https://link.springer.com/content/pdf/10.1007/s11127-007-9193-6.pdf \cite{Hammond2007}
 
-https://link.springer.com/article/10.1007/s11127-017-0494-0
+https://link.springer.com/article/10.1007/s11127-017-0494-0 \cite{BoudreauEtAl2018}
 
-Decathlon scoring: %https://research.ou.nl/en/publications/decathlon-towards-a-balanced-and-sustainable-performance-assessme
+Decathlon scoring: %https://research.ou.nl/en/publications/decathlon-towards-a-balanced-and-sustainable-performance-assessme \cit{Westera2006}
 
@@
 
+\subsections{Simulations}
+\subsubsection{uniform ranks}
+ - Distributions of points for qualifying and finals given uniform rankings. 
+ - Conditional probabilities.  
+ \subsubsections{Leave one out analysis} 
+ - For n finalists, how often is the ranking preserved when a single climber is removed.  - Independence of Irrelevant groups.  (IIG)
+
+\subsection{Data Analysis} 
+\subsubsections{Leave one out analysis} 
+ - How would do the ranks changes when a single particpant is removed.
+(i.e.
+What would happed if onyl 5 people made the finals instead of 6 and performance was the exact same?)  (Independence of Irrelevant Alternatives.
+(IIA))
+ - Examples where you remove the fifth ranked climber from the competition and the medals changed.
+
+ \subsubsection{Correlations}
+ - So bouldering and lead as highly correlatd with each other. 
+ - As a result bouldering and lead are more highly correlated with overall standings.  
+ - Speed climbers are going to suffer under this format.  
+ - Speed climbers are getting screwed!
+
+ \subsubsection{Rank Product statistic}
+- What does Hannah have?
+
+ Is there manipulability???!?!?!?
+
+outline of paper: 
+Simulation stuff.  
+Distribution of scores under random ranking. 
+Probability of advancement condtional on winnin/losing first event given random ranks.
+
+How different is the rank vs the product.  
+Borda count
+Voting properties?
+
+\section{Proposed better scoring method}
+ - Rank sum
+ - Weighted ranks (sum or product)
+ -
+
 \section{Conclusion}
+- Speed climbers are getting screwed.  
+- There is a great dependence on irrelevent party.
 

### 2021-07-13 manuscript/manuscript.Rnw
Manuscript update

@@
 
-\subsections{Simulations}
+\subsection{Simulations}
 \subsubsection{uniform ranks}
+
  - Distributions of points for qualifying and finals given uniform rankings.
+
  - Conditional probabilities.
- \subsubsections{Leave one out analysis} 
- - For n finalists, how often is the ranking preserved when a single climber is removed.  - Independence of Irrelevant groups.  (IIG)
+
+ \subsubsection{Leave one out analysis}
+
+ - For n finalists, how often is the ranking preserved when a single climber is removed.
+
+ - Independence of Irrelevant groups.  (IIG)
 
 \subsection{Data Analysis} 
-\subsubsections{Leave one out analysis} 
+\subsubsection{Leave one out analysis}
+
  - How would do the ranks changes when a single particpant is removed.
@@
 (IIA))
+
  - Examples where you remove the fifth ranked climber from the competition and the medals changed.
@@
  \subsubsection{Correlations}
+
  - So bouldering and lead as highly correlatd with each other.
+
  - As a result bouldering and lead are more highly correlated with overall standings.
+
  - Speed climbers are going to suffer under this format.
+
  - Speed climbers are getting screwed!
@@
  \subsubsection{Rank Product statistic}
+
 - What does Hannah have?
@@
 
-outline of paper: 
-Simulation stuff.  
-Distribution of scores under random ranking. 
-Probability of advancement condtional on winnin/losing first event given random ranks.
-
-How different is the rank vs the product.  
-Borda count
-Voting properties?
+% outline of paper: 
+% Simulation stuff.  
+% Distribution of scores under random ranking. 
+% Probability of advancement condtional on winnin/losing first event given random ranks.  
+% 
+% How different is the rank vs the product.  
+% Borda count
+% Voting properties?
 
@@
  - Rank sum
+
  - Weighted ranks (sum or product)
- -
 
 \section{Conclusion}
+
 - Speed climbers are getting screwed.
+
 - There is a great dependence on irrelevent party.

### 2021-07-13 manuscript/manuscript.tex
Manuscript update

@@
 Speed climbing takes place on a standardized course and competitors try to reach the top of the course as fast as possible.
+In the Olympics, speed climbing is being contested in a head-to-head format with ranks determined by how far a competitor advances in the bracket.
 In bouldering, contestants have a fixed amount of time to complete as many courses as they can.
@@
 In lead climbing, each competitor only gets one attempt and when they fall their attempt is over.
+%(https://www.cnet.com/news/climbing-at-the-tokyo-olympics-everything-you-need-to-know/)
 
@@
 In the Olympic sport climbing, there are 20 competitors at the start (in both men's and women's).
-All 20 competitors compete in each of the three events, and their performances in each event are ranked.
-A competitors final score is then computed as the product of their ranks in the three events.
+All 20 competitors compete in each of the three events in the qualification round, and their performances in each event are ranked from 1 to 20.
+A competitors final score is then computed as the product of their ranks in the three events and the lower product is better.
 Specifically,
@@
 
-where $R^S_i$, $R^B_i$, and $R^L_i$
+where $R^S_i$, $R^B_i$, and $R^L_i$ are the ranks of the $i$-th competitor in speed, bouldering, and lead, respectively.
+
+The top 8 (I think it's 8) finishers in the qualification round advance to the finals where the 8 remaining climbers again compete in all three events, they are again ranked from 1 to 8, and their final score is the product of these three ranks in the final.
+Whoever has the lowest product of ranks in the final wins the gold medal.
+
+This type of scoring system heavily rewards high finishes and relatively ignores poor finishes.
+For instance, if climber A finished 1st, 20th, and 20th and climber B finished 10th, 10th, and 10th, climber B would have a score of 1000 whereas climber A would have a much better score of 400, despie finishing last in 2 out of 3 of the events.
+
+\section{Other scoring methods}
+
+\section{Rank Product}
+https://www.sciencedirect.com/science/article/pii/S0014579304009354
 
@@
 
+\subsection{Simulations}
+\subsubsection{uniform ranks}
+
+ - Distributions of points for qualifying and finals given uniform rankings.
+
+ - Conditional probabilities.
+
+ \subsubsection{Leave one out analysis}
+
+ - For n finalists, how often is the ranking preserved when a single climber is removed.
+
+ - Independence of Irrelevant groups.  (IIG)
+
+\subsection{Data Analysis} 
+\subsubsection{Leave one out analysis}
+
+ - How would do the ranks changes when a single particpant is removed.
+(i.e.
+What would happed if onyl 5 people made the finals instead of 6 and performance was the exact same?) (Independence of Irrelevant Alternatives.
+(IIA))
+
+ - Examples where you remove the fifth ranked climber from the competition and the medals changed.
+
+ \subsubsection{Correlations}
+
+ - So bouldering and lead as highly correlatd with each other.
+
+ - As a result bouldering and lead are more highly correlated with overall standings.
+
+ - Speed climbers are going to suffer under this format.
+
+ - Speed climbers are getting screwed!
+
+ \subsubsection{Rank Product statistic}
+
+- What does Hannah have?
+
+Is there manipulability???!?!?!?
+
+% outline of paper: 
+% Simulation stuff.  
+% Distribution of scores under random ranking. 
+% Probability of advancement condtional on winnin/losing first event given random ranks.  
+% 
+% How different is the rank vs the product.  
+% Borda count
+% Voting properties?
+
+\section{Proposed better scoring method}
+ - Rank sum
+
+ - Weighted ranks (sum or product)
+
 \section{Conclusion}
+
+- Speed climbers are getting screwed.
+
+- There is a great dependence on irrelevent party.
 

### 2021-07-14 R/correlation_events.R
Add more data

@@ -1,0 +1,34 @@
+wq <- read_csv("data/youth_olympics_2018/women_qual.csv")
+mq <- read_csv("data/youth_olympics_2018/men_qual.csv")
+
+wf <- read_csv("data/youth_olympics_2018/women_final.csv")
+mf <- read_csv("data/youth_olympics_2018/men_final.csv")
+
+library(GGally)
+
+wq %>% 
+  select(speed, bould, lead) %>% 
+  ggpairs(diag = "blank",
+          upper = list(continuous = wrap("cor", method = "kendall")),
+          axisLabels = "internal") +
+  ggtitle("Kendall's rank correlations - Women's Qualification")
+
+mq %>% 
+  select(speed, bould, lead) %>% 
+  ggpairs(upper = list(continuous = wrap("cor", method = "kendall")))
+
+wf %>% 
+  select(speed, bould, lead) %>% 
+  ggpairs(upper = list(continuous = wrap("cor", method = "kendall")))
+
+mf %>% 
+  select(speed, bould, lead) %>% 
+  ggpairs(upper = list(continuous = wrap("cor", method = "kendall")))
+
+
+wq %>% 
+  bind_rows(mq, wf, mf) %>% 
+  select(speed, bould, lead) %>% 
+  ggpairs(upper = list(continuous = wrap("cor", method = "kendall")))
+
+

### 2021-07-14 data/asiad_2018/scraping.R

@@ -1,0 +1,46 @@
+# Source:
+# http://www.ocagames.com/orb/files/4/104/AG2018_OfficialResultBook_Sport%20Climbing_v1.0.pdf
+
+library(tidyverse)
+library(pdftools)
+txt <- pdf_text("data/asiad_2018/climbing_asiad_2018.pdf")
+# each index is a page
+
+# 
+
+
+# quick look at correlations
+# men's qualification
+
+rank <- 1:23
+
+speed <- c(6, 10, 9, 17, 5, 3,
+           1, 2, 8, 4, 11, 13, 15, 12, 19,
+           7, 16, 14, 18, 20, 22, 21, 23)
+
+bould <- c(1, 2, 3, 6, 8, 13,
+           19, 14, 4, 9, 7, 12, 5, 10, 11,
+           16, 17, 18, 15, 20, 22, 22, 22)
+
+lead <- c(1, 4, 3, 2, 7, 8,
+          18, 14, 13, 16, 10, 5, 11, 9, 6,
+          15, 12, 17, 21, 22, 19, 20, 23)
+
+cor(data.frame(speed, bould, lead))
+cor(rank, speed, method = "kendall")
+cor(rank, bould, method = "kendall")
+cor(rank, lead, method = "kendall")
+
+# women's qualification
+
+rank <- 1:20
+speed <- c(8, 6, 11, 1, 5, 9, 12, 3, 2, 4,
+           10, 16, 7, 15, 14, 13, 17, 20, 18, 19)
+bould <- c(2, 1, 3.5, 11, 6, 7, 3.5, 9, 15, 14, 10,
+           5, 13, 8, 12, 16, 17, 18, 19.5, 19.5)
+lead <- c(1, 3, 2, 16, 8, 4, 6, 11, 12, 7, 5,
+          9, 10, 13, 14, 15, 18, 17, 19, 20)
+cor(data.frame(speed, bould, lead))
+cor(rank, speed, method = "kendall")
+cor(rank, bould, method = "kendall")
+cor(rank, lead, method = "kendall")

### 2021-07-14 data/youth_olympics_2018/scraping.R

@@ -1,3 +1,6 @@
+# Source:
+# https://library.olympic.org/Default/digitalCollection/DigitalCollectionAttachmentDownloadHandler.ashx?parentDocumentId=177522&documentId=177531
+
 library(tidyverse)
 library(pdftools)
 txt <- pdf_text("data/youth_olympics_2018/climbing_youth_olympics_2018.pdf")

### 2021-07-14 R/correlation_events.R
Correlations between final rank and events

@@ -7,6 +7,10 @@ mf <- read_csv("data/youth_olympics_2018/men_final.csv")
 library(GGally)
 
 wq %>% 
+  select(rank, speed, bould, lead) %>%
+  cor(method = "kendall")
+
+wq %>% 
   select(speed, bould, lead) %>% 
   ggpairs(diag = "blank",
           upper = list(continuous = wrap("cor", method = "kendall")),
@@ -14,21 +18,19 @@ wq %>%
   ggtitle("Kendall's rank correlations - Women's Qualification")
 
 mq %>% 
-  select(speed, bould, lead) %>% 
-  ggpairs(upper = list(continuous = wrap("cor", method = "kendall")))
+  select(rank, speed, bould, lead) %>%
+  cor(method = "kendall")
 
-wf %>% 
+mq %>% 
   select(speed, bould, lead) %>% 
-  ggpairs(upper = list(continuous = wrap("cor", method = "kendall")))
+  ggpairs(diag = "blank",
+          upper = list(continuous = wrap("cor", method = "kendall")),
+          axisLabels = "internal") +
+  ggtitle("Kendall's rank correlations - Women's Qualification")
 
-mf %>% 
-  select(speed, bould, lead) %>% 
-  ggpairs(upper = list(continuous = wrap("cor", method = "kendall")))
 
 
-wq %>% 
-  bind_rows(mq, wf, mf) %>% 
-  select(speed, bould, lead) %>% 
-  ggpairs(upper = list(continuous = wrap("cor", method = "kendall")))
+
+
 
 

### 2021-07-14 data/asiad_2018/scraping.R

@@ -26,10 +26,7 @@ lead <- c(1, 4, 3, 2, 7, 8,
           18, 14, 13, 16, 10, 5, 11, 9, 6,
           15, 12, 17, 21, 22, 19, 20, 23)
 
-cor(data.frame(speed, bould, lead))
-cor(rank, speed, method = "kendall")
-cor(rank, bould, method = "kendall")
-cor(rank, lead, method = "kendall")
+cor(cbind(rank, speed, bould, lead), method = "kendall")
 
 # women's qualification
 
@@ -40,7 +37,6 @@ bould <- c(2, 1, 3.5, 11, 6, 7, 3.5, 9, 15, 14, 10,
            5, 13, 8, 12, 16, 17, 18, 19.5, 19.5)
 lead <- c(1, 3, 2, 16, 8, 4, 6, 11, 12, 7, 5,
           9, 10, 13, 14, 15, 18, 17, 19, 20)
-cor(data.frame(speed, bould, lead))
-cor(rank, speed, method = "kendall")
-cor(rank, bould, method = "kendall")
-cor(rank, lead, method = "kendall")
+
+cor(cbind(rank, speed, bould, lead), method = "kendall")
+

### 2021-07-15 R/drop_rerank.R
Update code

@@ -92,3 +92,46 @@ rerank_df <- women_qual %>%
   mutate(rank = row_number()) %>% 
   ungroup()
 
+
+# Simulation
+
+# In what percentage of the (8!)^3 orderings would the winner change 
+# if one of the 8 competitors was dropped and there were only 7 instead of 8?
+
+library(tidyverse)
+library(combinat)
+
+nplay <- 4
+perms <- crossing(e1 = permn(nplay), 
+                  e2 = permn(nplay), 
+                  e3 = permn(nplay))
+
+
+get_perm <- function(x) {
+  unnest_perm <- perms[x, ] %>%
+    unnest(cols = c(e1, e2, e3)) %>% 
+    mutate(total = e1 * e2 * e3,
+           initial_rank = rank(total, ties.method = "random"))
+  
+  unnest_perm %>%
+    mutate(drop = 0) %>%
+    bind_rows(unnest_perm %>% slice(-1) %>% mutate(drop = 1)) %>%
+    bind_rows(unnest_perm %>% slice(-2) %>% mutate(drop = 2)) %>%
+    bind_rows(unnest_perm %>% slice(-3) %>% mutate(drop = 3)) %>%
+    bind_rows(unnest_perm %>% slice(-4) %>% mutate(drop = 4)) %>%
+    mutate(perm = x)
+}
+
+
+perms_df <- tibble()
+for (row in 1:nrow(perms)) {
+  perms_df <- perms_df %>% 
+    bind_rows(get_perm(row))
+}
+
+f <- perms_df %>% 
+  filter(drop != 0) %>% 
+  mutate(total = e1 * e2 * e3) %>% 
+  group_by(perm, drop) %>% 
+  mutate(new_rank = rank(total, ties.method = "random"),
+         initial_rank_ordered = rank(initial_rank))

### 2021-07-16 stat_template/APA/wileyNJD-APA.tex
Add journal template

\documentclass[APA,LATO1COL]{WileyNJD-v2}

\articletype{Article Type}%

\received{26 April 2016}
\revised{6 June 2016}
\accepted{6 June 2016}

\raggedbottom

\begin{document}

\title{This is the sample article title\protect\thanks{This is an example for title footnote.}}

\author[1]{Author One*}

\author[2,3]{Author Two}

\author[3]{Author Three}

\authormark{AUTHOR ONE \textsc{et al}}

\address[1]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}

\address[2]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}

\address[3]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}

\corres{*Corresponding author name, This is sample corresponding address. \email{authorone@gmail.com}}

\presentaddress{This is sample for present address text this is sample for present address text}

\abstract[Summary]{This is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text.}

\keywords{keyword1, keyword2, keyword3, keyword4}

\jnlcitation{\cname{%
\author{Williams K.}, 
\author{B. Hoskins}, 
\author{R. Lee}, 
\author{G. Masato}, and 
\author{T. Woollings}} (\cyear{2016}), 
\ctitle{A regime analysis of Atlantic winter jet variability applied to evaluate HadGEM3-GC2}, \cjournal{Q.J.R. Meteorol. Soc.}, \cvol{2017;00:1--6}.}

\maketitle

\footnotetext{\textbf{Abbreviations:} ANA, anti-nuclear antibodies; APC, antigen-presenting cells; IRF, interferon regulatory factor}

\section{Sample for first level head}\label{sec1}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit~\citet{Rothermel1997} Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

\begin{eqnarray}
s(nT_{s}) &= &s(t)\times \sum\limits_{n=0}^{N-1} \delta (t-nT_{s}) \xleftrightarrow{\mathrm{DFT}}  S \left(\frac{m}{NT_{s}}\right) \nonumber\\
&= &\frac{1}{N} \sum\limits_{n=0}^{N-1} \sum\limits_{k=-N/2}^{N/2-1} s_{k} e^{\mathrm{j}2\pi k\Delta fnT_{s}} e^{-j\frac{2\pi}{N}mn}
\end{eqnarray}

\section{Sample for another first level head}\label{sec2}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus viverra
fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa ac
quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo~\cite{Rothermel1998} Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Example for bibliography citations cite~\citep{Elbaum2002}, cites~\cite{Allen2011,Yoo2007}

Quisque ullamcorper placerat ipsum. Cras nibh~\cite{Yoo2007,Schulz2012} Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.

\begin{figure}[t]
\centerline{\includegraphics[width=342pt,height=9pc,draft]{empty}}
\caption{This is the sample figure caption.\label{fig1}}
\end{figure}

Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor.
Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.

\begin{figure*}
\centerline{\includegraphics[width=342pt,height=9pc,draft]{empty}}
\caption{This is the sample figure caption.\label{fig2}}
\end{figure*}

\subsection{Example for second level head}

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit, dictum
in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam vestibulum
fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi. Pellentesque
eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut mi mi, lacinia
sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui. Mauris nibh leo,
facilisis non, adipiscing quis, ultrices a, dui.

Morbi luctus, wisi viverra faucibus pretium, nibh est placerat odio, nec commodo wisi enim eget quam. Quisque
libero justo, consectetuer a, feugiat vitae, porttitor eu, libero. Suspendisse sed mauris vitae elit sollicitudin malesuada.

Maecenas ultricies eros sit amet ante. Ut venenatis velit. Maecenas sed mi eget dui varius euismod. Phasellus aliquet
volutpat odio. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Pellentesque sit
amet pede ac sem eleifend consectetuer. Nullam elementum, urna vel imperdiet sodales, elit ipsum pharetra ligula,
ac pretium ante justo a nulla. Curabitur tristique arcu eu metus. Vestibulum lectus. Proin mauris. Proin eu nunc eu
urna hendrerit faucibus. Aliquam auctor, pede consequat laoreet varius, eros tellus scelerisque quam, pellentesque
hendrerit ipsum dolor sed augue. Nulla nec lacus.

\begin{quote}
This is an example~\citep{Elbaum2002,Blanchard2015} for quote text. This is an example for quote text. This is an example for quote text. This is an example for quote text~\cite{Blanchard2015}. This is an example for quote text. This is an example for quote text. This is an example for quote text. This is an example for quote text. This is an example for quote text. This is an example for quote text~\cite{Blanchard2015}. This is an example for quote text. This is an example for quote text. This is an example for quote text. 
\end{quote}

\section{Sample for next first level head}\label{sec3}

\subsection{Example for another second level head}

Suspendisse vitae elit. Aliquam arcu neque, ornare in, ullamcorper quis, commodo eu, libero. Fusce sagittis erat at
erat tristique mollis. Maecenas sapien libero, molestie et, lobortis in, sodales eget, dui. Morbi ultrices rutrum lorem.
Nam elementum ullamcorper leo. Morbi dui. Aliquam sagittis. Nunc placerat. Pellentesque tristique sodales est.
Maecenas imperdiet lacinia velit. Cras non urna. Morbi eros pede, suscipit ac, varius vel, egestas non, eros. Praesent
malesuada, diam id pretium elementum, eros sem dictum tortor, vel consectetuer odio sem sed wisi.

Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
Aenean vel lectus. Nunc imperdiet justo nec dolor.

\subsection{Second level head text}

Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.

\subsubsection{Third level head text}

Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra tortor. Sed fringilla
justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue est scelerisque
sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas urna mi, suscipit
in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.

Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
Vivamus eu dolor.

Nulla in ipsum. Praesent eros nulla, congue vitae, euismod ut, commodo a, wisi. Pellentesque habitant morbi
tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nonummy magna non leo. Sed felis erat,
ullamcorper in, dictum non, ultricies ut, lectus. Proin vel arcu a odio lobortis euismod. Vestibulum ante ipsum primis
in faucibus orci luctus et ultrices posuere cubilia Curae; Proin ut est. Aliquam odio. Pellentesque massa turpis, cursus
eu, euismod nec, tempor congue, nulla. Duis viverra gravida mauris. Cras tincidunt. Curabitur eros ligula, varius ut,
pulvinar in, cursus faucibus, augue.

\begin{boxtext}
\section*{Example of Boxtext}%
This is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext.
\end{boxtext}

\paragraph{Fourth level head text}

Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
Aenean vel lectus. Nunc imperdiet justo nec dolor.

Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.

\subparagraph{Fifth level head text}

Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra
tortor. Sed fringilla justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue
est scelerisque sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas
urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.

Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
Vivamus eu dolor.

in faucibus orci luctus et ultrices posuere cubilia Curae; Proin ut est. Aliquam odio. Pellentesque massa turpis, cursus
eu, euismod nec, tempor congue, nulla. Duis viverra gravida mauris. Cras tincidunt. Curabitur eros ligula, varius ut,
pulvinar in, cursus faucibus, augue.

Curabitur tellus magna, porttitor a, commodo a, commodo in, tortor. Donec interdum. Praesent scelerisque. Mae-
cenas posuere sodales odio. Vivamus metus lacus, varius quis, imperdiet quis, rhoncus a, turpis. Etiam ligula arcu,
elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel,
nibh. Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit. Etiam congue neque id dolor.

Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
Integer porta neque vitae massa. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo. Nulla nisl. Vestibulum
sed nisl eu sapien cursus rutrum.

Nulla non mauris vitae wisi posuere convallis. Sed eu nulla nec eros scelerisque pharetra. Nullam varius. Etiam
dignissim elementum metus. Vestibulum faucibus, metus sit amet mattis rhoncus, sapien dui laoreet odio, nec ultricies
nibh augue a enim. Fusce in ligula. Quisque at magna et nulla commodo consequat. Proin accumsan imperdiet sem.
Nunc porta. Donec feugiat mi at justo. Phasellus facilisis ipsum quis ante. In ac elit eget ipsum pharetra faucibus.
Maecenas viverra nulla in massa.

Nulla in ipsum. Praesent eros nulla, congue vitae, euismod ut, commodo a, wisi. Pellentesque habitant morbi
tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nonummy magna non leo. Sed felis erat,
ullamcorper in, dictum non, ultricies ut, lectus. Proin vel arcu a odio lobortis euismod. Vestibulum ante ipsum primis

\begin{center}
\begin{table*}[t]%
\caption{This is sample table caption.\label{tab1}}
\centering
\begin{tabular*}{500pt}{@{\extracolsep\fill}lccD{.}{.}{3}c@{\extracolsep\fill}}
\toprule
&\multicolumn{2}{@{}c@{}}{\textbf{Spanned heading\tnote{1}}} & \multicolumn{2}{@{}c@{}}{\textbf{Spanned heading\tnote{2}}} \\\cmidrule{2-3}\cmidrule{4-5}
\textbf{col1 head} & \textbf{col2 head}  & \textbf{col3 head}  & \multicolumn{1}{@{}l@{}}{\textbf{col4 head}}  & \textbf{col5 head}   \\
\midrule
col1 text & col2 text  & col3 text  & 12.34  & col5 text\tnote{1}   \\
col1 text & col2 text  & col3 text  & 1.62  & col5 text\tnote{2}   \\
col1 text & col2 text  & col3 text  & 51.809  & col5 text   \\
\bottomrule
\end{tabular*}
\begin{tablenotes}%%[341pt]
\item Source: Example for table source text.
\item[1] Example for a first table footnote.
\item[2] Example for a second table footnote.
\end{tablenotes}
\end{table*}
\end{center}

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.

\begin{center}
\begin{table}[t]%
\centering
\caption{This is sample table caption.\label{tab2}}%
\begin{tabular*}{500pt}{@{\extracolsep\fill}lcccc@{\extracolsep\fill}}
\toprule
\textbf{col1 head} & \textbf{col2 head}  & \textbf{col3 head}  & \textbf{col4 head}  & \textbf{col5 head} \\
\midrule
col1 text & col2 text  & col3 text  & col4 text  & col5 text\tnote{$\dagger$}   \\
col1 text & col2 text  & col3 text  & col4 text  & col5 text   \\
col1 text & col2 text  & col3 text  & col4 text  & col5 text\tnote{$\ddagger$}   \\
\bottomrule
\end{tabular*}
\begin{tablenotes}
\item Source: Example for table source text.
\item[$\dagger$] Example for a first table footnote.
\item[$\ddagger$] Example for a second table footnote.
\end{tablenotes}
\end{table}
\end{center}

Below is the example~\cite{Rothermel1998,Yoo2007,Schulz2012} for bulleted list. Below is the example for bulleted list. Below is the example for bulleted list. Below is the example for bulleted list. Below is the example for bulleted list. Below is the example for bulleted list\footnote{This is an example for footnote.}:
\begin{itemize}
\item bulleted list entry sample bulleted list entry~\cite{Allen2011} sample list entry text. 
\item bulleted list entry sample bulleted list entry. bulleted list entry sample bulleted list entry. bulleted list entry sample bulleted list entry.
\item bulleted list entry sample bulleted list entry~\cite{Ballen2011} bulleted list entry sample bulleted list entry~\citet{Allen2011} sample list entry text.  bulleted list entry sample bulleted list entry.
\item sample list entry text. sample list entry text.  
\end{itemize}

Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.

Below is the sample for description list. Below is the example for description list. Below is the example for description list. Below is the example for description list. Below is the example for description list. Below is the example for description list:\\[12pt]

\noindent\textbf{Description sample:}

\begin{description}
\item[first entry] description text. description text. description text. description text. description text. description text. description text. 
\item[second long entry] description text. description text. description text. description text. description text. description text. description text. 
\item[third entry] description text. description text. description text. description text. description text. 
\item[fourth entry] description text. description text. 
\end{description}

\noindent\textbf{Numbered list items sample:}

\begin{enumerate}[1.]
\item First level numbered list entry. sample numbered list entry.

\item First numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry.

\begin{enumerate}[a.]
\item Second level alpabetical list entry. Second level alpabetical list entry. Second level alpabetical list entry~\citet{Allen2011} Second level alpabetical list entry.

\item Second level alpabetical list entry. Second level alpabetical list entry~\citet{Schulz2012,Allen2011,Ballen2011}.

\begin{enumerate}[ii.]
\item Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry.

\item Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry~\cite{Yoo2007}.
\end{enumerate}

\item Second level alpabetical list entry. Second level alpabetical list entry~\cite{Elbaum2002}.
\end{enumerate}

\item First level numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry.

\item Another first level numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry. 
\end{enumerate}

\noindent\textbf{un-numbered list items sample:}

\begin{enumerate}[]
\item Sample unnumberd list text..
\item Sample unnumberd list text.
\item sample unnumberd list text. 
\item Sample unnumberd list text.
\end{enumerate}

\section{Examples for enunciations}\label{sec4}

\begin{theorem}[Theorem subhead]\label{thm1}
Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. 
\end{theorem}

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.

\begin{proposition}
Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. 
\end{proposition}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

\begin{definition}[Definition sub head]
Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text.
\end{definition}

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor.
Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.

\begin{proof}
Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. 
\end{proof}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

\begin{proof}[Proof of Theorem~\ref{thm1}]
Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. 
\end{proof}

Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra tortor. Sed fringilla
justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue est scelerisque
sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas urna mi, suscipit
in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut Curabitur tellus magna, porttitor a, commodo a, commodo in, tortor. Donec interdum. Praesent scelerisque. Mae-
cenas posuere sodales odio. Vivamus metus lacus, varius quis, imperdiet quis, rhoncus a, turpis. Etiam ligula arcu,
elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel,

\begin{sidewaystable}%[h]
\caption{Sideways table caption. For decimal alignment refer column 4 to 9 in tabular* preamble.\label{tab3}}%
\begin{tabular*}{\textheight}{@{\extracolsep\fill}lccD{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}@{\extracolsep\fill}}%
\toprule
  & \textbf{col2 head} & \textbf{col3 head} & \multicolumn{1}{c}{\textbf{10}} &\multicolumn{1}{c}{\textbf{20}} &\multicolumn{1}{c}{\textbf{30}} &\multicolumn{1}{c}{\textbf{10}} &\multicolumn{1}{c}{\textbf{20}} &\multicolumn{1}{c}{\textbf{30}} \\
\midrule
  &col2 text &col3 text &0.7568&1.0530&1.2642&0.9919&1.3541&1.6108 \\
  & &col2 text &12.5701 &19.6603&25.6809&18.0689&28.4865&37.3011 \\
3 &col2 text  & col3 text &0.7426&1.0393&1.2507&0.9095&1.2524&1.4958 \\
  & &col3 text &12.8008&19.9620&26.0324&16.6347&26.0843&34.0765 \\
  & col2 text & col3 text &0.7285&1.0257&1.2374&0.8195&1.1407&1.3691\tnote{*} \\
  & & col3 text &13.0360&20.2690&26.3895&15.0812&23.4932&30.6060\tnote{\dagger} \\
\bottomrule
\end{tabular*}
\begin{tablenotes}%%[\textheight]
\item[*] First sideways table footnote. Sideways table footnote. Sideways table footnote. Sideways table footnote.
\item[$\dagger$] Second sideways table footnote. Sideways table footnote. Sideways table footnote. Sideways table footnote.
\end{tablenotes}
\end{sidewaystable}

\begin{sidewaysfigure}
\centerline{\includegraphics[width=542pt,height=9pc,draft]{empty}}
\caption{Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption.\label{fig3}}
\end{sidewaysfigure}

nibh. Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit. Etiam congue neque id dolor.

\begin{algorithm}
\caption{Pseudocode for our algorithm}\label{alg1}
\begin{algorithmic}
  \For each frame
  \For water particles $f_{i}$
  \State compute fluid flow~\cite{Rothermel1997}
  \State compute fluid--solid interaction~\cite{Allen2011}
  \State apply adhesion and surface tension~\cite{Ballen2011}
  \EndFor
   \For solid particles $s_{i}$
   \For neighboring water particles $f_{j}$
   \State compute virtual water film \\(see Section~\ref{sec3})
   \EndFor
   \EndFor
   \For solid particles $s_{i}$
   \For neighboring water particles $f_{j}$
   \State compute growth direction vector \\(see Section~\ref{sec2})
   \EndFor
   \EndFor
   \For solid particles $s_{i}$
   \For neighboring water particles $f_{j}$
   \State compute $F_{\theta}$ (see Section~\ref{sec1})
   \State compute $CE(s_{i},f_{j})$ \\(see Section~\ref{sec3})
   \If $CE(b_{i}, f_{j})$ $>$ glaze threshold
   \State $j$th water particle's phase $\Leftarrow$ ICE
   \EndIf
   \If $CE(c_{i}, f_{j})$ $>$ icicle threshold
   \State $j$th water particle's phase $\Leftarrow$ ICE
   \EndIf
   \EndFor
   \EndFor
  \EndFor
\end{algorithmic}
\end{algorithm}

Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
Integer porta neque vitae massa~\cite{Rothermel1997,Elbaum2002}. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo~\cite{Elbaum2002} Nulla nisl. Vestibulum
sed nisl eu sapien cursus rutrum.

Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit. Etiam congue neque id dolor.

Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
Integer porta neque vitae massa. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo. Nulla nisl. Vestibulum
sed nisl eu sapien cursus rutrum.

\begin{equation}\label{eq23}
 \|\tilde{X}(k)\|^2
 =\frac{\left\|\sum\limits_{i=1}^{p}\tilde{Y}_i(k)+\sum\limits_{j=1}^{q}\tilde{Z}_j(k) \right\|^2}{(p+q)^2}
 \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.
\end{equation}

Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
Aenean vel lectus. Nunc imperdiet justo nec dolor.

Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.

\begin{equation}\label{eq24}
 \|\tilde{X}(k)\|^2
 =\frac{\left\|\sum\limits_{i=1}^{p}\tilde{Y}_i(k)+\sum\limits_{j=1}^{q}\tilde{Z}_j(k) \right\|^2}{(p+q)^2}
 \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.
\end{equation}

Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra
tortor. Sed fringilla justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue
est scelerisque sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas
urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.

Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
Vivamus eu dolor.

\section{Conclusions}\label{sec5}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

%\backmatter

\section*{Acknowledgments}
This is acknowledgment text~\cite{Elbaum2002}. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here.

\subsection*{Author contributions}

This is an author contribution text. This is an author contribution text. This is an author contribution text. This is an author contribution text. This is an author contribution text.

\subsection*{Financial disclosure}

None reported.

\subsection*{Conflict of interest}

The authors declare no potential conflict of interests.

\section*{Supporting information}

The following supporting information is available as part of the online article:

\noindent
\textbf{Figure S1.}
{500{\uns}hPa geopotential anomalies for GC2C calculated against the ERA Interim reanalysis. The period is 1989--2008.}

\noindent
\textbf{Figure S2.}
{The SST anomalies for GC2C calculated against the observations (OIsst).}

\appendix

\section{Section title of first appendix\label{app1}}

Use \verb+\begin{verbatim}...\end{verbatim}+ for program codes without math. Use \verb+\begin{alltt}...\end{alltt}+ for program codes with math. Based on the text provided inside the optional argument of \verb+\begin{code}[Psecode|Listing|Box|Code|+\hfill\break \verb+Specification|Procedure|Sourcecode|Program]...+ \verb+\end{code}+ tag corresponding boxed like floats are generated. Also note that \verb+\begin{code}[Code|Listing]...+ \verb+\end{code}+ tag with either Code or Listing text as optional argument text are set with computer modern typewriter font.  All other code environments are set with normal text font. Refer below example:

\begin{lstlisting}[caption={Descriptive Caption Text},label=DescriptiveLabel]
for i:=maxint to 0 do
begin
{ do nothing }
end;
Write('Case insensitive ');
WritE('Pascal keywords.');
\end{lstlisting}

\subsection{Subsection title of first appendix\label{app1.1a}}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

\subsubsection{Subsection title of first appendix\label{app1.1.1a}}

\noindent\textbf{Unnumbered figure}

\begin{center}
\includegraphics[width=7pc,height=8pc,draft]{empty}
\end{center}

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.

Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

\section{Section title of second appendix\label{app2}}%

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.

Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

%== Figure 4 ==
%% Example for figure inside appendix
\begin{figure}[t]
\centerline{\includegraphics[height=10pc,width=78mm,draft]{empty}}
\caption{This is an example for appendix figure.\label{fig5}}
\end{figure}

\subsection{Subsection title of second appendix\label{app2.1a}}

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
fermentum vel, eleifend faucibus, vehicula eu, lacus.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.

\subsubsection{Subsection title of second appendix\label{app2.1.1a}}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

\begin{center}
\begin{table}[b]%
\centering
\caption{This is an example of Appendix table showing food requirements of army, navy and airforce.\label{tab4}}%
\begin{tabular*}{300pt}{@{\extracolsep\fill}lcc@{\extracolsep\fill}}%
\toprule
\textbf{col1 head} & \textbf{col2 head} & \textbf{col3 head} \\
\midrule
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text\\
\bottomrule
\end{tabular*}
\end{table}
\end{center}

Example for an equation inside appendix
\begin{equation}
\mathcal{L}\quad \mathbf{\mathcal{L}} = i \bar{\psi} \gamma^\mu D_\mu \psi - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi\label{eq25}
\end{equation}

\section{Example of another appendix section\label{app3}}%

This is sample for paragraph text this is sample for paragraph text  this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

\begin{equation}
\mathcal{L} = i \bar{\psi} \gamma^\mu D_\mu \psi
    - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi
\label{eq26}
\end{equation}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

\begin{center}
\begin{tabular*}{300pt}{@{\extracolsep\fill}lcc@{\extracolsep\fill}}%
\toprule
\textbf{col1 head} & \textbf{col2 head} & \textbf{col3 head} \\
\midrule
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
\bottomrule
\end{tabular*}
\end{center}

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, evehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.

\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
\bibliography{wileyNJD-APA}%

\section*{Author Biography}

\begin{biography}{\includegraphics[width=60pt,height=70pt,draft]{empty}}{\textbf{Author Name.} This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
\end{biography}

\end{document}

### 2021-07-16 manuscript/manuscript.Rnw
Minor changes

@@
 \section{Introduction}
-The 2020 Olympics (held in 2021) will feature the debut of five new sports (https://olympics.com/ioc/news/ioc-approves-five-new-sports-for-olympic-games-tokyo-2020): surfing, skateboarding, karate, baseball/softball (not really new, but absent since 2008) and sport climbing.
+The 2020 Summer Olympics (held in 2021) will feature the debut of five new sports (https://olympics.com/ioc/news/ioc-approves-five-new-sports-for-olympic-games-tokyo-2020): surfing, skateboarding, karate, baseball/softball (not really new, but absent since 2008) and sport climbing.
 Sport climbing is particularly interesting as the International Olympic Committee (IOC) awarded only two sets of medals for the sport, one for men's and one for women's.
@@
 Winners are determined based on who completes the most courses and ties are broken based on who had the fewest attempts.
-Ties are further broken by the competitor acheived the most ``zone holds", which are holds approximately half way through each course.
+Ties are further broken by the competitor achieved the most ``zone holds", which are holds approximately half way through each course.
 Finally, in lead climbing a competitor gets one point for each hold that they reach, so whoever reaches the highest point on the wall is the winner.
@@
 $$
-Score_i = R^S_i \times R^B_i \times R^L_i
+Score_i = R^S_i\times R^B_i\times R^L_i
 $$,
@@
 
-Cross country scoring: %https://link.springer.com/content/pdf/10.1007/s11127-007-9193-6.pdf \cite{Hammond2007}
+Cross country scoring: %https://link.springer.com/content/pdf/10.1007/s11127-007-9193-6.pdf\cite{Hammond2007}
 
-https://link.springer.com/article/10.1007/s11127-017-0494-0 \cite{BoudreauEtAl2018}
+https://link.springer.com/article/10.1007/s11127-017-0494-0\cite{BoudreauEtAl2018}
 
-Decathlon scoring: %https://research.ou.nl/en/publications/decathlon-towards-a-balanced-and-sustainable-performance-assessme \cit{Westera2006}
+Decathlon scoring: %https://research.ou.nl/en/publications/decathlon-towards-a-balanced-and-sustainable-performance-assessme\cit{Westera2006}
 
@@
 \subsection{Simulations}
-\subsubsection{uniform ranks}
+\subsubsection{Uniform ranks}
 
@@
 
- - So bouldering and lead as highly correlatd with each other.
+- So bouldering and lead as highly correlated with each other.
 
@@
 \section{Proposed better scoring method}
+
 - Rank sum

### 2021-07-16 R/drop_rerank.R

@@ -43,7 +43,6 @@ rerank_df %>%
        x = "")
 
 # animation just for fun
-
 library(gganimate)
 
 a <- rerank_df %>% 
@@ -93,6 +92,8 @@ rerank_df <- women_qual %>%
   ungroup()
 
 
+#____________________________________________________________________________
+
 # Simulation
 
 # In what percentage of the (8!)^3 orderings would the winner change 
@@ -101,7 +102,7 @@ rerank_df <- women_qual %>%
 library(tidyverse)
 library(combinat)
 
-nplay <- 4
+nplay <- 6
 perms <- crossing(e1 = permn(nplay), 
                   e2 = permn(nplay), 
                   e3 = permn(nplay))
@@ -111,7 +112,7 @@ get_perm <- function(x) {
   unnest_perm <- perms[x, ] %>%
     unnest(cols = c(e1, e2, e3)) %>% 
     mutate(total = e1 * e2 * e3,
-           initial_rank = rank(total, ties.method = "random"))
+           initial_rank = rank(total))
   
   unnest_perm %>%
     mutate(drop = 0) %>%
@@ -119,12 +120,15 @@ get_perm <- function(x) {
     bind_rows(unnest_perm %>% slice(-2) %>% mutate(drop = 2)) %>%
     bind_rows(unnest_perm %>% slice(-3) %>% mutate(drop = 3)) %>%
     bind_rows(unnest_perm %>% slice(-4) %>% mutate(drop = 4)) %>%
+    bind_rows(unnest_perm %>% slice(-5) %>% mutate(drop = 5)) %>%
+    bind_rows(unnest_perm %>% slice(-6) %>% mutate(drop = 6)) %>%
     mutate(perm = x)
 }
 
 
 perms_df <- tibble()
 for (row in 1:nrow(perms)) {
+  print(row)
   perms_df <- perms_df %>% 
     bind_rows(get_perm(row))
 }
@@ -135,3 +139,41 @@ f <- perms_df %>%
   group_by(perm, drop) %>% 
   mutate(new_rank = rank(total, ties.method = "random"),
          initial_rank_ordered = rank(initial_rank))
+
+
+# dealing with ties
+
+
+qual <- bind_rows(qual_sim) %>% 
+  mutate(total = e1 * e2 * e3) %>% 
[…]

### 2021-07-17 R/rank_product_md.Rmd

---
title: "Rank Product Tests"
author: "Hannah Butler"
date: "7/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

## Finals Data
# you'll probably have to change these paths to get this to work
load("rp_data.rdata")
```

# About the Rank Product 
The Rank-Product test is a nonparametric test developed with the use in replicated gene expression experiments in mind. The idea is that we have $k$ sets of rankings for $n$ genes and want to know if a gene is differentially expressed. Outside of biology, this test also has applications in statistical meta-analysis and feature selection.

The rank-product $RP$ statistic is simply computed as the geometric mean of the rankings $r$ across the $k$ replicates for item $i = 1, 2, \dots, n$:

$$ RP_i = \Big( \prod_{j=1}^{k} r_{ij} \Big)^{1/k} $$

# Application to Olympic Sport Climbing
We believe this nonparametric test can be usefully applied to scoring data for the combined sport climbing events. We consider each of the 3 events a "replicate" for each climber, giving us $k=3$ sets of rankings. We then compute the rank product statistic for each climber, and compare it to a distribution of corresponding simulated rank products generated by uniformly and independently drawing a permutation of ranks for each event. Our test is to challenge this assumption that a climber's ranking is randomly achieved with equal odds.

## Function to Generate Rank Permutations

```{r}
assign_ranks <- function(c = 8) {
  # randomly assign ranks in 3 events & compute total
  # ranking in an event is considered to be independent from ranking in another event
  output <- data.frame(climber = sample(LETTERS, c, replace = FALSE)
                       , speed = sample(1:c, c, replace = FALSE)
                       , bould = sample(1:c, c, replace = FALSE)
                       , lead = sample(1:c, c, replace = FALSE)
                       ) %>%
    mutate(total = speed*bould*lead) %>%
    arrange(total)

  return(output)
}
```

# Computing the Rank-Product For Finalists
We start with the lower $n$ case

In the Combined-Event World Cup Climbing competitions, there are $n = 6$ competitors. We compute the geometric mean of their rankings for the three events:

## Men

```{r, echo = FALSE}
knitr::kable(m_fin_rp)
```

## Women

```{r, echo = FALSE}
knitr::kable(f_fin_rp)
```

```{r}
set.seed(80085)
n_perm <- 1000
# Rank Product for Finals (n = 6, k = 3)
fin_perms <- lapply(1:n_perm, function(x) assign_ranks(6) %>% mutate(rp = total^(1/3))
                    )

# list of rank-products for each total rank
rp_sets <- lapply(1:6, function(r) sapply(fin_perms, function(x) x$rp[r])
                  )
```

## Men's Finals
```{r}
E_m <- sapply(1:6, function(x) mean(m_fin_rp$rp[x] >= rp_sets[[x]]) # false discovery rate: /x
              )
```

```{r, echo = FALSE}
data.frame(last = m_fin_rp$last
           , first = m_fin_rp$first
           , rank = m_fin_rp$rank
           , rp = m_fin_rp$rp
           , pval = E_m
           ) %>% 
  knitr::kable()
```

## Women's Finals
```{r}
E_f <- sapply(1:6, function(x) mean(f_fin_rp$rp[x] >= rp_sets[[x]]) # false discovery rate:/x
              )
```

```{r, echo = FALSE}
data.frame(last = f_fin_rp$last
           , first = f_fin_rp$first
           , rank = f_fin_rp$rank
           , rp = f_fin_rp$rp
           , pval = E_f
           ) %>% 
  knitr::kable()
```

# Computing the Rank-Product For Qualifiers

### 2021-07-17 R/rank_product_md.Rmd
RP test for youth qualifiers

@@
 # Computing the Rank-Product For Qualifiers
+```{r}
+n_perm <- 1000
+# Rank Product for Finals (n = 6, k = 3)
+fin_perms <- lapply(1:n_perm, function(x) assign_ranks(21) %>% mutate(rp = total^(1/3))
+                    )
+
+# list of rank-products for each total rank
+rp_sets <- lapply(1:21, function(r) sapply(fin_perms, function(x) x$rp[r])
+                  )
+```
+
+## Men's Qualifiers
+```{r}
+E_m <- sapply(1:21, function(x) mean(m_qual_rp$rp[x] >= rp_sets[[x]]) # false discovery rate: /x
+              )
+```
+
+```{r, echo = FALSE}
+data.frame(last = m_qual_rp$last
+           , first = m_qual_rp$first
+           , rank = m_qual_rp$rank
+           , rp = m_qual_rp$rp
+           , pval = E_m
+           ) %>% 
+  knitr::kable()
+```
+
+## Women's Qualifiers
+```{r}
+E_f <- sapply(1:21, function(x) mean(f_qual_rp$rp[x] >= rp_sets[[x]]) # false discovery rate: /x
+              )
+```
+
+```{r, echo = FALSE}
+data.frame(last = f_qual_rp$last
+           , first = f_qual_rp$first
+           , rank = f_qual_rp$rank
+           , rp = f_qual_rp$rp
+           , pval = E_f
+           ) %>% 
+  knitr::kable()
+```

### 2021-07-18 R/displace_medalist.R
simulation results of medalist displacement/reordering when removing 1 competitor

@@ -3,11 +3,12 @@
 # Olympic Sport Climbing Scoring
 ##### Description #####
 # estimate the probability of a medalist being displaced if another competitor drops out
-
+# estimate the probability of a change in rankings if a competitor is removed
 ##### Notes ######
 # write code to deal with ties 
 # tiebreaker: determined by number of events with highest rank (super rudimentary, update)
 # in case of continued tie it doesn't matter, given the random nature of the ranks
+# how often does someone who outperformed in 2 events have a lower rank?
 ##### Libraries #####
 library(tidyverse)
 library(ggplot2)
@@ -16,12 +17,17 @@ library(ggplot2)
 ##### VISUALIZE #####
 # just look at this part
 load("R/medal_displacement_probabilities.rdata")
+load("R/top3_displace_probs.rdata")
 
 displacement_probs %>%
   ggplot() +
   geom_density(aes(x = displ_prob, fill =  medal), alpha = 0.3) +
   facet_wrap(~dropped_rank)
 
+displace_probs %>%
+  ggplot() +
+  geom_density(aes(x = displ_prob), alpha = 0.3) +
+  facet_wrap(~dropped_rank)
 ##### Define Functions #####
 assign_ranks <- function(c = 8) {
   # randomly assign ranks in 3 events & compute total
@@ -33,10 +39,10 @@ assign_ranks <- function(c = 8) {
                        ) %>%
     mutate(total = speed*bould*lead) %>%
     arrange(total) %>% # some questionable tie-breaking code
-    mutate(medal_rank_speed = ifelse(total %in% total[1:3] , rank(speed[1:3]), 0)
-           , medal_rank_bould = ifelse(total %in% total[1:3] , rank(bould[1:3]), 0)
-           , medal_rank_lead = ifelse(total %in% total[1:3] , rank(lead[1:3]), 0)
-           , tiebreaker = medal_rank_speed + medal_rank_bould + medal_rank_lead
+    mutate(m_rank_speed = ifelse(total %in% total[1:3] , rank(speed[1:3]), 0)
+           , m_rank_bould = ifelse(total %in% total[1:3] , rank(bould[1:3]), 0)
+           , m_rank_lead = ifelse(total %in% total[1:3] , rank(lead[1:3]), 0)
+           , tiebreaker = m_rank_speed + m_rank_bould + m_rank_lead
            ) %>%
     arrange(total
             , tiebreaker
[…]

### 2021-07-20 manuscript/manuscript.tex
Minor changes
rank-product test intro

@@
 
+\section{Data}
+
+World Championship 2018, men's and women's
+
+Asian Games 2018
+
 \section{Results}
@@
 
-\subsubsection{Rank Product statistic}
+\subsubsection{Rank Product Statistic}
 
-- What does Hannah have?
+The Rank-Product test is a nonparametric test developed with the use in replicated gene expression experiments in mind.
+The idea is that there are $k$ sets of rankings for $n$ genes and there is interest in knowing if a gene is differentially expressed.
+(A gene is declared differentially expressed if an observed difference or change in read counts or expression levels between two experimental conditions is statistically significant [ https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4827276/ ]).
+For example, a gene may be differentially expressed if it is consistently highly ranked across replicates in a microarray experiment.
+\\
+\\
+Outside of biology, this test also has applications in statistical meta-analysis and feature selection.
+\\
+\\
+Because of the unique nature of scoring in Olympic sport climbing (and hence, combined climbing competitions), we believe this test can be applied to identify "differentially expressed" climbers.
+That is, we are interested in knowing if a climber's performance is unlikely under the assumption of uniform random chance.
+In this scenario, each event (speed, bouldering, lead) is considered a replicate so that $k=3$.
+Because interest is typically in the highly ranked items, we will only be considering finals data, so that $n=6$.
+(In the Olympic events, there are 8 finalist slots).
+\\
+\\
+If the rankings across replicates are balanced, then the rank-product statistic for climber $i$ can be computed as the geometric mean across rankings $r_{ij}$:
+
+$$ RP_i = \Big( \prod_{j=1}^k r_{ij} \Big)^{1/k} $$
+
+Simulating a large number of ranking permutations, we can estimate a null distribution of rank-products which the climbers' own rank-products can be compared to.
+
+\textbf{Results}
 

### 2021-07-20 R/remove_climber_sim.Rmd
write up for removing climber simulations

---
title: "Remove Climber"
author: "Hannah Butler"
date: "7/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)

# load data
load("medal_displacement_probabilities.rdata")
load("top3_displace_probs.rdata")
load("medal_displacement_probabilities.rdata")
```

## Intro
A scoring system that multiplies the ranks across events is interesting because it does not possess the property that it is invariant to reordering when a competitor is removed. Examples of this can be found in existing combined climbing event data, where when a climber with a total rank of $r > 1$, higher ranks change order.

We can test this in simulation to see how often we would expect there to be a reordering in competitors when a competitor of a lower rank is removed. In particular, we can estimate the probability that the set of finalists will change if a non-finalist is removed from the rankings, and we can estimate the probability of medalists being displaced if a non-medalist is removed.

## Methodology
As it is the relevant upcoming event, we will focus the parameters of the simulation on those of sport climbing 2020 Tokyo Olympics (taking place in August, 2021). Namely, the qualifying round will consist of 40 individuals (20 men, 20 women) and the finals will consist of the 12(16?) (6(8?) men, 6(8?) women) climbers with the highest ranking in the qualifying round. For our simulations, we will naively assume that the ranks in all three events are assigned randomly and independently. (see the section on rank-product statistic which tests this assumption).

## Results

```{r, echo = FALSE}
# probability that medalists are displaced (changed or reordered)

displace_probs %>%
  group_by(dropped_rank) %>%
  rename("Dropped Rank" = dropped_rank) %>%
  summarize("# Simulations" = n()
            , "Mean Probability" = mean(displ_prob)
            , "Standard Deviation" = sd(displ_prob)
            ) %>%
  knitr::kable()

```

On average, regardless of which rank competitor is removed, there is a significant chance that the medalists will permute or change. This can be refined to specific medalist placement

```{r, echo = FALSE}
displacement_probs %>%
  group_by(medal, dropped_rank) %>%
  rename("Medal" = medal
         , "Dropped Rank" = dropped_rank) %>%
  summarize("# Simulations" = n()
            , "Mean Probability" = mean(displ_prob)
            , "Standard_Deviation" = sd(displ_prob)
            ) %>%
  knitr::kable()
```

### How often would a finalist be replaced
### How often does someone who wasn't a medalist before become a medalist 
### How often do we see a flip (two consecutive climbers switch places), jump (a climber jumps up more than one rank), drop (a climber drops down more than one rank)

### 2021-07-20 manuscript/draft_quang/draft.Rmd
Draft, simulation and correlation sections

---
title: \textbf{Title}
author: Names
date: \today
output:
  pdf_document: 
    number_sections: yes
    df_print: kable
    fig_caption: yes
    keep_tex: yes
abstract: This manuscript... The main results are...
fontsize: 11pt
---

<!-- | Author 1^[Corresponding author: Name, Dept, Univ, Email, etc.] , Author 2^[Name, etc.] , and Author 3^[Name, etc.] -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE,
                      comment = "",
                      fig.width = 4,
                      fig.height = 3)
```

# Introduction

# Data and Methods

# Results

## Simulations: Uniform Ranks

We conducted a simulation study to examine the rankings and scoring for climbers in both qualification and final rounds. For each round, we performed 10000 simulations, and this was accomplished by randomly assigning the ranks of each event to every participant, with the assumption that the ranks are uniformly distributed. After the completion of the simulations, we calculated the final scores for every simulated round, as well as the final standings for the climbing athletes. This data would then enable us to answer questions about various topics, including the distributions of scores for qualifying and final rounds, and the probabilities of advancing to the finals or winning a medal, given certain conditions.

```{r}
library(tidyverse)
library(knitr)
library(xtable)
theme_set(theme_bw())
```

```{r}
set.seed(1)
nsim <- 10000
nplay <- 20
qual_sim <- list()
for (i in 1:nsim) {
  qual_sim[[i]] <-
    bind_cols(
      player = 1:nplay,
      e1 = sample(1:nplay, replace = FALSE),
      e2 = sample(1:nplay, replace = FALSE),
      e3 = sample(1:nplay, replace = FALSE)
    ) %>%
    mutate(sim = i)
}

qual <- bind_rows(qual_sim) %>% 
  mutate(qual_score = e1 * e2 * e3) %>% 
  group_by(sim) %>% 
  mutate(qual_rank = rank(qual_score, ties.method = "random")) %>%
  ungroup()
```

```{r, fig.cap = "Distribution of qualification scores"}
qual %>%
  ggplot(aes(qual_score)) +
  geom_histogram(bins = 20, color = "white") + 
  labs(x = "Qualification Score",
       y = "Frequency") +
  theme(axis.title = element_text(size = 9),
        axis.text = element_text(size = 8))
```

Conditional probability

```{r, eval = FALSE}
qual %>%
  filter(e1 == 1 & qual_rank <= 8) %>%
  nrow()
```

From our simulations, the probability of advancing given winning the first event is 0.9951, and the probability of advancing to the final round for an athlete, given that they were the winner of at least one event, is 0.9947.

Table 1 shows the average score of climbers that finished in the top 8 of the qualifying round and hence advanced to the final round.

```{r}
qual %>% 
  group_by(qual_rank) %>% 
  summarize(avg_adv_score = mean(qual_score)) %>% 
  kable(caption = "Average score for each qualifying rank")
```

## Correlations

We collected data on major climbing competitions from 2018 to 2020, including the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Asian Games. We are interested in looking at the relationships between the ranks of the individual events and the final standings, and we computed Kendall’s Tau (Kendall Rank Correlation Coefficient).

Table is a correlation matrix between the ranks for the 2018 Youth Olympics Women's Qualification.

It is evidently clear that there is a strong and positive correlation between the ranks of bouldering and lead climbing, and as a results, the standings of these two events are highly correlated with the final rankings. On the other hand, the relationship is not as strong for speed climbing. Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those that are specialized in the other two concentrations.

```{r}
wq <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2018_youth_olympics/women_qual.csv")
# mq <- read_csv("data/2018_youth_olympics/men_qual.csv")
# 
# wf <- read_csv("data/2018_youth_olympics/women_final.csv")
# mf <- read_csv("data/2018_youth_olympics/men_final.csv")
library(GGally)
```

```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Women's Qualification"}
wq %>% 
  select(speed, bould, lead) %>% 
  ggpairs(diag = "blank",
          upper = list(continuous = wrap("cor", method = "kendall")))
```

```{r}
wq %>% 
  select(rank, speed:lead) %>% 
  cor(method = "kendall") %>% 
  data.frame()
```

### 2021-07-20 manuscript/draft_quang/draft.tex
Draft, simulation and correlation sections

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdfauthor={Names},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{\textbf{Title}}
\author{Names}
\date{\today}

\begin{document}
\maketitle
\begin{abstract}
This manuscript\ldots{} The main results are\ldots{}
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{data-and-methods}{%
\section{Data and Methods}\label{data-and-methods}}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{simulations-uniform-ranks}{%
\subsection{Simulations: Uniform
Ranks}\label{simulations-uniform-ranks}}

We conducted a simulation study to examine the rankings and scoring for
climbers in both qualification and final rounds. For each round, we
performed 10000 simulations, and this was accomplished by randomly
assigning the ranks of each event to every participant, with the
assumption that the ranks are uniformly distributed. After the
completion of the simulations, we calculated the final scores for every
simulated round, as well as the final standings for the climbing
athletes. This data would then enable us to answer questions about
various topics, including the distributions of scores for qualifying and
final rounds, and the probabilities of advancing to the finals or
winning a medal, given certain conditions.

\begin{figure}
\centering
\includegraphics{draft_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{Distribution of qualification scores}
\end{figure}

Conditional probability

From our simulations, the probability of advancing given winning the
first event is 0.9951, and the probability of advancing to the final
round for an athlete, given that they were the winner of at least one
event, is 0.9947.

Table 1 shows the average score of climbers that finished in the top 8
of the qualifying round and hence advanced to the final round.

\begin{longtable}[]{@{}rr@{}}
\caption{Average score for each qualifying rank}\tabularnewline
\toprule
qual\_rank & avg\_adv\_score \\
\midrule
\endfirsthead
\toprule
qual\_rank & avg\_adv\_score \\
\midrule
\endhead
1 & 36.0187 \\
2 & 73.6111 \\
3 & 115.3954 \\
4 & 162.2263 \\
5 & 216.0041 \\
6 & 278.1649 \\
7 & 350.3272 \\
8 & 434.5932 \\
9 & 532.1383 \\
10 & 642.3298 \\
11 & 771.0376 \\
12 & 919.1239 \\
13 & 1091.2140 \\
14 & 1291.9224 \\
15 & 1536.2010 \\
16 & 1834.3433 \\
17 & 2211.7246 \\
18 & 2695.6190 \\
19 & 3399.5684 \\
20 & 4585.3123 \\
\bottomrule
\end{longtable}

\hypertarget{correlations}{%
\subsection{Correlations}\label{correlations}}

We collected data on major climbing competitions from 2018 to 2020,
including the 2020 Continental Championships of Europe, Africa, Oceania,
Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and
2018 Asian Games. We are interested in looking at the relationships
between the ranks of the individual events and the final standings, and
we computed Kendall's Tau (Kendall Rank Correlation Coefficient).

Table is a correlation matrix between the ranks for the 2018 Youth
Olympics Women's Qualification.

It is evidently clear that there is a strong and positive correlation
between the ranks of bouldering and lead climbing, and as a results, the
standings of these two events are highly correlated with the final
rankings. On the other hand, the relationship is not as strong for speed
climbing. Thus, speed climbers are facing a huge disadvantage in this
scoring system, compared to those that are specialized in the other two
concentrations.

\begin{figure}
\centering
\includegraphics{draft_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{Kendall's rank correlations - 2018 World Championship, Women's
Qualification}
\end{figure}

\begin{longtable}[]{@{}lrrrr@{}}
\toprule
& rank & speed & bould & lead \\
\midrule
\endhead
rank & 1.0000000 & 0.3619048 & 0.6666667 & 0.6000000 \\
speed & 0.3619048 & 1.0000000 & 0.1238095 & 0.0952381 \\
bould & 0.6666667 & 0.1238095 & 1.0000000 & 0.5333333 \\
lead & 0.6000000 & 0.0952381 & 0.5333333 & 1.0000000 \\
\bottomrule
\end{longtable}

\end{document}

### 2021-07-21 R/displace_medalist.R
drop qualifier -> change in finalists

@@ -1,34 +1,37 @@
 # Hannah Butler
 # 07-09-2021
 # Olympic Sport Climbing Scoring
-##### Description #####
+#################### DESCRIPTION #########################
 # estimate the probability of a medalist being displaced if another competitor drops out
 # estimate the probability of a change in rankings if a competitor is removed
-##### Notes ######
+#################### NOTES ###############################
 # write code to deal with ties 
-# tiebreaker: determined by number of events with highest rank (super rudimentary, update)
+# tiebreaker: determined by sum of medalist rankings (super rudimentary, update)
 # in case of continued tie it doesn't matter, given the random nature of the ranks
 # how often does someone who outperformed in 2 events have a lower rank?
-##### Libraries #####
+#################### LIBRARIES ###########################
 library(tidyverse)
 library(ggplot2)
-##### set constants #####
-
-##### VISUALIZE #####
+#################### SET CONSTANTS #######################
+#
+#################### VISUALIZE ###########################
 # just look at this part
 load("R/medal_displacement_probabilities.rdata")
 load("R/top3_displace_probs.rdata")
 
+# densities for probability of displacement, by medal & dropped rank
 displacement_probs %>%
   ggplot() +
   geom_density(aes(x = displ_prob, fill =  medal), alpha = 0.3) +
   facet_wrap(~dropped_rank)
 
+# densities for probability of displacement, by dropped rank
 displace_probs %>%
   ggplot() +
-  geom_density(aes(x = displ_prob), alpha = 0.3) +
-  facet_wrap(~dropped_rank)
-##### Define Functions #####
+  geom_density(aes(x = displ_prob, group = dropped_rank, fill = as.factor(dropped_rank)), alpha = 0.3)
+
+#################### DEFINE FUNCTIONS ####################
+#>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>> Assign Ranks 
 assign_ranks <- function(c = 8) {
   # randomly assign ranks in 3 events & compute total
   # ranking in an event is considered to be independent from ranking in another event
@@ -50,7 +53,7 @@ assign_ranks <- function(c = 8) {
   
   return(output)
 }
-
[…]

### 2021-07-26 manuscript/Rmd/draft.Rmd
Create Rmd manuscript folder

---
title: Climbing

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Author 1
  thanks: 
  affiliation:

- name: Author 2
  affiliation:

- name: Author 3
  affiliation:

keywords:
- sport climbing
- scoring system

abstract: |
  This manuscript... The main results are... 
# bibliography: bib_example.bib
output: rticles::asa_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE,
                      comment = "",
                      fig.width = 5,
                      fig.height = 4,
                      results = "asis")
options(xtable.comment = FALSE,
        xtable.caption.placement = "top", 
        xtable.include.rownames = FALSE)
```

# Introduction

Copy from Rnw + Edit

# Data and Methods

# Results

## Simulations: Uniform Ranks

```{r}
library(tidyverse)
library(knitr)
library(xtable)
theme_set(theme_bw())
```

```{r}
set.seed(1)
nsim <- 10000
nplay <- 20
qual_sim <- list()
for (i in 1:nsim) {
  qual_sim[[i]] <-
    bind_cols(
      player = 1:nplay,
      e1 = sample(1:nplay, replace = FALSE),
      e2 = sample(1:nplay, replace = FALSE),
      e3 = sample(1:nplay, replace = FALSE)
    ) %>%
    mutate(sim = i)
}

qual <- bind_rows(qual_sim) %>% 
  mutate(qual_score = e1 * e2 * e3) %>% 
  group_by(sim) %>% 
  mutate(qual_rank = rank(qual_score, ties.method = "random")) %>%
  ungroup()
```

```{r}
set.seed(1)
nsim <- 10000
nplay <- 8
final_sim <- list()
for (i in 1:nsim) {
  final_sim[[i]] <-
    bind_cols(
      player = 1:nplay,
      e1 = sample(1:nplay, replace = FALSE),
      e2 = sample(1:nplay, replace = FALSE),
      e3 = sample(1:nplay, replace = FALSE)
    ) %>%
    mutate(sim = i)
}

final <- bind_rows(final_sim)

final <- final %>% 
  mutate(final_score = e1 * e2 * e3) %>% 
  group_by(sim) %>% 
  mutate(final_rank = rank(final_score, ties.method = "random")) %>%
  ungroup()
```

We conducted a simulation study to examine the rankings and scoring for climbers in both qualification and final rounds. For each round, we performed 10000 simulations, and this was accomplished by randomly assigning the ranks of each event to every participant, with the assumption that the ranks are uniformly distributed. After the completion of the simulations, we calculated the final scores for every simulated round, as well as the final standings for the climbing athletes. This data would then enable us to answer questions about various topics, including the distributions of scores for qualifying and final rounds, and the probabilities of advancing to the finals or winning a medal, given certain conditions.

```{r, fig.cap = "Distribution of qualification scores"}
qual %>%
  ggplot(aes(qual_score)) +
  geom_histogram(bins = 20, color = "white") + 
  labs(x = "Qualification Score",
       y = "Frequency")
  # theme(axis.title = element_text(size = 9),
  #       axis.text = element_text(size = 8))
```

For the qualification round, a climber is almost guaranteed to make the final round if they win the first event (with a 99.51% chance of advancing) or if they win at least one of the three climbing concentrations (99.48%). On the other hand, finishing last in the first event or in any event would certainly hurt an athlete's chance of finishing in the top 8, as the probabilities of a climber advancing given they finish last in the first and in any event are 0.1830 and 0.1885, respectively. In addition, the average score for qualification positions 1 to 8 are displayed in Table 1. We notice that on average, the minimum score that one should aim for in order to move on to the final round is 435 (for 8th rank).

```{r}
qual %>% 
  group_by(qual_rank) %>% 
  summarize(avg_adv_score = mean(qual_score)) %>% 
  filter(qual_rank <= 8) %>% 
  xtable(caption = "Average score for each qualifying rank") %>% 
  print()
```

Regarding the finals, a climber is very likely to earn a medal (finish in the top 3) if they win the first event (83.03% chance) or any event (85.01%). In order to obtain a climbing medal, the average score for getting gold, silver, and bronze are 9.6748, 20.4143, and 33.2648, respectively (Table 2). A notable trend we observe for both qualification and final rounds is as the rank increases, the distribution of the scores becomes more spread out. (Figure..., facet)

```{r}
final %>% 
  group_by(final_rank) %>% 
  summarize(avg_score = mean(final_score)) %>% 
  xtable(caption = "Average score of medalists") %>% 
  print()
```

```{r, fig.cap = "Boxplots of scoring distribution for every qualification rank"}
qual %>% 
  mutate(qual_rank = factor(qual_rank)) %>% 
  ggplot(aes(x = qual_score, y = qual_rank)) +
  geom_boxplot()
```

## Drawbacks of the scoring system

### Why product of ranks instead of sum

We performed the same tasks as we did for products

Probabilities of advancing, winning medals are all lower

Most notably for qualifications,

Obvious that the average score between the ranks are closer to each other

The amount of variability doesn't seem to be different as the rank increases

### Speed climbing vs lead and bouldering

We collected data on major climbing competitions from 2018 to 2020, including the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Asian Games. We are interested in looking at the relationships between the ranks of the individual events and the final standings, and we computed Kendall’s Tau (Kendall Rank Correlation Coefficient).

Figure 3 is a scatterplot and correlation matrix between the ranks for the 2018 Youth Olympics Women's Qualification.

It is evidently clear that there is a strong and positive correlation between the ranks of bouldering and lead climbing, and as a results, the standings of these two events are highly correlated with the final rankings. On the other hand, the relationship is not as strong for speed climbing. Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those that are specialized in the other two concentrations.

```{r}
wq <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2018_youth_olympics/women_qual.csv")
library(GGally)
```

```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Women's Qualification"}
wq %>% 
  select(rank, speed, bould, lead) %>% 
  ggpairs(diag = "blank",
          upper = list(continuous = wrap("cor", method = "kendall")))
```

```{r}
# wq %>% 
#   select(rank, speed:lead) %>% 
#   cor(method = "kendall") %>% 
#   data.frame()
```

### Drop and re-rank

### 2021-07-26 manuscript/Rmd/draft.tex
Create Rmd manuscript folder

--- 2021-07-26 manuscript/Rmd/draft.Rmd
@@
----
-title: Climbing
+% !TeX program = pdfLaTeX
+\documentclass[12pt]{article}
+\usepackage{amsmath}
+\usepackage{graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage{natbib}
+\usepackage{textcomp}
+\usepackage[hyphens]{url} % not crucial - just used below for the URL
+\usepackage{hyperref}
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
-# to produce blinded version set to 1
-blinded: 0
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
 
-authors: 
-- name: Author 1
-  thanks: 
-  affiliation:
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
 
-- name: Author 2
-  affiliation:
+%% load any required packages here
 
-- name: Author 3
-  affiliation:
+% Pandoc citation processing
 
-keywords:
-- sport climbing
-- scoring system
+\begin{document}
 
-abstract: |
-  This manuscript... The main results are... 
-# bibliography: bib_example.bib
-output: rticles::asa_article
----
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, 
-                      message = FALSE, 
-                      warning = FALSE,
-                      comment = "",
-                      fig.width = 5,
-                      fig.height = 4,
-                      results = "asis")
-options(xtable.comment = FALSE,
-        xtable.caption.placement = "top", 
-        xtable.include.rownames = FALSE)
-```
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-# Introduction
+\if0\blind
+{
+  \title{\bf Climbing}
+
+  \author{
+        Author 1 \\
+    \\
+     and \\     Author 2 \\
+    \\
+     and \\     Author 3 \\
+    \\
+      }
+  \maketitle
+} \fi
+
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Climbing}
+  \end{center}
+  \medskip
+} \fi
+
+\bigskip
+\begin{abstract}
+This manuscript\ldots{} The main results are\ldots{}
+\end{abstract}
+
+\noindent%
+{\it Keywords:} sport climbing, scoring system
+\vfill
+
+\newpage
+\spacingset{1.45} % DON'T change the spacing!
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-# Data and Methods
+\hypertarget{data-and-methods}{%
+\section{Data and Methods}\label{data-and-methods}}
 
-# Results
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
-## Simulations: Uniform Ranks
+\hypertarget{simulations-uniform-ranks}{%
+\subsection{Simulations: Uniform
+Ranks}\label{simulations-uniform-ranks}}
 
-```{r}
-library(tidyverse)
-library(knitr)
-library(xtable)
-theme_set(theme_bw())
-```
+We conducted a simulation study to examine the rankings and scoring for
+climbers in both qualification and final rounds. For each round, we
+performed 10000 simulations, and this was accomplished by randomly
+assigning the ranks of each event to every participant, with the
+assumption that the ranks are uniformly distributed. After the
+completion of the simulations, we calculated the final scores for every
+simulated round, as well as the final standings for the climbing
+athletes. This data would then enable us to answer questions about
+various topics, including the distributions of scores for qualifying and
+final rounds, and the probabilities of advancing to the finals or
+winning a medal, given certain conditions.
 
-```{r}
-set.seed(1)
-nsim <- 10000
-nplay <- 20
-qual_sim <- list()
-for (i in 1:nsim) {
-  qual_sim[[i]] <-
-    bind_cols(
-      player = 1:nplay,
-      e1 = sample(1:nplay, replace = FALSE),
-      e2 = sample(1:nplay, replace = FALSE),
-      e3 = sample(1:nplay, replace = FALSE)
-    ) %>%
-    mutate(sim = i)
-}
+\begin{figure}
+\centering
+\includegraphics{draft_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{Distribution of qualification scores}
+\end{figure}
 
-qual <- bind_rows(qual_sim) %>% 
-  mutate(qual_score = e1 * e2 * e3) %>% 
-  group_by(sim) %>% 
-  mutate(qual_rank = rank(qual_score, ties.method = "random")) %>%
-  ungroup()
-```
+For the qualification round, a climber is almost guaranteed to make the
+final round if they win the first event (with a 99.51\% chance of
+advancing) or if they win at least one of the three climbing
+concentrations (99.48\%). On the other hand, finishing last in the first
+event or in any event would certainly hurt an athlete's chance of
+finishing in the top 8, as the probabilities of a climber advancing
+given they finish last in the first and in any event are 0.1830 and
+0.1885, respectively. In addition, the average score for qualification
+positions 1 to 8 are displayed in Table 1. We notice that on average,
+the minimum score that one should aim for in order to move on to the
+final round is 435 (for 8th rank).
 
-```{r}
-set.seed(1)
-nsim <- 10000
-nplay <- 8
-final_sim <- list()
-for (i in 1:nsim) {
-  final_sim[[i]] <-
-    bind_cols(
-      player = 1:nplay,
-      e1 = sample(1:nplay, replace = FALSE),
-      e2 = sample(1:nplay, replace = FALSE),
-      e3 = sample(1:nplay, replace = FALSE)
-    ) %>%
-    mutate(sim = i)
-}
+\begin{table}[ht]
+\centering
+\caption{Average score for each qualifying rank} 
+\begin{tabular}{rr}
+  \hline
+qual\_rank & avg\_adv\_score \\ 
+  \hline
+  1 & 36.02 \\ 
+    2 & 73.61 \\ 
+    3 & 115.40 \\ 
+    4 & 162.23 \\ 
+    5 & 216.00 \\ 
+    6 & 278.16 \\ 
+    7 & 350.33 \\ 
+    8 & 434.59 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
-final <- bind_rows(final_sim)
+Regarding the finals, a climber is very likely to earn a medal (finish
+in the top 3) if they win the first event (83.03\% chance) or any event
+(85.01\%). In order to obtain a climbing medal, the average score for
+getting gold, silver, and bronze are 9.6748, 20.4143, and 33.2648,
+respectively (Table 2). A notable trend we observe for both
+qualification and final rounds is as the rank increases, the
+distribution of the scores becomes more spread out. (Figure\ldots,
+facet)
 
-final <- final %>% 
-  mutate(final_score = e1 * e2 * e3) %>% 
-  group_by(sim) %>% 
-  mutate(final_rank = rank(final_score, ties.method = "random")) %>%
-  ungroup()
-```
+\begin{table}[ht]
+\centering
+\caption{Average score of medalists} 
+\begin{tabular}{rr}
+  \hline
+final\_rank & avg\_score \\ 
+  \hline
+  1 & 9.67 \\ 
+    2 & 20.41 \\ 
+    3 & 33.26 \\ 
+    4 & 50.59 \\ 
+    5 & 74.76 \\ 
+    6 & 110.05 \\ 
+    7 & 164.43 \\ 
+    8 & 265.78 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
-We conducted a simulation study to examine the rankings and scoring for climbers in both qualification and final rounds.
-For each round, we performed 10000 simulations, and this was accomplished by randomly assigning the ranks of each event to every participant, with the assumption that the ranks are uniformly distributed.
-After the completion of the simulations, we calculated the final scores for every simulated round, as well as the final standings for the climbing athletes.
-This data would then enable us to answer questions about various topics, including the distributions of scores for qualifying and final rounds, and the probabilities of advancing to the finals or winning a medal, given certain conditions.
+\begin{figure}
+\centering
+\includegraphics{draft_files/figure-latex/unnamed-chunk-7-1.pdf}
+\caption{Boxplots of scoring distribution for every qualification rank}
+\end{figure}
 
-```{r, fig.cap = "Distribution of qualification scores"}
-qual %>%
-  ggplot(aes(qual_score)) +
-  geom_histogram(bins = 20, color = "white") + 
-  labs(x = "Qualification Score",
-       y = "Frequency")
-  # theme(axis.title = element_text(size = 9),
-  #       axis.text = element_text(size = 8))
-```
+\hypertarget{drawbacks-of-the-scoring-system}{%
+\subsection{Drawbacks of the scoring
+system}\label{drawbacks-of-the-scoring-system}}
 
-For the qualification round, a climber is almost guaranteed to make the final round if they win the first event (with a 99.51% chance of advancing) or if they win at least one of the three climbing concentrations (99.48%).
-On the other hand, finishing last in the first event or in any event would certainly hurt an athlete's chance of finishing in the top 8, as the probabilities of a climber advancing given they finish last in the first and in any event are 0.1830 and 0.1885, respectively.
-In addition, the average score for qualification positions 1 to 8 are displayed in Table 1.
-We notice that on average, the minimum score that one should aim for in order to move on to the final round is 435 (for 8th rank).
-
-```{r}
-qual %>% 
-  group_by(qual_rank) %>% 
-  summarize(avg_adv_score = mean(qual_score)) %>% 
-  filter(qual_rank <= 8) %>% 
-  xtable(caption = "Average score for each qualifying rank") %>% 
-  print()
-```
-
-Regarding the finals, a climber is very likely to earn a medal (finish in the top 3) if they win the first event (83.03% chance) or any event (85.01%).
-In order to obtain a climbing medal, the average score for getting gold, silver, and bronze are 9.6748, 20.4143, and 33.2648, respectively (Table 2).
-A notable trend we observe for both qualification and final rounds is as the rank increases, the distribution of the scores becomes more spread out.
-(Figure..., facet)
-
-```{r}
-final %>% 
-  group_by(final_rank) %>% 
-  summarize(avg_score = mean(final_score)) %>% 
-  xtable(caption = "Average score of medalists") %>% 
-  print()
-```
-
-```{r, fig.cap = "Boxplots of scoring distribution for every qualification rank"}
-qual %>% 
-  mutate(qual_rank = factor(qual_rank)) %>% 
-  ggplot(aes(x = qual_score, y = qual_rank)) +
-  geom_boxplot()
-```
-
-## Drawbacks of the scoring system
-
-### Why product of ranks instead of sum
+\hypertarget{why-product-of-ranks-instead-of-sum}{%
+\subsubsection{Why product of ranks instead of
+sum}\label{why-product-of-ranks-instead-of-sum}}
 
@@
 
-Obvious that the average score between the ranks are closer to each other
+Obvious that the average score between the ranks are closer to each
+other
 
-The amount of variability doesn't seem to be different as the rank increases
+The amount of variability doesn't seem to be different as the rank
+increases
 
-### Speed climbing vs lead and bouldering
+\hypertarget{speed-climbing-vs-lead-and-bouldering}{%
+\subsubsection{Speed climbing vs lead and
+bouldering}\label{speed-climbing-vs-lead-and-bouldering}}
 
-We collected data on major climbing competitions from 2018 to 2020, including the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Asian Games.
-We are interested in looking at the relationships between the ranks of the individual events and the final standings, and we computed Kendall’s Tau (Kendall Rank Correlation Coefficient).
+We collected data on major climbing competitions from 2018 to 2020,
+including the 2020 Continental Championships of Europe, Africa, Oceania,
+Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and
+2018 Asian Games. We are interested in looking at the relationships
+between the ranks of the individual events and the final standings, and
+we computed Kendall's Tau (Kendall Rank Correlation Coefficient).
 
-Figure 3 is a scatterplot and correlation matrix between the ranks for the 2018 Youth Olympics Women's Qualification.
+Figure 3 is a scatterplot and correlation matrix between the ranks for
+the 2018 Youth Olympics Women's Qualification.
 
-It is evidently clear that there is a strong and positive correlation between the ranks of bouldering and lead climbing, and as a results, the standings of these two events are highly correlated with the final rankings.
-On the other hand, the relationship is not as strong for speed climbing.
-Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those that are specialized in the other two concentrations.
+It is evidently clear that there is a strong and positive correlation
+between the ranks of bouldering and lead climbing, and as a results, the
+standings of these two events are highly correlated with the final
+rankings. On the other hand, the relationship is not as strong for speed
+climbing. Thus, speed climbers are facing a huge disadvantage in this
+scoring system, compared to those that are specialized in the other two
+concentrations.
 
-```{r}
-wq <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2018_youth_olympics/women_qual.csv")
-library(GGally)
-```
+\begin{figure}
+\centering
+\includegraphics{draft_files/figure-latex/unnamed-chunk-9-1.pdf}
+\caption{Kendall's rank correlations - 2018 World Championship, Women's
+Qualification}
+\end{figure}
 
-```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Women's Qualification"}
-wq %>% 
-  select(rank, speed, bould, lead) %>% 
-  ggpairs(diag = "blank",
-          upper = list(continuous = wrap("cor", method = "kendall")))
-```
+\hypertarget{drop-and-re-rank}{%
+\subsubsection{Drop and re-rank}\label{drop-and-re-rank}}
 
-```{r}
-# wq %>% 
-#   select(rank, speed:lead) %>% 
-#   cor(method = "kendall") %>% 
-#   data.frame()
-```
+\bibliographystyle{agsm}
+\bibliography{}
 
-### Drop and re-rank
+\end{document}

### 2021-07-26 manuscript/Rmd/draft.Rmd
Create Rmd manuscript folder

@@
                       comment = "",
-                      fig.width = 5,
-                      fig.height = 4,
+                      # fig.width = 5,
+                      # fig.height = 4,
                       results = "asis")
@@
 
-Copy from Rnw + Edit
+(Copy from Rnw + Edit)
+
+## Sport Climbing
+
+## Other Scoring Methods
 
 # Data and Methods
+
+We collected data on major climbing competitions from 2018 to 2020, including the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Asian Games.
 
@@
 
-We collected data on major climbing competitions from 2018 to 2020, including the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Asian Games.
-We are interested in looking at the relationships between the ranks of the individual events and the final standings, and we computed Kendall’s Tau (Kendall Rank Correlation Coefficient).
+Using past competitions data, w are interested in looking at the relationships between the ranks of the individual events and the final standings, and we computed Kendall’s Tau (Kendall Rank Correlation Coefficient).
 
@@
 ### Drop and re-rank
+
+A single climber excluded changes things drastically, espcially order of medalists.
+
+The cases where someone behind you drops out and your ranking changes.
+
+```{r}
+women_final <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2018_youth_olympics/women_final.csv")
+rerank <- list()
+for (i in 1:nrow(women_final)){
+  rerank[[i]] <- women_final[-i,] %>%
+    mutate(rank_drop = i)
+}
+
+rerank_df <- women_final %>% 
+  mutate(rank_drop = 0) %>% 
+  bind_rows(rerank) %>% 
+  group_by(rank_drop) %>% 
+  mutate(speed = rank(speed),
+         bould = rank(bould),
+         lead = rank(lead),
+         total = speed * bould * lead) %>% 
+  arrange(total, .by_group = TRUE) %>% 
+  ungroup() %>% 
+  group_by(rank_drop, total) %>% 
+  mutate(speed_tb = ifelse(speed < lag(speed), 1, 0),
+         bould_tb = ifelse(bould < lag(bould), 1, 0),
+         lead_tb = ifelse(lead < lag(lead), 1, 0),
+         tb = speed_tb + bould_tb + lead_tb,
+         tb = ifelse(is.na(tb), 1, tb)) %>% 
+  ungroup() %>% 
+  group_by(rank_drop) %>% 
+  arrange(total, -tb, .by_group = TRUE) %>% 
+  mutate(rank = row_number())
+```
+
+```{r, fig.cap = "Original rankings and rankings after each rank is dropped"}
+rerank_df %>% 
+  mutate(last = fct_reorder(last, -rank)) %>% 
+  # filter(rank_drop %in% c(0, 5)) %>% 
+  ggplot(aes(x = last, y = total)) +
+  geom_col() +
+  geom_text(aes(label = rank), hjust = -0.1, size = 3) +
+  coord_flip() +
+  facet_wrap(~ rank_drop) +
+  labs(y = "Score",
+       x = "Climber")
+```

### 2021-07-26 manuscript/Rmd/draft.tex
Create Rmd manuscript folder

@@
 
-Copy from Rnw + Edit
+(Copy from Rnw + Edit)
+
+\hypertarget{sport-climbing}{%
+\subsection{Sport Climbing}\label{sport-climbing}}
+
+\hypertarget{other-scoring-methods}{%
+\subsection{Other Scoring Methods}\label{other-scoring-methods}}
 
@@
 \section{Data and Methods}\label{data-and-methods}}
+
+We collected data on major climbing competitions from 2018 to 2020,
+including the 2020 Continental Championships of Europe, Africa, Oceania,
+Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and
+2018 Asian Games.
 
@@
 
-We collected data on major climbing competitions from 2018 to 2020,
-including the 2020 Continental Championships of Europe, Africa, Oceania,
-Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and
-2018 Asian Games. We are interested in looking at the relationships
-between the ranks of the individual events and the final standings, and
-we computed Kendall's Tau (Kendall Rank Correlation Coefficient).
+Using past competitions data, w are interested in looking at the
+relationships between the ranks of the individual events and the final
+standings, and we computed Kendall's Tau (Kendall Rank Correlation
+Coefficient).
 
@@
 
+A single climber excluded changes things drastically, espcially order of
+medalists.
+
+The cases where someone behind you drops out and your ranking changes.
+
+\begin{figure}
+\centering
+\includegraphics{draft_files/figure-latex/unnamed-chunk-12-1.pdf}
+\caption{Original rankings and rankings after each rank is dropped}
+\end{figure}
+
 \bibliographystyle{agsm}

### 2021-07-27 manuscript/Rmd/draft.tex
Update Rmd

--- 2021-07-26 manuscript/Rmd/draft.tex
@@
 
-Copy from Rnw + Edit
+(Copy from Rnw + Edit)
+
+\hypertarget{sport-climbing}{%
+\subsection{Sport Climbing}\label{sport-climbing}}
+
+\hypertarget{other-scoring-systems}{%
+\subsection{Other Scoring Systems}\label{other-scoring-systems}}
 
@@
 
-\hypertarget{why-product-of-ranks-instead-of-sum}{%
-\subsubsection{Why product of ranks instead of
-sum}\label{why-product-of-ranks-instead-of-sum}}
+\hypertarget{sum-of-ranks-instead-of-product}{%
+\subsubsection{Sum of ranks instead of
+product}\label{sum-of-ranks-instead-of-product}}
 
@@
 
-We collected data on major climbing competitions from 2018 to 2020,
-including the 2020 Continental Championships of Europe, Africa, Oceania,
-Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and
-2018 Asian Games. We are interested in looking at the relationships
-between the ranks of the individual events and the final standings, and
-we computed Kendall's Tau (Kendall Rank Correlation Coefficient).
-
-Figure 3 is a scatterplot and correlation matrix between the ranks for
-the 2018 Youth Olympics Women's Qualification.
-
-It is evidently clear that there is a strong and positive correlation
-between the ranks of bouldering and lead climbing, and as a results, the
-standings of these two events are highly correlated with the final
-rankings. On the other hand, the relationship is not as strong for speed
+For our analysis on the relationship between the rankings of the events
+and the final result, we used data from the 2018 Youth Olympics Women's
+Qualification. Figure 3 is a scatterplot and correlation matrix between
+the ranks of the individual events and the final standings, with
+Kendall's Tau (Kendall Rank Correlation Coefficient) as our measure of
+ordinal association between the quantities. It is evidently clear that
+there is a strong and positive correlation between the ranks of
+bouldering and lead climbing, and as a results, the standings of these
+two events are highly correlated with the final rankings. On the other
+hand, the correlation with the final rank is not as strong for speed
 climbing. Thus, speed climbers are facing a huge disadvantage in this
@@
 concentrations.
+
+This trend also holds for most of the past competitions.
 
@@
 
+A single climber excluded changes things drastically, especially order
+of medalists.
+
+The cases where someone behind you drops out and your ranking changes.
+
+Example from 2018 youth, women's final dropping ranks 3 and 5 change the
+medalist order
+
+\begin{figure}
+\centering
+\includegraphics{draft_files/figure-latex/unnamed-chunk-11-1.pdf}
+\caption{Original rankings and rankings after each rank is dropped}
+\end{figure}
+
 \bibliographystyle{agsm}

### 2021-07-27 manuscript/Rmd/draft.Rmd

@@
 
-## Other Scoring Methods
+## Other Scoring Systems
 
@@
 
-### Why product of ranks instead of sum
+### Sum of ranks instead of product
 
@@
 
-Using past competitions data, w are interested in looking at the relationships between the ranks of the individual events and the final standings, and we computed Kendall’s Tau (Kendall Rank Correlation Coefficient).
+For our analysis on the relationship between the rankings of the events and the final result, we used data from the 2018 Youth Olympics Women's Qualification.
+Figure 3 is a scatterplot and correlation matrix between the ranks of the individual events and the final standings, with Kendall’s Tau (Kendall Rank Correlation Coefficient) as our measure of ordinal association between the quantities.
+It is evidently clear that there is a strong and positive correlation between the ranks of bouldering and lead climbing, and as a results, the standings of these two events are highly correlated with the final rankings.
+On the other hand, the correlation with the final rank is not as strong for speed climbing.
+Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those that are specialized in the other two concentrations.
 
-Figure 3 is a scatterplot and correlation matrix between the ranks for the 2018 Youth Olympics Women's Qualification.
-
-It is evidently clear that there is a strong and positive correlation between the ranks of bouldering and lead climbing, and as a results, the standings of these two events are highly correlated with the final rankings.
-On the other hand, the relationship is not as strong for speed climbing.
-Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those that are specialized in the other two concentrations.
+This trend also holds for most of the past competitions.
 
@@
 
-```{r}
-# wq %>% 
-#   select(rank, speed:lead) %>% 
-#   cor(method = "kendall") %>% 
-#   data.frame()
-```
-
 ### Drop and re-rank
 
-A single climber excluded changes things drastically, espcially order of medalists.
+A single climber excluded changes things drastically, especially order of medalists.
 
 The cases where someone behind you drops out and your ranking changes.
+
+Example from 2018 youth, women's final
+dropping ranks 3 and 5 change the medalist order
 

### 2021-07-27 R/rank_product_md.Rmd
fixed coding error

--- 2021-07-17 R/rank_product_md.Rmd
@@
 ## Finals Data
-# you'll probably have to change these paths to get this to work
 load("rp_data.rdata")
@@
 The idea is that we have $k$ sets of rankings for $n$ genes and want to know if a gene is differentially expressed.
+(A gene is declared differentially expressed if an observed difference or change in read counts or expression levels between two experimental conditions is statistically significant https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4827276/ ).
 Outside of biology, this test also has applications in statistical meta-analysis and feature selection.
 
-The rank-product $RP$ statistic is simply computed as the geometric mean of the rankings $r$ across the $k$ replicates for item $i = 1, 2, \dots, n$:
+If the rankings are balanced (that is, if each item has an equal number of replicates), then the rank-product $RP$ statistic for item $i$ can simply be computed as the geometric mean of the rankings $r_{ij}$ across the $k$ replicates for item $i = 1, \dots, n$ and replicate $j = 1,\dots, k$ :
 
 $$ RP_i = \Big( \prod_{j=1}^{k} r_{ij} \Big)^{1/k} $$
+
+https://febs.onlinelibrary.wiley.com/doi/epdf/10.1016/j.febslet.2004.07.055
+
+## Determining Significance Level
 
@@
 
-## Function to Generate Rank Permutations
-
+## Function to Generate Rank-Product Permutations
 ```{r}
 assign_ranks <- function(c = 8) {
-  # randomly assign ranks in 3 events & compute total
+  # randomly assign ranks in 3 events & compute product across events
   # ranking in an event is considered to be independent from ranking in another event
-  output <- data.frame(climber = sample(LETTERS, c, replace = FALSE)
-                       , speed = sample(1:c, c, replace = FALSE)
-                       , bould = sample(1:c, c, replace = FALSE)
-                       , lead = sample(1:c, c, replace = FALSE)
+  output <- cbind(sample(1:c, c, replace = FALSE)
+                  , sample(1:c, c, replace = FALSE)
+                  , sample(1:c, c, replace = FALSE)
                   ) %>%
-    mutate(total = speed*bould*lead) %>%
-    arrange(total)
+    apply(1, prod)
 
@@
 
-# Computing the Rank-Product For Finalists
-We start with the lower $n$ case
+```{r}
+compare_rp <- function(rp_real, rp_perm) {
+  # count how many of the permuted rank-products are smaller than the real rank-product
+  sum(rp_real >= rp_perm) %>%
+  return()
+}
 
-In the Combined-Event World Cup Climbing competitions, there are $n = 6$ competitors. We compute the geometric mean of their rankings for the three events:
-
-## Men
-
-```{r, echo = FALSE}
-knitr::kable(m_fin_rp)
-```
-
-## Women
-
-```{r, echo = FALSE}
-knitr::kable(f_fin_rp)
-```
-
-```{r}
-set.seed(80085)
-n_perm <- 1000
-# Rank Product for Finals (n = 6, k = 3)
-fin_perms <- lapply(1:n_perm, function(x) assign_ranks(6) %>% mutate(rp = total^(1/3))
+nsim <- 1000
+nqual <- nrow(m_qual_rp)
+rp_sims <- replicate(nsim, assign_ranks(nqual))
+df <- data.frame(first = rep(NA, nqual)
+                 , last = rep(NA, nqual)
+                 , speed = rep(NA, nqual)
+                 , bould = rep(NA, nqual)
+                 , lead = rep(NA, nqual)
+                 , total = rep(NA, nqual)
+                 , count_above = rep(NA, nqual)
+                 , E = rep(NA, nqual)
+                 , q = rep(NA, nqual)
                  )
 
-# list of rank-products for each total rank
-rp_sets <- lapply(1:6, function(r) sapply(fin_perms, function(x) x$rp[r])
+for (i in 1:nqual) {
+  t <- m_qual_rp$total[i]                                            # experimental/real score
+  #cat(m_qual_rp$first[i], m_qual_rp$last[i], " ", t, "\n")
+  s <-  compare_rp(t, rp_sims) %>% sum()                             # count which simulated rp values <= real rp
+  e <- s/nsim                                                        # avg expected value
+  q <- e/i                                                           # q value (false discovery rate)
+  #print(q)
+  df[i,] <- c(m_qual_rp$first[i]
+              , m_qual_rp$last[i]
+              , m_qual_rp$speed[i]
+              , m_qual_rp$bould[i]
+              , m_qual_rp$lead[i]
+              , m_qual_rp$total[i]
+              , s
+              , e
+              , q
               )
+}
+
+df %>% knitr::kable()
 ```
-
-## Men's Finals
-```{r}
-E_m <- sapply(1:6, function(x) mean(m_fin_rp$rp[x] >= rp_sets[[x]]) # false discovery rate: /x
-              )
-```
-
-```{r, echo = FALSE}
-data.frame(last = m_fin_rp$last
-           , first = m_fin_rp$first
-           , rank = m_fin_rp$rank
-           , rp = m_fin_rp$rp
-           , pval = E_m
-           ) %>% 
-  knitr::kable()
-```
-
-## Women's Finals
-```{r}
-E_f <- sapply(1:6, function(x) mean(f_fin_rp$rp[x] >= rp_sets[[x]]) # false discovery rate:/x
-              )
-```
-
-```{r, echo = FALSE}
-data.frame(last = f_fin_rp$last
-           , first = f_fin_rp$first
-           , rank = f_fin_rp$rank
-           , rp = f_fin_rp$rp
-           , pval = E_f
-           ) %>% 
-  knitr::kable()
-```
-
-# Computing the Rank-Product For Qualifiers

### 2021-07-27 R/rank_product.R

@@ -10,10 +10,10 @@
 ##### Libraries #####
 library(tidyverse)
 ##### Set Constants #####
-m_qual_path <- "data/youth_olympics_2018/men_qual.csv"
-m_fin_path <- "data/youth_olympics_2018/men_final.csv"
-f_qual_path <- "data/youth_olympics_2018/women_qual.csv"
-f_fin_path <- "data/youth_olympics_2018/women_final.csv"
+m_qual_path <- "data/2018_youth_olympics/men_qual.csv"
+m_fin_path <- "data/2018_youth_olympics/men_final.csv"
+f_qual_path <- "data/2018_youth_olympics/women_qual.csv"
+f_fin_path <- "data/2018_youth_olympics/women_final.csv"
 ##### Define Functions #####
 geo_mean <- function(x, l = length(x)) {
   geomean <- prod(x)^(1/l)

### 2021-07-27 R/rank_product_md.Rmd
added 2019 world championships

@@
 load("rp_data.rdata")
+load(file = "world_2019_data.rdata")
+
+set.seed(80085)
 ```
@@
 }
-```
 
-```{r}
 compare_rp <- function(rp_real, rp_perm) {
@@
 }
+```
 
+```{r, echo = FALSE}
 nsim <- 1000
@@
 rp_sims <- replicate(nsim, assign_ranks(nqual))
+```
+
+# 2018 Youth Olympics
+
+## Men's qualifiers
+```{r, echo = FALSE}
 df <- data.frame(first = rep(NA, nqual)
@@
 ```
+
+## Women's qualifiers
+```{r, echo =  FALSE}
+df_f <- data.frame(first = rep(NA, nqual)
+                 , last = rep(NA, nqual)
+                 , speed = rep(NA, nqual)
+                 , bould = rep(NA, nqual)
+                 , lead = rep(NA, nqual)
+                 , total = rep(NA, nqual)
+                 , count_above = rep(NA, nqual)
+                 , E = rep(NA, nqual)
+                 , q = rep(NA, nqual)
+                 )
+
+for (i in 1:nqual) {
+  t <- f_qual_rp$total[i]                                            # experimental/real score
+  #cat(m_qual_rp$first[i], m_qual_rp$last[i], " ", t, "\n")
+  s <-  compare_rp(t, rp_sims) %>% sum()                             # count which simulated rp values <= real rp
+  e <- s/nsim                                                        # avg expected value
+  q <- e/i                                                           # q value (false discovery rate)
+  #print(q)
+  df_f[i,] <- c(f_qual_rp$first[i]
+              , f_qual_rp$last[i]
+              , f_qual_rp$speed[i]
+              , f_qual_rp$bould[i]
+              , f_qual_rp$lead[i]
+              , f_qual_rp$total[i]
+              , s
+              , e
+              , q
+              )
+}
+
+df_f %>% knitr::kable()
+```
+
+# 2019 World Championships
+
+```{r}
+nqual <- 20
+rp_sims <- replicate(nsim, assign_ranks(nqual))
+```
+
+## Men's Championships
+```{r, echo = FALSE}
+df <- data.frame(name = rep(NA, nqual)
+                 , speed = rep(NA, nqual)
+                 , bould = rep(NA, nqual)
+                 , lead = rep(NA, nqual)
+                 , total = rep(NA, nqual)
+                 , count_above = rep(NA, nqual)
+                 , E = rep(NA, nqual)
+                 , q = rep(NA, nqual)
+                 )
+
+for (i in 1:nqual) {
+  t <- world_2019_m_qual$total[i]                                            # experimental/real score
+  #cat(m_qual_rp$first[i], m_qual_rp$last[i], " ", t, "\n")
+  s <-  compare_rp(t, rp_sims) %>% sum()                             # count which simulated rp values <= real rp
+  e <- s/nsim                                                        # avg expected value
+  q <- e/i                                                           # q value (false discovery rate)
+  #print(q)
+  df[i,] <- c(world_2019_m_qual$name[i]
+              , world_2019_m_qual$speed[i]
+              , world_2019_m_qual$bould[i]
+              , world_2019_m_qual$lead[i]
+              , world_2019_m_qual$total[i]
+              , s
+              , e
+              , q
+              )
+}
+
+df %>% knitr::kable()
+```
+
+## Women's Championships
+```{r, echo = FALSE}
+df <- data.frame(name = rep(NA, nqual)
+                 , speed = rep(NA, nqual)
+                 , bould = rep(NA, nqual)
+                 , lead = rep(NA, nqual)
+                 , total = rep(NA, nqual)
+                 , count_above = rep(NA, nqual)
+                 , E = rep(NA, nqual)
+                 , q = rep(NA, nqual)
+                 )
+
+for (i in 1:nqual) {
+  t <- world_2019_w_qual$total[i]                                            # experimental/real score
+  #cat(m_qual_rp$first[i], m_qual_rp$last[i], " ", t, "\n")
+  s <-  compare_rp(t, rp_sims) %>% sum()                             # count which simulated rp values <= real rp
+  e <- s/nsim                                                        # avg expected value
+  q <- e/i                                                           # q value (false discovery rate)
+  #print(q)
+  df[i,] <- c(world_2019_w_qual$name[i]
+              , world_2019_w_qual$speed[i]
+              , world_2019_w_qual$bould[i]
+              , world_2019_w_qual$lead[i]
+              , world_2019_w_qual$total[i]
+              , s
+              , e
+              , q
+              )
+}
+
+df %>% knitr::kable()
+```

### 2021-07-28 manuscript/Rmd/draft.Rmd
Update Rmd

@@
 ---
-title: Climbing
+title: An Examination of Sport Climbing Scoring System
 
@@
 - name: Author 1
-  thanks: 
+  thanks: "Corresponding author: Name. Email:"
   affiliation:
@@
 - name: Author 2
+  thanks: 
   affiliation:
@@
 abstract: |
-  This manuscript... The main results are... 
+  The purpose of this paper is to investigate sport climbing, a new sport featured in the 2020 Tokyo Summer Olympics. Simulation. Data Analysis: drop and re-rank, correlation
 # bibliography: bib_example.bib
@@
 ---
+
+\normalsize
 
@@
 
-# Introduction
+# Introduction (sport climbing, other scoring system, mention rank product)
 
-(Copy from Rnw + Edit)
+In 2016, the International Olympic Committee announced the addition of five new sports to the 2020 Summer Olympics in Tokyo, Japan, which would then reschedule for 2021 due to the impact of the COVID-19 global pandemic.
+The five new features to Tokyo 2020's competitions program include baseball/softball, karate, skateboard, sports climbing, and surfing.
+One of the new sports, sport climbing, got our attention, specifically because of its unique scoring system and the fact that only one set of medals is awarded for each gender.
 
-## Sport Climbing
+Sport climbing at the 2020 Tokyo Olympics consists of three disciplines: speed climbing, bouldering, and lead climbing.
+Speed climbing takes place on a standardized course and competitors try to reach the top of the course as fast as possible.
+For Tokyo 2020, speed climbing is being contested in a head-to-head format with ranks determined by how far a competitor advances in the bracket.
+In bouldering, contestants have a fixed amount of time to complete as many courses as they can.
+Winners are determined based on who completes the most courses and ties are broken based on who had the fewest attempts.
+Ties are further broken by the competitor achieved the most "zone holds", which are holds approximately halfway through each course.
+Finally, in lead climbing, an athlete gets one point for each hold that they reach, so whoever reaches the highest point on the wall is the winner.
+Each lead climber only gets one attempt and when they fall their attempt is over.
+These three different climbing disciplines demand different sets of skills and, often, athletes specialize in a single event.
+However, since only one set of Olympic medals is awarded to sport climbing, rather than choosing only one of these disciplines to include in the Olympics, all three events were chosen to be included as a sort of climbing triathlon.
 
-## Other Scoring Systems
+In the Olympic sport climbing, there are 20 competitors at the start (in both men's and women's).
+All 20 competitors compete in each of the three events in the qualification round, and their performances in each event are ranked from 1 to 20.
+A competitors final score is then computed as the product of their ranks in the three events and the lower product is better.
+Specifically, $$Score_i = R^S_i\times R^B_i\times R^L_i,$$ where $R^S_i$, $R^B_i$, and $R^L_i$ are the ranks of the $i$-th competitor in speed climbing, bouldering, and lead climbing, respectively.
+
+The top 8 finishers in the qualification round advance to the finals where they once again compete in all three events, they are again ranked from 1 to 8, and their final score is the product of these three ranks in the final.
+Whoever has the lowest product of ranks in the final wins the gold medal.
+This type of scoring system heavily rewards high finishes and relatively ignores poor finishes.
+For instance, if climber A finished 1st, 20th, and 20th and climber B finished 10th, 10th, and 10th, climber B would have a score of 1000 whereas climber A would have a much better score of 400, despite finishing last in 2 out of 3 of the events.
+
+Heavily criticized
+
+Other sports scoring methods
 
@@
 
+## Data Collection
+
 We collected data on major climbing competitions from 2018 to 2020, including the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Asian Games.
+
+## EDA
 
@@
 
-## Simulations: Uniform Ranks
+## Simulations
 
@@
 In order to obtain a climbing medal, the average score for getting gold, silver, and bronze are 9.6748, 20.4143, and 33.2648, respectively (Table 2).
-A notable trend we observe for both qualification and final rounds is as the rank increases, the distribution of the scores becomes more spread out.
-(Figure..., facet)
+
+<!-- A notable trend we observe for both qualification and final rounds is as the rank increases, the distribution of the scores becomes more spread out. (Figure..., facet) -->
 
@@
 
-## Drawbacks of the scoring system
+## Data Analysis
 
-### Sum of ranks instead of product
+<!-- ### Sum of ranks instead of product -->
 
-We performed the same tasks as we did for products
+<!-- We performed the same tasks as we did for products -->
 
-Probabilities of advancing, winning medals are all lower
+<!-- Probabilities of advancing, winning medals are all lower -->
 
-Most notably for qualifications,
+<!-- Most notably for qualifications,  -->
 
-Obvious that the average score between the ranks are closer to each other
+<!-- Obvious that the average score between the ranks are closer to each other -->
 
-The amount of variability doesn't seem to be different as the rank increases
+<!-- The amount of variability doesn't seem to be different as the rank increases -->
 
@@
   ggplot(aes(x = last, y = total)) +
-  geom_col() +
-  geom_text(aes(label = rank), hjust = -0.1, size = 3) +
+  geom_col(show.legend = FALSE) +
+  geom_text(aes(label = rank), hjust = -0.2, size = 3) +
   coord_flip() +
   facet_wrap(~ rank_drop) +
+  expand_limits(y = 65) +
   labs(y = "Score",

### 2021-07-28 manuscript/Rmd/draft.tex
Update Rmd

@@
 {
-  \title{\bf Climbing}
+  \title{\bf An Examination of Sport Climbing Scoring System}
 
   \author{
-        Author 1 \\
+        Author 1 \thanks{Corresponding author: Name. Email:} \\
     \\
@@
   \begin{center}
-    {\LARGE\bf Climbing}
+    {\LARGE\bf An Examination of Sport Climbing Scoring System}
   \end{center}
@@
 \begin{abstract}
-This manuscript\ldots{} The main results are\ldots{}
+The purpose of this paper is to investigate sport climbing, a new sport
+featured in the 2020 Tokyo Summer Olympics. Simulation. Data Analysis:
+drop and re-rank, correlation
 \end{abstract}
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\normalsize
 
-(Copy from Rnw + Edit)
+\hypertarget{introduction-sport-climbing-other-scoring-system-mention-rank-product}{%
+\section{Introduction (sport climbing, other scoring system, mention
+rank
+product)}\label{introduction-sport-climbing-other-scoring-system-mention-rank-product}}
 
-\hypertarget{sport-climbing}{%
-\subsection{Sport Climbing}\label{sport-climbing}}
+In 2016, the International Olympic Committee announced the addition of
+five new sports to the 2020 Summer Olympics in Tokyo, Japan, which would
+then reschedule for 2021 due to the impact of the COVID-19 global
+pandemic. The five new features to Tokyo 2020's competitions program
+include baseball/softball, karate, skateboard, sports climbing, and
+surfing. One of the new sports, sport climbing, got our attention,
+specifically because of its unique scoring system and the fact that only
+one set of medals is awarded for each gender.
 
-\hypertarget{other-scoring-systems}{%
-\subsection{Other Scoring Systems}\label{other-scoring-systems}}
+Sport climbing at the 2020 Tokyo Olympics consists of three disciplines:
+speed climbing, bouldering, and lead climbing. Speed climbing takes
+place on a standardized course and competitors try to reach the top of
+the course as fast as possible. For Tokyo 2020, speed climbing is being
+contested in a head-to-head format with ranks determined by how far a
+competitor advances in the bracket. In bouldering, contestants have a
+fixed amount of time to complete as many courses as they can. Winners
+are determined based on who completes the most courses and ties are
+broken based on who had the fewest attempts. Ties are further broken by
+the competitor achieved the most ``zone holds'', which are holds
+approximately halfway through each course. Finally, in lead climbing, an
+athlete gets one point for each hold that they reach, so whoever reaches
+the highest point on the wall is the winner. Each lead climber only gets
+one attempt and when they fall their attempt is over. These three
+different climbing disciplines demand different sets of skills and,
+often, athletes specialize in a single event. However, since only one
+set of Olympic medals is awarded to sport climbing, rather than choosing
+only one of these disciplines to include in the Olympics, all three
+events were chosen to be included as a sort of climbing triathlon.
+
+In the Olympic sport climbing, there are 20 competitors at the start (in
+both men's and women's). All 20 competitors compete in each of the three
+events in the qualification round, and their performances in each event
+are ranked from 1 to 20. A competitors final score is then computed as
+the product of their ranks in the three events and the lower product is
+better. Specifically, \[Score_i = R^S_i\times R^B_i\times R^L_i,\] where
+\(R^S_i\), \(R^B_i\), and \(R^L_i\) are the ranks of the \(i\)-th
+competitor in speed climbing, bouldering, and lead climbing,
+respectively.
+
+The top 8 finishers in the qualification round advance to the finals
+where they once again compete in all three events, they are again ranked
+from 1 to 8, and their final score is the product of these three ranks
+in the final. Whoever has the lowest product of ranks in the final wins
+the gold medal. This type of scoring system heavily rewards high
+finishes and relatively ignores poor finishes. For instance, if climber
+A finished 1st, 20th, and 20th and climber B finished 10th, 10th, and
+10th, climber B would have a score of 1000 whereas climber A would have
+a much better score of 400, despite finishing last in 2 out of 3 of the
+events.
+
+Heavily criticized
+
+Other sports scoring methods
 
@@
 \section{Data and Methods}\label{data-and-methods}}
+
+\hypertarget{data-collection}{%
+\subsection{Data Collection}\label{data-collection}}
 
@@
 
+\hypertarget{eda}{%
+\subsection{EDA}\label{eda}}
+
 \hypertarget{results}{%
@@
 
-\hypertarget{simulations-uniform-ranks}{%
-\subsection{Simulations: Uniform
-Ranks}\label{simulations-uniform-ranks}}
+\hypertarget{simulations}{%
+\subsection{Simulations}\label{simulations}}
 
@@
 getting gold, silver, and bronze are 9.6748, 20.4143, and 33.2648,
-respectively (Table 2). A notable trend we observe for both
-qualification and final rounds is as the rank increases, the
-distribution of the scores becomes more spread out. (Figure\ldots,
-facet)
+respectively (Table 2).
 
@@
 
-\hypertarget{drawbacks-of-the-scoring-system}{%
-\subsection{Drawbacks of the scoring
-system}\label{drawbacks-of-the-scoring-system}}
-
-\hypertarget{sum-of-ranks-instead-of-product}{%
-\subsubsection{Sum of ranks instead of
-product}\label{sum-of-ranks-instead-of-product}}
-
-We performed the same tasks as we did for products
-
-Probabilities of advancing, winning medals are all lower
-
-Most notably for qualifications,
-
-Obvious that the average score between the ranks are closer to each
-other
-
-The amount of variability doesn't seem to be different as the rank
-increases
+\hypertarget{data-analysis}{%
+\subsection{Data Analysis}\label{data-analysis}}
 

### 2021-07-29 manuscript/Rmd/draft.Rmd
Update Rmd

@@
 ---
-title: An Examination of Sport Climbing Scoring System
+title: An Examination of Sport Climbing Competition Format and Scoring System
 
@@
 - scoring system
+- 2020 Summer Olympics
 
 abstract: |
-  The purpose of this paper is to investigate sport climbing, a new sport featured in the 2020 Tokyo Summer Olympics. Simulation. Data Analysis: drop and re-rank, correlation
+  The purpose of this paper is to investigate sport climbing, one of the sports making its debut on the Olympics stage at Tokyo 2020.
+In particular, we take a closer look at the controversial competition format and scoring system of sport climbing.
+
+  ... Simulation. Data Analysis: drop and re-rank, correlations
 # bibliography: bib_example.bib
@@
 ---
-
-\normalsize
 
@@
                       comment = "",
-                      # fig.width = 5,
-                      # fig.height = 4,
+                      fig.height = 4,
+                      fig.width = 5,
                       results = "asis")
@@
 
-# Introduction (sport climbing, other scoring system, mention rank product)
+# Introduction
 
@@
 A competitors final score is then computed as the product of their ranks in the three events and the lower product is better.
-Specifically, $$Score_i = R^S_i\times R^B_i\times R^L_i,$$ where $R^S_i$, $R^B_i$, and $R^L_i$ are the ranks of the $i$-th competitor in speed climbing, bouldering, and lead climbing, respectively.
+Specifically,
+\begin{equation}
+Score_i = R^S_i\times R^B_i\times R^L_i,
+\end{equation} 
+where $R^S_i$, $R^B_i$, and $R^L_i$ are the ranks of the $i$-th competitor in speed climbing, bouldering, and lead climbing, respectively.
 
@@
 
+As a side note, rank product, genes
+
 # Data and Methods
 
-## Data Collection
+We collected data on major climbing competitions from 2018 to 2020, including the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Youth Olympics.
+Data were obtained from various sources, including the event websites, Wikipedia, and the International Federation of Sport Climbing (IFSC).
+The main attributes of our datasets are the name and nationality of the climbers; bib number (for some competitions); the finishing place of climbers in speed climbing, bouldering, and lead climbing; the total score (which equals the product of event ranks); and the final rank.
+We utilize this data to compute the correlations between the event ranks and final table position, as well as to look at how often the final orderings change if one athlete is dropped and the ranks for each discipline are re-computed.
 
-We collected data on major climbing competitions from 2018 to 2020, including the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Asian Games.
+EDA
 
-## EDA
+Corr
+
+Drop 1
 
@@
 For each round, we performed 10000 simulations, and this was accomplished by randomly assigning the ranks of each event to every participant, with the assumption that the ranks are uniformly distributed.
-After the completion of the simulations, we calculated the final scores for every simulated round, as well as the final standings for the climbing athletes.
-This data would then enable us to answer questions about various topics, including the distributions of scores for qualifying and final rounds, and the probabilities of advancing to the finals or winning a medal, given certain conditions.
+After the completion of the simulations, we calculated the total scores for every simulated round, as well as the final standings for the climbing athletes.
+Table
+and
+Figure
+are numerical and visual summaries of final scores obtaining from our simulated data for the qualification round.
 
-```{r, fig.cap = "Distribution of qualification scores"}
+```{r}
+mosaic::favstats(qual$qual_score)[-9] %>% 
+  xtable(caption = "Descriptive statistics of simulated qualification scores") %>% 
+  print()
+```
+
+```{r, fig.cap = "Histogram of qualification scores"}
 qual %>%
@@
 
+The simulation results allow to answer questions about various topics, including the distributions of scores for qualifying and final rounds, and the probabilities of advancing to the finals or winning a medal, given certain conditions.
 For the qualification round, a climber is almost guaranteed to make the final round if they win the first event (with a 99.51% chance of advancing) or if they win at least one of the three climbing concentrations (99.48%).
@@
 
-```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Women's Qualification"}
+```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Women's Qualification", fig.width = 6}
 wq %>% 
@@
   ggpairs(diag = "blank",
-          upper = list(continuous = wrap("cor", method = "kendall")))
+          axisLabels = "internal",
+          upper = list(continuous = wrap("cor", method = "kendall", stars = FALSE))) +
+  theme(panel.grid.major = element_blank())
 ```
@@
 ```{r}
-women_final <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2018_youth_olympics/women_final.csv")
+wf <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2018_youth_olympics/women_final.csv")
+mf <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2018_youth_olympics/men_final.csv")
+
+drop_rerank <- function(df) {
   rerank <- list()
-for (i in 1:nrow(women_final)){
-  rerank[[i]] <- women_final[-i,] %>%
+  for (i in 1:nrow(df)) {
+    rerank[[i]] <- df[-i, ] %>%
       mutate(rank_drop = i)
@@
 
-rerank_df <- women_final %>% 
+  rerank_df <- df %>%
     mutate(rank_drop = 0) %>%
@@
     group_by(rank_drop) %>%
-  mutate(speed = rank(speed),
+    mutate(
+      speed = rank(speed),
       bould = rank(bould),
       lead = rank(lead),
-         total = speed * bould * lead) %>% 
+      total = speed * bould * lead
+    ) %>%
     arrange(total, .by_group = TRUE) %>%
@@
     group_by(rank_drop, total) %>%
-  mutate(speed_tb = ifelse(speed < lag(speed), 1, 0),
+    mutate(
+      speed_tb = ifelse(speed < lag(speed), 1, 0),
       bould_tb = ifelse(bould < lag(bould), 1, 0),
@@
       tb = speed_tb + bould_tb + lead_tb,
-         tb = ifelse(is.na(tb), 1, tb)) %>% 
+      tb = ifelse(is.na(tb), 1, tb)
+    ) %>%
     ungroup() %>%
     group_by(rank_drop) %>%
-  arrange(total, -tb, .by_group = TRUE) %>% 
-  mutate(rank = row_number())
+    arrange(total,-tb, .by_group = TRUE) %>%
+    mutate(rank = row_number(),
+           last = str_to_title(last))
+
+  return(rerank_df)
+}
 ```
 
-```{r, fig.cap = "Original rankings and rankings after each rank is dropped"}
-rerank_df %>% 
-  mutate(last = fct_reorder(last, -rank)) %>% 
-  # filter(rank_drop %in% c(0, 5)) %>% 
-  ggplot(aes(x = last, y = total)) +
+```{r, fig.cap = "This figure shows the rankings of the Men's Sport Climbing Final at the 2018 Youth Olympics.
+Each panel represents the dropped rank, with 0 being the original final standings.
+The cases where a change in ordering happened are highlighted.", fig.width = 7}
+drop_rerank(mf) %>% 
+  mutate(last = fct_reorder(last, -rank),
+         rank = as.factor(rank),
+         rank_change = ifelse(
+           rank_drop %in% c(0, 4) | 
+             rank_drop == 1 & rank %in% 4:5 |
+              rank_drop == 2 & rank %in% c(1, 5) |
+                rank_drop %in% c(3, 5) & rank %in% c(1, 2, 5) |
+                  rank_drop == 6 & rank %in% 1:3, 
+           "no", "yes")) %>% 
+  ggplot(aes(x = last, y = total, fill = rank_change)) +
   geom_col(show.legend = FALSE) +
@@
   coord_flip() +
-  facet_wrap(~ rank_drop) +
-  expand_limits(y = 65) +
+  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
+            data = ~ filter(., rank_drop %in% c(1, 2, 3, 5, 6)), 
+            color = "black", size = 1.5, fill = NA, inherit.aes = FALSE) +
+  facet_wrap(~ rank_drop, nrow = 2) +
+  expand_limits(y = 102, x = 0:7) +
+  scale_fill_manual(values = c("grey", "chocolate")) +
   labs(y = "Score",
-       x = "Climber")
+       x = "Climber") +
+  theme_minimal() +
+  theme(axis.ticks = element_blank(),
+        panel.grid.minor = element_blank())
 ```
+
+```{r, fig.cap = "This figure shows the final rankings of 2018 Youth Olympics Women's Final.
+Each panel represents the dropped rank, with 0 being the original final standings.
+The cases where a change in ordering happened are highlighted.", fig.width = 7}
+drop_rerank(wf) %>% 
+  mutate(last = fct_reorder(last, -rank),
+         rank = as.factor(rank),
+         rank_change = ifelse(
+           rank_drop %in% c(0, 1, 4, 6) |
+             rank_drop == 2 & rank %in% 1:3 |
+              rank_drop == 3 & rank %in% c(1, 4, 5) |
+                rank_drop == 5 & rank %in% c(1, 5),
+           "no", "yes")) %>%
+  ggplot(aes(x = last, y = total, fill = rank_change)) +
+  geom_col(show.legend = FALSE) +
+  geom_text(aes(label = rank), hjust = -0.2, size = 3) +
+  coord_flip() +
+  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
+            data = ~ filter(., rank_drop %in% c(2, 3, 5)), 
+            color = "black", size = 1.5, fill = NA, inherit.aes = FALSE) +
+  facet_wrap(~ rank_drop, nrow = 2, ncol = 4) +
+  expand_limits(y = 62, x = 0:7) +
+  scale_fill_manual(values = c("grey", "chocolate")) +
+  labs(y = "Score",
+       x = "Climber") +
+  theme_minimal() +
+  theme(axis.ticks = element_blank(),
+        panel.grid.minor = element_blank())
+```

### 2021-07-29 manuscript/Rmd/draft.tex
Update Rmd

@@
 \usepackage{textcomp}
+\usepackage{float}
 \usepackage[hyphens]{url} % not crucial - just used below for the URL
@@
 {
-  \title{\bf An Examination of Sport Climbing Scoring System}
+  \title{\bf An Examination of Sport Climbing Competition Format and
+Scoring System}
 
@@
   \begin{center}
-    {\LARGE\bf An Examination of Sport Climbing Scoring System}
+    {\LARGE\bf An Examination of Sport Climbing Competition Format and
+Scoring System}
   \end{center}
@@
 \begin{abstract}
-The purpose of this paper is to investigate sport climbing, a new sport
-featured in the 2020 Tokyo Summer Olympics. Simulation. Data Analysis:
-drop and re-rank, correlation
+The purpose of this paper is to investigate sport climbing, one of the
+sports making its debut on the Olympics stage at Tokyo 2020. In
+particular, we take a closer look at the controversial competition
+format and scoring system of sport climbing.
+
+\ldots{} Simulation. Data Analysis: drop and re-rank, correlations
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} sport climbing, scoring system
+{\it Keywords:} sport climbing, scoring system, 2020 Summer Olympics
 \vfill
@@
 
-\normalsize
-
-\hypertarget{introduction-sport-climbing-other-scoring-system-mention-rank-product}{%
-\section{Introduction (sport climbing, other scoring system, mention
-rank
-product)}\label{introduction-sport-climbing-other-scoring-system-mention-rank-product}}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 the product of their ranks in the three events and the lower product is
-better. Specifically, \[Score_i = R^S_i\times R^B_i\times R^L_i,\] where
-\(R^S_i\), \(R^B_i\), and \(R^L_i\) are the ranks of the \(i\)-th
-competitor in speed climbing, bouldering, and lead climbing,
-respectively.
+better. Specifically, \begin{equation}
+Score_i = R^S_i\times R^B_i\times R^L_i,
+\end{equation} where \(R^S_i\), \(R^B_i\), and \(R^L_i\) are the ranks
+of the \(i\)-th competitor in speed climbing, bouldering, and lead
+climbing, respectively.
 
@@
 
+As a side note, rank product, genes
+
 \hypertarget{data-and-methods}{%
 \section{Data and Methods}\label{data-and-methods}}
-
-\hypertarget{data-collection}{%
-\subsection{Data Collection}\label{data-collection}}
 
@@
 Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and
-2018 Asian Games.
+2018 Youth Olympics. Data were obtained from various sources, including
+the event websites, Wikipedia, and the International Federation of Sport
+Climbing (IFSC). The main attributes of our datasets are the name and
+nationality of the climbers; bib number (for some competitions); the
+finishing place of climbers in speed climbing, bouldering, and lead
+climbing; the total score (which equals the product of event ranks); and
+the final rank. We utilize this data to compute the correlations between
+the event ranks and final table position, as well as to look at how
+often the final orderings change if one athlete is dropped and the ranks
+for each discipline are re-computed.
 
-\hypertarget{eda}{%
-\subsection{EDA}\label{eda}}
+EDA
+
+Corr
+
+Drop 1
 
@@
 assumption that the ranks are uniformly distributed. After the
-completion of the simulations, we calculated the final scores for every
+completion of the simulations, we calculated the total scores for every
 simulated round, as well as the final standings for the climbing
-athletes. This data would then enable us to answer questions about
-various topics, including the distributions of scores for qualifying and
-final rounds, and the probabilities of advancing to the finals or
-winning a medal, given certain conditions.
+athletes. Table and Figure are numerical and visual summaries of final
+scores obtaining from our simulated data for the qualification round.
 
-\begin{figure}
+\begin{table}[ht]
 \centering
-\includegraphics{draft_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{Distribution of qualification scores}
+\caption{Descriptive statistics of simulated qualification scores} 
+\begin{tabular}{rrrrrrrr}
+  \hline
+min & Q1 & median & Q3 & max & mean & sd & n \\ 
+  \hline
+1.00 & 240.00 & 684.00 & 1638.00 & 8000.00 & 1158.84 & 1273.48 & 200000 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{figure}[H]
+\centering
+\includegraphics{draft_files/figure-latex/unnamed-chunk-5-1.pdf}
+\caption{Histogram of qualification scores}
 \end{figure}
 
-For the qualification round, a climber is almost guaranteed to make the
-final round if they win the first event (with a 99.51\% chance of
-advancing) or if they win at least one of the three climbing
-concentrations (99.48\%). On the other hand, finishing last in the first
-event or in any event would certainly hurt an athlete's chance of
-finishing in the top 8, as the probabilities of a climber advancing
-given they finish last in the first and in any event are 0.1830 and
-0.1885, respectively. In addition, the average score for qualification
-positions 1 to 8 are displayed in Table 1. We notice that on average,
-the minimum score that one should aim for in order to move on to the
-final round is 435 (for 8th rank).
+The simulation results allow to answer questions about various topics,
+including the distributions of scores for qualifying and final rounds,
+and the probabilities of advancing to the finals or winning a medal,
+given certain conditions. For the qualification round, a climber is
+almost guaranteed to make the final round if they win the first event
+(with a 99.51\% chance of advancing) or if they win at least one of the
+three climbing concentrations (99.48\%). On the other hand, finishing
+last in the first event or in any event would certainly hurt an
+athlete's chance of finishing in the top 8, as the probabilities of a
+climber advancing given they finish last in the first and in any event
+are 0.1830 and 0.1885, respectively. In addition, the average score for
+qualification positions 1 to 8 are displayed in Table 1. We notice that
+on average, the minimum score that one should aim for in order to move
+on to the final round is 435 (for 8th rank).
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 \centering
-\includegraphics{draft_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{draft_files/figure-latex/unnamed-chunk-8-1.pdf}
 \caption{Boxplots of scoring distribution for every qualification rank}
@@
 
-\begin{figure}
+\begin{figure}[H]
 \centering
-\includegraphics{draft_files/figure-latex/unnamed-chunk-9-1.pdf}
+\includegraphics{draft_files/figure-latex/unnamed-chunk-10-1.pdf}
 \caption{Kendall's rank correlations - 2018 World Championship, Women's
@@
 
-\begin{figure}
+\begin{figure}[H]
 \centering
-\includegraphics{draft_files/figure-latex/unnamed-chunk-11-1.pdf}
-\caption{Original rankings and rankings after each rank is dropped}
+\includegraphics{draft_files/figure-latex/unnamed-chunk-12-1.pdf}
+\caption{This figure shows the rankings of the Men's Sport Climbing
+Final at the 2018 Youth Olympics. Each panel represents the dropped
+rank, with 0 being the original final standings. The cases where a
+change in ordering happened are highlighted.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics{draft_files/figure-latex/unnamed-chunk-13-1.pdf}
+\caption{This figure shows the final rankings of 2018 Youth Olympics
+Women's Final. Each panel represents the dropped rank, with 0 being the
+original final standings. The cases where a change in ordering happened
+are highlighted.}
 \end{figure}

### 2021-07-29 manuscript/Rmd/draft.Rmd
Update Rmd + code

@@
   The purpose of this paper is to investigate sport climbing, one of the sports making its debut on the Olympics stage at Tokyo 2020.
-In particular, we take a closer look at the controversial competition format and scoring system of sport climbing.
-
-  ... Simulation. Data Analysis: drop and re-rank, correlations
+In particular, we take a closer look at the controversial competition format and scoring system of sport climbing...
+Simulation.
+Data Analysis: drop and re-rank, correlations
 # bibliography: bib_example.bib
@@
 
-In the Olympic sport climbing, there are 20 competitors at the start (in both men's and women's).
-All 20 competitors compete in each of the three events in the qualification round, and their performances in each event are ranked from 1 to 20.
-A competitors final score is then computed as the product of their ranks in the three events and the lower product is better.
-Specifically,
+At the 2020 Summer Olympics, both men's and women's sport climbing competitions begin with 20 climbers who has previously qualified for the Olympics from qualifying events held in 2019 and 2020.
+All 20 athletes compete in each of the three disciplines in the qualification round, and their performances in each concentration are ranked from 1 to 20.
+A competitor's total score is then computed as the product of their ranks in the three events and the lower product is better; specifically,
 \begin{equation}
@@
 
-The top 8 finishers in the qualification round advance to the finals where they once again compete in all three events, they are again ranked from 1 to 8, and their final score is the product of these three ranks in the final.
-Whoever has the lowest product of ranks in the final wins the gold medal.
+The top 8 finishers in the qualification round advance to the finals where they once again compete in all three of speed climbing, bouldering, and lead climbing.
+The total score for each climber in the final stage is determined by multiplying their ranks in each discipline, similar to the qualification round.
+This implies that the climber with the lowest product of ranks in the final wins the gold medal.
 This type of scoring system heavily rewards high finishes and relatively ignores poor finishes.
@@
 
-As a side note, rank product, genes
+The manuscript is outlined as follows.
+We first begin with some descriptions of the data and methods in Section 2.
+Our analyses and results are then presented in Sections 3 and 4.
+Finally, in Section 5, we summarize up of main findings and give a discussion to close out the paper.
 
@@
 
-We collected data on major climbing competitions from 2018 to 2020, including the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Youth Olympics.
+The first sets of data that we rely on are historical results from major climbing competitions in the past.
+We collected data on climbing contests that took place between 2018 and 2020, where the combined format was used to determine the scores and ranks of climbers.
+The events include the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Youth Olympics.
 Data were obtained from various sources, including the event websites, Wikipedia, and the International Federation of Sport Climbing (IFSC).
@@
 
-EDA
+Our second data source comes from a simulation study that we conducted, with the purpose of examining the rankings and scoring for climbers in both qualification and final rounds.
+For each round, we performed 10000 simulations, and this was accomplished by randomly assigning the ranks of each event to every participant, with the assumption that the ranks are uniformly distributed.
+After the completion of the simulations, we calculated the total scores for every simulated round, as well as the final standings for the climbing athletes.
+The simulation results allow to answer questions about various topics, including the distributions of scores for qualifying and final rounds, and the probabilities of advancing to the finals or winning a medal, given certain conditions.
 
-Corr
-
-Drop 1
-
-# Results
-
-## Simulations
+Table 1 and Figure 1 are numerical and visual summaries of climbing total score obtained from our simulated data for the qualification round.
 
@@
 qual <- bind_rows(qual_sim) %>% 
-  mutate(qual_score = e1 * e2 * e3) %>% 
+  mutate(score = e1 * e2 * e3) %>% 
   group_by(sim) %>% 
-  mutate(qual_rank = rank(qual_score, ties.method = "random")) %>%
+  mutate(rank = rank(score, ties.method = "random")) %>%
   ungroup()
@@
 final <- final %>% 
-  mutate(final_score = e1 * e2 * e3) %>% 
+  mutate(score = e1 * e2 * e3) %>% 
   group_by(sim) %>% 
-  mutate(final_rank = rank(final_score, ties.method = "random")) %>%
+  mutate(rank = rank(score, ties.method = "random")) %>%
   ungroup()
+
+qual_final <- qual %>% 
+  mutate(round = "Qualification") %>% 
+  bind_rows(mutate(final, round = "Final")) %>% 
+  mutate(round = fct_rev(round))
 ```
 
-We conducted a simulation study to examine the rankings and scoring for climbers in both qualification and final rounds.
-For each round, we performed 10000 simulations, and this was accomplished by randomly assigning the ranks of each event to every participant, with the assumption that the ranks are uniformly distributed.
-After the completion of the simulations, we calculated the total scores for every simulated round, as well as the final standings for the climbing athletes.
-Table
-and
-Figure
-are numerical and visual summaries of final scores obtaining from our simulated data for the qualification round.
-
 ```{r}
-mosaic::favstats(qual$qual_score)[-9] %>% 
-  xtable(caption = "Descriptive statistics of simulated qualification scores") %>% 
+mosaic::favstats(score ~ round, data = qual_final)[-10] %>% 
+  xtable(caption = "Descriptive statistics of simulated scores for qualification and final rounds") %>% 
   print()
@@
 
-```{r, fig.cap = "Histogram of qualification scores"}
-qual %>%
-  ggplot(aes(qual_score)) +
-  geom_histogram(bins = 20, color = "white") + 
-  labs(x = "Qualification Score",
+```{r, fig.cap = "Histogram of simulated scores for qualification and final rounds", fig.width = 6.4, fig.height = 3.2}
+qual_final %>%
+  ggplot(aes(score)) +
+  geom_histogram(bins = 20, fill = "lightgray", color = "black") + 
+  facet_wrap(~ round, scales = "free") +
+  labs(x = "Score",
        y = "Frequency")
-  # theme(axis.title = element_text(size = 9),
-  #       axis.text = element_text(size = 8))
 ```
 
-The simulation results allow to answer questions about various topics, including the distributions of scores for qualifying and final rounds, and the probabilities of advancing to the finals or winning a medal, given certain conditions.
+# Simulations
+
 For the qualification round, a climber is almost guaranteed to make the final round if they win the first event (with a 99.51% chance of advancing) or if they win at least one of the three climbing concentrations (99.48%).
@@
 qual %>% 
-  group_by(qual_rank) %>% 
-  summarize(avg_adv_score = mean(qual_score)) %>% 
-  filter(qual_rank <= 8) %>% 
+  group_by(rank) %>% 
+  summarize(avg_adv_score = mean(score)) %>% 
+  filter(rank <= 8) %>% 
   xtable(caption = "Average score for each qualifying rank") %>% 
@@
 final %>% 
-  group_by(final_rank) %>% 
-  summarize(avg_score = mean(final_score)) %>% 
+  group_by(rank) %>% 
+  summarize(avg_score = mean(score)) %>% 
   xtable(caption = "Average score of medalists") %>% 
@@
 
-```{r, fig.cap = "Boxplots of scoring distribution for every qualification rank"}
+```{r, fig.cap = "Boxplots of scoring distribution for every qualification rank", eval = FALSE}
 qual %>% 
-  mutate(qual_rank = factor(qual_rank)) %>% 
-  ggplot(aes(x = qual_score, y = qual_rank)) +
+  mutate(rank = factor(rank)) %>% 
+  ggplot(aes(x = score, y = rank)) +
   geom_boxplot()
@@
 
-## Data Analysis
+# Data Analysis
+
+Corr
+
+Drop 1
 
@@
 
-### Speed climbing vs lead and bouldering
+## Speed climbing vs lead and bouldering
 
@@
 
-```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Women's Qualification", fig.width = 6}
+```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Women's Qualification"}
 wq %>% 
@@
 
-### Drop and re-rank
+## Drop and re-rank
 

### 2021-07-29 manuscript/Rmd/draft.tex
Update Rmd + code

@@
 \usepackage{textcomp}
-\usepackage{float}
 \usepackage[hyphens]{url} % not crucial - just used below for the URL
@@
 particular, we take a closer look at the controversial competition
-format and scoring system of sport climbing.
-
-\ldots{} Simulation. Data Analysis: drop and re-rank, correlations
+format and scoring system of sport climbing\ldots{} Simulation. Data
+Analysis: drop and re-rank, correlations
 \end{abstract}
@@
 
-In the Olympic sport climbing, there are 20 competitors at the start (in
-both men's and women's). All 20 competitors compete in each of the three
-events in the qualification round, and their performances in each event
-are ranked from 1 to 20. A competitors final score is then computed as
-the product of their ranks in the three events and the lower product is
-better. Specifically, \begin{equation}
+At the 2020 Summer Olympics, both men's and women's sport climbing
+competitions begin with 20 climbers who has previously qualified for the
+Olympics from qualifying events held in 2019 and 2020. All 20 athletes
+compete in each of the three disciplines in the qualification round, and
+their performances in each concentration are ranked from 1 to 20. A
+competitor's total score is then computed as the product of their ranks
+in the three events and the lower product is better; specifically,
+\begin{equation}
 Score_i = R^S_i\times R^B_i\times R^L_i,
@@
 The top 8 finishers in the qualification round advance to the finals
-where they once again compete in all three events, they are again ranked
-from 1 to 8, and their final score is the product of these three ranks
-in the final. Whoever has the lowest product of ranks in the final wins
-the gold medal. This type of scoring system heavily rewards high
-finishes and relatively ignores poor finishes. For instance, if climber
-A finished 1st, 20th, and 20th and climber B finished 10th, 10th, and
-10th, climber B would have a score of 1000 whereas climber A would have
-a much better score of 400, despite finishing last in 2 out of 3 of the
-events.
+where they once again compete in all three of speed climbing,
+bouldering, and lead climbing. The total score for each climber in the
+final stage is determined by multiplying their ranks in each discipline,
+similar to the qualification round. This implies that the climber with
+the lowest product of ranks in the final wins the gold medal. This type
+of scoring system heavily rewards high finishes and relatively ignores
+poor finishes. For instance, if climber A finished 1st, 20th, and 20th
+and climber B finished 10th, 10th, and 10th, climber B would have a
+score of 1000 whereas climber A would have a much better score of 400,
+despite finishing last in 2 out of 3 of the events.
 
@@
 
-As a side note, rank product, genes
+The manuscript is outlined as follows. We first begin with some
+descriptions of the data and methods in Section 2. Our analyses and
+results are then presented in Sections 3 and 4. Finally, in Section 5,
+we summarize up of main findings and give a discussion to close out the
+paper.
 
@@
 
-We collected data on major climbing competitions from 2018 to 2020,
-including the 2020 Continental Championships of Europe, Africa, Oceania,
-Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and
-2018 Youth Olympics. Data were obtained from various sources, including
-the event websites, Wikipedia, and the International Federation of Sport
-Climbing (IFSC). The main attributes of our datasets are the name and
-nationality of the climbers; bib number (for some competitions); the
-finishing place of climbers in speed climbing, bouldering, and lead
-climbing; the total score (which equals the product of event ranks); and
-the final rank. We utilize this data to compute the correlations between
-the event ranks and final table position, as well as to look at how
-often the final orderings change if one athlete is dropped and the ranks
-for each discipline are re-computed.
+The first sets of data that we rely on are historical results from major
+climbing competitions in the past. We collected data on climbing
+contests that took place between 2018 and 2020, where the combined
+format was used to determine the scores and ranks of climbers. The
+events include the 2020 Continental Championships of Europe, Africa,
+Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian
+Games; and 2018 Youth Olympics. Data were obtained from various sources,
+including the event websites, Wikipedia, and the International
+Federation of Sport Climbing (IFSC). The main attributes of our datasets
+are the name and nationality of the climbers; bib number (for some
+competitions); the finishing place of climbers in speed climbing,
+bouldering, and lead climbing; the total score (which equals the product
+of event ranks); and the final rank. We utilize this data to compute the
+correlations between the event ranks and final table position, as well
+as to look at how often the final orderings change if one athlete is
+dropped and the ranks for each discipline are re-computed.
 
-EDA
+Our second data source comes from a simulation study that we conducted,
+with the purpose of examining the rankings and scoring for climbers in
+both qualification and final rounds. For each round, we performed 10000
+simulations, and this was accomplished by randomly assigning the ranks
+of each event to every participant, with the assumption that the ranks
+are uniformly distributed. After the completion of the simulations, we
+calculated the total scores for every simulated round, as well as the
+final standings for the climbing athletes. The simulation results allow
+to answer questions about various topics, including the distributions of
+scores for qualifying and final rounds, and the probabilities of
+advancing to the finals or winning a medal, given certain conditions.
 
-Corr
-
-Drop 1
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{simulations}{%
-\subsection{Simulations}\label{simulations}}
-
-We conducted a simulation study to examine the rankings and scoring for
-climbers in both qualification and final rounds. For each round, we
-performed 10000 simulations, and this was accomplished by randomly
-assigning the ranks of each event to every participant, with the
-assumption that the ranks are uniformly distributed. After the
-completion of the simulations, we calculated the total scores for every
-simulated round, as well as the final standings for the climbing
-athletes. Table and Figure are numerical and visual summaries of final
-scores obtaining from our simulated data for the qualification round.
+Table 1 and Figure 1 are numerical and visual summaries of climbing
+total score obtained from our simulated data for the qualification
+round.
 
@@
 \centering
-\caption{Descriptive statistics of simulated qualification scores} 
-\begin{tabular}{rrrrrrrr}
+\caption{Descriptive statistics of simulated scores for qualification and final rounds} 
+\begin{tabular}{lrrrrrrrr}
   \hline
-min & Q1 & median & Q3 & max & mean & sd & n \\ 
+round & min & Q1 & median & Q3 & max & mean & sd & n \\ 
   \hline
-1.00 & 240.00 & 684.00 & 1638.00 & 8000.00 & 1158.84 & 1273.48 & 200000 \\ 
+Qualification & 1.00 & 240.00 & 684.00 & 1638.00 & 8000.00 & 1158.84 & 1273.48 & 200000 \\ 
+  Final & 1.00 & 24.00 & 60.00 & 126.00 & 512.00 & 91.12 & 91.35 & 80000 \\ 
    \hline
@@
 
-\begin{figure}[H]
+\begin{figure}
 \centering
 \includegraphics{draft_files/figure-latex/unnamed-chunk-5-1.pdf}
-\caption{Histogram of qualification scores}
+\caption{Histogram of simulated scores for qualification and final
+rounds}
 \end{figure}
 
-The simulation results allow to answer questions about various topics,
-including the distributions of scores for qualifying and final rounds,
-and the probabilities of advancing to the finals or winning a medal,
-given certain conditions. For the qualification round, a climber is
-almost guaranteed to make the final round if they win the first event
-(with a 99.51\% chance of advancing) or if they win at least one of the
-three climbing concentrations (99.48\%). On the other hand, finishing
-last in the first event or in any event would certainly hurt an
-athlete's chance of finishing in the top 8, as the probabilities of a
-climber advancing given they finish last in the first and in any event
-are 0.1830 and 0.1885, respectively. In addition, the average score for
-qualification positions 1 to 8 are displayed in Table 1. We notice that
-on average, the minimum score that one should aim for in order to move
-on to the final round is 435 (for 8th rank).
+\hypertarget{simulations}{%
+\section{Simulations}\label{simulations}}
+
+For the qualification round, a climber is almost guaranteed to make the
+final round if they win the first event (with a 99.51\% chance of
+advancing) or if they win at least one of the three climbing
+concentrations (99.48\%). On the other hand, finishing last in the first
+event or in any event would certainly hurt an athlete's chance of
+finishing in the top 8, as the probabilities of a climber advancing
+given they finish last in the first and in any event are 0.1830 and
+0.1885, respectively. In addition, the average score for qualification
+positions 1 to 8 are displayed in Table 1. We notice that on average,
+the minimum score that one should aim for in order to move on to the
+final round is 435 (for 8th rank).
 
@@
   \hline
-qual\_rank & avg\_adv\_score \\ 
+rank & avg\_adv\_score \\ 
   \hline
@@
   \hline
-final\_rank & avg\_score \\ 
+rank & avg\_score \\ 
   \hline
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics{draft_files/figure-latex/unnamed-chunk-8-1.pdf}
-\caption{Boxplots of scoring distribution for every qualification rank}
-\end{figure}
+\hypertarget{data-analysis}{%
+\section{Data Analysis}\label{data-analysis}}
 
-\hypertarget{data-analysis}{%
-\subsection{Data Analysis}\label{data-analysis}}
+Corr
+
+Drop 1
 
 \hypertarget{speed-climbing-vs-lead-and-bouldering}{%
-\subsubsection{Speed climbing vs lead and
+\subsection{Speed climbing vs lead and
 bouldering}\label{speed-climbing-vs-lead-and-bouldering}}
@@
 
-\begin{figure}[H]
+\begin{figure}
 \centering
@@
 \hypertarget{drop-and-re-rank}{%
-\subsubsection{Drop and re-rank}\label{drop-and-re-rank}}
+\subsection{Drop and re-rank}\label{drop-and-re-rank}}
 

### 2021-07-29 R/LOOA_real.R

@@ -90,90 +90,3 @@ rerank_df <- women_qual %>%
   arrange(total, .by_group = TRUE) %>% 
   mutate(rank = row_number()) %>% 
   ungroup()
-
-
-#____________________________________________________________________________
-
-# Simulation
-
-# In what percentage of the (8!)^3 orderings would the winner change 
-# if one of the 8 competitors was dropped and there were only 7 instead of 8?
-
-library(tidyverse)
-library(combinat)
-
-nplay <- 6
-perms <- crossing(e1 = permn(nplay), 
-                  e2 = permn(nplay), 
-                  e3 = permn(nplay))
-
-
-get_perm <- function(x) {
-  unnest_perm <- perms[x, ] %>%
-    unnest(cols = c(e1, e2, e3)) %>% 
-    mutate(total = e1 * e2 * e3,
-           initial_rank = rank(total))
-  
-  unnest_perm %>%
-    mutate(drop = 0) %>%
-    bind_rows(unnest_perm %>% slice(-1) %>% mutate(drop = 1)) %>%
-    bind_rows(unnest_perm %>% slice(-2) %>% mutate(drop = 2)) %>%
-    bind_rows(unnest_perm %>% slice(-3) %>% mutate(drop = 3)) %>%
-    bind_rows(unnest_perm %>% slice(-4) %>% mutate(drop = 4)) %>%
-    bind_rows(unnest_perm %>% slice(-5) %>% mutate(drop = 5)) %>%
-    bind_rows(unnest_perm %>% slice(-6) %>% mutate(drop = 6)) %>%
-    mutate(perm = x)
-}
-
-
-perms_df <- tibble()
-for (row in 1:nrow(perms)) {
-  print(row)
-  perms_df <- perms_df %>% 
-    bind_rows(get_perm(row))
-}
-
-f <- perms_df %>% 
-  filter(drop != 0) %>% 
-  mutate(total = e1 * e2 * e3) %>% 
-  group_by(perm, drop) %>% 
-  mutate(new_rank = rank(total, ties.method = "random"),
-         initial_rank_ordered = rank(initial_rank))
-
-
-# dealing with ties
-
-
-qual <- bind_rows(qual_sim) %>% 
-  mutate(total = e1 * e2 * e3) %>% 
-  group_by(sim) %>% 
[…]

### 2021-07-29 R/LOOA_sim.R

@@ -1,0 +1,92 @@
+# In what percentage of the (6!)^3 orderings would the winner change 
+# if one of the 6 competitors was dropped and there were only 5 instead of 6?
+
+library(tidyverse)
+library(combinat)
+
+nplay <- 5
+perms <- crossing(e1 = permn(nplay), 
+                  e2 = permn(nplay), 
+                  e3 = permn(nplay))
+# perms is a tibble of lists
+
+# expand out
+perms_expand <- perms %>% 
+  mutate(perm = row_number()) %>% 
+  unnest(cols = c(e1, e2, e3))
+
+perms_expand %>% 
+  mutate(total = e1 * e2 * e3) %>% 
+  group_by(perm) %>% 
+  mutate(rank = rank(total)) %>%  # need to fix ties
+  ungroup()
+
+
+
+# get_perm <- function(x) {
+#   unnest_perm <- perms[x, ] %>%
+#     unnest(cols = c(e1, e2, e3)) %>% 
+#     mutate(total = e1 * e2 * e3,
+#            initial_rank = rank(total))
+#   
+#   unnest_perm %>%
+#     mutate(drop = 0) %>%
+#     bind_rows(unnest_perm %>% slice(-1) %>% mutate(drop = 1)) %>%
+#     bind_rows(unnest_perm %>% slice(-2) %>% mutate(drop = 2)) %>%
+#     bind_rows(unnest_perm %>% slice(-3) %>% mutate(drop = 3)) %>%
+#     bind_rows(unnest_perm %>% slice(-4) %>% mutate(drop = 4)) %>%
+#     bind_rows(unnest_perm %>% slice(-5) %>% mutate(drop = 5)) %>%
+#     bind_rows(unnest_perm %>% slice(-6) %>% mutate(drop = 6)) %>%
+#     mutate(perm = x)
+# }
+# 
+# 
+# perms_df <- tibble()
+# for (row in 1:nrow(perms)) {
+#   perms_df <- perms_df %>% 
+#     bind_rows(get_perm(row))
+# }
+# 
+# f <- perms_df %>% 
+#   filter(drop != 0) %>% 
+#   mutate(total = e1 * e2 * e3) %>% 
+#   group_by(perm, drop) %>% 
+#   mutate(new_rank = rank(total, ties.method = "random"),
[…]

### 2021-07-29 R/LOOA_sim.R
Leave one out simulation

@@ -15,13 +15,11 @@ perms_expand <- perms %>%
   mutate(perm = row_number()) %>% 
   unnest(cols = c(e1, e2, e3))
 
-perms_expand %>% 
+perms_expand <- perms_expand %>% 
   mutate(total = e1 * e2 * e3) %>% 
   group_by(perm) %>% 
   mutate(rank = rank(total)) %>%  # need to fix ties
   ungroup()
-
-
 
 # get_perm <- function(x) {
 #   unnest_perm <- perms[x, ] %>%

### 2021-07-29 manuscript/Rmd/draft.Rmd

@@
 Our analyses and results are then presented in Sections 3 and 4.
-Finally, in Section 5, we summarize up of main findings and give a discussion to close out the paper.
+Finally, in Section 5, we summarize up our main findings and provide a discussion to close out the paper.
 
@@
 
-Table 1 and Figure 1 are numerical and visual summaries of climbing total score obtained from our simulated data for the qualification round.
+Table 1 and Figure 1 are numerical and visual summaries of climbing total score obtained from our simulated data for the qualification and final rounds.
 
@@
 
+## Uniform Ranks
+
 For the qualification round, a climber is almost guaranteed to make the final round if they win the first event (with a 99.51% chance of advancing) or if they win at least one of the three climbing concentrations (99.48%).
@@
 
+## Leave-one-climber-out
+
+6!
+
 # Data Analysis
-
-Corr
-
-Drop 1
 
@@
 
-## Speed climbing vs lead and bouldering
+## Correlations
 
@@
 
-## Drop and re-rank
+## Leave-one-climber-out Analysis
 
@@
 
-```{r, fig.cap = "This figure shows the rankings of the Men's Sport Climbing Final at the 2018 Youth Olympics.
-Each panel represents the dropped rank, with 0 being the original final standings.
-The cases where a change in ordering happened are highlighted.", fig.width = 7}
-drop_rerank(mf) %>% 
+```{r}
+mf_gg <- drop_rerank(mf) %>% 
   mutate(last = fct_reorder(last, -rank),
@@
   scale_fill_manual(values = c("grey", "chocolate")) +
-  labs(y = "Score",
-       x = "Climber") +
+  labs(x = "Men", y = "") +
   theme_minimal() +
@@
         panel.grid.minor = element_blank())
-```
 
-```{r, fig.cap = "This figure shows the final rankings of 2018 Youth Olympics Women's Final.
-Each panel represents the dropped rank, with 0 being the original final standings.
-The cases where a change in ordering happened are highlighted.", fig.width = 7}
-drop_rerank(wf) %>% 
+wf_gg <- drop_rerank(wf) %>% 
   mutate(last = fct_reorder(last, -rank),
@@
   scale_fill_manual(values = c("grey", "chocolate")) +
-  labs(y = "Score",
-       x = "Climber") +
+  labs(y = "\nScore", x = "Women") +
   theme_minimal() +
@@
 ```
+
+```{r, fig.cap = "This figure illustrates the changes to the final rankings of the 2018 Youth Olympics Men's and Women's Finals when we leave out a single climber.
+For each gender competition, each panel represents the rank of the drop-out athlete, with 0 being the original final results.
+Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar", fig.width = 6.5, fig.height = 7.5}
+library(patchwork)
+mf_gg + wf_gg +
+  plot_layout(nrow = 2)
+```

### 2021-07-29 manuscript/Rmd/draft.tex
Update Rmd

@@
 \usepackage{hyperref}
+\usepackage{caption}
+\captionsetup[table]{skip=3pt}
+\usepackage{float}
 \providecommand{\tightlist}{%
@@
 results are then presented in Sections 3 and 4. Finally, in Section 5,
-we summarize up of main findings and give a discussion to close out the
-paper.
+we summarize up our main findings and provide a discussion to close out
+the paper.
 
@@
 Table 1 and Figure 1 are numerical and visual summaries of climbing
-total score obtained from our simulated data for the qualification
-round.
+total score obtained from our simulated data for the qualification and
+final rounds.
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 \centering
@@
 \section{Simulations}\label{simulations}}
+
+\hypertarget{uniform-ranks}{%
+\subsection{Uniform Ranks}\label{uniform-ranks}}
 
@@
 
+\hypertarget{leave-one-climber-out}{%
+\subsection{Leave-one-climber-out}\label{leave-one-climber-out}}
+
+6!
+
 \hypertarget{data-analysis}{%
@@
 
-Corr
-
-Drop 1
-
-\hypertarget{speed-climbing-vs-lead-and-bouldering}{%
-\subsection{Speed climbing vs lead and
-bouldering}\label{speed-climbing-vs-lead-and-bouldering}}
+\hypertarget{correlations}{%
+\subsection{Correlations}\label{correlations}}
 
@@
 
-\hypertarget{drop-and-re-rank}{%
-\subsection{Drop and re-rank}\label{drop-and-re-rank}}
+\hypertarget{leave-one-climber-out-analysis}{%
+\subsection{Leave-one-climber-out
+Analysis}\label{leave-one-climber-out-analysis}}
 
@@
 The cases where someone behind you drops out and your ranking changes.
-
 Example from 2018 youth, women's final dropping ranks 3 and 5 change the
@@
 
-\begin{figure}
-\centering
-\includegraphics{draft_files/figure-latex/unnamed-chunk-12-1.pdf}
-\caption{This figure shows the rankings of the Men's Sport Climbing
-Final at the 2018 Youth Olympics. Each panel represents the dropped
-rank, with 0 being the original final standings. The cases where a
-change in ordering happened are highlighted.}
-\end{figure}
-
-\begin{figure}
+\begin{figure}[H]
 \centering
 \includegraphics{draft_files/figure-latex/unnamed-chunk-13-1.pdf}
-\caption{This figure shows the final rankings of 2018 Youth Olympics
-Women's Final. Each panel represents the dropped rank, with 0 being the
-original final standings. The cases where a change in ordering happened
-are highlighted.}
+\caption{This figure illustrates the changes to the final rankings of
+the 2018 Youth Olympics Men's and Women's Finals when we leave out a
+single climber. For each gender competition, each panel represents the
+rank of the drop-out athlete, with 0 being the original final results.
+Each case with a change in rank orderings is highlighted by a black
+panel border, and any player with a rank change is represented by a
+red-filled bar}
 \end{figure}

### 2021-07-29 manuscript/Rmd/draft.tex
3.2 added intro

@@
 
-6!
+We would expect that if a climber with a high combined ranking is removed from the list of competitors, the placement of lower-ranked climbers will shift.
+In the most trivial case, if we were to remove the 3rd place climber, we would see 4th place climber shift to 3rd, 5th place climber shift to 4th, etc.
+It is maybe less trivial that when a climber is removed from the list, the placement of their higher-ranking competitors may also shift.
+In other words, there is a non-zero probability of seeing a change in the placement - particularly of medalists - regardless of the fact that no changes occurred in the performance of the remaining climbers.
+Real examples are given in section 4.2.
+In this section, we conduct simulations to analyze the probability of seeing a reordering of medalists when one non-medalist is removed from the list of finalists.
 

### 2021-07-29 manuscript/Rmd/draft.Rmd
3.2 added intro
Update Rmd

@@
 
-The first sets of data that we rely on are historical results from major climbing competitions in the past.
-We collected data on climbing contests that took place between 2018 and 2020, where the combined format was used to determine the scores and ranks of climbers.
-The events include the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Youth Olympics.
-Data were obtained from various sources, including the event websites, Wikipedia, and the International Federation of Sport Climbing (IFSC).
-The main attributes of our datasets are the name and nationality of the climbers; bib number (for some competitions); the finishing place of climbers in speed climbing, bouldering, and lead climbing; the total score (which equals the product of event ranks); and the final rank.
-We utilize this data to compute the correlations between the event ranks and final table position, as well as to look at how often the final orderings change if one athlete is dropped and the ranks for each discipline are re-computed.
-
-Our second data source comes from a simulation study that we conducted, with the purpose of examining the rankings and scoring for climbers in both qualification and final rounds.
+Our first sets of data come from a simulation study that we conducted, with the purpose of examining the rankings and scoring for climbers in both qualification and final rounds.
 For each round, we performed 10000 simulations, and this was accomplished by randomly assigning the ranks of each event to every participant, with the assumption that the ranks are uniformly distributed.
@@
 library(xtable)
-theme_set(theme_bw())
+theme_set(theme_minimal())
 ```
@@
 
+<!-- A notable trend we observe for both qualification and final rounds is as the rank increases, the distribution of the scores becomes more spread out. (Figure..., facet) -->
+
+Additionally, we rely historical results from major climbing competitions in recent years.
+We collected data on climbing contests that took place between 2018 and 2020, where the combined format was used to determine the scores and ranks of climbers.
+The events include the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Youth Olympics.
+Data were obtained from various sources, including the event websites, Wikipedia, and the International Federation of Sport Climbing (IFSC).
+The main attributes of our datasets are the name and nationality of the climbers; bib number (for some competitions); the finishing place of climbers in speed climbing, bouldering, and lead climbing; the total score (which equals the product of event ranks); and the final rank.
+We utilize this data to compute the correlations between the event ranks and final table position, as well as to look at how often the final orderings change if one athlete is dropped and the ranks for each discipline are re-computed.
+
 # Simulations
@@
 
-For the qualification round, a climber is almost guaranteed to make the final round if they win the first event (with a 99.51% chance of advancing) or if they win at least one of the three climbing concentrations (99.48%).
+In this section, we discuss the results of our simulations described in Section \ref{data-and-methods}.
+For the qualification round, our simulation study shows that a climber is almost guaranteed to make the final round if they win the first event (with a 99.51% chance of advancing) or if they win at least one of the three climbing concentrations (99.48%).
 On the other hand, finishing last in the first event or in any event would certainly hurt an athlete's chance of finishing in the top 8, as the probabilities of a climber advancing given they finish last in the first and in any event are 0.1830 and 0.1885, respectively.
-In addition, the average score for qualification positions 1 to 8 are displayed in Table 1.
-We notice that on average, the minimum score that one should aim for in order to move on to the final round is 435 (for 8th rank).
+In addition, the average score for qualification positions 1 to 8 are displayed in Table 2.
+We notice that on average, the minimum score that one should aim for in order to move on to the final round is about 435 (for 8th rank).
 
@@
 qual %>% 
-  group_by(rank) %>% 
-  summarize(avg_adv_score = mean(score)) %>% 
-  filter(rank <= 8) %>% 
-  xtable(caption = "Average score for each qualifying rank") %>% 
+  rename(Rank = rank) %>% 
+  group_by(Rank) %>% 
+  summarize(`Average score` = mean(score)) %>% 
+  filter(Rank <= 10) %>% 
+  xtable(caption = "The average scores for the top 10 qualification ranks according to our simulations. A climber will secure a finalist spot if they finish in the top 8.") %>% 
   print()
@@
 
-Regarding the finals, a climber is very likely to earn a medal (finish in the top 3) if they win the first event (83.03% chance) or any event (85.01%).
-In order to obtain a climbing medal, the average score for getting gold, silver, and bronze are 9.6748, 20.4143, and 33.2648, respectively (Table 2).
-
-<!-- A notable trend we observe for both qualification and final rounds is as the rank increases, the distribution of the scores becomes more spread out. (Figure..., facet) -->
+Regarding the finals, a climber is very likely to finish in the top 3 and hence earn a medal if they win the first event (83.03% chance) or any event (85.01% chance).
+Furthermore, according to our final simulations, in order to obtain a climbing medal, the average scores that put an athlete in position to receive gold, silver, and bronze medals are 9.6748, 20.4143, and 33.2648, respectively (see Table 3).
 
@@
 final %>% 
-  group_by(rank) %>% 
-  summarize(avg_score = mean(score)) %>% 
-  xtable(caption = "Average score of medalists") %>% 
+  rename(Rank = rank) %>% 
+  group_by(Rank) %>% 
+  summarize(`Average score` = mean(score)) %>% 
+  xtable(caption = "The average scores for all final ranks according to our simulations. Ranks 1, 2, and 3 are table positions that guarantee medalist status for climbers.") %>% 
   print()
-```
-
-```{r, fig.cap = "Boxplots of scoring distribution for every qualification rank", eval = FALSE}
-qual %>% 
-  mutate(rank = factor(rank)) %>% 
-  ggplot(aes(x = score, y = rank)) +
-  geom_boxplot()
 ```
@@
 
-<!-- ### Sum of ranks instead of product -->
-
-<!-- We performed the same tasks as we did for products -->
-
-<!-- Probabilities of advancing, winning medals are all lower -->
-
-<!-- Most notably for qualifications,  -->
-
-<!-- Obvious that the average score between the ranks are closer to each other -->
-
-<!-- The amount of variability doesn't seem to be different as the rank increases -->
-
-## Correlations
+## Rank Correlation
 
 For our analysis on the relationship between the rankings of the events and the final result, we used data from the 2018 Youth Olympics Women's Qualification.
-Figure 3 is a scatterplot and correlation matrix between the ranks of the individual events and the final standings, with Kendall’s Tau (Kendall Rank Correlation Coefficient) as our measure of ordinal association between the quantities.
-It is evidently clear that there is a strong and positive correlation between the ranks of bouldering and lead climbing, and as a results, the standings of these two events are highly correlated with the final rankings.
+Figure 2 is a scatterplot and correlation matrix between the ranks of the individual events and the final standings, with Kendall’s Tau (Kendall Rank Correlation Coefficient) as our measure of ordinal association between the quantities.
+It is evidently clear that there exists a strong and positive correlation between the ranks of bouldering and lead climbing, and as a results, the standings of these two events are highly correlated with the final rankings.
 On the other hand, the correlation with the final rank is not as strong for speed climbing.
 Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those that are specialized in the other two concentrations.
-
-This trend also holds for most of the past competitions.
+<!-- This trend also holds for most of the past competitions. -->
 
@@
           upper = list(continuous = wrap("cor", method = "kendall", stars = FALSE))) +
+  theme_bw() +
   theme(panel.grid.major = element_blank())
@@
 
+Another interesting question that we are interested in investigating is "What would happen to the rankings if a single climber is removed?".
+
+This illustrates the idea of IIA
+
+We once again make use of data from the 2018 Youth Olympics for this analysis, but this time we examine the final round of both men's and women's competitions.
+
+Figure 3 shows the plots
+
 A single climber excluded changes things drastically, especially order of medalists.
 
-The cases where someone behind you drops out and your ranking changes.
-
-Example from 2018 youth, women's final
-dropping ranks 3 and 5 change the medalist order
+The cases where someone behind you drops out and your ranking changes. Figure
 
@@
   labs(x = "Men", y = "") +
-  theme_minimal() +
   theme(axis.ticks = element_blank(),
@@
   labs(y = "\nScore", x = "Women") +
-  theme_minimal() +
   theme(axis.ticks = element_blank(),

### 2021-07-29 manuscript/Rmd/draft.tex
Update Rmd

@@
 \usepackage{hyperref}
-\usepackage{caption}
-\captionsetup[table]{skip=3pt}
-\usepackage{float}
 \providecommand{\tightlist}{%
@@
 
-The first sets of data that we rely on are historical results from major
-climbing competitions in the past. We collected data on climbing
-contests that took place between 2018 and 2020, where the combined
-format was used to determine the scores and ranks of climbers. The
-events include the 2020 Continental Championships of Europe, Africa,
-Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian
-Games; and 2018 Youth Olympics. Data were obtained from various sources,
-including the event websites, Wikipedia, and the International
-Federation of Sport Climbing (IFSC). The main attributes of our datasets
-are the name and nationality of the climbers; bib number (for some
-competitions); the finishing place of climbers in speed climbing,
-bouldering, and lead climbing; the total score (which equals the product
-of event ranks); and the final rank. We utilize this data to compute the
-correlations between the event ranks and final table position, as well
-as to look at how often the final orderings change if one athlete is
-dropped and the ranks for each discipline are re-computed.
-
-Our second data source comes from a simulation study that we conducted,
+Our first sets of data come from a simulation study that we conducted,
 with the purpose of examining the rankings and scoring for climbers in
@@
 \end{figure}
+
+Additionally, we rely historical results from major climbing
+competitions in recent years. We collected data on climbing contests
+that took place between 2018 and 2020, where the combined format was
+used to determine the scores and ranks of climbers. The events include
+the 2020 Continental Championships of Europe, Africa, Oceania,
+Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and
+2018 Youth Olympics. Data were obtained from various sources, including
+the event websites, Wikipedia, and the International Federation of Sport
+Climbing (IFSC). The main attributes of our datasets are the name and
+nationality of the climbers; bib number (for some competitions); the
+finishing place of climbers in speed climbing, bouldering, and lead
+climbing; the total score (which equals the product of event ranks); and
+the final rank. We utilize this data to compute the correlations between
+the event ranks and final table position, as well as to look at how
+often the final orderings change if one athlete is dropped and the ranks
+for each discipline are re-computed.
 
@@
 
-For the qualification round, a climber is almost guaranteed to make the
+In this section, we discuss the results of our simulations described in
+Section \ref{data-and-methods}. For the qualification round, our
+simulation study shows that a climber is almost guaranteed to make the
 final round if they win the first event (with a 99.51\% chance of
@@
 0.1885, respectively. In addition, the average score for qualification
-positions 1 to 8 are displayed in Table 1. We notice that on average,
+positions 1 to 8 are displayed in Table 2. We notice that on average,
 the minimum score that one should aim for in order to move on to the
-final round is 435 (for 8th rank).
+final round is about 435 (for 8th rank).
 
@@
 \centering
-\caption{Average score for each qualifying rank} 
+\caption{The average scores for the top 10 qualification ranks according to our simulations. A climber will secure a finalist spot if they finish in the top 8.} 
 \begin{tabular}{rr}
   \hline
-rank & avg\_adv\_score \\ 
+Rank & Average score \\ 
   \hline
@@
     8 & 434.59 \\ 
+    9 & 532.14 \\ 
+   10 & 642.33 \\ 
    \hline
@@
 
-Regarding the finals, a climber is very likely to earn a medal (finish
-in the top 3) if they win the first event (83.03\% chance) or any event
-(85.01\%). In order to obtain a climbing medal, the average score for
-getting gold, silver, and bronze are 9.6748, 20.4143, and 33.2648,
-respectively (Table 2).
+Regarding the finals, a climber is very likely to finish in the top 3
+and hence earn a medal if they win the first event (83.03\% chance) or
+any event (85.01\% chance). Furthermore, according to our final
+simulations, in order to obtain a climbing medal, the average scores
+that put an athlete in position to receive gold, silver, and bronze
+medals are 9.6748, 20.4143, and 33.2648, respectively (see Table 3).
 
@@
 \centering
-\caption{Average score of medalists} 
+\caption{The average scores for all final ranks according to our simulations. Ranks 1, 2, and 3 are table positions that guarantee medalist status for climbers.} 
 \begin{tabular}{rr}
   \hline
-rank & avg\_score \\ 
+Rank & Average score \\ 
   \hline
@@
 
-We would expect that if a climber with a high combined ranking is removed from the list of competitors, the placement of lower-ranked climbers will shift.
-In the most trivial case, if we were to remove the 3rd place climber, we would see 4th place climber shift to 3rd, 5th place climber shift to 4th, etc.
-It is maybe less trivial that when a climber is removed from the list, the placement of their higher-ranking competitors may also shift.
-In other words, there is a non-zero probability of seeing a change in the placement - particularly of medalists - regardless of the fact that no changes occurred in the performance of the remaining climbers.
-Real examples are given in section 4.2.
-In this section, we conduct simulations to analyze the probability of seeing a reordering of medalists when one non-medalist is removed from the list of finalists.
+We would expect that if a climber with a high combined ranking is
+removed from the list of competitors, the placement of lower-ranked
+climbers will shift. In the most trivial case, if we were to remove the
+3rd place climber, we would see 4th place climber shift to 3rd, 5th
+place climber shift to 4th, etc. It is maybe less trivial that when a
+climber is removed from the list, the placement of their higher-ranking
+competitors may also shift. In other words, there is a non-zero
+probability of seeing a change in the placement - particularly of
+medalists - regardless of the fact that no changes occurred in the
+performance of the remaining climbers. Real examples are given in
+section 4.2. In this section, we conduct simulations to analyze the
+probability of seeing a reordering of medalists when one non-medalist is
+removed from the list of finalists.
 
@@
 
-\hypertarget{correlations}{%
-\subsection{Correlations}\label{correlations}}
+\hypertarget{rank-correlation}{%
+\subsection{Rank Correlation}\label{rank-correlation}}
 
@@
 and the final result, we used data from the 2018 Youth Olympics Women's
-Qualification. Figure 3 is a scatterplot and correlation matrix between
+Qualification. Figure 2 is a scatterplot and correlation matrix between
 the ranks of the individual events and the final standings, with
@@
 ordinal association between the quantities. It is evidently clear that
-there is a strong and positive correlation between the ranks of
+there exists a strong and positive correlation between the ranks of
 bouldering and lead climbing, and as a results, the standings of these
@@
 
-This trend also holds for most of the past competitions.
-
-\begin{figure}[H]
+\begin{figure}
 \centering
-\includegraphics{draft_files/figure-latex/unnamed-chunk-10-1.pdf}
+\includegraphics{draft_files/figure-latex/unnamed-chunk-9-1.pdf}
 \caption{Kendall's rank correlations - 2018 World Championship, Women's
@@
 
+Another interesting question that we are interested in investigating is
+``What would happen to the rankings if a single climber is removed?''.
+
+This illustrates the idea of IIA
+
+We once again make use of data from the 2018 Youth Olympics for this
+analysis, but this time we examine the final round of both men's and
+women's competitions.
+
+Figure 3 shows the plots
+
 A single climber excluded changes things drastically, especially order
@@
 The cases where someone behind you drops out and your ranking changes.
-Example from 2018 youth, women's final dropping ranks 3 and 5 change the
-medalist order
+Figure
 
-\begin{figure}[H]
+\begin{figure}
 \centering
-\includegraphics{draft_files/figure-latex/unnamed-chunk-13-1.pdf}
+\includegraphics{draft_files/figure-latex/unnamed-chunk-12-1.pdf}
 \caption{This figure illustrates the changes to the final rankings of

### 2021-07-29 manuscript/Rmd/draft.Rmd
Add author info
Fix commit/merge issues

@@
 authors: 
-- name: Author 1
-  thanks: "Corresponding author: Name. Email:"
-  affiliation:
+- name: Quang Nguyen
+  affiliation: LUC
 
-- name: Author 2
-  thanks:
-  affiliation:
+- name: Hannah Butler
+  affiliation: CSU
 
-- name: Author 3
-  affiliation:
+- name: Gregory Matthews
+  affiliation: LUC
 
@@
 In addition, the average score for qualification positions 1 to 8 are displayed in Table 2.
-We notice that on average, the minimum score that one should aim for in order to move on to the final round is about 435 (for 8th rank).
+We notice that on average, the minimum score that one should aim for in order to move on to the final round is about 434 (for 8th rank).
 
@@
 Regarding the finals, a climber is very likely to finish in the top 3 and hence earn a medal if they win the first event (83.03% chance) or any event (85.01% chance).
-Furthermore, according to our final simulations, in order to obtain a climbing medal, the average scores that put an athlete in position to receive gold, silver, and bronze medals are 9.6748, 20.4143, and 33.2648, respectively (see Table 3).
+Furthermore, according to our final simulations, in order to obtain a climbing medal, the average scores (rounded down) that put an athlete in position to receive gold, silver, and bronze medals are 9, 20, and 33, respectively (see Table 3).
 
@@
 
-```{r, fig.cap = "This figure illustrates the changes to the final rankings of the 2018 Youth Olympics Men's and Women's Finals when we leave out a single climber.
+```{r, fig.cap = "This figure illustrates the changes to the final rankings of the 2018 Youth Olympics Men's and Women's Finals when we leave out one climber.
 For each gender competition, each panel represents the rank of the drop-out athlete, with 0 being the original final results.

### 2021-07-29 manuscript/Rmd/draft.tex
Add author info
Fix commit/merge issues

@@
   \author{
-        Author 1 \thanks{Corresponding author: Name. Email:} \\
-    \\
-     and \\     Author 2 \\
-    \\
-     and \\     Author 3 \\
-    \\
+        Quang Nguyen \\
+    LUC\\
+     and \\     Hannah Butler \\
+    CSU\\
+     and \\     Gregory Matthews \\
+    LUC\\
       }
@@
 the minimum score that one should aim for in order to move on to the
-final round is about 435 (for 8th rank).
+final round is about 434 (for 8th rank).
 
@@
 simulations, in order to obtain a climbing medal, the average scores
-that put an athlete in position to receive gold, silver, and bronze
-medals are 9.6748, 20.4143, and 33.2648, respectively (see Table 3).
+(rounded down) that put an athlete in position to receive gold, silver,
+and bronze medals are 9, 20, and 33, respectively (see Table 3).
 
@@
 \caption{This figure illustrates the changes to the final rankings of
-the 2018 Youth Olympics Men's and Women's Finals when we leave out a
-single climber. For each gender competition, each panel represents the
-rank of the drop-out athlete, with 0 being the original final results.
-Each case with a change in rank orderings is highlighted by a black
-panel border, and any player with a rank change is represented by a
-red-filled bar}
+the 2018 Youth Olympics Men's and Women's Finals when we leave out one
+climber. For each gender competition, each panel represents the rank of
+the drop-out athlete, with 0 being the original final results. Each case
+with a change in rank orderings is highlighted by a black panel border,
+and any player with a rank change is represented by a red-filled bar}
 \end{figure}

### 2021-07-30 manuscript/Rmd/draft.Rmd
Add bib file
Bed time

@@
 Data Analysis: drop and re-rank, correlations
-# bibliography: bib_example.bib
+bibliography: references.bib
 output: rticles::asa_article
@@
 In 2016, the International Olympic Committee announced the addition of five new sports to the 2020 Summer Olympics in Tokyo, Japan, which would then reschedule for 2021 due to the impact of the COVID-19 global pandemic.
-The five new features to Tokyo 2020's competitions program include baseball/softball, karate, skateboard, sports climbing, and surfing.
+The five new features to Tokyo 2020's competitions program include baseball/softball, karate, skateboard, sports climbing, and surfing [@ioc2016].
 One of the new sports, sport climbing, got our attention, specifically because of its unique scoring system and the fact that only one set of medals is awarded for each gender.
@@
 
-At the 2020 Summer Olympics, both men's and women's sport climbing competitions begin with 20 climbers who has previously qualified for the Olympics from qualifying events held in 2019 and 2020.
+The decision to combine the three climbing events and only award one set of medals for both men's and women's events in the Olympics has received a large amount of criticism from climbing athletes all over the world.
+In a series of interviews conducted by Climbing Magazine in 2016, a number of climbers shared their thoughts and concerns about the new Olympics climbing format.
+Climber Lynn Hill compared the idea of combining speed climbing, bouldering, and lead climbing to "asking a middle distance runner to compete in the sprint".
+She then added "Speed climbing is a sport within our sport".
+Other climbers also hold the same opinion as Hill regarding speed climbing, using words and phrases like "bogus", "a bummer", "less than ideal", "not in support", and "cheesy and unfair" to describe the new combined competition format.
+Courtney Woods stated "Speed climbers will have the biggest disadvantage because their realm isn’t based on difficult movements".
+Mike Doyle believed "Honestly, the people that will suffer the most are the ones that focus only on speed climbing.
+Those skills/abilities don’t transfer as well to the other disciplines".
+Most climbers also expressed their hope for a change in the format in future tournaments, with some calling for each discipline to have its own set of medals.
+See @blanchard2016 for more information on these interviews.
+
+In terms of participation and scoring, both sport climbing competitions for male and female at the 2020 Summer Olympics begin with 20 climbers who has previously qualified for the Olympics from qualifying events held in 2019 and 2020.
 All 20 athletes compete in each of the three disciplines in the qualification round, and their performances in each concentration are ranked from 1 to 20.
-A competitor's total score is then computed as the product of their ranks in the three events and the lower product is better; specifically,
+A competitor's combined score is computed as the product of their ranks in the three events; specifically,
 \begin{equation}
@@
 
-The top 8 finishers in the qualification round advance to the finals where they once again compete in all three of speed climbing, bouldering, and lead climbing.
-The total score for each climber in the final stage is determined by multiplying their ranks in each discipline, similar to the qualification round.
+The top 8 lowest-scoring climbers in the qualification round advance to the finals where they once again compete in all three of speed climbing, bouldering, and lead climbing.
+The total score for each person in the final stage is determined by multiplying their ranks in each discipline, similar to the qualification round.
 This implies that the climber with the lowest product of ranks in the final wins the gold medal.
@@
 
-Heavily criticized
-
 Other sports scoring methods
 
+https://www.climbing.com/competition/olympics/a-guide-to-the-olympic-climbing-format/
+
+Soccer points: @haugen2008
+
+Cross country scoring: @hammond2007, @boudreau2018
+
+Decathlon scoring: @westera2006
+
 The manuscript is outlined as follows.
-We first begin with some descriptions of the data and methods in Section 2.
-Our analyses and results are then presented in Sections 3 and 4.
-Finally, in Section 5, we summarize up our main findings and provide a discussion to close out the paper.
+We first begin with some descriptions of the data and methods in Section \ref{data-and-methods}.
+Our analyses and results are then presented in Sections \ref{simulations} and \ref{data-analysis}.
+Finally, in Section \ref{conclusion-and-discussion}, we summarize up our main findings and provide a discussion to close out the paper.
 
@@
 The simulation results allow to answer questions about various topics, including the distributions of scores for qualifying and final rounds, and the probabilities of advancing to the finals or winning a medal, given certain conditions.
-
-Table 1 and Figure 1 are numerical and visual summaries of climbing total score obtained from our simulated data for the qualification and final rounds.
+First of all, Table 1 and Figure 1 are numerical and visual summaries of climbing total score obtained from our simulated data for the qualification and final rounds.
 
@@
 library(tidyverse)
-library(knitr)
 library(xtable)
@@
 
+Brief paragraph on IIA
+
 # Simulations
@@
 For our analysis on the relationship between the rankings of the events and the final result, we used data from the 2018 Youth Olympics Women's Qualification.
-Figure 2 is a scatterplot and correlation matrix between the ranks of the individual events and the final standings, with Kendall’s Tau (Kendall Rank Correlation Coefficient) as our measure of ordinal association between the quantities.
+Figure 2 is a scatterplot and correlation matrix between the ranks of the individual events and the final standings, with Kendall’s Tau [@kendall1938] as our measure of ordinal association between the quantities.
 It is evidently clear that there exists a strong and positive correlation between the ranks of bouldering and lead climbing, and as a results, the standings of these two events are highly correlated with the final rankings.
@@
 ```{r}
+library(GGally)
+mq <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2018_youth_olympics/men_qual.csv")
 wq <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2018_youth_olympics/women_qual.csv")
-library(GGally)
 ```
 
-```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Women's Qualification"}
-wq %>% 
-  select(rank, speed, bould, lead) %>% 
+```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications", fig.height=8, fig.width=6}
+mcor <- mq %>% 
+  select(Overall = rank, Speed = speed, Bouldering = bould, Lead = lead) %>% 
   ggpairs(diag = "blank",
@@
           upper = list(continuous = wrap("cor", method = "kendall", stars = FALSE))) +
-  theme_bw() +
-  theme(panel.grid.major = element_blank())
+  theme(panel.grid.major = element_blank()) +
+  labs(y = "Men\n")
+
+wcor <- wq %>% 
+  select(Overall = rank, Speed = speed, Bouldering = bould, Lead = lead) %>% 
+  ggpairs(diag = "blank",
+          axisLabels = "internal",
+          upper = list(continuous = wrap("cor", method = "kendall", stars = FALSE))) +
+  theme(panel.grid.major = element_blank())  +
+  labs(y = "Women\n")
+
+library(cowplot)
+
+plot_grid(
+  ggmatrix_gtable(mcor),
+  ggmatrix_gtable(wcor),
+  nrow = 2
+)
 ```
@@
   expand_limits(y = 62, x = 0:7) +
-  scale_fill_manual(values = c("grey", "chocolate")) +
+  scale_fill_manual(values = c("grey", "red")) +
   labs(y = "\nScore", x = "Women") +
@@
 ```
+
+# Conclusion and Discussion
+
+Overall

### 2021-07-30 manuscript/Rmd/draft.tex
Add bib file
Bed time

@@
 include baseball/softball, karate, skateboard, sports climbing, and
-surfing. One of the new sports, sport climbing, got our attention,
-specifically because of its unique scoring system and the fact that only
-one set of medals is awarded for each gender.
+surfing \citep{ioc2016}. One of the new sports, sport climbing, got our
+attention, specifically because of its unique scoring system and the
+fact that only one set of medals is awarded for each gender.
 
@@
 
-At the 2020 Summer Olympics, both men's and women's sport climbing
-competitions begin with 20 climbers who has previously qualified for the
-Olympics from qualifying events held in 2019 and 2020. All 20 athletes
-compete in each of the three disciplines in the qualification round, and
-their performances in each concentration are ranked from 1 to 20. A
-competitor's total score is then computed as the product of their ranks
-in the three events and the lower product is better; specifically,
-\begin{equation}
+The decision to combine the three climbing events and only award one set
+of medals for both men's and women's events in the Olympics has received
+a large amount of criticism from climbing athletes all over the world.
+In a series of interviews conducted by Climbing Magazine in 2016, a
+number of climbers shared their thoughts and concerns about the new
+Olympics climbing format. Climber Lynn Hill compared the idea of
+combining speed climbing, bouldering, and lead climbing to ``asking a
+middle distance runner to compete in the sprint''. She then added
+``Speed climbing is a sport within our sport''. Other climbers also hold
+the same opinion as Hill regarding speed climbing, using words and
+phrases like ``bogus'', ``a bummer'', ``less than ideal'', ``not in
+support'', and ``cheesy and unfair'' to describe the new combined
+competition format. Courtney Woods stated ``Speed climbers will have the
+biggest disadvantage because their realm isn't based on difficult
+movements''. Mike Doyle believed ``Honestly, the people that will suffer
+the most are the ones that focus only on speed climbing. Those
+skills/abilities don't transfer as well to the other disciplines''. Most
+climbers also expressed their hope for a change in the format in future
+tournaments, with some calling for each discipline to have its own set
+of medals. See \citet{blanchard2016} for more information on these
+interviews.
+
+In terms of participation and scoring, both sport climbing competitions
+for male and female at the 2020 Summer Olympics begin with 20 climbers
+who has previously qualified for the Olympics from qualifying events
+held in 2019 and 2020. All 20 athletes compete in each of the three
+disciplines in the qualification round, and their performances in each
+concentration are ranked from 1 to 20. A competitor's combined score is
+computed as the product of their ranks in the three events;
+specifically, \begin{equation}
 Score_i = R^S_i\times R^B_i\times R^L_i,
@@
 
-The top 8 finishers in the qualification round advance to the finals
-where they once again compete in all three of speed climbing,
-bouldering, and lead climbing. The total score for each climber in the
+The top 8 lowest-scoring climbers in the qualification round advance to
+the finals where they once again compete in all three of speed climbing,
+bouldering, and lead climbing. The total score for each person in the
 final stage is determined by multiplying their ranks in each discipline,
@@
 
-Heavily criticized
-
 Other sports scoring methods
 
+\url{https://www.climbing.com/competition/olympics/a-guide-to-the-olympic-climbing-format/}
+
+Soccer points: \citet{haugen2008}
+
+Cross country scoring: \citet{hammond2007}, \citet{boudreau2018}
+
+Decathlon scoring: \citet{westera2006}
+
 The manuscript is outlined as follows. We first begin with some
-descriptions of the data and methods in Section 2. Our analyses and
-results are then presented in Sections 3 and 4. Finally, in Section 5,
-we summarize up our main findings and provide a discussion to close out
-the paper.
+descriptions of the data and methods in Section \ref{data-and-methods}.
+Our analyses and results are then presented in Sections
+\ref{simulations} and \ref{data-analysis}. Finally, in Section
+\ref{conclusion-and-discussion}, we summarize up our main findings and
+provide a discussion to close out the paper.
 
@@
 advancing to the finals or winning a medal, given certain conditions.
-
-Table 1 and Figure 1 are numerical and visual summaries of climbing
-total score obtained from our simulated data for the qualification and
-final rounds.
+First of all, Table 1 and Figure 1 are numerical and visual summaries of
+climbing total score obtained from our simulated data for the
+qualification and final rounds.
 
@@
 for each discipline are re-computed.
+
+Brief paragraph on IIA
 
@@
 the ranks of the individual events and the final standings, with
-Kendall's Tau (Kendall Rank Correlation Coefficient) as our measure of
-ordinal association between the quantities. It is evidently clear that
-there exists a strong and positive correlation between the ranks of
-bouldering and lead climbing, and as a results, the standings of these
-two events are highly correlated with the final rankings. On the other
-hand, the correlation with the final rank is not as strong for speed
-climbing. Thus, speed climbers are facing a huge disadvantage in this
-scoring system, compared to those that are specialized in the other two
-concentrations.
+Kendall's Tau \citep{kendall1938} as our measure of ordinal association
+between the quantities. It is evidently clear that there exists a strong
+and positive correlation between the ranks of bouldering and lead
+climbing, and as a results, the standings of these two events are highly
+correlated with the final rankings. On the other hand, the correlation
+with the final rank is not as strong for speed climbing. Thus, speed
+climbers are facing a huge disadvantage in this scoring system, compared
+to those that are specialized in the other two concentrations.
 
-\begin{figure}
+\begin{figure}[H]
 \centering
 \includegraphics{draft_files/figure-latex/unnamed-chunk-9-1.pdf}
-\caption{Kendall's rank correlations - 2018 World Championship, Women's
-Qualification}
+\caption{Kendall's rank correlations - 2018 World Championship, Men's
+and Women's Qualifications}
 \end{figure}
@@
 
+\hypertarget{conclusion-and-discussion}{%
+\section{Conclusion and Discussion}\label{conclusion-and-discussion}}
+
+Overall
+
 \bibliographystyle{agsm}
-\bibliography{}
+\bibliography{references.bib}
 

### 2021-07-30 manuscript/Rmd/draft.Rmd
Bed time

@@
 ```{r setup, include=FALSE}
+# install.packages(c("rticles", "xtable", "GGally", "patchwork", "cowplot"))
 knitr::opts_chunk$set(echo = FALSE, 
@@
 Overall
+
+# Supplementary Material {-}
+
+All of the materials related to this manuscript are publicly available on GitHub at \newline https://github.com/qntkhvn/climbing.

### 2021-07-30 manuscript/Rmd/draft.Rmd
Rise and shine

@@
 
-Other sports scoring methods
+<!-- # https://www.climbing.com/competition/olympics/a-guide-to-the-olympic-climbing-format/ -->
 
-https://www.climbing.com/competition/olympics/a-guide-to-the-olympic-climbing-format/
+Paragraph on scoring methods in other sports: Soccer points: @haugen2008; Cross country scoring: @hammond2007, @boudreau2018; Decathlon scoring: @westera2006
 
-Soccer points: @haugen2008
-
-Cross country scoring: @hammond2007, @boudreau2018
-
-Decathlon scoring: @westera2006
+As a side note, rank product, genes
 
@@
 
-Brief paragraph on IIA
+Paragraph on IIA/Social Choice
+
+Suppose we have 3 people A, B, and C participating in a competition.
+If A finishes in the first place and C is later disqualified and removed, A should still win.
+If the original winner (A) loses the modified competition (with C removed), then the Independence of Irrelevant Alternatives has been violated.
+
+Economics, psychology, probability theory. Connect with cross country in previous section
+
+First mentioned by Arrow (1951) https://cowles.yale.edu/sites/default/files/files/pub/mon/m12-all.pdf
+
+Luce (1959) http://www.scholarpedia.org/article/Luce%27s_choice_axiom
+
+and Luce (1977) https://www.imbs.uci.edu/files/personnel/luce/pre1990/1977/Luce_JMP_1977a.pdf
+
+Ray (1973) https://www.jstor.org/stable/1913820
 
@@
 In this section, we discuss the results of our simulations described in Section \ref{data-and-methods}.
-For the qualification round, our simulation study shows that a climber is almost guaranteed to make the final round if they win the first event (with a 99.51% chance of advancing) or if they win at least one of the three climbing concentrations (99.48%).
+For the qualification round, our simulation study shows that a climber is almost guaranteed to make the final round if they win the first event or if they win at least one of the three climbing concentrations  (99.51% and 99.48% change of advancing, respectively).
 On the other hand, finishing last in the first event or in any event would certainly hurt an athlete's chance of finishing in the top 8, as the probabilities of a climber advancing given they finish last in the first and in any event are 0.1830 and 0.1885, respectively.
@@
 
+Another interesting question that we are interested in investigating is "What would happen to the rankings if a single climber is removed?".
+This is related to the concept of independence of irrelevant alternatives (IIA), as mentioned in Section \ref{data-and-methods}.
+
 We would expect that if a climber with a high combined ranking is removed from the list of competitors, the placement of lower-ranked climbers will shift.
@@
 In other words, there is a non-zero probability of seeing a change in the placement - particularly of medalists - regardless of the fact that no changes occurred in the performance of the remaining climbers.
-Real examples are given in section 4.2.
+Real examples are given in Section \ref{data-analysis}.
 In this section, we conduct simulations to analyze the probability of seeing a reordering of medalists when one non-medalist is removed from the list of finalists.
+
+<!-- In what percentage of the (6!)^3 orderings would the winner change if one of the 6 competitors was dropped and there were only 5 instead of 6? -->
 
@@
 Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those that are specialized in the other two concentrations.
-<!-- This trend also holds for most of the past competitions. -->
+This trend also holds for most of the past competitions.
+(add table in the appendix)
 
@@
 
-```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications", fig.height=8, fig.width=6}
+```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications", fig.height=8, fig.width=6.5}
 mcor <- mq %>% 
@@
 
-Another interesting question that we are interested in investigating is "What would happen to the rankings if a single climber is removed?".
-
-This illustrates the idea of IIA
-
+Next, we perform analysis on the situations where a climber is dropped from the original standings.
 We once again make use of data from the 2018 Youth Olympics for this analysis, but this time we examine the final round of both men's and women's competitions.
 
-Figure 3 shows the plots
+Figure 3 shows the modified versions of the rankings after each ranked climber is excluded for both gender events.
+We have clear evidence from this plot that removing a single climber changes the rankings drastically, especially order of medalists.
+One particular interesting case is where an athlete's position change when someone who originally finished behind them drops out.
+This situation is illustrated by panel 5 of the women's competition, where the fifth-place climber, Krasovskaia, was excluded; and Meul, whose final rank was fourth in real life, moved up to the second spot and claim the silver medal.
 
-A single climber excluded changes things drastically, especially order of medalists.
-
-The cases where someone behind you drops out and your ranking changes. Figure
+Likewise, dropping a higher-ranked athlete could also result in a major shake-up to the new orderings of players.
+This type of ranking change is demonstrated by panel 1 and 2 of the men's facet.
+After dropping the first ranked male contestant, Dohi; Schenk, who previously finished fourth and outside of the medalist group now has the lowest score and become the gold medal winner.
+A similar situation occurs in the case where the second best competitor (Tanaka) was removed, which results in a jump from fourth place to second place for Pan.
 
@@
 
+Speed climbers
+
+There is a great dependence on irrelevent party.
+
 # Supplementary Material {-}

### 2021-07-30 manuscript/Rmd/draft.tex
Bed time
Rise and shine

@@
 
-Other sports scoring methods
+Paragraph on scoring methods in other sports: Soccer points:
+\citet{haugen2008}; Cross country scoring: \citet{hammond2007},
+\citet{boudreau2018}; Decathlon scoring: \citet{westera2006}
 
-\url{https://www.climbing.com/competition/olympics/a-guide-to-the-olympic-climbing-format/}
-
-Soccer points: \citet{haugen2008}
-
-Cross country scoring: \citet{hammond2007}, \citet{boudreau2018}
-
-Decathlon scoring: \citet{westera2006}
+As a side note, rank product, genes
 
@@
 
-Brief paragraph on IIA
+Paragraph on IIA/Social Choice
+
+Suppose we have 3 people A, B, and C participating in a competition. If
+A finishes in the first place and C is later disqualified and removed, A
+should still win. If the original winner (A) loses the modified
+competition (with C removed), then the Independence of Irrelevant
+Alternatives has been violated.
+
+Economics, psychology, probability theory. Connect with cross country in
+previous section
+
+First mentioned by Arrow (1951)
+\url{https://cowles.yale.edu/sites/default/files/files/pub/mon/m12-all.pdf}
+
+Luce (1959)
+\url{http://www.scholarpedia.org/article/Luce\%27s_choice_axiom}
+
+and Luce (1977)
+\url{https://www.imbs.uci.edu/files/personnel/luce/pre1990/1977/Luce_JMP_1977a.pdf}
+
+Ray (1973) \url{https://www.jstor.org/stable/1913820}
 
@@
 simulation study shows that a climber is almost guaranteed to make the
-final round if they win the first event (with a 99.51\% chance of
-advancing) or if they win at least one of the three climbing
-concentrations (99.48\%). On the other hand, finishing last in the first
+final round if they win the first event or if they win at least one of
+the three climbing concentrations (99.51\% and 99.48\% change of
+advancing, respectively). On the other hand, finishing last in the first
 event or in any event would certainly hurt an athlete's chance of
@@
 
+Another interesting question that we are interested in investigating is
+``What would happen to the rankings if a single climber is removed?''.
+This is related to the concept of independence of irrelevant
+alternatives (IIA), as mentioned in Section \ref{data-and-methods}.
+
 We would expect that if a climber with a high combined ranking is
@@
 performance of the remaining climbers. Real examples are given in
-section 4.2. In this section, we conduct simulations to analyze the
-probability of seeing a reordering of medalists when one non-medalist is
-removed from the list of finalists.
+Section \ref{data-analysis}. In this section, we conduct simulations to
+analyze the probability of seeing a reordering of medalists when one
+non-medalist is removed from the list of finalists.
 
@@
 climbers are facing a huge disadvantage in this scoring system, compared
-to those that are specialized in the other two concentrations.
+to those that are specialized in the other two concentrations. This
+trend also holds for most of the past competitions. (add table in the
+appendix)
 
-\begin{figure}[H]
+\begin{figure}
 \centering
@@
 
-Another interesting question that we are interested in investigating is
-``What would happen to the rankings if a single climber is removed?''.
+Next, we perform analysis on the situations where a climber is dropped
+from the original standings. We once again make use of data from the
+2018 Youth Olympics for this analysis, but this time we examine the
+final round of both men's and women's competitions.
 
-This illustrates the idea of IIA
+Figure 3 shows the modified versions of the rankings after each ranked
+climber is excluded for both gender events. We have clear evidence from
+this plot that removing a single climber changes the rankings
+drastically, especially order of medalists. One particular interesting
+case is where an athlete's position change when someone who originally
+finished behind them drops out. This situation is illustrated by panel 5
+of the women's competition, where the fifth-place climber, Krasovskaia,
+was excluded; and Meul, whose final rank was fourth in real life, moved
+up to the second spot and claim the silver medal.
 
-We once again make use of data from the 2018 Youth Olympics for this
-analysis, but this time we examine the final round of both men's and
-women's competitions.
+Likewise, dropping a higher-ranked athlete could also result in a major
+shake-up to the new orderings of players. This type of ranking change is
+demonstrated by panel 1 and 2 of the men's facet. After dropping the
+first ranked male contestant, Dohi; Schenk, who previously finished
+fourth and outside of the medalist group now has the lowest score and
+become the gold medal winner. A similar situation occurs in the case
+where the second best competitor (Tanaka) was removed, which results in
+a jump from fourth place to second place for Pan.
 
-Figure 3 shows the plots
-
-A single climber excluded changes things drastically, especially order
-of medalists.
-
-The cases where someone behind you drops out and your ranking changes.
-Figure
-
-\begin{figure}[H]
+\begin{figure}
 \centering
@@
 
+Speed climbers
+
+There is a great dependence on irrelevent party.
+
+\hypertarget{supplementary-material}{%
+\section*{Supplementary Material}\label{supplementary-material}}
+\addcontentsline{toc}{section}{Supplementary Material}
+
+All of the materials related to this manuscript are publicly available
+on GitHub at \newline \url{https://github.com/qntkhvn/climbing}.
+
 \bibliographystyle{agsm}

### 2021-07-30 manuscript/Rmd/draft.Rmd
Update Rmd
f

@@
 In 2016, the International Olympic Committee announced the addition of five new sports to the 2020 Summer Olympics in Tokyo, Japan, which would then reschedule for 2021 due to the impact of the COVID-19 global pandemic.
-The five new features to Tokyo 2020's competitions program include baseball/softball, karate, skateboard, sports climbing, and surfing [@ioc2016].
+The five new events added to Tokyo 2020's competition program were baseball/softball, karate, skateboard, sports climbing, and surfing [@ioc2016].
 One of the new sports, sport climbing, got our attention, specifically because of its unique scoring system and the fact that only one set of medals is awarded for each gender.
@@
 This trend also holds for most of the past competitions.
-(add table in the appendix)
+(add table to the appendix)
 
@@
 
-There is a great dependence on irrelevent party.
+There is a great dependence on irrelevant party.
+
+Future work
 

### 2021-07-30 manuscript/Rmd/draft.Rmd
Rmd

--- 2021-07-30 manuscript/Rmd/draft.Rmd
@@
 - name: Quang Nguyen
-  affiliation: LUC
+  affiliation: \textit{Loyola University Chicago}
 
 - name: Hannah Butler
-  affiliation: CSU
+  affiliation: \textit{Colorado State University}
 
-- name: Gregory Matthews
-  affiliation: LUC
+- name: Gregory J. Matthews
+  affiliation: \textit{Loyola University Chicago}
 
@@
 It is evidently clear that there exists a strong and positive correlation between the ranks of bouldering and lead climbing, and as a results, the standings of these two events are highly correlated with the final rankings.
-On the other hand, the correlation with the final rank is not as strong for speed climbing.
+On the other hand, the correlation with the final rank is not as strong for speed climbing as the other two events.
 Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those that are specialized in the other two concentrations.
-This trend also holds for most of the past competitions.
-(add table in the appendix)
+This validates the concerns of climbers in the interview mentioned in Section \ref{introduction}
+
+This trend also holds for most of the past competitions. (add table to the appendix)
 

### 2021-07-30 manuscript/Rmd/draft.Rmd
Some edits.

--- 2021-07-30 manuscript/Rmd/draft.Rmd
@@
 
-In 2016, the International Olympic Committee announced the addition of five new sports to the 2020 Summer Olympics in Tokyo, Japan, which would then reschedule for 2021 due to the impact of the COVID-19 global pandemic.
-The five new events added to Tokyo 2020's competition program were baseball/softball, karate, skateboard, sports climbing, and surfing [@ioc2016].
-One of the new sports, sport climbing, got our attention, specifically because of its unique scoring system and the fact that only one set of medals is awarded for each gender.
+In 2016, the International Olympic Committee (IOC) announced the addition of five new sports to the 2020 Summer Olympics in Tokyo, Japan, which would then reschedule for 2021 due to the impact of the COVID-19 global pandemic.
+The five new events added to Tokyo 2020's competition program were baseball/softball, karate, skateboard, surfing, sports climbing [@ioc2016].
+One of the new sports, sport climbing is particularly interesting because of its scoring system which uses the product of ranks across three disciplines to determine the medalists, with the lowest rank product declared the winner.
 
-Sport climbing at the 2020 Tokyo Olympics consists of three disciplines: speed climbing, bouldering, and lead climbing.
+#While there are three main disciplines in sport climbing, the IOC only allotted one set of medals per gender.
+So rather than choosing only a single discipline, they chose to combine all three of the events together to create a combined sport climbing event.
+
+The three disciplines that comprise sport climbing at the 2020 Tokyo Olympics are: speed climbing, bouldering, and lead climbing.
 Speed climbing takes place on a standardized course and competitors try to reach the top of the course as fast as possible.
@@
 Each lead climber only gets one attempt and when they fall their attempt is over.
+
 These three different climbing disciplines demand different sets of skills and, often, athletes specialize in a single event.
@@
 
-The decision to combine the three climbing events and only award one set of medals for both men's and women's events in the Olympics has received a large amount of criticism from climbing athletes all over the world.
+The decision to combine the three climbing events and only award one set of medals for both men's and women's events in the Olympics has received a large amount of criticism from climbing athletes all over the world (CITATION NEEDED).
 In a series of interviews conducted by Climbing Magazine in 2016, a number of climbers shared their thoughts and concerns about the new Olympics climbing format.

### 2021-07-30 manuscript/Rmd/draft.Rmd

@@
 It is evidently clear that there exists a strong and positive correlation between the ranks of bouldering and lead climbing, and as a results, the standings of these two events are highly correlated with the final rankings.
-On the other hand, the correlation with the final rank is not as strong for speed climbing.
+On the other hand, the correlation with the final rank is not as strong for speed climbing as the other two events.
 Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those that are specialized in the other two concentrations.
-This trend also holds for most of the past competitions.
-(add table to the appendix)
+This validates the concerns of climbers in the interview mentioned in Section \ref{introduction}
+
+This trend also holds for most of the past competitions. (add table to the appendix)
 

### 2021-07-30 manuscript/Rmd/draft.tex
Update Rmd

@@
         Quang Nguyen \\
-    LUC\\
+    \textit{Loyola University Chicago}\\
      and \\     Hannah Butler \\
-    CSU\\
-     and \\     Gregory Matthews \\
-    LUC\\
+    \textit{Colorado State University}\\
+     and \\     Gregory J. Matthews \\
+    \textit{Loyola University Chicago}\\
       }
@@
 correlated with the final rankings. On the other hand, the correlation
-with the final rank is not as strong for speed climbing. Thus, speed
-climbers are facing a huge disadvantage in this scoring system, compared
-to those that are specialized in the other two concentrations. This
-trend also holds for most of the past competitions. (add table in the
-appendix)
+with the final rank is not as strong for speed climbing as the other two
+events. Thus, speed climbers are facing a huge disadvantage in this
+scoring system, compared to those that are specialized in the other two
+concentrations. This validates the concerns of climbers in the interview
+mentioned in Section \ref{introduction}
+
+This trend also holds for most of the past competitions. (add table to
+the appendix)
 

### 2021-07-30 manuscript/Rmd/draft.Rmd
More edits.

@@
 
+[//]: (I think this paragraph and the next should get moved somewhere else (or removed)  Maybe in the conclusions?
+Definitely for the blog post include this, but I'm not sure how much of this needs to be in the actual article)
 These three different climbing disciplines demand different sets of skills and, often, athletes specialize in a single event.
@@
 In a series of interviews conducted by Climbing Magazine in 2016, a number of climbers shared their thoughts and concerns about the new Olympics climbing format.
-Climber Lynn Hill compared the idea of combining speed climbing, bouldering, and lead climbing to "asking a middle distance runner to compete in the sprint".
+Climber Lynn Hill compared the idea of combining speed climbing, bouldering, and lead climbing to "asking a middle distance runner to compete in the sprint" (CITATION NEEDED).
 She then added "Speed climbing is a sport within our sport".
@@
 
-In terms of participation and scoring, both sport climbing competitions for male and female at the 2020 Summer Olympics begin with 20 climbers who has previously qualified for the Olympics from qualifying events held in 2019 and 2020.
+At the Tokyo Olympics, both sport climbing competitions for male and female at the 2020 Summer Olympics begin with 20 climbers who has previously qualified for the Olympics from qualifying events held in 2019 and 2020.
 All 20 athletes compete in each of the three disciplines in the qualification round, and their performances in each concentration are ranked from 1 to 20.
@@
 
-The top 8 lowest-scoring climbers in the qualification round advance to the finals where they once again compete in all three of speed climbing, bouldering, and lead climbing.
-The total score for each person in the final stage is determined by multiplying their ranks in each discipline, similar to the qualification round.
-This implies that the climber with the lowest product of ranks in the final wins the gold medal.
+The 8 climbers with the lowest score (i.e.
+lowest product of ranks across the three discplines) in the qualification round advance to the finals where they once again compete in all three disciplines.
+The final score for each person in the final stage is determined by multiplying their ranks in each discipline, similar to the qualification round, with the main difference being that there are only 8 competitors in the final as opposed to 20 in the qualification round.
+The climbers with the lowest, second lowest, and third lowest product of ranks in the final wins the gold, silver, and bronze medal, respectively.
+
 This type of scoring system heavily rewards high finishes and relatively ignores poor finishes.

### 2021-07-30 manuscript/Rmd/draft.Rmd
Rmd before lunch

@@
 The five new events added to Tokyo 2020's competition program were baseball/softball, karate, skateboard, surfing, sports climbing [@ioc2016].
-One of the new sports, sport climbing is particularly interesting because of its scoring system which uses the product of ranks across three disciplines to determine the medalists, with the lowest rank product declared the winner.
+One of the new sports, sport climbing, is particularly interesting because of its scoring system which uses the product of ranks across three disciplines to determine the medalists, with the lowest rank product declared the winner.
 
@@
 
-The three disciplines that comprise sport climbing at the 2020 Tokyo Olympics are: speed climbing, bouldering, and lead climbing.
+The three disciplines that comprise sport climbing at Tokyo 2020 are: speed climbing, bouldering, and lead climbing.
 Speed climbing takes place on a standardized course and competitors try to reach the top of the course as fast as possible.
@@
 
-At the Tokyo Olympics, both sport climbing competitions for male and female at the 2020 Summer Olympics begin with 20 climbers who has previously qualified for the Olympics from qualifying events held in 2019 and 2020.
+At the 2020 Summer Olympics, both sport climbing competitions for male and female at the begin with 20 climbers who has previously qualified for the Olympics from qualifying events held in 2019 and 2020.
 All 20 athletes compete in each of the three disciplines in the qualification round, and their performances in each concentration are ranked from 1 to 20.
@@
 One particular interesting case is where an athlete's position change when someone who originally finished behind them drops out.
-This situation is illustrated by panel 5 of the women's competition, where the fifth-place climber, Krasovskaia, was excluded; and Meul, whose final rank was fourth in real life, moved up to the second spot and claim the silver medal.
+This situation is illustrated by panel 5 of the women's competition, where the fifth-place climber, Krasovskaia, was excluded; and Meul, whose actual final rank was fourth, moved up to the second spot and claim the silver medal.
 

### 2021-07-30 manuscript/Rmd/draft.Rmd
f

--- 2021-07-30 manuscript/Rmd/draft.Rmd
@@
 <!-- # https://www.climbing.com/competition/olympics/a-guide-to-the-olympic-climbing-format/ -->
+
+To the best of our knowledge, we know of no sporting event, team or individual, that uses the product of ranks to determine an overall rankings.
+There are examples of team sports that use the sum of ranks to determine the winning team such as cross country, where the team with the lowest sum of ranks of the top five runners is determined to be the winner.
+For more detail on rank sum scoring, see @hammond2007 and @boudreau2018.
 

### 2021-07-30 manuscript/Rmd/draft.tex

@@
 
-In 2016, the International Olympic Committee announced the addition of
-five new sports to the 2020 Summer Olympics in Tokyo, Japan, which would
-then reschedule for 2021 due to the impact of the COVID-19 global
-pandemic. The five new features to Tokyo 2020's competitions program
-include baseball/softball, karate, skateboard, sports climbing, and
-surfing \citep{ioc2016}. One of the new sports, sport climbing, got our
-attention, specifically because of its unique scoring system and the
-fact that only one set of medals is awarded for each gender.
+In 2016, the International Olympic Committee (IOC) announced the
+addition of five new sports to the 2020 Summer Olympics in Tokyo, Japan,
+which would then reschedule for 2021 due to the impact of the COVID-19
+global pandemic. The five new events added to Tokyo 2020's competition
+program were baseball/softball, karate, skateboard, surfing, sports
+climbing \citep{ioc2016}. One of the new sports, sport climbing, is
+particularly interesting because of its scoring system which uses the
+product of ranks across three disciplines to determine the medalists,
+with the lowest rank product declared the winner.
 
-Sport climbing at the 2020 Tokyo Olympics consists of three disciplines:
+\#While there are three main disciplines in sport climbing, the IOC only
+allotted one set of medals per gender. So rather than choosing only a
+single discipline, they chose to combine all three of the events
+together to create a combined sport climbing event.
+
+The three disciplines that comprise sport climbing at Tokyo 2020 are:
 speed climbing, bouldering, and lead climbing. Speed climbing takes
@@
 the highest point on the wall is the winner. Each lead climber only gets
-one attempt and when they fall their attempt is over. These three
-different climbing disciplines demand different sets of skills and,
-often, athletes specialize in a single event. However, since only one
-set of Olympic medals is awarded to sport climbing, rather than choosing
-only one of these disciplines to include in the Olympics, all three
-events were chosen to be included as a sort of climbing triathlon.
+one attempt and when they fall their attempt is over.
+
+These three different climbing disciplines demand different sets of
+skills and, often, athletes specialize in a single event. However, since
+only one set of Olympic medals is awarded to sport climbing, rather than
+choosing only one of these disciplines to include in the Olympics, all
+three events were chosen to be included as a sort of climbing triathlon.
 
@@
 of medals for both men's and women's events in the Olympics has received
-a large amount of criticism from climbing athletes all over the world.
-In a series of interviews conducted by Climbing Magazine in 2016, a
-number of climbers shared their thoughts and concerns about the new
-Olympics climbing format. Climber Lynn Hill compared the idea of
-combining speed climbing, bouldering, and lead climbing to ``asking a
-middle distance runner to compete in the sprint''. She then added
-``Speed climbing is a sport within our sport''. Other climbers also hold
-the same opinion as Hill regarding speed climbing, using words and
-phrases like ``bogus'', ``a bummer'', ``less than ideal'', ``not in
-support'', and ``cheesy and unfair'' to describe the new combined
-competition format. Courtney Woods stated ``Speed climbers will have the
-biggest disadvantage because their realm isn't based on difficult
-movements''. Mike Doyle believed ``Honestly, the people that will suffer
-the most are the ones that focus only on speed climbing. Those
-skills/abilities don't transfer as well to the other disciplines''. Most
-climbers also expressed their hope for a change in the format in future
-tournaments, with some calling for each discipline to have its own set
-of medals. See \citet{blanchard2016} for more information on these
-interviews.
+a large amount of criticism from climbing athletes all over the world
+(CITATION NEEDED). In a series of interviews conducted by Climbing
+Magazine in 2016, a number of climbers shared their thoughts and
+concerns about the new Olympics climbing format. Climber Lynn Hill
+compared the idea of combining speed climbing, bouldering, and lead
+climbing to ``asking a middle distance runner to compete in the sprint''
+(CITATION NEEDED). She then added ``Speed climbing is a sport within our
+sport''. Other climbers also hold the same opinion as Hill regarding
+speed climbing, using words and phrases like ``bogus'', ``a bummer'',
+``less than ideal'', ``not in support'', and ``cheesy and unfair'' to
+describe the new combined competition format. Courtney Woods stated
+``Speed climbers will have the biggest disadvantage because their realm
+isn't based on difficult movements''. Mike Doyle believed ``Honestly,
+the people that will suffer the most are the ones that focus only on
+speed climbing. Those skills/abilities don't transfer as well to the
+other disciplines''. Most climbers also expressed their hope for a
+change in the format in future tournaments, with some calling for each
+discipline to have its own set of medals. See \citet{blanchard2016} for
+more information on these interviews.
 
-In terms of participation and scoring, both sport climbing competitions
-for male and female at the 2020 Summer Olympics begin with 20 climbers
-who has previously qualified for the Olympics from qualifying events
-held in 2019 and 2020. All 20 athletes compete in each of the three
-disciplines in the qualification round, and their performances in each
-concentration are ranked from 1 to 20. A competitor's combined score is
-computed as the product of their ranks in the three events;
-specifically, \begin{equation}
+At the 2020 Summer Olympics, both sport climbing competitions for male
+and female at the begin with 20 climbers who has previously qualified
+for the Olympics from qualifying events held in 2019 and 2020. All 20
+athletes compete in each of the three disciplines in the qualification
+round, and their performances in each concentration are ranked from 1 to
+20. A competitor's combined score is computed as the product of their
+ranks in the three events; specifically, \begin{equation}
 Score_i = R^S_i\times R^B_i\times R^L_i,
@@
 
-The top 8 lowest-scoring climbers in the qualification round advance to
-the finals where they once again compete in all three of speed climbing,
-bouldering, and lead climbing. The total score for each person in the
-final stage is determined by multiplying their ranks in each discipline,
-similar to the qualification round. This implies that the climber with
-the lowest product of ranks in the final wins the gold medal. This type
-of scoring system heavily rewards high finishes and relatively ignores
-poor finishes. For instance, if climber A finished 1st, 20th, and 20th
-and climber B finished 10th, 10th, and 10th, climber B would have a
-score of 1000 whereas climber A would have a much better score of 400,
-despite finishing last in 2 out of 3 of the events.
+The 8 climbers with the lowest score (i.e.~lowest product of ranks
+across the three discplines) in the qualification round advance to the
+finals where they once again compete in all three disciplines. The final
+score for each person in the final stage is determined by multiplying
+their ranks in each discipline, similar to the qualification round, with
+the main difference being that there are only 8 competitors in the final
+as opposed to 20 in the qualification round. The climbers with the
+lowest, second lowest, and third lowest product of ranks in the final
+wins the gold, silver, and bronze medal, respectively.
+
+This type of scoring system heavily rewards high finishes and relatively
+ignores poor finishes. For instance, if climber A finished 1st, 20th,
+and 20th and climber B finished 10th, 10th, and 10th, climber B would
+have a score of 1000 whereas climber A would have a much better score of
+400, despite finishing last in 2 out of 3 of the events.
 
@@
 of the women's competition, where the fifth-place climber, Krasovskaia,
-was excluded; and Meul, whose final rank was fourth in real life, moved
-up to the second spot and claim the silver medal.
+was excluded; and Meul, whose actual final rank was fourth, moved up to
+the second spot and claim the silver medal.
 

### 2021-07-30 manuscript/Rmd/draft.Rmd
Added two articles.  And some edits.
Some edits.

@@
 For more detail on rank sum scoring, see @hammond2007 and @boudreau2018.
+
+Some individual sports such as the decathlon and heptathlon rely on a sum of scores from the ten or seven events, however, these scores are not determined based on the ranks of the competitiors.
+That is a decathlete's score is entirely based upon their times, distances, and heights, and their overall score will be exactly the same if the times, distances, and heights remain the same regardless of the performance of other individuals.
+Full details of dacathlon scoring can be found in @westera2006.
+
+There are other individual competitions consisted of several events combined that do base their scoring on ranks, such as crossfit competitions.
+In each event, points are earned based on the competitor's rank in the event based on a scoring table.
+(https://games.crossfit.com/announcement/12940), and a competitors final score is based on the sum of their scores across all the events.
 

### 2021-07-30 manuscript/Rmd/draft.tex
Break

@@
 400, despite finishing last in 2 out of 3 of the events.
+
+To the best of our knowledge, we know of no sporting event, team or
+individual, that uses the product of ranks to determine an overall
+rankings. There are examples of team sports that use the sum of ranks to
+determine the winning team such as cross country, where the team with
+the lowest sum of ranks of the top five runners is determined to be the
+winner. For more detail on rank sum scoring, see \citet{hammond2007} and
+\citet{boudreau2018}.
+
+Some individual sports such as the decathlon and heptathlon rely on a
+sum of scores from the ten or seven events, however, these scores are
+not determined based on the ranks of the competitiors. That is a
+decathlete's score is entirely based upon their times, distances, and
+heights, and their overall score will be exactly the same if the times,
+distances, and heights remain the same regardless of the performance of
+other individuals. Full details of dacathlon scoring can be found in
+\citet{westera2006}.
+
+There are other individual competitions consisted of several events
+combined that do base their scoring on ranks, such as crossfit
+competitions. In each event, points are earned based on the competitor's
+rank in the event based on a scoring table.
+(\url{https://games.crossfit.com/announcement/12940}), and a competitors
+final score is based on the sum of their scores across all the events.
 

### 2021-07-31 manuscript/Rmd/draft.Rmd
Break
Late night

@@
 ```{r setup, include=FALSE}
-# install.packages(c("rticles", "xtable", "GGally", "patchwork", "cowplot"))
+# install.packages(c("rticles", "kableExtra", "GGally", "patchwork", "cowplot"))
 knitr::opts_chunk$set(echo = FALSE, 
@@
                       fig.width = 5,
-                      results = "asis")
-options(xtable.comment = FALSE,
-        xtable.caption.placement = "top", 
-        xtable.include.rownames = FALSE)
+                      fig.align = "center")
+options(scipen = 999)
 ```
@@
 
-[//]: (I think this paragraph and the next should get moved somewhere else (or removed)  Maybe in the conclusions?
+<!-- 
+(I think this paragraph and the next should get moved somewhere else (or removed)  Maybe in the conclusions?
 Definitely for the blog post include this, but I'm not sure how much of this needs to be in the actual article)
@@
 However, since only one set of Olympic medals is awarded to sport climbing, rather than choosing only one of these disciplines to include in the Olympics, all three events were chosen to be included as a sort of climbing triathlon.
+-->
 
@@
 The 8 climbers with the lowest score (i.e.
-lowest product of ranks across the three discplines) in the qualification round advance to the finals where they once again compete in all three disciplines.
+lowest product of ranks across the three disciplines) in the qualification round advance to the finals where they once again compete in all three disciplines.
 The final score for each person in the final stage is determined by multiplying their ranks in each discipline, similar to the qualification round, with the main difference being that there are only 8 competitors in the final as opposed to 20 in the qualification round.
@@
 
-Some individual sports such as the decathlon and heptathlon rely on a sum of scores from the ten or seven events, however, these scores are not determined based on the ranks of the competitiors.
+Some individual sports such as the decathlon and heptathlon rely on a sum of scores from the ten or seven events, however, these scores are not determined based on the ranks of the competitors.
 That is a decathlete's score is entirely based upon their times, distances, and heights, and their overall score will be exactly the same if the times, distances, and heights remain the same regardless of the performance of other individuals.
-Full details of dacathlon scoring can be found in @westera2006.
+Full details of decathlon scoring can be found in @westera2006.
 
@@
 library(tidyverse)
-library(xtable)
+library(kableExtra)
 theme_set(theme_minimal())
@@
 mosaic::favstats(score ~ round, data = qual_final)[-10] %>% 
-  xtable(caption = "Descriptive statistics of simulated scores for qualification and final rounds") %>% 
-  print()
+  kable(caption = "Descriptive statistics of simulated scores for qualification and final rounds",
+        booktabs = TRUE)
 ```
@@
   filter(Rank <= 10) %>% 
-  xtable(caption = "The average scores for the top 10 qualification ranks according to our simulations. A climber will secure a finalist spot if they finish in the top 8.") %>% 
-  print()
+  kable(caption = "The average scores for the top 10 qualification ranks according to our simulations. A climber will secure a finalist spot if they finish in the top 8.",
+        booktabs = TRUE)
 ```
@@
   summarize(`Average score` = mean(score)) %>% 
-  xtable(caption = "The average scores for all final ranks according to our simulations. Ranks 1, 2, and 3 are table positions that guarantee medalist status for climbers.") %>% 
-  print()
+  kable(caption = "The average scores for all final ranks according to our simulations. Ranks 1, 2, and 3 are table positions that guarantee medalist status for climbers.",
+        booktabs = TRUE)
 ```
@@
 Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those that are specialized in the other two concentrations.
-This validates the concerns of climbers in the interview mentioned in Section \ref{introduction}
-
-This trend also holds for most of the past competitions. (add table to the appendix)
+This validates the concerns of climbers in the interview mentioned in Section \ref{introduction}.
+We also have evidence for these correlations from the qualification rounds of the 2018 Asian Games (Table 4), 2019 World Championship (Table 5), 2020 European Championship (Table 6), and 2020 Pan-American Championships (Table 7).
 
@@
 
+Recommendations: climbers were right, need a change in format
+
 # Supplementary Material {-}
@@
 All of the materials related to this manuscript are publicly available on GitHub at \newline https://github.com/qntkhvn/climbing.
+
+```{r}
+print_cor <- function(event = "", 
+                      gender = c("men_qual", "women_qual"),
+                      title = "") {
+  path <- "https://raw.githubusercontent.com/qntkhvn/climbing/main/data/"
+  men <- read_csv(paste0(path, event, "/", gender[1], ".csv")) %>% 
+    select(Overall = rank, Speed = speed, Bouldering = bould, Lead = lead) %>% 
+    drop_na() %>% 
+    cor(method = "kendall") %>% 
+    as_tibble() %>% 
+    round(4)
+  women <- read_csv(paste0(path, event, "/", gender[2], ".csv")) %>% 
+    select(Overall = rank, Speed = speed, Bouldering = bould, Lead = lead) %>% 
+    drop_na() %>% 
+    cor(method = "kendall") %>% 
+    as_tibble() %>% 
+    round(4)
+  out <- cbind(men, women)
+  rownames(out) <- c("Overall", "Speed", "Bouldering", "Lead")
+  kable(out, caption = title, booktabs = TRUE, position = "H") %>% 
+    add_header_above(c("", "Men" = 4, "Women" = 4)) %>%
+    kable_styling(latex_options = c("repeat_header", "scale_down"))
+}
+```
+
+```{r}
+print_cor("2018_asiad", title = "This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games")
+```
+
+```{r}
+print_cor("2019_world", title = "This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships")
+```
+
+```{r}
+print_cor("2020_europe", title = "This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship")
+```
+
+```{r}
+print_cor("2020_pan_american", title = "This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship")
+```

### 2021-07-31 manuscript/Rmd/draft.tex
Late night

@@
 \usepackage{hyperref}
+\usepackage{caption}
+\captionsetup[figure]{skip=3pt}
+\usepackage{float}
 \providecommand{\tightlist}{%
@@
 % Pandoc citation processing
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 
-These three different climbing disciplines demand different sets of
-skills and, often, athletes specialize in a single event. However, since
-only one set of Olympic medals is awarded to sport climbing, rather than
-choosing only one of these disciplines to include in the Olympics, all
-three events were chosen to be included as a sort of climbing triathlon.
-
 The decision to combine the three climbing events and only award one set
@@
 The 8 climbers with the lowest score (i.e.~lowest product of ranks
-across the three discplines) in the qualification round advance to the
+across the three disciplines) in the qualification round advance to the
 finals where they once again compete in all three disciplines. The final
@@
 sum of scores from the ten or seven events, however, these scores are
-not determined based on the ranks of the competitiors. That is a
+not determined based on the ranks of the competitors. That is a
 decathlete's score is entirely based upon their times, distances, and
@@
 distances, and heights remain the same regardless of the performance of
-other individuals. Full details of dacathlon scoring can be found in
+other individuals. Full details of decathlon scoring can be found in
 \citet{westera2006}.
@@
 
-\begin{table}[ht]
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-4}Descriptive statistics of simulated scores for qualification and final rounds}
 \centering
-\caption{Descriptive statistics of simulated scores for qualification and final rounds} 
-\begin{tabular}{lrrrrrrrr}
-  \hline
-round & min & Q1 & median & Q3 & max & mean & sd & n \\ 
-  \hline
-Qualification & 1.00 & 240.00 & 684.00 & 1638.00 & 8000.00 & 1158.84 & 1273.48 & 200000 \\ 
-  Final & 1.00 & 24.00 & 60.00 & 126.00 & 512.00 & 91.12 & 91.35 & 80000 \\ 
-   \hline
+\begin{tabular}[t]{lrrrrrrrr}
+\toprule
+round & min & Q1 & median & Q3 & max & mean & sd & n\\
+\midrule
+Qualification & 1 & 240 & 684 & 1638 & 8000 & 1158.84378 & 1273.47884 & 200000\\
+Final & 1 & 24 & 60 & 126 & 512 & 91.12077 & 91.35293 & 80000\\
+\bottomrule
 \end{tabular}
@@
 
-\begin{figure}
-\centering
-\includegraphics{draft_files/figure-latex/unnamed-chunk-5-1.pdf}
-\caption{Histogram of simulated scores for qualification and final
-rounds}
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-5-1}
+
+}
+
+\caption{Histogram of simulated scores for qualification and final rounds}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 
-\begin{table}[ht]
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-6}The average scores for the top 10 qualification ranks according to our simulations. A climber will secure a finalist spot if they finish in the top 8.}
 \centering
-\caption{The average scores for the top 10 qualification ranks according to our simulations. A climber will secure a finalist spot if they finish in the top 8.} 
-\begin{tabular}{rr}
-  \hline
-Rank & Average score \\ 
-  \hline
-  1 & 36.02 \\ 
-    2 & 73.61 \\ 
-    3 & 115.40 \\ 
-    4 & 162.23 \\ 
-    5 & 216.00 \\ 
-    6 & 278.16 \\ 
-    7 & 350.33 \\ 
-    8 & 434.59 \\ 
-    9 & 532.14 \\ 
-   10 & 642.33 \\ 
-   \hline
+\begin{tabular}[t]{rr}
+\toprule
+Rank & Average score\\
+\midrule
+1 & 36.0187\\
+2 & 73.6111\\
+3 & 115.3954\\
+4 & 162.2263\\
+5 & 216.0041\\
+\addlinespace
+6 & 278.1649\\
+7 & 350.3272\\
+8 & 434.5932\\
+9 & 532.1383\\
+10 & 642.3298\\
+\bottomrule
 \end{tabular}
@@
 
-\begin{table}[ht]
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-7}The average scores for all final ranks according to our simulations. Ranks 1, 2, and 3 are table positions that guarantee medalist status for climbers.}
 \centering
-\caption{The average scores for all final ranks according to our simulations. Ranks 1, 2, and 3 are table positions that guarantee medalist status for climbers.} 
-\begin{tabular}{rr}
-  \hline
-Rank & Average score \\ 
-  \hline
-  1 & 9.67 \\ 
-    2 & 20.41 \\ 
-    3 & 33.26 \\ 
-    4 & 50.59 \\ 
-    5 & 74.76 \\ 
-    6 & 110.05 \\ 
-    7 & 164.43 \\ 
-    8 & 265.78 \\ 
-   \hline
+\begin{tabular}[t]{rr}
+\toprule
+Rank & Average score\\
+\midrule
+1 & 9.6748\\
+2 & 20.4143\\
+3 & 33.2648\\
+4 & 50.5933\\
+5 & 74.7620\\
+\addlinespace
+6 & 110.0536\\
+7 & 164.4274\\
+8 & 265.7760\\
+\bottomrule
 \end{tabular}
@@
 concentrations. This validates the concerns of climbers in the interview
-mentioned in Section \ref{introduction}
+mentioned in Section \ref{introduction}. We also have evidence for these
+correlations from the qualification rounds of the 2018 Asian Games
+(Table 4), 2019 World Championship (Table 5), 2020 European Championship
+(Table 6), and 2020 Pan-American Championships (Table 7).
 
-This trend also holds for most of the past competitions. (add table to
-the appendix)
+\begin{figure}[H]
 
-\begin{figure}
-\centering
-\includegraphics{draft_files/figure-latex/unnamed-chunk-9-1.pdf}
-\caption{Kendall's rank correlations - 2018 World Championship, Men's
-and Women's Qualifications}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-9-1}
+
+}
+
+\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-9}
 \end{figure}
@@
 
-\begin{figure}
-\centering
-\includegraphics{draft_files/figure-latex/unnamed-chunk-12-1.pdf}
-\caption{This figure illustrates the changes to the final rankings of
-the 2018 Youth Olympics Men's and Women's Finals when we leave out one
-climber. For each gender competition, each panel represents the rank of
-the drop-out athlete, with 0 being the original final results. Each case
-with a change in rank orderings is highlighted by a black panel border,
-and any player with a rank change is represented by a red-filled bar}
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-12-1}
+
+}
+
+\caption{This figure illustrates the changes to the final rankings of the 2018 Youth Olympics Men's and Women's Finals when we leave out one climber.
+For each gender competition, each panel represents the rank of the drop-out athlete, with 0 being the original final results.
+Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-12}
 \end{figure}
@@
 
+Recommendations: climbers were right, need a change in format
+
 \hypertarget{supplementary-material}{%
@@
 
+\newpage
+
 \bibliographystyle{agsm}
@@
 
+\newpage
+
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-14}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrrrrrrrr}
+\toprule
+\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
+\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
+  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
+\midrule
+Overall & 1.0000 & 0.5731 & 0.6760 & 0.6126 & 1.0000 & 0.5474 & 0.6878 & 0.7053\\
+Speed & 0.5731 & 1.0000 & 0.3022 & 0.2174 & 0.5474 & 1.0000 & 0.2540 & 0.2947\\
+Bouldering & 0.6760 & 0.3022 & 1.0000 & 0.6203 & 0.6878 & 0.2540 & 1.0000 & 0.6878\\
+Lead & 0.6126 & 0.2174 & 0.6203 & 1.0000 & 0.7053 & 0.2947 & 0.6878 & 1.0000\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-15}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrrrrrrrr}
+\toprule
+\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
+\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
+  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
+\midrule
+Overall & 1.0000 & 0.2632 & 0.3684 & 0.3474 & 1.0000 & 0.2316 & 0.5013 & 0.2526\\
+Speed & 0.2632 & 1.0000 & -0.0947 & -0.3263 & 0.2316 & 1.0000 & 0.0686 & -0.4526\\
+Bouldering & 0.3684 & -0.0947 & 1.0000 & 0.2000 & 0.5013 & 0.0686 & 1.0000 & 0.2902\\
+Lead & 0.3474 & -0.3263 & 0.2000 & 1.0000 & 0.2526 & -0.4526 & 0.2902 & 1.0000\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-16}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrrrrrrrr}
+\toprule
+\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
+\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
+  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
+\midrule
+Overall & 1.0000 & 0.0947 & 0.3810 & 0.2632 & 1.0000 & 0.1847 & 0.5358 & 0.5368\\
+Speed & 0.0947 & 1.0000 & -0.1587 & -0.4316 & 0.1847 & 1.0000 & -0.1915 & -0.0897\\
+Bouldering & 0.3810 & -0.1587 & 1.0000 & 0.2751 & 0.5358 & -0.1915 & 1.0000 & 0.5040\\
+Lead & 0.2632 & -0.4316 & 0.2751 & 1.0000 & 0.5368 & -0.0897 & 0.5040 & 1.0000\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-17}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrrrrrrrr}
+\toprule
+\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
+\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
+  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
+\midrule
+Overall & 1.0000 & 0.5906 & 0.5556 & 0.6491 & 1.0000 & 0.5789 & 0.6140 & 0.5906\\
+Speed & 0.5906 & 1.0000 & 0.2164 & 0.4035 & 0.5789 & 1.0000 & 0.2632 & 0.3099\\
+Bouldering & 0.5556 & 0.2164 & 1.0000 & 0.2749 & 0.6140 & 0.2632 & 1.0000 & 0.3918\\
+Lead & 0.6491 & 0.4035 & 0.2749 & 1.0000 & 0.5906 & 0.3099 & 0.3918 & 1.0000\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
 \end{document}

### 2021-07-31 manuscript/Rmd/draft.Rmd
Late night

@@
   kable(caption = "Descriptive statistics of simulated scores for qualification and final rounds",
-        booktabs = TRUE)
+        booktabs = TRUE,
+        position = "H")
 ```
@@
   kable(caption = "The average scores for the top 10 qualification ranks according to our simulations. A climber will secure a finalist spot if they finish in the top 8.",
-        booktabs = TRUE)
+        booktabs = TRUE,
+        position = "H")
 ```

### 2021-08-01 R/LOOA_real.R
Animation

@@ -70,23 +70,33 @@ a <- rerank_df %>%
 animate(a, nframes = 100, fps = 20, height = 480, width = 600, res = 95)
 
 
-# Women qualification
+library(cowplot)
 
-women_qual <- read_csv("data/youth_olympics_2018/women_qual.csv")
-rerank <- list()
-for (i in 1:nrow(women_qual)){
-  rerank[[i]] <- women_qual[-i,] %>%
-    mutate(rank_drop = i)
-}
+a <- drop_rerank(wf) %>% 
+  ggplot(aes(x = rank_drop, y = rank, color = last)) +
+  geom_line(aes(color = last, alpha = 1), size = 2) +
+  geom_point(aes(color = last, alpha = 1), size = 4) +
+  # geom_text(aes(x = 7, label = last), hjust = -0.1, alpha = 0.7, size = 6) +
+  # geom_text(aes(x = rank_drop, y = 6.5, label = as.character(rank_drop)),
+  #           alpha = 0.5,  col = "gray", size = 10) +
+  # geom_segment(aes(xend = max(rank_drop), yend = rank), linetype = 2, size = 0.7) +
+  # geom_text(aes(x = 0, y = 0, label = paste("Leaving out climber with rank", floor(rank_drop))), hjust = 0, size = 5.5) +
+  scale_y_reverse(breaks = 1:6) + 
+  scale_x_continuous(breaks = 0:6) +
+  theme_minimal_grid(font_size = 14, line_size = 0) +
+  theme(
+    legend.position = "none",
+    # axis.title = element_blank(),
+    # axis.text = element_blank(),
+    panel.grid.minor = element_blank(),
+    panel.grid.major = element_blank(),
+  ) +
+  transition_reveal(rank_drop) +
+  enter_fade() +
+  exit_fade() +
+  labs(x = "Excluded Rank",
+       y = "Final Rank",
+       title = "Leave-one-climber-out Modified Rankings\n2018 Youth Olympics - Women's Event")
 
-rerank_df <- women_qual %>% 
-  mutate(rank_drop = 0) %>% 
-  bind_rows(rerank) %>% 
-  group_by(rank_drop) %>% 
-  mutate(speed = rank(speed),
-         bould = rank(bould),
-         lead = rank(lead),
-         total = speed * bould * lead) %>% 
-  arrange(total, .by_group = TRUE) %>% 
-  mutate(rank = row_number()) %>% 
-  ungroup()
+
+animate(a, res = 90, duration = 6, width = 550)

### 2021-08-02 manuscript/Rmd/draft.Rmd
Paper abstract

@@
 - sport climbing
-- scoring system
+- rankings
 - 2020 Summer Olympics
@@
 abstract: |
-  The purpose of this paper is to investigate sport climbing, one of the sports making its debut on the Olympics stage at Tokyo 2020.
-In particular, we take a closer look at the controversial competition format and scoring system of sport climbing...
-Simulation.
-Data Analysis: drop and re-rank, correlations
+  The purpose of this paper is to investigate the controversial competition format and scoring system of sport climbing, which is one of the sports making its debut on the Olympics stage at Tokyo 2020.
+Climbing Olympians will participate in three disciplines (speed climbing, bouldering, and lead climbing) for a single set of medals, and a multiplied score of climbers' rankings in all three disciplines will be used to determine the overall standings.
+This work finds great evidence from historical data and simulations that speed climbing speed climbing should be separated from the combined format and have its own medals.
+Furthermore, it is shown that the product of ranks scoring system violates the independence of irrelevant alternatives property.
+
 bibliography: references.bib
@@
 
-The decision to combine the three climbing events and only award one set of medals for both men's and women's events in the Olympics has received a large amount of criticism from climbing athletes all over the world (CITATION NEEDED).
+The decision to combine the three climbing events and only award one set of medals for both men's and women's events in the Olympics has received a large amount of criticism from climbing athletes all over the world (CITATION NEEDED - I mentioned the article at the end of the paragrah).
 In a series of interviews conducted by Climbing Magazine in 2016, a number of climbers shared their thoughts and concerns about the new Olympics climbing format.

### 2021-08-02 manuscript/Rmd/draft.tex
Paper abstract

@@
 \usepackage{hyperref}
-\usepackage{caption}
-\captionsetup[figure]{skip=3pt}
-\usepackage{float}
 \providecommand{\tightlist}{%
@@
 \begin{abstract}
-The purpose of this paper is to investigate sport climbing, one of the
-sports making its debut on the Olympics stage at Tokyo 2020. In
-particular, we take a closer look at the controversial competition
-format and scoring system of sport climbing\ldots{} Simulation. Data
-Analysis: drop and re-rank, correlations
+The purpose of this paper is to investigate the controversial
+competition format and scoring system of sport climbing, which is one of
+the sports making its debut on the Olympics stage at Tokyo 2020.
+Climbing Olympians will participate in three disciplines (speed
+climbing, bouldering, and lead climbing) for a single set of medals, and
+a multiplied score of climbers' rankings in all three disciplines will
+be used to determine the overall standings. This work finds great
+evidence from historical data and simulations that speed climbing speed
+climbing should be separated from the combined format and have its own
+medals. Furthermore, it is shown that the product of ranks scoring
+system violates the independence of irrelevant alternatives property.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} sport climbing, scoring system, 2020 Summer Olympics
+{\it Keywords:} sport climbing, rankings, 2020 Summer Olympics
 \vfill
@@
 a large amount of criticism from climbing athletes all over the world
-(CITATION NEEDED). In a series of interviews conducted by Climbing
-Magazine in 2016, a number of climbers shared their thoughts and
-concerns about the new Olympics climbing format. Climber Lynn Hill
-compared the idea of combining speed climbing, bouldering, and lead
-climbing to ``asking a middle distance runner to compete in the sprint''
-(CITATION NEEDED). She then added ``Speed climbing is a sport within our
-sport''. Other climbers also hold the same opinion as Hill regarding
-speed climbing, using words and phrases like ``bogus'', ``a bummer'',
-``less than ideal'', ``not in support'', and ``cheesy and unfair'' to
-describe the new combined competition format. Courtney Woods stated
-``Speed climbers will have the biggest disadvantage because their realm
-isn't based on difficult movements''. Mike Doyle believed ``Honestly,
-the people that will suffer the most are the ones that focus only on
-speed climbing. Those skills/abilities don't transfer as well to the
-other disciplines''. Most climbers also expressed their hope for a
-change in the format in future tournaments, with some calling for each
-discipline to have its own set of medals. See \citet{blanchard2016} for
-more information on these interviews.
+(CITATION NEEDED - I mentioned the article at the end of the paragrah).
+In a series of interviews conducted by Climbing Magazine in 2016, a
+number of climbers shared their thoughts and concerns about the new
+Olympics climbing format. Climber Lynn Hill compared the idea of
+combining speed climbing, bouldering, and lead climbing to ``asking a
+middle distance runner to compete in the sprint'' (CITATION NEEDED). She
+then added ``Speed climbing is a sport within our sport''. Other
+climbers also hold the same opinion as Hill regarding speed climbing,
+using words and phrases like ``bogus'', ``a bummer'', ``less than
+ideal'', ``not in support'', and ``cheesy and unfair'' to describe the
+new combined competition format. Courtney Woods stated ``Speed climbers
+will have the biggest disadvantage because their realm isn't based on
+difficult movements''. Mike Doyle believed ``Honestly, the people that
+will suffer the most are the ones that focus only on speed climbing.
+Those skills/abilities don't transfer as well to the other
+disciplines''. Most climbers also expressed their hope for a change in
+the format in future tournaments, with some calling for each discipline
+to have its own set of medals. See \citet{blanchard2016} for more
+information on these interviews.
 
@@
 
-\begin{figure}[H]
+\begin{figure}
 
@@
 
-\begin{table}[H]
+\begin{table}
 
@@
 on GitHub at \newline \url{https://github.com/qntkhvn/climbing}.
-
-\newpage
-
-\bibliographystyle{agsm}
-\bibliography{references.bib}
-
-\newpage
 
@@
 
+\bibliographystyle{agsm}
+\bibliography{references.bib}
+
 \end{document}

### 2021-08-03 R/remove_climber_sim.Rmd
update remove-climber simulations rmd

@@
 ---
-title: "Remove Climber"
+title: "How Removing a Ranked Climber Changes the Winners"
 author: "Hannah Butler"
@@
 load("medal_displacement_probabilities.rdata")
+
+##### functions #####
+assign_ranks <- function(c = 8) {
+  # randomly assign ranks in 3 events & compute total
+  # ranking in an event is considered to be independent from ranking in another event
+  output <- data.frame(climber = sample(LETTERS, c, replace = FALSE)
+                       , speed = sample(1:c, c, replace = FALSE)
+                       , bould = sample(1:c, c, replace = FALSE)
+                       , lead = sample(1:c, c, replace = FALSE)
+                       ) %>%
+    mutate(total = speed*bould*lead) %>%
+    arrange(total) %>% # some questionable tie-breaking code
+    mutate(m_rank_speed = ifelse(total %in% total[1:3] , rank(speed[1:3]), 0)
+           , m_rank_bould = ifelse(total %in% total[1:3] , rank(bould[1:3]), 0)
+           , m_rank_lead = ifelse(total %in% total[1:3] , rank(lead[1:3]), 0)
+           , tiebreaker = m_rank_speed + m_rank_bould + m_rank_lead
+           ) %>%
+    arrange(total
+            , tiebreaker
+            )
+
+  return(output)
+}
+
+drop_climber <- function(results, r = 4) {
+  # drop a competitor and recompute rankings and totals
+  # r = which rank competitor to drop
+  ud_results <- results[-r, ] %>%
+    mutate(speed = rank(speed)
+           , bould = rank(bould)
+           , lead = rank(lead)
+           , total = speed*bould*lead
+           ) %>%
+    arrange(total)%>%
+    mutate(m_rank_speed = ifelse(total %in% total[1:3], rank(speed[1:3]), 0)
+           , m_rank_bould = ifelse(total %in% total[1:3], rank(bould[1:3]), 0)
+           , m_rank_lead = ifelse(total %in% total[1:3], rank(lead[1:3]), 0)
+           , tiebreaker = m_rank_speed + m_rank_bould + m_rank_lead
+           ) %>%
+    arrange(total
+            , tiebreaker
+            )
+  return(ud_results)
+}
+
+displace_medalist <- function(nsim = 1000, drop_rank = 4, nc = 8) {
+  # simulate nsim rankings
+  rank_perms <- lapply(1:nsim, function(x) assign_ranks(c = nc))
+  # drop ith competitor and recalculate ranks
+  dropout_ranks <- lapply(rank_perms, function(x) drop_climber(x, r = drop_rank))
+  # compare 1st-3rd place & calculate % not matched
+  pct_displ <- sapply(1:nsim, 
+                      function(x) {
+                        sum(rank_perms[[x]]$climber[1:3] != dropout_ranks[[x]]$climber[1:3]) > 0
+                      }
+                      ) %>%
+    sum()/nsim
+
+  return(pct_displ)
+}
+
+displace_climber <- function(nsim = 1000, drop_rank = 4, nc = 8) {
+  # simulate nsim rankings
+  rank_perms <- lapply(1:nsim, function(x) assign_ranks(c = nc))
+  # drop ith competitor and recalculate ranks
+  dropout_ranks <- lapply(rank_perms, function(x) drop_climber(x, r = drop_rank))
+  # compare 1st - (r-1)th place & calculate % not matched
+  pct_displ <- sapply(1:nsim, 
+                      function(x) {
+                        sum(rank_perms[[x]]$climber[1:(drop_rank - 1)] != dropout_ranks[[x]]$climber[1:(drop_rank - 1)]) > 0
+                      }
+                      ) %>%
+    sum()/nsim
+
+  return(pct_displ)
+}
 ```
 
-## Intro
+# Intro
 A scoring system that multiplies the ranks across events is interesting because it does not possess the property that it is invariant to reordering when a competitor is removed.
 Examples of this can be found in existing combined climbing event data, where when a climber with a total rank of $r > 1$, higher ranks change order.
+This has to do with the Independence of Irrelevant Alternatives (IIA) in Voting Systems, which will be discussed later.
 
@@
 
-## Methodology
+## Independence of Irrelevant Alternatives
+
+# Methodology
+
 As it is the relevant upcoming event, we will focus the parameters of the simulation on those of sport climbing 2020 Tokyo Olympics (taking place in August, 2021).
@@
 
-## Results
+# Results
+
+## Probability of Reordering in Higher Ranks
+We first look at the rate at which the ranks above that of which is dropped are reordered.
+```{r, echo = FALSE}
+load("reorder_all_above_probs.rdata")
+
+sum_dat1 <- reorder_all_above %>%
+  group_by(dropped_rank) %>%
+  rename("Dropped Rank" = dropped_rank) %>%
+  summarize("Simulations" = n()
+            , "Avg. Probability" = mean(displ_prob)
+            , "Median Probability" = median(displ_prob)
+            , "St. Deviation" = sd(displ_prob)
+            )
+
+knitr::kable(sum_dat1)
+```
+
+## Probability of Change in Finalists
+
+A potential event of interest is the replacement of one finalist with another, taking the chance away from the original finalist to win a medal in the finals round.
+Here we compute the probability that a finalist will be replaced with a non-finalist after removing another non-finalist from the rankings.
+Below is a table summarizing the results.
+
+```{r, echo=FALSE}
+load("drop_qualifier_sims.rdata")
+
+reorder_dat2 <- lapply(9:20
+                  , function(x) {
+                    data.frame(dropped_rank = x
+                               , displ_prob = pct_displace_finalists[[x-8]]
+                               )
+                    }
+                  ) %>%
+  do.call(rbind, .)
+
+sum_dat2 <- reorder_dat2 %>%
+  group_by(dropped_rank) %>%
+  rename("Dropped Rank" = dropped_rank) %>%
+  summarize("Simulations" = n()
+            , "Avg. Probability" = mean(displ_prob)
+            , "Median Probability" = median(displ_prob)
+            , "St. Deviation" = sd(displ_prob)
+            )
+
+knitr::kable(sum_dat2)
+
+```
+
+as can be seen, the probability is relatively low when the 9th place competitor is removed.
+My hypothesis is that there is enough separation between the original finalists and the 10th+ ranking competitors in the individual events that there isn't a high chance that the total rankings of the finalists can be overtaken by 10+th place competitors.
+Note: it might be worth computing the probability of getting a low ranking in any of the individual events, given that a competitor is a finalist.
+
+## Probability that Medalists are Reordered or Changed
+
+The second scenario of interest is if a medalist is displaced by another medalist, or even a non-medalist.
+Below are the results summarizing the probability of this scenario occurring under uniform random ranking conditions.
 
@@
   rename("Dropped Rank" = dropped_rank) %>%
-  summarize("# Simulations" = n()
-            , "Mean Probability" = mean(displ_prob)
-            , "Standard Deviation" = sd(displ_prob)
+  summarize("Simulations" = n()
+            , "Avg. Probability" = mean(displ_prob)
+            , "Median Probability" = median(displ_prob)
+            , "St. Deviation" = sd(displ_prob)
             ) %>%
@@
 
-On average, regardless of which rank competitor is removed, there is a significant chance that the medalists will permute or change. This can be refined to specific medalist placement
+# prob of displacing medalists as list of competitors grows
+```{r, eval=FALSE, include=FALSE}
+set.seed(80085)
+remove_c4 <- lapply(5:20, function(x) replicate(100, displace_climber(nsim = 100, drop_rank = 4, nc = x)))
 
-```{r, echo = FALSE}
-displacement_probs %>%
-  group_by(medal, dropped_rank) %>%
-  rename("Medal" = medal
-         , "Dropped Rank" = dropped_rank) %>%
-  summarize("# Simulations" = n()
-            , "Mean Probability" = mean(displ_prob)
-            , "Standard_Deviation" = sd(displ_prob)
+remove_c4_df <- lapply(5:20
+                  , function(x) {
+                    data.frame(n_climbers = x
+                               , dropped_rank = 4
+                               , displ_prob = remove_c4[[x-4]]
+                               )
+                    }
                   ) %>%
-  knitr::kable()
+  do.call(rbind, .)
+#save(remove_c4_df, "R/remove_c4_sims.rdata")
+
+remove_c5 <- lapply(5:20, function(x) replicate(100, displace_climber(nsim = 100, drop_rank = 5, nc = x)))
+
+remove_c5_df <- lapply(5:20
+                  , function(x) {
+                    data.frame(n_climbers = x
+                               , dropped_rank = 5
+                               , displ_prob = remove_c4[[x-4]]
+                               )
+                    }
+                  ) %>%
+  do.call(rbind, .)
+#save(remove_c5_df, "R/remove_c5_sims.rdata")
 ```
 
-### How often would a finalist be replaced
-### How often does someone who wasn't a medalist before become a medalist 
-### How often do we see a flip (two consecutive climbers switch places), jump (a climber jumps up more than one rank), drop (a climber drops down more than one rank)
+### Removing 4th Place
+```{r, echo = FALSE}
+load("remove_c4_sims.rdata")
+remove_c4_df %>%
+  ggplot(aes(x = n_climbers, y = displ_prob)) +
+  geom_boxplot(aes(group = n_climbers)) +
+  theme_minimal() +
+  ylab("Probability of Displacing Top 3") +
+  xlab("Number of Original Climbers")
+```
+
+# Questions: 
+- How often would a finalist be replaced
+- How often does someone who wasn't a medalist before become a medalist 
+- How often do we see a flip (two consecutive climbers switch places), jump (a climber jumps up more than one rank), drop (a climber drops down more than one rank)
+- how do the simulations compare to doing the same with actual data?
+
+# Appendix: R Functions
+
+**Randomly generate ranks for 3 events**
+```{r, eval=FALSE}
+assign_ranks <- function(c = 8) {
+  # randomly assign ranks in 3 events & compute total
+  # ranking in an event is considered to be independent from ranking in another event
+  output <- data.frame(climber = sample(LETTERS, c, replace = FALSE)
+                       , speed = sample(1:c, c, replace = FALSE)
+                       , bould = sample(1:c, c, replace = FALSE)
+                       , lead = sample(1:c, c, replace = FALSE)
+                       ) %>%
+    mutate(total = speed*bould*lead) %>%
+    arrange(total) %>% # some questionable tie-breaking code
+    mutate(m_rank_speed = ifelse(total %in% total[1:3] , rank(speed[1:3]), 0)
+           , m_rank_bould = ifelse(total %in% total[1:3] , rank(bould[1:3]), 0)
+           , m_rank_lead = ifelse(total %in% total[1:3] , rank(lead[1:3]), 0)
+           , tiebreaker = m_rank_speed + m_rank_bould + m_rank_lead
+           ) %>%
+    arrange(total
+            , tiebreaker
+            )
+
+  return(output)
+}
+```
+
+**Remove a climber from simulated data and rerank remaining**
+```{r, eval=FALSE}
+drop_climber <- function(results, r = 4) {
+  # drop a competitor and recompute rankings and totals
+  # r = which rank competitor to drop
+  ud_results <- results[-r, ] %>%
+    mutate(speed = rank(speed)
+           , bould = rank(bould)
+           , lead = rank(lead)
+           , total = speed*bould*lead
+           ) %>%
+    arrange(total)%>%
+    mutate(m_rank_speed = ifelse(total %in% total[1:3], rank(speed[1:3]), 0)
+           , m_rank_bould = ifelse(total %in% total[1:3], rank(bould[1:3]), 0)
+           , m_rank_lead = ifelse(total %in% total[1:3], rank(lead[1:3]), 0)
+           , tiebreaker = m_rank_speed + m_rank_bould + m_rank_lead
+           ) %>%
+    arrange(total
+            , tiebreaker
+            )
+  return(ud_results)
+}
+```
+
+**Simulate reranking of all climbers above dropped climber**
+```{r, eval=FALSE}
+displace_climber <- function(nsim = 1000, drop_rank = 4, nc = 8) {
+  # simulate nsim rankings
+  rank_perms <- lapply(1:nsim, function(x) assign_ranks(c = nc))
+  # drop ith competitor and recalculate ranks
+  dropout_ranks <- lapply(rank_perms, function(x) drop_climber(x, r = drop_rank))
+  # compare 1st - (r-1)th place & calculate % not matched
+  pct_displ <- sapply(1:nsim, 
+                      function(x) {
+                        sum(rank_perms[[x]]$climber[1:(drop_rank - 1)] != dropout_ranks[[x]]$climber[1:(drop_rank - 1)]) > 0
+                      }
+                      ) %>%
+    sum()/nsim
+
+  return(pct_displ)
+}
+```
+
+**Simulate reranking of medalists when a non-medalist is dropped**
+```{r, eval=FALSE}
+displace_medalist <- function(nsim = 1000, drop_rank = 4, nc = 8) {
+  # simulate nsim rankings
+  rank_perms <- lapply(1:nsim, function(x) assign_ranks(nc))
+  # drop ith competitor and recalculate ranks
+  dropout_ranks <- lapply(rank_perms, function(x) drop_climber(x, r = drop_rank))
+  # compare 1st-3rd place & calculate % not matched
+  pct_displ <- sapply(1:nsim, 
+                      function(x) {
+                        sum(rank_perms[[x]]$climber[1:3] != dropout_ranks[[x]]$climber[1:3]) > 0
+                      }
+                      ) %>%
+    sum()/nsim
+
+  return(pct_displ)
+}
+```
+
+**Simulate replacement of finalists when a non-finalist is dropped** 
+*permutations do not count*
+```{r, eval=FALSE}
+displace_finalist <- function(nsim = 100, drop_rank = 9) {
+  # simulate nsim rankings
+  rank_perms <- lapply(1:nsim, function(x) assign_ranks(c = 20))
+  # drop ith competitor and recalculate ranks
+  dropout_ranks <- lapply(rank_perms, function(x) drop_climber(x, r = drop_rank))
+  # compare 1st-3rd place & calculate % not matched
+  pct_displ <- sapply(1:nsim, 
+                      function(x) {
+                        sum(rank_perms[[x]]$climber[1:8] %in% dropout_ranks[[x]]$climber[1:8]) < 8
+                      }
+  ) %>%
+    sum()/nsim
+
+  return(pct_displ)
+}
+```

### 2021-08-04 manuscript/Rmd/draft.Rmd
Trying to finish

@@
 
-The decision to combine the three climbing events and only award one set of medals for both men's and women's events in the Olympics has received a large amount of criticism from climbing athletes all over the world (CITATION NEEDED - I mentioned the article at the end of the paragrah).
+The decision to combine the three climbing events and only award one set of medals for both men's and women's events in the Olympics has received a large amount of criticism from climbing athletes all over the world (CITATION NEEDED).
 In a series of interviews conducted by Climbing Magazine in 2016, a number of climbers shared their thoughts and concerns about the new Olympics climbing format.
@@
 For more detail on rank sum scoring, see @hammond2007 and @boudreau2018.
-
-Some individual sports such as the decathlon and heptathlon rely on a sum of scores from the ten or seven events, however, these scores are not determined based on the ranks of the competitors.
-That is a decathlete's score is entirely based upon their times, distances, and heights, and their overall score will be exactly the same if the times, distances, and heights remain the same regardless of the performance of other individuals.
+In addition, some individual sports such as the decathlon and heptathlon rely on a sum of scores from the ten or seven events, however, these scores are not determined based on the ranks of the competitors.
+That is, a decathlete's score is entirely based upon their times, distances, and heights, and their overall score will be exactly the same if the times, distances, and heights remain the same regardless of the performance of other individuals.
 Full details of decathlon scoring can be found in @westera2006.
 
-There are other individual competitions consisted of several events combined that do base their scoring on ranks, such as crossfit competitions.
+<!-- 
+here are other individual competitions consisted of several events combined that do base their scoring on ranks, such as crossfit competitions.
 In each event, points are earned based on the competitor's rank in the event based on a scoring table.
@@
 As a side note, rank product, genes 
+-->
 
@@
 ```{r}
-set.seed(1)
-nsim <- 10000
-nplay <- 20
-qual_sim <- list()
+climbing_sim <- function(nsim = 10000, nplay) {
+  sims <- list()
   for (i in 1:nsim) {
-  qual_sim[[i]] <-
+    sims[[i]] <-
       bind_cols(
@@
   }
-
-qual <- bind_rows(qual_sim) %>% 
-  mutate(score = e1 * e2 * e3) %>% 
-  group_by(sim) %>% 
-  mutate(rank = rank(score, ties.method = "random")) %>%
-  ungroup()
-```
-
-```{r}
-set.seed(1)
-nsim <- 10000
-nplay <- 8
-final_sim <- list()
-for (i in 1:nsim) {
-  final_sim[[i]] <-
-    bind_cols(
-      player = 1:nplay,
-      e1 = sample(1:nplay, replace = FALSE),
-      e2 = sample(1:nplay, replace = FALSE),
-      e3 = sample(1:nplay, replace = FALSE)
-    ) %>%
-    mutate(sim = i)
-}
-
-final <- bind_rows(final_sim)
-
-final <- final %>% 
+  results <- bind_rows(sims) %>% 
     mutate(score = e1 * e2 * e3) %>% 
@@
 
+  return(results)
+}
+
+set.seed(1)
+qual <- climbing_sim(nsim = 10000, nplay = 20)
+final <- climbing_sim(nsim = 10000, nplay = 8)
+```
+
+```{r}
 qual_final <- qual %>% 

### 2021-08-04 manuscript/Rmd/draft.tex
Trying to finish

@@
 a large amount of criticism from climbing athletes all over the world
-(CITATION NEEDED - I mentioned the article at the end of the paragrah).
-In a series of interviews conducted by Climbing Magazine in 2016, a
-number of climbers shared their thoughts and concerns about the new
-Olympics climbing format. Climber Lynn Hill compared the idea of
-combining speed climbing, bouldering, and lead climbing to ``asking a
-middle distance runner to compete in the sprint'' (CITATION NEEDED). She
-then added ``Speed climbing is a sport within our sport''. Other
-climbers also hold the same opinion as Hill regarding speed climbing,
-using words and phrases like ``bogus'', ``a bummer'', ``less than
-ideal'', ``not in support'', and ``cheesy and unfair'' to describe the
-new combined competition format. Courtney Woods stated ``Speed climbers
-will have the biggest disadvantage because their realm isn't based on
-difficult movements''. Mike Doyle believed ``Honestly, the people that
-will suffer the most are the ones that focus only on speed climbing.
-Those skills/abilities don't transfer as well to the other
-disciplines''. Most climbers also expressed their hope for a change in
-the format in future tournaments, with some calling for each discipline
-to have its own set of medals. See \citet{blanchard2016} for more
-information on these interviews.
+(CITATION NEEDED). In a series of interviews conducted by Climbing
+Magazine in 2016, a number of climbers shared their thoughts and
+concerns about the new Olympics climbing format. Climber Lynn Hill
+compared the idea of combining speed climbing, bouldering, and lead
+climbing to ``asking a middle distance runner to compete in the sprint''
+(CITATION NEEDED). She then added ``Speed climbing is a sport within our
+sport''. Other climbers also hold the same opinion as Hill regarding
+speed climbing, using words and phrases like ``bogus'', ``a bummer'',
+``less than ideal'', ``not in support'', and ``cheesy and unfair'' to
+describe the new combined competition format. Courtney Woods stated
+``Speed climbers will have the biggest disadvantage because their realm
+isn't based on difficult movements''. Mike Doyle believed ``Honestly,
+the people that will suffer the most are the ones that focus only on
+speed climbing. Those skills/abilities don't transfer as well to the
+other disciplines''. Most climbers also expressed their hope for a
+change in the format in future tournaments, with some calling for each
+discipline to have its own set of medals. See \citet{blanchard2016} for
+more information on these interviews.
 
@@
 winner. For more detail on rank sum scoring, see \citet{hammond2007} and
-\citet{boudreau2018}.
-
-Some individual sports such as the decathlon and heptathlon rely on a
-sum of scores from the ten or seven events, however, these scores are
-not determined based on the ranks of the competitors. That is a
-decathlete's score is entirely based upon their times, distances, and
-heights, and their overall score will be exactly the same if the times,
-distances, and heights remain the same regardless of the performance of
-other individuals. Full details of decathlon scoring can be found in
-\citet{westera2006}.
-
-There are other individual competitions consisted of several events
-combined that do base their scoring on ranks, such as crossfit
-competitions. In each event, points are earned based on the competitor's
-rank in the event based on a scoring table.
-(\url{https://games.crossfit.com/announcement/12940}), and a competitors
-final score is based on the sum of their scores across all the events.
-
-Paragraph on scoring methods in other sports: Soccer points:
-\citet{haugen2008}; Cross country scoring: \citet{hammond2007},
-\citet{boudreau2018}; Decathlon scoring: \citet{westera2006}
-
-As a side note, rank product, genes
+\citet{boudreau2018}. In addition, some individual sports such as the
+decathlon and heptathlon rely on a sum of scores from the ten or seven
+events, however, these scores are not determined based on the ranks of
+the competitors. That is, a decathlete's score is entirely based upon
+their times, distances, and heights, and their overall score will be
+exactly the same if the times, distances, and heights remain the same
+regardless of the performance of other individuals. Full details of
+decathlon scoring can be found in \citet{westera2006}.
 
@@
 Qualification & 1 & 240 & 684 & 1638 & 8000 & 1158.84378 & 1273.47884 & 200000\\
-Final & 1 & 24 & 60 & 126 & 512 & 91.12077 & 91.35293 & 80000\\
+Final & 1 & 24 & 60 & 126 & 512 & 91.19935 & 91.11065 & 80000\\
 \bottomrule
@@
 \midrule
-1 & 9.6748\\
-2 & 20.4143\\
-3 & 33.2648\\
-4 & 50.5933\\
-5 & 74.7620\\
+1 & 9.6687\\
+2 & 20.1844\\
+3 & 33.3658\\
+4 & 50.5734\\
+5 & 75.2499\\
 \addlinespace
-6 & 110.0536\\
-7 & 164.4274\\
-8 & 265.7760\\
+6 & 110.7070\\
+7 & 164.8258\\
+8 & 265.0198\\
 \bottomrule

### 2021-08-04 R/tokyo.R

@@ -1,0 +1,220 @@
+
+
+Overall <- 1:20
+Speed <- c(3, 2, 6, 12, 18, 7, 1, 10, 19, 5, 4, 11, 16, 20, 8, 9, 14, 15, 13, 17)
+Bouldering <- c(1, 2, 5, 7, 3, 14, 18, 11, 6, 10, 17, 9, 4, 8, 13, 16, 15, 12, 19.5, 19.5)
+Lead <- c(11, 14, 2, 1, 4, 3, 20, 5, 6, 16, 13, 9, 15, 7, 12, 10, 8, 17, 18, 19)
+
+
+tibble(Overall, Speed, Bouldering, Lead) %>% 
+  ggpairs(diag = "blank",
+          axisLabels = "none",
+          upper = list(continuous = wrap("cor", method = "kendall", stars = FALSE))) +
+  theme_light() + ggtitle("2020 Olympics - Men's Qualification")
+
+
+tibble(Overall, Speed, Bouldering, Lead) %>% 
+  # mutate(Score = Speed * Bouldering * Lead)
+  pivot_longer(!Overall) %>% 
+  ggplot(aes(x = name, y = value)) +
+  geom_bar(stat = "identity", position = "dodge")
+  
+
+
+
+library(tidyverse)
+library(rvest)
+library(janitor)
+library(viridis)
+
+
+url <- "https://en.wikipedia.org/wiki/Sport_climbing_at_the_2020_Summer_Olympics_%E2%80%93_Men%27s"
+
+
+
+mq <- url %>% 
+  read_html() %>% 
+  html_table() %>% 
+  .[[3]] %>% 
+  row_to_names(row_number = 1) %>% 
+  clean_names() 
+
+mq %>% 
+  select(climber,
+         overall = rank, 
+         speed = cp,
+         bouldering = cp_2,
+         lead = cp_3,
+         total) %>% 
+  select(climber, S = speed, B = bouldering, L = lead) %>% 
+  pivot_longer(!climber, names_to = "event", values_to = "rank") %>% 
+  mutate(climber = fct_rev(factor(climber, levels = mq$climber))) %>% 
+  ggplot(aes(x = event, y = climber, fill = as.numeric(rank))) +
+  geom_tile(color = "white", size = 0.1) +
+  scale_fill_viridis(breaks = c(1, 5, 10, 15, 20), direction = -1) +
+  coord_equal() +
+  labs(fill = "Rank")
+
+
+
+df <- mq %>% 
+  select(climber,
+         speed = cp,
[…]

### 2021-08-04 manuscript/Rmd/draft.Rmd
Intro edit

@@
   The purpose of this paper is to investigate the controversial competition format and scoring system of sport climbing, which is one of the sports making its debut on the Olympics stage at Tokyo 2020.
-Climbing Olympians will participate in three disciplines (speed climbing, bouldering, and lead climbing) for a single set of medals, and a multiplied score of climbers' rankings in all three disciplines will be used to determine the overall standings.
+Climbing Olympians will participate in three disciplines: speed climbing, bouldering, and lead climbing for a single set of medals, and a multiplied score of climbers' rankings in all three disciplines will be used to determine the overall standings.
 This work finds great evidence from historical data and simulations that speed climbing speed climbing should be separated from the combined format and have its own medals.
@@
 
-#While there are three main disciplines in sport climbing, the IOC only allotted one set of medals per gender.
+<!-- 
+While there are three main disciplines in sport climbing, the IOC only allotted one set of medals per gender.
 So rather than choosing only a single discipline, they chose to combine all three of the events together to create a combined sport climbing event.
+-->
 
@@
 
-The decision to combine the three climbing events and only award one set of medals for both men's and women's events in the Olympics has received a large amount of criticism from climbing athletes all over the world (CITATION NEEDED).
+The decision to combine the three climbing events and only award one set of medals for both men's and women's events in the Olympics has received a large amount of criticism from climbing athletes all over the world.
 In a series of interviews conducted by Climbing Magazine in 2016, a number of climbers shared their thoughts and concerns about the new Olympics climbing format.
-Climber Lynn Hill compared the idea of combining speed climbing, bouldering, and lead climbing to "asking a middle distance runner to compete in the sprint" (CITATION NEEDED).
+Climber Lynn Hill compared the idea of combining speed climbing, bouldering, and lead climbing to "asking a middle distance runner to compete in the sprint".
 She then added "Speed climbing is a sport within our sport".
@@
 The climbers with the lowest, second lowest, and third lowest product of ranks in the final wins the gold, silver, and bronze medal, respectively.
-
 This type of scoring system heavily rewards high finishes and relatively ignores poor finishes.

### 2021-08-04 manuscript/Rmd/draft.tex
Intro edit

@@
 the sports making its debut on the Olympics stage at Tokyo 2020.
-Climbing Olympians will participate in three disciplines (speed
-climbing, bouldering, and lead climbing) for a single set of medals, and
+Climbing Olympians will participate in three disciplines: speed
+climbing, bouldering, and lead climbing for a single set of medals, and
 a multiplied score of climbers' rankings in all three disciplines will
@@
 
-\#While there are three main disciplines in sport climbing, the IOC only
-allotted one set of medals per gender. So rather than choosing only a
-single discipline, they chose to combine all three of the events
-together to create a combined sport climbing event.
-
 The three disciplines that comprise sport climbing at Tokyo 2020 are:
@@
 of medals for both men's and women's events in the Olympics has received
-a large amount of criticism from climbing athletes all over the world
-(CITATION NEEDED). In a series of interviews conducted by Climbing
-Magazine in 2016, a number of climbers shared their thoughts and
-concerns about the new Olympics climbing format. Climber Lynn Hill
-compared the idea of combining speed climbing, bouldering, and lead
-climbing to ``asking a middle distance runner to compete in the sprint''
-(CITATION NEEDED). She then added ``Speed climbing is a sport within our
-sport''. Other climbers also hold the same opinion as Hill regarding
-speed climbing, using words and phrases like ``bogus'', ``a bummer'',
-``less than ideal'', ``not in support'', and ``cheesy and unfair'' to
-describe the new combined competition format. Courtney Woods stated
-``Speed climbers will have the biggest disadvantage because their realm
-isn't based on difficult movements''. Mike Doyle believed ``Honestly,
-the people that will suffer the most are the ones that focus only on
-speed climbing. Those skills/abilities don't transfer as well to the
-other disciplines''. Most climbers also expressed their hope for a
-change in the format in future tournaments, with some calling for each
-discipline to have its own set of medals. See \citet{blanchard2016} for
-more information on these interviews.
+a large amount of criticism from climbing athletes all over the world.
+In a series of interviews conducted by Climbing Magazine in 2016, a
+number of climbers shared their thoughts and concerns about the new
+Olympics climbing format. Climber Lynn Hill compared the idea of
+combining speed climbing, bouldering, and lead climbing to ``asking a
+middle distance runner to compete in the sprint''. She then added
+``Speed climbing is a sport within our sport''. Other climbers also hold
+the same opinion as Hill regarding speed climbing, using words and
+phrases like ``bogus'', ``a bummer'', ``less than ideal'', ``not in
+support'', and ``cheesy and unfair'' to describe the new combined
+competition format. Courtney Woods stated ``Speed climbers will have the
+biggest disadvantage because their realm isn't based on difficult
+movements''. Mike Doyle believed ``Honestly, the people that will suffer
+the most are the ones that focus only on speed climbing. Those
+skills/abilities don't transfer as well to the other disciplines''. Most
+climbers also expressed their hope for a change in the format in future
+tournaments, with some calling for each discipline to have its own set
+of medals. See \citet{blanchard2016} for more information on these
+interviews.
 
@@
 lowest, second lowest, and third lowest product of ranks in the final
-wins the gold, silver, and bronze medal, respectively.
-
-This type of scoring system heavily rewards high finishes and relatively
-ignores poor finishes. For instance, if climber A finished 1st, 20th,
-and 20th and climber B finished 10th, 10th, and 10th, climber B would
-have a score of 1000 whereas climber A would have a much better score of
-400, despite finishing last in 2 out of 3 of the events.
+wins the gold, silver, and bronze medal, respectively. This type of
+scoring system heavily rewards high finishes and relatively ignores poor
+finishes. For instance, if climber A finished 1st, 20th, and 20th and
+climber B finished 10th, 10th, and 10th, climber B would have a score of
+1000 whereas climber A would have a much better score of 400, despite
+finishing last in 2 out of 3 of the events.
 

### 2021-08-04 R/remove_climber_sim.Rmd
update remove1 sims + add tokyo2020 qualification data

@@
 My hypothesis is that there is enough separation between the original finalists and the 10th+ ranking competitors in the individual events that there isn't a high chance that the total rankings of the finalists can be overtaken by 10+th place competitors.
-Note: it might be worth computing the probability of getting a low ranking in any of the individual events, given that a competitor is a finalist.
+Note: it might be worth computing the probability of getting a low ranking in any of the individual events, given that a competitor is a finalist (ie: $P(r_1 > 8 \text{ or } r_2 > 8 \text{ or } r_3 > 8 \text{ }|\text{ } R_T \leq 8)$).
 
@@
 
-remove_c5 <- lapply(5:20, function(x) replicate(100, displace_climber(nsim = 100, drop_rank = 5, nc = x)))
+remove_c5 <- lapply(5:20, function(x) replicate(100, displace_medalist(nsim = 100, drop_rank = 5, nc = x)))
 
@@
                                , dropped_rank = 5
-                               , displ_prob = remove_c4[[x-4]]
+                               , displ_prob = remove_c5[[x-4]]
                                )
@@
   do.call(rbind, .)
-#save(remove_c5_df, "R/remove_c5_sims.rdata")
+#save(remove_c5_df, file = "R/remove_c5_sims.rdata")
+
+remove_c6 <- lapply(6:20, function(x) replicate(100, displace_medalist(nsim = 100, drop_rank = 6, nc = x)))
+remove_c6_df <- lapply(6:20
+                  , function(x) {
+                    data.frame(n_climbers = x
+                               , dropped_rank = 6
+                               , displ_prob = remove_c6[[x-5]]
+                               )
+                    }
+                  ) %>%
+  do.call(rbind, .)
 ```
@@
 ### Removing 4th Place
-```{r, echo = FALSE}
+```{r, include=FALSE}
 load("remove_c4_sims.rdata")
@@
 
+```{r, echo = FALSE}
+load("remove_c5_sims.rdata")
+load("remove_c6_sims.rdata")
+remove_c456 <- rbind(remove_c4_df, remove_c5_df, remove_c6_df)
+remove_c456 %>%
+  group_by(n_climbers, dropped_rank) %>%
+  summarize(av_disp_prob = mean(displ_prob)) %>%
+  ggplot(aes(x = n_climbers, y = av_disp_prob)) +
+  geom_line(aes(color = as.factor(dropped_rank))) +
+  theme_minimal()
+```
+
 # Questions: 
-- How often would a finalist be replaced
-- How often does someone who wasn't a medalist before become a medalist 
 - How often do we see a flip (two consecutive climbers switch places), jump (a climber jumps up more than one rank), drop (a climber drops down more than one rank)

### 2021-08-04 manuscript/Rmd/draft.Rmd
update draft

@@
 
-<!-- In what percentage of the (6!)^3 orderings would the winner change if one of the 6 competitors was dropped and there were only 5 instead of 6? -->
+### Probability of Reordering in Higher Ranks
+We first look at the rate at which the ranks above that of which is dropped are reordered.
+```{r, echo = FALSE}
+load("draft_files/reorder_all_above_probs.rdata")
+
+sum_dat1 <- reorder_all_above %>%
+  group_by(dropped_rank) %>%
+  rename("Dropped Rank" = dropped_rank) %>%
+  summarize("Simulations" = n()
+            , "Avg. Probability" = mean(displ_prob)
+            , "Median Probability" = median(displ_prob)
+            , "St. Deviation" = sd(displ_prob)
+            )
+
+kable(sum_dat1)
+```
+
+### Probability of Change in Finalists
+
+A potential event of interest is the replacement of one finalist with another, taking the chance away from the original finalist to win a medal in the finals round.
+Here we compute the probability that a finalist will be replaced with a non-finalist after removing another non-finalist from the rankings.
+Below is a table summarizing the results.
+
+```{r, echo=FALSE}
+load("draft_files/drop_qualifier_sims.rdata")
+
+reorder_dat2 <- lapply(9:20
+                  , function(x) {
+                    data.frame(dropped_rank = x
+                               , displ_prob = pct_displace_finalists[[x-8]]
+                               )
+                    }
+                  ) %>%
+  do.call(rbind, .)
+
+sum_dat2 <- reorder_dat2 %>%
+  group_by(dropped_rank) %>%
+  rename("Dropped Rank" = dropped_rank) %>%
+  summarize("Simulations" = n()
+            , "Avg. Probability" = mean(displ_prob)
+            , "Median Probability" = median(displ_prob)
+            , "St. Deviation" = sd(displ_prob)
+            )
+
+kable(sum_dat2)
+
+```
+
+as can be seen, the probability is relatively low when the 9th place competitor is removed.
+My hypothesis is that there is enough separation between the original finalists and the 10th+ ranking competitors in the individual events that there isn't a high chance that the total rankings of the finalists can be overtaken by 10+th place competitors.
+Note: it might be worth computing the probability of getting a low ranking in any of the individual events, given that a competitor is a finalist (ie: $P(r_1 > 8 \text{ or } r_2 > 8 \text{ or } r_3 > 8 \text{ }|\text{ } R_T \leq 8)$).
+
+### Probability that Medalists are Reordered or Changed
+
+The second scenario of interest is if a medalist is displaced by another medalist, or even a non-medalist.
+Below are the results summarizing the probability of this scenario occurring under uniform random ranking conditions.
+
+```{r, echo = FALSE}
+# probability that medalists are displaced (changed or reordered)
+load("draft_files/top3_displace_probs.rdata")
+
+displace_probs %>%
+  group_by(dropped_rank) %>%
+  rename("Dropped Rank" = dropped_rank) %>%
+  summarize("Simulations" = n()
+            , "Avg. Probability" = mean(displ_prob)
+            , "Median Probability" = median(displ_prob)
+            , "St. Deviation" = sd(displ_prob)
+            ) %>%
+ kable()
+
+```
 

### 2021-08-04 manuscript/Rmd/draft.tex
update draft

@@
 {
-  \title{\bf An Examination of Sport Climbing Competition Format and
-Scoring System}
+  \title{\bf An Examination of Sport Climbing Competition Format and Scoring System}
 
@@
   \begin{center}
-    {\LARGE\bf An Examination of Sport Climbing Competition Format and
-Scoring System}
+    {\LARGE\bf An Examination of Sport Climbing Competition Format and Scoring System}
   \end{center}
@@
 \midrule
-Qualification & 1 & 240 & 684 & 1638 & 8000 & 1158.84378 & 1273.47884 & 200000\\
-Final & 1 & 24 & 60 & 126 & 512 & 91.19935 & 91.11065 & 80000\\
+Qualification & 1 & 240 & 684 & 1638 & 8000 & 1158.23077 & 1269.92888 & 200000\\
+Final & 1 & 24 & 60 & 126 & 512 & 91.02296 & 90.68192 & 80000\\
 \bottomrule
@@
 \midrule
-1 & 36.0187\\
-2 & 73.6111\\
-3 & 115.3954\\
-4 & 162.2263\\
-5 & 216.0041\\
+1 & 35.8986\\
+2 & 73.6254\\
+3 & 115.0981\\
+4 & 162.6459\\
+5 & 216.5088\\
 \addlinespace
-6 & 278.1649\\
-7 & 350.3272\\
-8 & 434.5932\\
-9 & 532.1383\\
-10 & 642.3298\\
+6 & 278.7528\\
+7 & 351.4263\\
+8 & 435.0092\\
+9 & 532.5390\\
+10 & 642.8740\\
 \bottomrule
@@
 \midrule
-1 & 9.6687\\
-2 & 20.1844\\
-3 & 33.3658\\
-4 & 50.5734\\
-5 & 75.2499\\
+1 & 9.6823\\
+2 & 20.3044\\
+3 & 33.3976\\
+4 & 50.5062\\
+5 & 74.9904\\
 \addlinespace
-6 & 110.7070\\
-7 & 164.8258\\
-8 & 265.0198\\
+6 & 110.3614\\
+7 & 164.9222\\
+8 & 264.0192\\
 \bottomrule
@@
 non-medalist is removed from the list of finalists.
+
+\hypertarget{probability-of-reordering-in-higher-ranks}{%
+\subsubsection{Probability of Reordering in Higher
+Ranks}\label{probability-of-reordering-in-higher-ranks}}
+
+We first look at the rate at which the ranks above that of which is
+dropped are reordered.
+
+\begin{tabular}{r|r|r|r|r}
+\hline
+Dropped Rank & Simulations & Avg. Probability & Median Probability & St. Deviation\\
+\hline
+4 & 100 & 0.25869 & 0.2590 & 0.0145433\\
+\hline
+5 & 100 & 0.34032 & 0.3405 & 0.0151263\\
+\hline
+6 & 100 & 0.38012 & 0.3820 & 0.0152430\\
+\hline
+7 & 100 & 0.40636 & 0.4060 & 0.0149582\\
+\hline
+8 & 100 & 0.39439 & 0.3930 & 0.0151023\\
+\hline
+\end{tabular}
+
+\hypertarget{probability-of-change-in-finalists}{%
+\subsubsection{Probability of Change in
+Finalists}\label{probability-of-change-in-finalists}}
+
+A potential event of interest is the replacement of one finalist with
+another, taking the chance away from the original finalist to win a
+medal in the finals round. Here we compute the probability that a
+finalist will be replaced with a non-finalist after removing another
+non-finalist from the rankings. Below is a table summarizing the
+results.
+
+\begin{tabular}{r|r|r|r|r}
+\hline
+Dropped Rank & Simulations & Avg. Probability & Median Probability & St. Deviation\\
+\hline
+9 & 100 & 0.0505 & 0.050 & 0.0217597\\
+\hline
+10 & 100 & 0.1538 & 0.150 & 0.0318386\\
+\hline
+11 & 100 & 0.1603 & 0.160 & 0.0343616\\
+\hline
+12 & 100 & 0.1416 & 0.140 & 0.0353259\\
+\hline
+13 & 100 & 0.1365 & 0.140 & 0.0336162\\
+\hline
+14 & 100 & 0.1217 & 0.120 & 0.0344965\\
+\hline
+15 & 100 & 0.1213 & 0.120 & 0.0308353\\
+\hline
+16 & 100 & 0.1159 & 0.120 & 0.0306197\\
+\hline
+17 & 100 & 0.1086 & 0.105 & 0.0314312\\
+\hline
+18 & 100 & 0.1026 & 0.100 & 0.0329867\\
+\hline
+19 & 100 & 0.0886 & 0.090 & 0.0291641\\
+\hline
+20 & 100 & 0.0700 & 0.070 & 0.0258590\\
+\hline
+\end{tabular}
+
+as can be seen, the probability is relatively low when the 9th place
+competitor is removed. My hypothesis is that there is enough separation
+between the original finalists and the 10th+ ranking competitors in the
+individual events that there isn't a high chance that the total rankings
+of the finalists can be overtaken by 10+th place competitors. Note: it
+might be worth computing the probability of getting a low ranking in any
+of the individual events, given that a competitor is a finalist (ie:
+\(P(r_1 > 8 \text{ or } r_2 > 8 \text{ or } r_3 > 8 \text{ }|\text{ } R_T \leq 8)\)).
+
+\hypertarget{probability-that-medalists-are-reordered-or-changed}{%
+\subsubsection{Probability that Medalists are Reordered or
+Changed}\label{probability-that-medalists-are-reordered-or-changed}}
+
+The second scenario of interest is if a medalist is displaced by another
+medalist, or even a non-medalist. Below are the results summarizing the
+probability of this scenario occurring under uniform random ranking
+conditions.
+
+\begin{tabular}{r|r|r|r|r}
+\hline
+Dropped Rank & Simulations & Avg. Probability & Median Probability & St. Deviation\\
+\hline
+4 & 100 & 0.25872 & 0.2580 & 0.0137106\\
+\hline
+5 & 100 & 0.31988 & 0.3190 & 0.0161416\\
+\hline
+6 & 100 & 0.28741 & 0.2870 & 0.0127969\\
+\hline
+7 & 100 & 0.25631 & 0.2545 & 0.0145301\\
+\hline
+8 & 100 & 0.20631 & 0.2050 & 0.0120283\\
+\hline
+\end{tabular}
 
@@
 
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-9-1}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-12-1}
 
@@
 
-\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-9}
+\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-12}
 \end{figure}
@@
 
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-12-1}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-15-1}
 
@@
 For each gender competition, each panel represents the rank of the drop-out athlete, with 0 being the original final results.
-Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-12}
+Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-15}
 \end{figure}
@@
 
-\caption{\label{tab:unnamed-chunk-14}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
+\caption{\label{tab:unnamed-chunk-17}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-15}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
+\caption{\label{tab:unnamed-chunk-18}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-16}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
+\caption{\label{tab:unnamed-chunk-19}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-17}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
+\caption{\label{tab:unnamed-chunk-20}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
 \centering

### 2021-08-04 manuscript/Rmd/draft.Rmd
Add simulation tables and plots

--- 2021-08-04 manuscript/Rmd/draft.Rmd
@@
         booktabs = TRUE,
-        position = "H")
+        position = "H",
+        digits = 2)
 ```
@@
 In this section, we discuss the results of our simulations described in Section \ref{data-and-methods}.
-For the qualification round, our simulation study shows that a climber is almost guaranteed to make the final round if they win the first event or if they win at least one of the three climbing concentrations  (99.51% and 99.48% change of advancing, respectively).
+We first look at the chances of advancing to the final round for a qualifier, given that 1) they finish first in any of the three disciplines, and 2) they win the first event (which happens to be speed climbing in reality).
+Our simulation study shows that a climber is almost guaranteed to make the final round if they win at least one of the three climbing concentrations (99.51%, see Table) or if they win the opening event (99.48%, see Table).
+After winning the first event, a climber is also more likely to finish first overall (30%, see Table) than any other positions in the overall qualification rankings.
+Therefore, this product of ranks scoring format highly rewards climbers who can get off to a strong start, which is a major strategy climbing athletes should take note of.
+
+In addition, the average score for qualification positions 1 to 8 are displayed in Table X.
+We notice that on average, the minimum score that one should aim for in order to move on to the final round is about 434 (for 8th rank).
+
+```{r}
+win_any_qual <- qual %>% 
+  filter(e1 == 1 | e2 == 1 | e3 == 1) %>% 
+  count(rank) %>% 
+  mutate(prob = n / sum(n),
+         cum_prob = cumsum(prob))
+win_first_qual <- qual %>% 
+  filter(e1 == 1) %>% 
+  count(rank) %>% 
+  mutate(prob = n / sum(n),
+         cum_prob = cumsum(prob))
+
+cbind(win_any_qual, win_first_qual) %>% 
+  kable(col.names = rep(c("Rank", "Count", "Probability", "Cumulative"), 2),
+        caption = "title", 
+        booktabs = TRUE, 
+        position = "H",
+        digits = 4) %>% 
+  add_header_above(c("Win Any" = 4, "Win First" = 4)) %>%
+  kable_styling(latex_options = c("repeat_header", "scale_down"))
+```
+
+```{r, fig.cap="Probability of finishing at every rank given", fig.width = 6.4, fig.height = 3.2}
+win_any_qual %>% 
+  mutate(type = "Win Any") %>% 
+  bind_rows(mutate(win_first_qual, type = "Win First")) %>% 
+  mutate(finalist = ifelse(rank %in% 1:8, "Yes", "No")) %>% 
+  ggplot(aes(rank, weight = prob, fill = finalist)) +
+  geom_bar() +
+  geom_vline(xintercept = 8.5) +
+  scale_x_continuous() +
+  labs(x = "Rank",
+       y = "Probability",
+       fill = "Finalist") +
+  facet_wrap(~ type)
+```
+
+<!--
 On the other hand, finishing last in the first event or in any event would certainly hurt an athlete's chance of finishing in the top 8, as the probabilities of a climber advancing given they finish last in the first and in any event are 0.1830 and 0.1885, respectively.
-In addition, the average score for qualification positions 1 to 8 are displayed in Table 2.
-We notice that on average, the minimum score that one should aim for in order to move on to the final round is about 434 (for 8th rank).
+-->
 
@@
 Furthermore, according to our final simulations, in order to obtain a climbing medal, the average scores (rounded down) that put an athlete in position to receive gold, silver, and bronze medals are 9, 20, and 33, respectively (see Table 3).
+
+```{r}
+win_any_final <- final %>% 
+  filter(e1 == 1 | e2 == 1 | e3 == 1) %>% 
+  count(rank) %>% 
+  mutate(prob = n / sum(n),
+         cum_prob = cumsum(prob))
+win_first_final <- final %>% 
+  filter(e1 == 1) %>% 
+  count(rank) %>% 
+  mutate(prob = n / sum(n),
+         cum_prob = cumsum(prob))
+
+cbind(win_any_final, win_first_final) %>% 
+  kable(col.names = rep(c("Rank", "Count", "Probability", "Cumulative"), 2),
+        caption = "title", 
+        booktabs = TRUE, 
+        position = "H",
+        digits = 4) %>% 
+  add_header_above(c("Win Any" = 4, "Win First" = 4)) %>%
+  kable_styling(latex_options = c("repeat_header", "scale_down"))
+```
+
+```{r, fig.cap="Probability of finishing at every rank given", fig.width = 6.4, fig.height = 3.2}
+win_any_final %>% 
+  mutate(type = "Win Any") %>% 
+  bind_rows(mutate(win_first_final, type = "Win First")) %>% 
+  mutate(medalist = ifelse(rank %in% 1:3, "Yes", "No")) %>% 
+  ggplot(aes(rank, weight = prob, fill = medalist)) +
+  geom_bar() +
+  geom_vline(xintercept = 3.5) +
+  scale_x_continuous() +
+  labs(x = "Rank",
+       y = "Probability",
+       fill = "Medalist") +
+  facet_wrap(~ type)
+```
 

### 2021-08-04 manuscript/Rmd/draft.tex
Add simulation tables and plots

--- 2021-08-04 manuscript/Rmd/draft.tex
@@
 \midrule
-Qualification & 1 & 240 & 684 & 1638 & 8000 & 1158.84378 & 1273.47884 & 200000\\
-Final & 1 & 24 & 60 & 126 & 512 & 91.19935 & 91.11065 & 80000\\
+Qualification & 1 & 240 & 684 & 1638 & 8000 & 1158.84 & 1273.48 & 200000\\
+Final & 1 & 24 & 60 & 126 & 512 & 91.20 & 91.11 & 80000\\
 \bottomrule
@@
 In this section, we discuss the results of our simulations described in
-Section \ref{data-and-methods}. For the qualification round, our
-simulation study shows that a climber is almost guaranteed to make the
-final round if they win the first event or if they win at least one of
-the three climbing concentrations (99.51\% and 99.48\% change of
-advancing, respectively). On the other hand, finishing last in the first
-event or in any event would certainly hurt an athlete's chance of
-finishing in the top 8, as the probabilities of a climber advancing
-given they finish last in the first and in any event are 0.1830 and
-0.1885, respectively. In addition, the average score for qualification
-positions 1 to 8 are displayed in Table 2. We notice that on average,
-the minimum score that one should aim for in order to move on to the
-final round is about 434 (for 8th rank).
+Section \ref{data-and-methods}. We first look at the chances of
+advancing to the final round for a qualifier, given that 1) they finish
+first in any of the three disciplines, and 2) they win the first event
+(which happens to be speed climbing in reality). Our simulation study
+shows that a climber is almost guaranteed to make the final round if
+they win at least one of the three climbing concentrations (99.51\%, see
+Table) or if they win the opening event (99.48\%, see Table). After
+winning the first event, a climber is also more likely to finish first
+overall (30\%, see Table) than any other positions in the overall
+qualification rankings. Therefore, this product of ranks scoring format
+highly rewards climbers who can get off to a strong start, which is a
+major strategy climbing athletes should take note of.
+
+In addition, the average score for qualification positions 1 to 8 are
+displayed in Table X. We notice that on average, the minimum score that
+one should aim for in order to move on to the final round is about 434
+(for 8th rank).
 
@@
 
-\caption{\label{tab:unnamed-chunk-6}The average scores for the top 10 qualification ranks according to our simulations. A climber will secure a finalist spot if they finish in the top 8.}
+\caption{\label{tab:unnamed-chunk-6}title}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{rrrrrrrr}
+\toprule
+\multicolumn{4}{c}{Win Any} & \multicolumn{4}{c}{Win First} \\
+\cmidrule(l{3pt}r{3pt}){1-4} \cmidrule(l{3pt}r{3pt}){5-8}
+Rank & Count & Probability & Cumulative & Rank & Count & Probability & Cumulative\\
+\midrule
+1 & 7890 & 0.2762 & 0.2762 & 1 & 2999 & 0.2999 & 0.2999\\
+2 & 5943 & 0.2080 & 0.4842 & 2 & 1963 & 0.1963 & 0.4962\\
+3 & 4644 & 0.1626 & 0.6467 & 3 & 1559 & 0.1559 & 0.6521\\
+4 & 3651 & 0.1278 & 0.7745 & 4 & 1269 & 0.1269 & 0.7790\\
+5 & 2766 & 0.0968 & 0.8714 & 5 & 957 & 0.0957 & 0.8747\\
+\addlinespace
+6 & 1960 & 0.0686 & 0.9400 & 6 & 686 & 0.0686 & 0.9433\\
+7 & 1118 & 0.0391 & 0.9791 & 7 & 363 & 0.0363 & 0.9796\\
+8 & 448 & 0.0157 & 0.9948 & 8 & 155 & 0.0155 & 0.9951\\
+9 & 127 & 0.0044 & 0.9992 & 9 & 42 & 0.0042 & 0.9993\\
+10 & 19 & 0.0007 & 0.9999 & 10 & 6 & 0.0006 & 0.9999\\
+\addlinespace
+11 & 3 & 0.0001 & 1.0000 & 11 & 1 & 0.0001 & 1.0000\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{figure}
+
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-7-1}
+
+}
+
+\caption{Probability of finishing at every rank given}\label{fig:unnamed-chunk-7}
+\end{figure}
+
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-8}The average scores for the top 10 qualification ranks according to our simulations. A climber will secure a finalist spot if they finish in the top 8.}
 \centering
@@
 
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-9}title}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{rrrrrrrr}
+\toprule
+\multicolumn{4}{c}{Win Any} & \multicolumn{4}{c}{Win First} \\
+\cmidrule(l{3pt}r{3pt}){1-4} \cmidrule(l{3pt}r{3pt}){5-8}
+Rank & Count & Probability & Cumulative & Rank & Count & Probability & Cumulative\\
+\midrule
+1 & 9277 & 0.3512 & 0.3512 & 1 & 4295 & 0.4295 & 0.4295\\
+2 & 7409 & 0.2805 & 0.6317 & 2 & 2505 & 0.2505 & 0.6800\\
+3 & 5257 & 0.1990 & 0.8308 & 3 & 1702 & 0.1702 & 0.8502\\
+4 & 3078 & 0.1165 & 0.9473 & 4 & 1020 & 0.1020 & 0.9522\\
+5 & 1187 & 0.0449 & 0.9922 & 5 & 393 & 0.0393 & 0.9915\\
+\addlinespace
+6 & 198 & 0.0075 & 0.9997 & 6 & 83 & 0.0083 & 0.9998\\
+7 & 7 & 0.0003 & 1.0000 & 7 & 2 & 0.0002 & 1.0000\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{figure}
+
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-10-1}
+
+}
+
+\caption{Probability of finishing at every rank given}\label{fig:unnamed-chunk-10}
+\end{figure}
+
 \begin{table}
 
-\caption{\label{tab:unnamed-chunk-7}The average scores for all final ranks according to our simulations. Ranks 1, 2, and 3 are table positions that guarantee medalist status for climbers.}
+\caption{\label{tab:unnamed-chunk-11}The average scores for all final ranks according to our simulations. Ranks 1, 2, and 3 are table positions that guarantee medalist status for climbers.}
 \centering
@@
 
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-9-1}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-13-1}
 
@@
 
-\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-9}
+\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-13}
 \end{figure}
@@
 
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-12-1}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-16-1}
 
@@
 For each gender competition, each panel represents the rank of the drop-out athlete, with 0 being the original final results.
-Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-12}
+Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-16}
 \end{figure}
@@
 
-\caption{\label{tab:unnamed-chunk-14}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
+\caption{\label{tab:unnamed-chunk-18}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-15}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
+\caption{\label{tab:unnamed-chunk-19}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-16}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
+\caption{\label{tab:unnamed-chunk-20}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-17}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
+\caption{\label{tab:unnamed-chunk-21}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
 \centering

### 2021-08-04 manuscript/Rmd/draft.tex

@@
 
+\hypertarget{probability-of-reordering-in-higher-ranks}{%
+\subsubsection{Probability of Reordering in Higher
+Ranks}\label{probability-of-reordering-in-higher-ranks}}
+
+We first look at the rate at which the ranks above that of which is
+dropped are reordered.
+
+\begin{tabular}{r|r|r|r|r}
+\hline
+Dropped Rank & Simulations & Avg. Probability & Median Probability & St. Deviation\\
+\hline
+4 & 100 & 0.25869 & 0.2590 & 0.0145433\\
+\hline
+5 & 100 & 0.34032 & 0.3405 & 0.0151263\\
+\hline
+6 & 100 & 0.38012 & 0.3820 & 0.0152430\\
+\hline
+7 & 100 & 0.40636 & 0.4060 & 0.0149582\\
+\hline
+8 & 100 & 0.39439 & 0.3930 & 0.0151023\\
+\hline
+\end{tabular}
+
+\hypertarget{probability-of-change-in-finalists}{%
+\subsubsection{Probability of Change in
+Finalists}\label{probability-of-change-in-finalists}}
+
+A potential event of interest is the replacement of one finalist with
+another, taking the chance away from the original finalist to win a
+medal in the finals round. Here we compute the probability that a
+finalist will be replaced with a non-finalist after removing another
+non-finalist from the rankings. Below is a table summarizing the
+results.
+
+\begin{tabular}{r|r|r|r|r}
+\hline
+Dropped Rank & Simulations & Avg. Probability & Median Probability & St. Deviation\\
+\hline
+9 & 100 & 0.0505 & 0.050 & 0.0217597\\
+\hline
+10 & 100 & 0.1538 & 0.150 & 0.0318386\\
+\hline
+11 & 100 & 0.1603 & 0.160 & 0.0343616\\
+\hline
+12 & 100 & 0.1416 & 0.140 & 0.0353259\\
+\hline
+13 & 100 & 0.1365 & 0.140 & 0.0336162\\
+\hline
+14 & 100 & 0.1217 & 0.120 & 0.0344965\\
+\hline
+15 & 100 & 0.1213 & 0.120 & 0.0308353\\
+\hline
+16 & 100 & 0.1159 & 0.120 & 0.0306197\\
+\hline
+17 & 100 & 0.1086 & 0.105 & 0.0314312\\
+\hline
+18 & 100 & 0.1026 & 0.100 & 0.0329867\\
+\hline
+19 & 100 & 0.0886 & 0.090 & 0.0291641\\
+\hline
+20 & 100 & 0.0700 & 0.070 & 0.0258590\\
+\hline
+\end{tabular}
+
+as can be seen, the probability is relatively low when the 9th place
+competitor is removed. My hypothesis is that there is enough separation
+between the original finalists and the 10th+ ranking competitors in the
+individual events that there isn't a high chance that the total rankings
+of the finalists can be overtaken by 10+th place competitors. Note: it
+might be worth computing the probability of getting a low ranking in any
+of the individual events, given that a competitor is a finalist (ie:
+\(P(r_1 > 8 \text{ or } r_2 > 8 \text{ or } r_3 > 8 \text{ }|\text{ } R_T \leq 8)\)).
+
+\hypertarget{probability-that-medalists-are-reordered-or-changed}{%
+\subsubsection{Probability that Medalists are Reordered or
+Changed}\label{probability-that-medalists-are-reordered-or-changed}}
+
+The second scenario of interest is if a medalist is displaced by another
+medalist, or even a non-medalist. Below are the results summarizing the
+probability of this scenario occurring under uniform random ranking
+conditions.
+
+\begin{tabular}{r|r|r|r|r}
+\hline
+Dropped Rank & Simulations & Avg. Probability & Median Probability & St. Deviation\\
+\hline
+4 & 100 & 0.25872 & 0.2580 & 0.0137106\\
+\hline
+5 & 100 & 0.31988 & 0.3190 & 0.0161416\\
+\hline
+6 & 100 & 0.28741 & 0.2870 & 0.0127969\\
+\hline
+7 & 100 & 0.25631 & 0.2545 & 0.0145301\\
+\hline
+8 & 100 & 0.20631 & 0.2050 & 0.0120283\\
+\hline
+\end{tabular}
+
 \hypertarget{data-analysis}{%
@@
 
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-13-1}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-16-1}
 
@@
 
-\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-13}
+\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-16}
 \end{figure}
@@
 
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-16-1}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-19-1}
 
@@
 For each gender competition, each panel represents the rank of the drop-out athlete, with 0 being the original final results.
-Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-16}
+Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-19}
 \end{figure}
@@
 
-\caption{\label{tab:unnamed-chunk-18}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
+\caption{\label{tab:unnamed-chunk-21}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-19}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
+\caption{\label{tab:unnamed-chunk-22}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-20}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
+\caption{\label{tab:unnamed-chunk-23}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-21}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
+\caption{\label{tab:unnamed-chunk-24}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
 \centering

### 2021-08-04 manuscript/Rmd/draft.Rmd
Merge and commit
Table captions

@@
 
-<!-- # https://www.climbing.com/competition/olympics/a-guide-to-the-olympic-climbing-format/ -->
+<!-- https://www.climbing.com/competition/olympics/a-guide-to-the-olympic-climbing-format/ -->
 
@@
 <!-- 
-here are other individual competitions consisted of several events combined that do base their scoring on ranks, such as crossfit competitions.
+There are other individual competitions consisted of several events combined that do base their scoring on ranks, such as crossfit competitions.
 In each event, points are earned based on the competitor's rank in the event based on a scoring table.
@@
 We first look at the chances of advancing to the final round for a qualifier, given that 1) they finish first in any of the three disciplines, and 2) they win the first event (which happens to be speed climbing in reality).
-Our simulation study shows that a climber is almost guaranteed to make the final round if they win at least one of the three climbing concentrations (99.51%, see Table) or if they win the opening event (99.48%, see Table).
-After winning the first event, a climber is also more likely to finish first overall (30%, see Table) than any other positions in the overall qualification rankings.
+Our simulation study shows that a climber is almost guaranteed to make the final round if they win at least one of the three climbing concentrations or if they win the opening event, as the probabilities of advancing for these two situations are 0.9948 and 0.9951, respectively (see Table 2).
+Moreover, after winning the first event, a climber is also more likely to finish first overall (see Table 2) than any other positions in the overall qualification rankings.
 Therefore, this product of ranks scoring format highly rewards climbers who can get off to a strong start, which is a major strategy climbing athletes should take note of.
@@
   kable(col.names = rep(c("Rank", "Count", "Probability", "Cumulative"), 2),
-        caption = "title", 
+        caption = "This table shows every possible qualification rank obtained from simulations for a climber given that a they win at least one discipline or finish first in the first discipline.
+For each rank, information on the number of times that climbers finish at the given rank, the probability of finishing at exactly the given rank, and the probability of finishing at or below the given rank are included.",
         booktabs = TRUE, 
@@
   kable(col.names = rep(c("Rank", "Count", "Probability", "Cumulative"), 2),
-        caption = "title", 
+        caption = "This table shows every possible final rank obtained from simulations for a climber given that a they win at least one discipline or finish first in the first discipline.
+For each rank, information on the number of times that climbers finish at the given rank, the probability of finishing at exactly the given rank, and the probability of finishing at or below the given rank are included.",
         booktabs = TRUE, 
@@
 ### Probability of Reordering in Higher Ranks
+
 We first look at the rate at which the ranks above that of which is dropped are reordered.
+
 ```{r, echo = FALSE}

### 2021-08-04 manuscript/Rmd/draft.tex
Table captions

@@
 shows that a climber is almost guaranteed to make the final round if
-they win at least one of the three climbing concentrations (99.51\%, see
-Table) or if they win the opening event (99.48\%, see Table). After
-winning the first event, a climber is also more likely to finish first
-overall (30\%, see Table) than any other positions in the overall
+they win at least one of the three climbing concentrations or if they
+win the opening event, as the probabilities of advancing for these two
+situations are 0.9948 and 0.9951, respectively (see Table 2). Moreover,
+after winning the first event, a climber is also more likely to finish
+first overall (see Table 2) than any other positions in the overall
 qualification rankings. Therefore, this product of ranks scoring format
@@
 
-\caption{\label{tab:unnamed-chunk-6}title}
+\caption{\label{tab:unnamed-chunk-6}This table shows every possible qualification rank obtained from simulations for a climber given that a they win at least one discipline or finish first in the first discipline.
+For each rank, information on the number of times that climbers finish at the given rank, the probability of finishing at exactly the given rank, and the probability of finishing at or below the given rank are included.}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-9}title}
+\caption{\label{tab:unnamed-chunk-9}This table shows every possible final rank obtained from simulations for a climber given that a they win at least one discipline or finish first in the first discipline.
+For each rank, information on the number of times that climbers finish at the given rank, the probability of finishing at exactly the given rank, and the probability of finishing at or below the given rank are included.}
 \centering

### 2021-08-04 manuscript/Rmd/draft.Rmd
Change theme

@@
 library(kableExtra)
-theme_set(theme_minimal())
+theme_set(theme_bw())
 ```
@@
 
-```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications", fig.height=8, fig.width=6.5}
+```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications", fig.height=8, fig.width=5}
 mcor <- mq %>% 
@@
   ggpairs(diag = "blank",
-          axisLabels = "internal",
-          upper = list(continuous = wrap("cor", method = "kendall", stars = FALSE))) +
-  theme(panel.grid.major = element_blank()) +
+          lower = list(continuous = wrap("points", size = 1)),
+          upper = list(continuous = wrap("cor", method = "kendall", stars = FALSE, size = 3))) +
+  theme(panel.grid.major = element_blank(),
+        axis.text = element_text(size = 6)) +
   labs(y = "Men\n")
@@
   ggpairs(diag = "blank",
-          axisLabels = "internal",
-          upper = list(continuous = wrap("cor", method = "kendall", stars = FALSE))) +
-  theme(panel.grid.major = element_blank())  +
+          lower = list(continuous = wrap("points", size = 1)),
+          upper = list(continuous = wrap("cor", method = "kendall", stars = FALSE, size = 3))) +
+  theme(panel.grid.major = element_blank(),
+        axis.text = element_text(size = 6))  +
   labs(y = "Women\n")

### 2021-08-05 manuscript/Rmd/draft.Rmd
Plenty to edit still

@@
   The purpose of this paper is to investigate the controversial competition format and scoring system of sport climbing, which is one of the sports making its debut on the Olympics stage at Tokyo 2020.
-Climbing Olympians will participate in three disciplines: speed climbing, bouldering, and lead climbing for a single set of medals, and a multiplied score of climbers' rankings in all three disciplines will be used to determine the overall standings.
-This work finds great evidence from historical data and simulations that speed climbing speed climbing should be separated from the combined format and have its own medals.
+Climbing Olympians are required to participate in three disciplines: speed climbing, bouldering, and lead climbing for only a single set of medals, and a multiplied score of climbers' rankings in all three disciplines will be used to determine the overall standings.
+This work finds great evidence from historical data and simulations that the current competition format is putting speed climbers at a disadvantage.
 Furthermore, it is shown that the product of ranks scoring system violates the independence of irrelevant alternatives property.
@@
 
-The 8 climbers with the lowest score (i.e.
-lowest product of ranks across the three disciplines) in the qualification round advance to the finals where they once again compete in all three disciplines.
+The 8 climbers with the lowest score (in terms of product of ranks across the three disciplines) in the qualification round advance to the finals where they once again compete in all three disciplines.
 The final score for each person in the final stage is determined by multiplying their ranks in each discipline, similar to the qualification round, with the main difference being that there are only 8 competitors in the final as opposed to 20 in the qualification round.
@@
 To the best of our knowledge, we know of no sporting event, team or individual, that uses the product of ranks to determine an overall rankings.
-There are examples of team sports that use the sum of ranks to determine the winning team such as cross country, where the team with the lowest sum of ranks of the top five runners is determined to be the winner.
-For more detail on rank sum scoring, see @hammond2007 and @boudreau2018.
+There are examples of team sports that use the sum of ranks to determine the winning team such as cross country, where the team with the lowest sum of ranks of the top five runners is determined to be the winner [@hammond2007; @boudreau2018].
 In addition, some individual sports such as the decathlon and heptathlon rely on a sum of scores from the ten or seven events, however, these scores are not determined based on the ranks of the competitors.
-That is, a decathlete's score is entirely based upon their times, distances, and heights, and their overall score will be exactly the same if the times, distances, and heights remain the same regardless of the performance of other individuals.
-Full details of decathlon scoring can be found in @westera2006.
+That is, a decathlete's score is entirely based upon their times, distances, and heights, and their overall score will be exactly the same if the times, distances, and heights remain the same regardless of the performance of other individuals [@westera2006].
+Furthermore, there are other individual competitions consisted of several events combined that do base their scoring on ranks, such as crossfit competitions.
+In each event, points are earned based on the competitor's rank in the event based on a scoring table, and a competitors final score is based on the sum of their scores across all the events [@crossfit2021].
 
 <!--
-There are other individual competitions consisted of several events combined that do base their scoring on ranks, such as crossfit competitions.
-In each event, points are earned based on the competitor's rank in the event based on a scoring table.
-(https://games.crossfit.com/announcement/12940), and a competitors final score is based on the sum of their scores across all the events.
+Paragraph on scoring methods in other sports: Soccer points: @haugen2008; Cross country scoring: @hammond2007, @boudreau2018; Decathlon scoring: @westera2006
+As a side note, rank product, genes
 
-Paragraph on scoring methods in other sports: Soccer points: @haugen2008; Cross country scoring: @hammond2007, @boudreau2018; Decathlon scoring: @westera2006
-
-As a side note, rank product, genes 
+In particular, we show how this format is putting speed climbers at a disadvantage, as well as address the 
 -->
 
+In this paper, we attempt to use simple statistical analysis to address the limitations of sport climbing's combined competition format and ranking system.
 The manuscript is outlined as follows.
@@
 The main attributes of our datasets are the name and nationality of the climbers; bib number (for some competitions); the finishing place of climbers in speed climbing, bouldering, and lead climbing; the total score (which equals the product of event ranks); and the final rank.
-We utilize this data to compute the correlations between the event ranks and final table position, as well as to look at how often the final orderings change if one athlete is dropped and the ranks for each discipline are re-computed.
+We utilize this data to compute the correlations between the event ranks and final table position, as well as to look at how often the final orderings change if one athlete is dropped and the ranks for each discipline as well as the total score are re-calculated.
 
-Paragraph on IIA/Social Choice
+Since we are interested in applying the "leave-one-climber-out" type of analysis to sport climbing, there is a connection between this situation and the idea of independence of irrelevant alternatives (IIA).
+The IIA criterion is a property of a voting system which states that after a winner is determined, if one of the losing candidates drops out and the votes are recounted, there should not be a change in the winner.
+First mentioned by @arrow1951, the IIA condition is also known as Luce's choice axiom [@luce1959] in probability theory, and it has had a number of applications in the fields of decision theory, economics, and psychology over the years.
 
-Suppose we have 3 people A, B, and C participating in a competition.
+We notice a link between the concept of IIA and the topic of ranking system in sports.
+As an illustration, suppose we have 3 players A, B, and C participating in a competition.
 If A finishes in the first place and C is later disqualified and removed, A should still win.
 If the original winner (A) loses the modified competition (with C removed), then the Independence of Irrelevant Alternatives has been violated.
+For our particular case, this sort of analysis can be helpful in examining the results of a climbing competition, specifically how a disqualification can affect the standings of medalists in the final round.
 
-Economics, psychology, probability theory. Connect with cross country in previous section
-
+<!--
 First mentioned by Arrow (1951) https://cowles.yale.edu/sites/default/files/files/pub/mon/m12-all.pdf
@@
 Ray (1973) https://www.jstor.org/stable/1913820
+-->
 
@@
 
-In addition, the average score for qualification positions 1 to 8 are displayed in Table X.
-We notice that on average, the minimum score that one should aim for in order to move on to the final round is about 434 (for 8th rank).
+In addition, the average score for qualification positions 1 to 8 are displayed in Table 3.
+We observe that on average, the minimum score that one should aim for in order to move on to the final round is about 434 (for 8th rank).
 
@@
 
-```{r, fig.cap="Probability of finishing at every rank given", fig.width = 6.4, fig.height = 3.2}
+```{r, fig.cap="Distribution plots of qualification ranks, given 1) winning any event and 2) winning first event", fig.width = 6.4, fig.height = 3.2}
 win_any_qual %>% 
@@
   geom_vline(xintercept = 8.5) +
-  scale_x_continuous() +
+  scale_x_continuous(breaks = 1:11) +
   labs(x = "Rank",
@@
 
-Regarding the finals, a climber is very likely to finish in the top 3 and hence earn a medal if they win the first event (83.03% chance) or any event (85.01% chance).
-Furthermore, according to our final simulations, in order to obtain a climbing medal, the average scores (rounded down) that put an athlete in position to receive gold, silver, and bronze medals are 9, 20, and 33, respectively (see Table 3).
+Regarding the finals, Table 4 indicates that a climber is very likely to finish in the top 3 and hence earn a medal if they win the first event (83.03% chance) or any event (85.01% chance).
+Furthermore, according to our final simulations, in order to obtain a climbing medal, the average scores (rounded down) that put an athlete in position to receive gold, silver, and bronze medals are 9, 20, and 33, respectively (see Table 5).
 
@@
 
-```{r, fig.cap="Probability of finishing at every rank given", fig.width = 6.4, fig.height = 3.2}
+```{r, fig.cap="Distribution plots of final ranks, given 1) winning any event and 2) wiining first event", fig.width = 6.4, fig.height = 3.2}
 win_any_final %>% 
@@
   geom_vline(xintercept = 3.5) +
-  scale_x_continuous() +
+  scale_x_continuous(breaks = 1:7) +
   labs(x = "Rank",
@@
 
-sum_dat1 <- reorder_all_above %>%
+reorder_all_above %>%
   group_by(dropped_rank) %>%
   rename("Dropped Rank" = dropped_rank) %>%
-  summarize("Simulations" = n()
-            , "Avg. Probability" = mean(displ_prob)
-            , "Median Probability" = median(displ_prob)
-            , "St. Deviation" = sd(displ_prob)
-            )
-
-kable(sum_dat1)
+  summarize("Simulations" = n(), 
+            "Avg. Probability" = mean(displ_prob), 
+            "Median Probability" = median(displ_prob), 
+            "St. Deviation" = sd(displ_prob)) %>% 
+  kable(booktabs = TRUE)
 ```
@@
 
-reorder_dat2 <- lapply(9:20
-                  , function(x) {
-                    data.frame(dropped_rank = x
-                               , displ_prob = pct_displace_finalists[[x-8]]
-                               )
-                    }
-                  ) %>%
-  do.call(rbind, .)
+reorder_dat2 <- lapply(9:20, 
+                       function(x) {
+                         data.frame(dropped_rank = x, 
+                                    displ_prob = pct_displace_finalists[[x - 8]])}) %>%
+  bind_rows()
 
-sum_dat2 <- reorder_dat2 %>%
+reorder_dat2 %>%
   group_by(dropped_rank) %>%
   rename("Dropped Rank" = dropped_rank) %>%
-  summarize("Simulations" = n()
-            , "Avg. Probability" = mean(displ_prob)
-            , "Median Probability" = median(displ_prob)
-            , "St. Deviation" = sd(displ_prob)
-            )
-
-kable(sum_dat2)
-
+  summarize("Simulations" = n(), 
+            "Avg. Probability" = mean(displ_prob), 
+            "Median Probability" = median(displ_prob),
+            "St. Deviation" = sd(displ_prob)) %>% 
+  kable(booktabs = TRUE)
 ```
@@
   rename("Dropped Rank" = dropped_rank) %>%
-  summarize("Simulations" = n()
-            , "Avg. Probability" = mean(displ_prob)
-            , "Median Probability" = median(displ_prob)
-            , "St. Deviation" = sd(displ_prob)
-            ) %>%
- kable()
+  summarize("Simulations" = n(), 
+            "Avg. Probability" = mean(displ_prob), 
+            "Median Probability" = median(displ_prob), 
+            "St. Deviation" = sd(displ_prob)) %>%
+ kable(booktabs = TRUE)
 
@@
 This situation is illustrated by panel 5 of the women's competition, where the fifth-place climber, Krasovskaia, was excluded; and Meul, whose actual final rank was fourth, moved up to the second spot and claim the silver medal.
+Furthermore, this implies that the ranking system of Olympics sport climbing violates the IIA criterion.
 
@@
   facet_wrap(~ rank_drop, nrow = 2) +
-  expand_limits(y = 102, x = 0:7) +
+  expand_limits(y = 104, x = 0:7) +
   scale_fill_manual(values = c("grey", "chocolate")) +

### 2021-08-05 manuscript/Rmd/draft.tex
Plenty to edit still

@@
 the sports making its debut on the Olympics stage at Tokyo 2020.
-Climbing Olympians will participate in three disciplines: speed
-climbing, bouldering, and lead climbing for a single set of medals, and
-a multiplied score of climbers' rankings in all three disciplines will
-be used to determine the overall standings. This work finds great
-evidence from historical data and simulations that speed climbing speed
-climbing should be separated from the combined format and have its own
-medals. Furthermore, it is shown that the product of ranks scoring
-system violates the independence of irrelevant alternatives property.
+Climbing Olympians are required to participate in three disciplines:
+speed climbing, bouldering, and lead climbing for only a single set of
+medals, and a multiplied score of climbers' rankings in all three
+disciplines will be used to determine the overall standings. This work
+finds great evidence from historical data and simulations that the
+current competition format is putting speed climbers at a disadvantage.
+Furthermore, it is shown that the product of ranks scoring system
+violates the independence of irrelevant alternatives property.
 \end{abstract}
@@
 
-The 8 climbers with the lowest score (i.e.~lowest product of ranks
+The 8 climbers with the lowest score (in terms of product of ranks
 across the three disciplines) in the qualification round advance to the
@@
 the lowest sum of ranks of the top five runners is determined to be the
-winner. For more detail on rank sum scoring, see \citet{hammond2007} and
-\citet{boudreau2018}. In addition, some individual sports such as the
-decathlon and heptathlon rely on a sum of scores from the ten or seven
-events, however, these scores are not determined based on the ranks of
-the competitors. That is, a decathlete's score is entirely based upon
-their times, distances, and heights, and their overall score will be
-exactly the same if the times, distances, and heights remain the same
-regardless of the performance of other individuals. Full details of
-decathlon scoring can be found in \citet{westera2006}.
+winner \citep{hammond2007, boudreau2018}. In addition, some individual
+sports such as the decathlon and heptathlon rely on a sum of scores from
+the ten or seven events, however, these scores are not determined based
+on the ranks of the competitors. That is, a decathlete's score is
+entirely based upon their times, distances, and heights, and their
+overall score will be exactly the same if the times, distances, and
+heights remain the same regardless of the performance of other
+individuals \citep{westera2006}. Furthermore, there are other individual
+competitions consisted of several events combined that do base their
+scoring on ranks, such as crossfit competitions. In each event, points
+are earned based on the competitor's rank in the event based on a
+scoring table, and a competitors final score is based on the sum of
+their scores across all the events \citep{crossfit2021}.
 
-The manuscript is outlined as follows. We first begin with some
-descriptions of the data and methods in Section \ref{data-and-methods}.
-Our analyses and results are then presented in Sections
-\ref{simulations} and \ref{data-analysis}. Finally, in Section
+In this paper, we attempt to use simple statistical analysis to address
+the limitations of sport climbing's combined competition format and
+ranking system. The manuscript is outlined as follows. We first begin
+with some descriptions of the data and methods in Section
+\ref{data-and-methods}. Our analyses and results are then presented in
+Sections \ref{simulations} and \ref{data-analysis}. Finally, in Section
 \ref{conclusion-and-discussion}, we summarize up our main findings and
@@
 often the final orderings change if one athlete is dropped and the ranks
-for each discipline are re-computed.
+for each discipline as well as the total score are re-calculated.
 
-Paragraph on IIA/Social Choice
+Since we are interested in applying the ``leave-one-climber-out'' type
+of analysis to sport climbing, there is a connection between this
+situation and the idea of independence of irrelevant alternatives (IIA).
+The IIA criterion is a property of a voting system which states that
+after a winner is determined, if one of the losing candidates drops out
+and the votes are recounted, there should not be a change in the winner.
+First mentioned by \citet{arrow1951}, the IIA condition is also known as
+Luce's choice axiom \citep{luce1959} in probability theory, and it has
+had a number of applications in the fields of decision theory,
+economics, and psychology over the years.
 
-Suppose we have 3 people A, B, and C participating in a competition. If
-A finishes in the first place and C is later disqualified and removed, A
-should still win. If the original winner (A) loses the modified
-competition (with C removed), then the Independence of Irrelevant
-Alternatives has been violated.
-
-Economics, psychology, probability theory. Connect with cross country in
-previous section
-
-First mentioned by Arrow (1951)
-\url{https://cowles.yale.edu/sites/default/files/files/pub/mon/m12-all.pdf}
-
-Luce (1959)
-\url{http://www.scholarpedia.org/article/Luce\%27s_choice_axiom}
-
-and Luce (1977)
-\url{https://www.imbs.uci.edu/files/personnel/luce/pre1990/1977/Luce_JMP_1977a.pdf}
-
-Ray (1973) \url{https://www.jstor.org/stable/1913820}
+We notice a link between the concept of IIA and the topic of ranking
+system in sports. As an illustration, suppose we have 3 players A, B,
+and C participating in a competition. If A finishes in the first place
+and C is later disqualified and removed, A should still win. If the
+original winner (A) loses the modified competition (with C removed),
+then the Independence of Irrelevant Alternatives has been violated. For
+our particular case, this sort of analysis can be helpful in examining
+the results of a climbing competition, specifically how a
+disqualification can affect the standings of medalists in the final
+round.
 
@@
 In addition, the average score for qualification positions 1 to 8 are
-displayed in Table X. We notice that on average, the minimum score that
+displayed in Table 3. We observe that on average, the minimum score that
 one should aim for in order to move on to the final round is about 434
@@
 
-\caption{Probability of finishing at every rank given}\label{fig:unnamed-chunk-7}
+\caption{Distribution plots of qualification ranks, given 1) winning any event and 2) winning first event}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 
-Regarding the finals, a climber is very likely to finish in the top 3
-and hence earn a medal if they win the first event (83.03\% chance) or
-any event (85.01\% chance). Furthermore, according to our final
-simulations, in order to obtain a climbing medal, the average scores
-(rounded down) that put an athlete in position to receive gold, silver,
-and bronze medals are 9, 20, and 33, respectively (see Table 3).
+Regarding the finals, Table 4 indicates that a climber is very likely to
+finish in the top 3 and hence earn a medal if they win the first event
+(83.03\% chance) or any event (85.01\% chance). Furthermore, according
+to our final simulations, in order to obtain a climbing medal, the
+average scores (rounded down) that put an athlete in position to receive
+gold, silver, and bronze medals are 9, 20, and 33, respectively (see
+Table 5).
 
@@
 
-\caption{Probability of finishing at every rank given}\label{fig:unnamed-chunk-10}
+\caption{Distribution plots of final ranks, given 1) winning any event and 2) wiining first event}\label{fig:unnamed-chunk-10}
 \end{figure}
@@
 
-\begin{tabular}{r|r|r|r|r}
-\hline
+\begin{tabular}{rrrrr}
+\toprule
 Dropped Rank & Simulations & Avg. Probability & Median Probability & St. Deviation\\
-\hline
+\midrule
 4 & 100 & 0.25869 & 0.2590 & 0.0145433\\
-\hline
 5 & 100 & 0.34032 & 0.3405 & 0.0151263\\
-\hline
 6 & 100 & 0.38012 & 0.3820 & 0.0152430\\
-\hline
 7 & 100 & 0.40636 & 0.4060 & 0.0149582\\
-\hline
 8 & 100 & 0.39439 & 0.3930 & 0.0151023\\
-\hline
+\bottomrule
 \end{tabular}
@@
 
-\begin{tabular}{r|r|r|r|r}
-\hline
+\begin{tabular}{rrrrr}
+\toprule
 Dropped Rank & Simulations & Avg. Probability & Median Probability & St. Deviation\\
-\hline
+\midrule
 9 & 100 & 0.0505 & 0.050 & 0.0217597\\
-\hline
 10 & 100 & 0.1538 & 0.150 & 0.0318386\\
-\hline
 11 & 100 & 0.1603 & 0.160 & 0.0343616\\
-\hline
 12 & 100 & 0.1416 & 0.140 & 0.0353259\\
-\hline
 13 & 100 & 0.1365 & 0.140 & 0.0336162\\
-\hline
+\addlinespace
 14 & 100 & 0.1217 & 0.120 & 0.0344965\\
-\hline
 15 & 100 & 0.1213 & 0.120 & 0.0308353\\
-\hline
 16 & 100 & 0.1159 & 0.120 & 0.0306197\\
-\hline
 17 & 100 & 0.1086 & 0.105 & 0.0314312\\
-\hline
 18 & 100 & 0.1026 & 0.100 & 0.0329867\\
-\hline
+\addlinespace
 19 & 100 & 0.0886 & 0.090 & 0.0291641\\
-\hline
 20 & 100 & 0.0700 & 0.070 & 0.0258590\\
-\hline
+\bottomrule
 \end{tabular}
@@
 
-\begin{tabular}{r|r|r|r|r}
-\hline
+\begin{tabular}{rrrrr}
+\toprule
 Dropped Rank & Simulations & Avg. Probability & Median Probability & St. Deviation\\
-\hline
+\midrule
 4 & 100 & 0.25872 & 0.2580 & 0.0137106\\
-\hline
 5 & 100 & 0.31988 & 0.3190 & 0.0161416\\
-\hline
 6 & 100 & 0.28741 & 0.2870 & 0.0127969\\
-\hline
 7 & 100 & 0.25631 & 0.2545 & 0.0145301\\
-\hline
 8 & 100 & 0.20631 & 0.2050 & 0.0120283\\
-\hline
+\bottomrule
 \end{tabular}
@@
 was excluded; and Meul, whose actual final rank was fourth, moved up to
-the second spot and claim the silver medal.
+the second spot and claim the silver medal. Furthermore, this implies
+that the ranking system of Olympics sport climbing violates the IIA
+criterion.
 

### 2021-08-05 R/tokyo.R
Tokyo data scraping

@@ -32,12 +32,22 @@ url <- "https://en.wikipedia.org/wiki/Sport_climbing_at_the_2020_Summer_Olympics
 
 
 
-mq <- url %>% 
-  read_html() %>% 
-  html_table() %>% 
-  .[[3]] %>% 
-  row_to_names(row_number = 1) %>% 
-  clean_names() 
+# mq <- url %>% 
+#   read_html() %>% 
+#   html_table() %>% 
+#   .[[3]] %>% 
+#   row_to_names(row_number = 1) %>% 
+#   clean_names() 
+
+mf <- url %>%
+  read_html() %>%
+  html_table() %>%
+  .[[7]] %>%
+  row_to_names(row_number = 1) %>%
+  clean_names() %>% 
+  filter(climber != "Climber") %>% 
+  mutate(rank = 1:8)
+
 
 mq %>% 
   select(climber,
@@ -57,11 +67,12 @@ mq %>%
 
 
 
-df <- mq %>% 
+df <- mf %>% 
   select(climber,
          speed = cp,
          bould = cp_2,
-         lead = cp_3)
+         lead = cp_3) %>% 
+  filter(climber != "Bassa Mawem")
 
 rerank <- list()
 for (i in 1:nrow(df)) {
@@ -115,9 +126,9 @@ rerank_df %>%
   geom_col(show.legend = FALSE) +
   geom_text(aes(label = rank), hjust = -0.2, size = 3) +
   coord_flip() +
-  # geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
-  #           data = ~ filter(., rank_drop %in% c(1, 2, 3, 5, 6)), 
-  #           color = "black", size = 1.5, fill = NA, inherit.aes = FALSE) +
+  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
+            data = ~ filter(., rank_drop %in% 1:6),
+            color = "black", size = 1.5, fill = NA, inherit.aes = FALSE) +
   facet_wrap(~ rank_drop, nrow = 2) +
   expand_limits(y = 150, x = 0:7) +
   # scale_fill_manual(values = c("grey", "chocolate")) +

### 2021-08-05 R/remove_climber_sim.Rmd
simulations added

@@
 Note: it might be worth computing the probability of getting a low ranking in any of the individual events, given that a competitor is a finalist (ie: $P(r_1 > 8 \text{ or } r_2 > 8 \text{ or } r_3 > 8 \text{ }|\text{ } R_T \leq 8)$).
+
+## UPDATE
+Olympic Finalist Bassa Mawem pulled out of the sport climbing finals due to an injury in his left arm.
+Reuters reported that Megos, originally 9th in the total qualification rankings would not be replacing him.
+We suspect that this might be part of the unclear format of the Olympic Sport Climbing format.
+However, according to the climber's rankings, Megos technically should not have been next in line to fill in the missing finalist spot.
+Upon removing B.
+Mawem from the list and reranking the remaining competitors, we find that Chon would have moved into 8th place with a 640 combined score and Megos would have remained in 9th with 648.
+B.
+Mawem originally ranked first in the speed climbing, and Megos did comparatively poorly with 19th place.
+Both performed better than Mawem in bouldering and lead, leaving those rankings unchanged.
+
+Prior, we considered a change in finalists if a non-finalist was removed.
+It is clear now though that the more interesting scenario is would would replace an original finalist who is removed from the rankings.
+Should it be the original 9th place competitor, or does it make more sense with such a scoring system to rerank?
+Surely Chon should be rewarded for his superior performance in speed compared to Megos with the chance to win a medal?
+
+The next set of simulations will evaluate the probability that a removed finalist is replaced with someone other than the original 9th place competitor.
+
+```{r, eval = FALSE, include = FALSE}
+displace_finalist <- function(nsim = 100, drop_rank = 9) {
+  # simulate nsim rankings
+  rank_perms <- lapply(1:nsim, function(x) assign_ranks(c = 20))
+  # drop ith competitor and recalculate ranks
+  dropout_ranks <- lapply(rank_perms, function(x) drop_climber(x, r = drop_rank))
+  # compare 1st-3rd place & calculate % not matched
+  pct_displ <- sapply(1:nsim, 
+                      function(x) {
+                        rank_perms[[x]]$climber[9] %in% dropout_ranks[[x]]$climber[1:8]
+                      }
+  ) %>%
+    sum()/nsim
+
+  return(pct_displ)
+}
+
+pct_displace_finalists <- lapply(1:8, function(x) replicate(100, displace_finalist(drop_rank = x)))
+
+nineth_to_finals <- lapply(1:8
+                  , function(x) {
+                    data.frame(dropped_rank = x, nineth_to_fin = pct_displace_finalists[[x]])
+                    }
+                  ) %>%
+  do.call(rbind, .)
+
+#save(nineth_to_finals, file = "nineth_to_finals.rdata")
+```
+
+```{r, echo = FALSE}
+load("nineth_to_finals.rdata")
+
+nineth_to_finals %>%
+  ggplot(aes(x = dropped_rank, y = nineth_to_fin)) +
+  geom_boxplot(aes(group = dropped_rank)) +
+  theme_minimal() +
+  xlab("Dropped Finalist Rank") +
+  ylab("Probability that 9th Place Moves to Finals")
+```
 

### 2021-08-06 manuscript/Rmd/draft.Rmd

@@
 
+Olympic Finalist Bassa Mawem pulled out of the sport climbing finals due to an injury in his left arm.
+Reuters reported that Megos, originally 9th in the total qualification rankings would not be replacing him.
+We suspect that this might be part of the unclear format of the Olympic Sport Climbing format.
+However, according to the climber's rankings, Megos technically should not have been next in line to fill in the missing finalist spot.
+Upon removing B.
+Mawem from the list and reranking the remaining competitors, we find that Chon would have moved into 8th place with a 640 combined score and Megos would have remained in 9th with 648.
+B.
+Mawem originally ranked first in the speed climbing, and Megos did comparatively poorly with 19th place.
+Both performed better than Mawem in bouldering and lead, leaving those rankings unchanged.
+
+Prior, we considered a change in finalists if a non-finalist was removed.
+It is clear now though that the more interesting scenario is would would replace an original finalist who is removed from the rankings.
+Should it be the original 9th place competitor, or does it make more sense with such a scoring system to rerank?
+Surely Chon should be rewarded for his superior performance in speed compared to Megos with the chance to win a medal?
+
+The next set of simulations will evaluate the probability that a removed finalist is replaced with someone other than the original 9th place competitor.
+
+```{r, echo = FALSE}
+load("draft_files/nineth_to_finals.rdata")
+
+nineth_to_finals %>%
+  ggplot(aes(x = dropped_rank, y = nineth_to_fin)) +
+  geom_boxplot(aes(group = dropped_rank)) +
+  theme_minimal() +
+  xlab("Dropped Finalist Rank") +
+  ylab("Probability that 9th Place Moves to Finals")
+```
+
 ### Probability that Medalists are Reordered or Changed

### 2021-08-06 R/tokyo.R
Update Tokyo scraping script

@@ -1,52 +1,76 @@
 
-
-Overall <- 1:20
-Speed <- c(3, 2, 6, 12, 18, 7, 1, 10, 19, 5, 4, 11, 16, 20, 8, 9, 14, 15, 13, 17)
-Bouldering <- c(1, 2, 5, 7, 3, 14, 18, 11, 6, 10, 17, 9, 4, 8, 13, 16, 15, 12, 19.5, 19.5)
-Lead <- c(11, 14, 2, 1, 4, 3, 20, 5, 6, 16, 13, 9, 15, 7, 12, 10, 8, 17, 18, 19)
-
-
-tibble(Overall, Speed, Bouldering, Lead) %>% 
-  ggpairs(diag = "blank",
-          axisLabels = "none",
-          upper = list(continuous = wrap("cor", method = "kendall", stars = FALSE))) +
-  theme_light() + ggtitle("2020 Olympics - Men's Qualification")
-
-
-tibble(Overall, Speed, Bouldering, Lead) %>% 
-  # mutate(Score = Speed * Bouldering * Lead)
-  pivot_longer(!Overall) %>% 
-  ggplot(aes(x = name, y = value)) +
-  geom_bar(stat = "identity", position = "dodge")
   
-
+# DATA
 
 
 library(tidyverse)
 library(rvest)
 library(janitor)
-library(viridis)
 
 
-url <- "https://en.wikipedia.org/wiki/Sport_climbing_at_the_2020_Summer_Olympics_%E2%80%93_Men%27s"
+men_url <- "https://en.wikipedia.org/wiki/Sport_climbing_at_the_2020_Summer_Olympics_%E2%80%93_Men%27s"
 
-
-
-# mq <- url %>% 
-#   read_html() %>% 
-#   html_table() %>% 
-#   .[[3]] %>% 
-#   row_to_names(row_number = 1) %>% 
-#   clean_names() 
-
-mf <- url %>%
+mq <- men_url %>%
   read_html() %>%
   html_table() %>%
   .[[7]] %>%
   row_to_names(row_number = 1) %>%
   clean_names() %>% 
+  rename(overall = rank, 
+         speed = cp,
+         bouldering = cp_2,
+         lead = cp_3,
+         total = total_11)
+  
+
+mf <- men_url %>%
+  read_html() %>%
+  html_table() %>%
+  .[[8]] %>%
+  row_to_names(row_number = 1) %>%
+  clean_names() %>% 
   filter(climber != "Climber") %>% 
-  mutate(rank = 1:8)
[…]

### 2021-08-06 R/tokyo.R
Tokyo data

@@ -1,6 +1,6 @@
 
   
-# DATA
+# DATA scraping
 
 
 library(tidyverse)

### 2021-08-06 R/pca.R
PCA

@@ -1,0 +1,87 @@
+library(tidyverse)
+library(broom)
+theme_set(theme_light())
+
+wq <- read_csv("data/2020_olympics/wq.csv")
+wq_cleaned <- wq %>%
+  mutate(
+    a = as.numeric(str_remove_all(a, "[A-z]")),
+    b = as.numeric(str_remove_all(b, "[A-z]")),
+    speed_best = ifelse(is.na(a), b,
+                        ifelse(is.na(b), a,
+                               ifelse(a < b, a, b))),
+    bould_tops = as.numeric(str_sub(results, 1, 1)),
+    lead_hr = ifelse(
+      str_detect(hr, "\\+"),
+      as.numeric(str_sub(hr, 1, nchar(hr) - 1)) + 0.5,
+      as.numeric(hr)
+    )
+  ) %>%
+  select(climber, speed_best, bould_tops, lead_hr) %>%
+  column_to_rownames(var = "climber")
+
+# pca fit
+wq_pca <- wq_cleaned %>%
+  prcomp(scale = TRUE)
+
+# plot
+
+library(ggfortify)
+wq_pca %>% 
+  autoplot(label = TRUE, 
+           label.size = 2,
+           loadings = TRUE,
+           loadings.label = TRUE, 
+           loadings.label.size = 3)
+
+
+# percent variability
+wq_pca %>% 
+  tidy(matrix = "eigenvalues")
+
+wq_pca %>% 
+  tidy(matrix = "eigenvalues") %>% 
+  ggplot(aes(PC, percent)) +
+  geom_col()
+
+# scores and loadings
+wq_pca %>% 
+  tidy(matrix = "scores")
+
+wq_pca %>% 
+  tidy(matrix = "loadings")
+
+# how each variable contributes to each PC
+wq_pca %>% 
+  tidy(matrix = "loadings") %>% 
+  ggplot(aes(value, column)) +
+  facet_wrap(~ PC) +
+  geom_col()
+
+wq_pca %>% 
+  tidy(matrix = "rotation")
+
+# attach pca fit to original data
+wq_pca %>% 
+  augment(wq_cleaned)
+
+
+
+# https://clauswilke.com/blog/2020/09/07/pca-tidyverse-style/
+
+# arrow_style <- arrow(
+#   angle = 20, ends = "first", type = "closed", length = grid::unit(8, "pt")
+# )
+# 
+# wq_pca %>%
[…]

### 2021-08-06 R/tokyo.R

@@ -65,7 +65,7 @@ wf <- women_url %>%
          lead = cp_3)
 
 
-write_csv(wf, "data/2020_olympics/wf.csv")
+write_csv(mf, "data/2020_olympics/mf.csv")
 
 
 

### 2021-08-06 R/pca.R
PCA

@@ -20,12 +20,13 @@ wq_cleaned <- wq %>%
   select(climber, speed_best, bould_tops, lead_hr) %>%
   column_to_rownames(var = "climber")
 
+# SOME EDA
+
 # pca fit
 wq_pca <- wq_cleaned %>%
   prcomp(scale = TRUE)
 
 # plot
-
 library(ggfortify)
 wq_pca %>% 
   autoplot(label = TRUE, 

### 2021-08-06 R/pca.R
PCA

@@ -2,8 +2,28 @@ library(tidyverse)
 library(broom)
 theme_set(theme_light())
 
+# Women's qual
 wq <- read_csv("data/2020_olympics/wq.csv")
 wq_cleaned <- wq %>%
+  mutate(
+    a = as.numeric(str_remove_all(a, "[A-z]")),
+    b = as.numeric(str_remove_all(b, "[A-z]")),
+    speed_best = ifelse(is.na(a), b,
+                        ifelse(is.na(b), a,
+                               ifelse(a < b, a, b))),
+    bould_tops = as.numeric(str_sub(results, 1, 1)),
+    lead_hr = ifelse(
+      str_detect(hr, "\\+"),
+      as.numeric(str_sub(hr, 1, nchar(hr) - 1)) + 0.5,
+      as.numeric(hr)
+    )
+  ) %>%
+  select(climber, speed_best, bould_tops, lead_hr) %>%
+  column_to_rownames(var = "climber")
+
+# Men's qual
+mq <- read_csv("data/2020_olympics/mq.csv")
+mq_cleaned <- mq %>%
   mutate(
     a = as.numeric(str_remove_all(a, "[A-z]")),
     b = as.numeric(str_remove_all(b, "[A-z]")),
@@ -65,24 +85,3 @@ wq_pca %>%
 # attach pca fit to original data
 wq_pca %>% 
   augment(wq_cleaned)
-
-
-
-# https://clauswilke.com/blog/2020/09/07/pca-tidyverse-style/
-
-# arrow_style <- arrow(
-#   angle = 20, ends = "first", type = "closed", length = grid::unit(8, "pt")
-# )
-# 
-# wq_pca %>%
-#   tidy(matrix = "rotation") %>%
-#   pivot_wider(names_from = "PC", names_prefix = "PC", values_from = "value") %>%
-#   ggplot(aes(PC1, PC2)) +
-#   geom_segment(xend = 0, yend = 0, arrow = arrow_style) +
-#   geom_text(
-#     aes(label = column),
-#     hjust = 1, nudge_x = -0.02, 
-#     color = "#904C2F"
-#   ) +
-#   coord_fixed() +
-#   xlim(-1.25, .5) + ylim(-.5, 1)

### 2021-08-06 R/tokyo_EDA.R
Clean up Tokyo scripts

@@ -1,77 +1,4 @@
-
-  
-# DATA scraping
-
-
-library(tidyverse)
-library(rvest)
-library(janitor)
-
-
-men_url <- "https://en.wikipedia.org/wiki/Sport_climbing_at_the_2020_Summer_Olympics_%E2%80%93_Men%27s"
-
-mq <- men_url %>%
-  read_html() %>%
-  html_table() %>%
-  .[[7]] %>%
-  row_to_names(row_number = 1) %>%
-  clean_names() %>% 
-  rename(overall = rank, 
-         speed = cp,
-         bouldering = cp_2,
-         lead = cp_3,
-         total = total_11)
-  
-
-mf <- men_url %>%
-  read_html() %>%
-  html_table() %>%
-  .[[8]] %>%
-  row_to_names(row_number = 1) %>%
-  clean_names() %>% 
-  filter(climber != "Climber") %>% 
-  mutate(rank = 1:8) %>% 
-  rename(overall = rank, 
-         speed = cp,
-         bouldering = cp_2,
-         lead = cp_3,
-         total = total_16)
-
-women_url <- "https://en.wikipedia.org/wiki/Sport_climbing_at_the_2020_Summer_Olympics_%E2%80%93_Women%27s_combined"
-
-wq <- women_url %>% 
-  read_html() %>% 
-  html_table() %>% 
-  .[[7]] %>% 
-  row_to_names(row_number = 1) %>% 
-  clean_names() %>% 
-  rename(overall = rank, 
-         speed = cp,
-         bouldering = cp_2,
-         lead = cp_3,
-         total = total_10)
-
-wf <- women_url %>% 
-  read_html() %>% 
-  html_table() %>% 
-  .[[8]] %>% 
-  row_to_names(row_number = 1) %>% 
-  clean_names() %>% 
-  filter(climber != "Climber") %>% 
-  mutate(rank = 1:8) %>% 
-  rename(overall = rank, 
-         speed = cp,
-         bouldering = cp_2,
-         lead = cp_3)
-
-
-write_csv(mf, "data/2020_olympics/mf.csv")
-
-
-
-
-#####
-
+# load data from data folder first
 
 mq %>% 
   select(climber,
@@ -198,7 +125,7 @@ wf %>%
           axisLabels = "none",
           upper = list(continuous = wrap("cor", method = "kendall", stars = FALSE))) +
   theme_light() + ggtitle("2020 Olympics - Women's Qualification")
-  
+
 
 rerank <- list()
 for (i in 1:nrow(df)) {

### 2021-08-13 R/climber_swaps.Rmd
results after swapping of first/second place in each event

---
title: "swaps"
author: "Hannah Butler"
date: "8/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
load(file = "2020_olympics.rdata")
```

### Men's Olympic Finals

```{r, echo = FALSE}
mens_finals_events <- mens_finals %>%
  select(climber, og_speed = speed, og_bould = bouldering, og_lead = lead)

mens_swaps12 <- mens_finals_events %>%
  mutate(swap_speed = ifelse(climber == "Alberto Ginés López", 2, og_speed)
         , swap_speed = ifelse(climber == "Tomoa Narasaki", 1, swap_speed)
         , swap_bould = ifelse(climber == "Nathaniel Coleman", 2, og_bould)
         , swap_bould = ifelse(climber == "Mickaël Mawem", 1, swap_bould)
         , swap_lead = ifelse(climber == "Jakob Schubert", 2, og_lead)
         , swap_lead = ifelse(climber == "Adam Ondra", 1, swap_lead)
         ) %>%
  pivot_longer(og_speed:swap_lead) %>%
  rename(swap = name) %>%
  mutate(event = sub(".*_", "", swap)) %>%
  mutate(swap = ifelse(substr(swap, 1,2) == "og", "original", sub(".*_", "", swap))) %>%
  pivot_wider(names_from = event, values_from = value) %>%
  group_by(climber) %>%
  mutate(speed = ifelse(is.na(speed), speed[swap == "original"], speed)
         , bould = ifelse(is.na(bould), bould[swap == "original"], bould)
         , lead = ifelse(is.na(lead), lead[swap == "original"], lead)
         ) %>%
  ungroup() %>%
  group_by(swap) %>%
  mutate(total = speed*bould*lead
         , total_rank = rank(total)
         ) %>%
  arrange(swap, total_rank)

mens_swaps12 %>% 
  filter(swap == "original") %>%
  knitr::kable()

mens_swaps12 %>% 
  filter(swap == "speed") %>%
  knitr::kable()

mens_swaps12 %>% 
  filter(swap == "bould") %>%
  knitr::kable()

mens_swaps12 %>% 
  filter(swap == "lead") %>%
  knitr::kable()
```

### Women's Olympic Finals

```{r, echo = FALSE}
womens_finals_events <- womens_finals %>%
  select(climber, og_speed = speed, og_bould = bouldering, og_lead = lead)

womens_swaps12 <- womens_finals_events %>%
  mutate(swap_speed = ifelse(climber == "Aleksandra Mirosław", 2, og_speed)
         , swap_speed = ifelse(climber == "Anouck Jaubert", 1, swap_speed)
         , swap_bould = ifelse(climber == "Janja Garnbret", 2, og_bould)
         , swap_bould = ifelse(climber == "Brooke Raboutou", 1, swap_bould)
         , swap_lead = ifelse(climber == "Janja Garnbret", 2, og_lead)
         , swap_lead = ifelse(climber == "Seo Chae-hyun", 1, swap_lead)
         ) %>%
  pivot_longer(og_speed:swap_lead) %>%
  rename(swap = name) %>%
  mutate(event = sub(".*_", "", swap)) %>%
  mutate(swap = ifelse(substr(swap, 1,2) == "og", "original", sub(".*_", "", swap))) %>%
  pivot_wider(names_from = event, values_from = value) %>%
  group_by(climber) %>%
  mutate(speed = ifelse(is.na(speed), speed[swap == "original"], speed)
         , bould = ifelse(is.na(bould), bould[swap == "original"], bould)
         , lead = ifelse(is.na(lead), lead[swap == "original"], lead)
         ) %>%
  ungroup() %>%
  group_by(swap) %>%
  mutate(total = speed*bould*lead
         , total_rank = rank(total)
         ) %>%
  arrange(swap, total_rank)

womens_swaps12 %>% 
  filter(swap == "original") %>%
  knitr::kable()

womens_swaps12 %>% 
  filter(swap == "speed") %>%
  knitr::kable()

womens_swaps12 %>% 
  filter(swap == "bould") %>%
  knitr::kable()

womens_swaps12 %>% 
  filter(swap == "lead") %>%
  knitr::kable()
```

### How unstable are the medalists in their positions?

### 2021-08-13 R/climber_swaps.Rmd
start some sort of sensitivity analysis of total ranking based on event

@@
 
-### How unstable are the medalists in their positions?
+### How volatile are the scores depending on the event?
+
+```{r}
+# swaps 2 scores
+swap_cons <- function(n, n1, n2) {
+  if (n == n1) {
+    return(n2)
+  } else if (n == n2) {
+    return(n1)
+  } else {
+    return(n)
+  }
+}
+```
+
+below is a table summarizing the results for all consecutive ranking swaps in each event, holding the other 2 events constant.
+
+```{r, echo = FALSE}
+mens_swaps_all <- 
+  mens_swaps12 %>% 
+  filter(swap == "original") %>%
+  group_by(climber) %>%
+  mutate(speed_12 = swap_cons(speed, 1, 2)
+         , speed_23 = swap_cons(speed, 2, 3)
+         , speed_34 = swap_cons(speed, 3, 4)
+         , speed_45 = swap_cons(speed, 4, 5)
+         , speed_56 = swap_cons(speed, 5, 6)
+         , speed_67 = swap_cons(speed, 6, 7)
+         ) %>%
+  mutate(bould_12 = swap_cons(bould, 1, 2)
+         , bould_23 = swap_cons(bould, 2, 3)
+         , bould_34 = swap_cons(bould, 3, 4)
+         , bould_45 = swap_cons(bould, 4, 5)
+         , bould_56 = swap_cons(bould, 5, 6)
+         , bould_67 = swap_cons(bould, 6, 7)
+         ) %>%
+  mutate(lead_12 = swap_cons(lead, 1, 2)
+         , lead_23 = swap_cons(lead, 2, 3)
+         , lead_34 = swap_cons(lead, 3, 4)
+         , lead_45 = swap_cons(lead, 4, 5)
+         , lead_56 = swap_cons(lead, 5, 6)
+         , lead_67 = swap_cons(lead, 6, 7)
+         ) %>%
+  ungroup() %>%
+  select(-c(swap, total_rank, total)) %>%
+  pivot_longer(speed:lead_67) %>%
+  mutate(event = sub("_.*", "", name)
+         , swap = ifelse(name %in% c("speed", "bould", "lead"), "original", name)
+         ) %>%
+  select(-name) %>%
+  pivot_wider(names_from = event, values_from = value) %>%
+  group_by(climber) %>%
+  mutate(speed = ifelse(is.na(speed), speed[swap == "original"], speed)
+         , bould = ifelse(is.na(bould), bould[swap == "original"], bould)
+         , lead = ifelse(is.na(lead), lead[swap == "original"], lead)
+         ) %>%
+  ungroup() %>%
+  mutate(total_score = speed*bould*lead) %>%
+  group_by(swap) %>%
+  mutate(total_rank = rank(total_score)) %>%
+  ungroup() %>%
+  arrange(swap, total_rank)  %>%
+  separate(swap, into = c("swap", "position"), sep = "_")
+
+mens_swaps_all %>% knitr::kable()
+```

### 2021-08-23 manuscript/Rmd/draft.Rmd
Start PCA

@@
 
+Paragraph on rank swapping
+
+Paragraph on PCA
+
 <!--
@@
 We first look at the rate at which the ranks above that of which is dropped are reordered.
+
+(Need a more detailed description)
 
@@
 
+## Rank Swapping Analysis
+
+## Principal Component Analysis
+
+Olympic results
+
+```{r}
+# Women's qual
+wq <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2020_olympics/wq.csv")
+wq_cleaned <- wq %>%
+  mutate(
+    a = as.numeric(str_remove_all(a, "[A-z]")),
+    b = as.numeric(str_remove_all(b, "[A-z]")),
+    speed_best = ifelse(is.na(a), b,
+                        ifelse(is.na(b), a,
+                               ifelse(a < b, a, b))),
+    bould_tops = as.numeric(str_sub(results, 1, 1)),
+    lead_hr = ifelse(
+      str_detect(hr, "\\+"),
+      as.numeric(str_sub(hr, 1, nchar(hr) - 1)) + 0.5,
+      as.numeric(hr)
+    )
+  ) %>%
+  select(climber, speed_best, bould_tops, lead_hr) %>%
+  column_to_rownames(var = "climber")
+
+# Men's qual
+mq <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2020_olympics/mq.csv")
+mq_cleaned <- mq %>%
+  mutate(
+    a = as.numeric(str_remove_all(a, "[A-z]")),
+    b = as.numeric(str_remove_all(b, "[A-z]")),
+    speed_best = ifelse(is.na(a), b,
+                        ifelse(is.na(b), a,
+                               ifelse(a < b, a, b))),
+    bould_tops = as.numeric(str_sub(results, 1, 1)),
+    lead_hr = ifelse(
+      str_detect(hr, "\\+"),
+      as.numeric(str_sub(hr, 1, nchar(hr) - 1)) + 0.5,
+      as.numeric(hr)
+    )
+  ) %>%
+  select(climber, speed_best, bould_tops, lead_hr) %>%
+  column_to_rownames(var = "climber")
+```
+
+```{r}
+library(broom)
+```
+
+```{r, fig.cap= "Women's Qualification PCA", fig.width = 6, fig.height = 5}
+# pca fit
+wq_pca <- wq_cleaned %>%
+  prcomp(scale = TRUE)
+
+# plot
+library(ggfortify)
+wq_pca %>% 
+  autoplot(label = TRUE, 
+           label.size = 3,
+           label.repel = TRUE,
+           loadings = TRUE,
+           loadings.label = TRUE, 
+           loadings.label.size = 5)
+```
+
+```{r, fig.cap= "Men's Qualification PCA", fig.width = 6, fig.height = 5}
+# pca fit
+mq_pca <- mq_cleaned %>%
+  prcomp(scale = TRUE)
+
+# plot
+library(ggfortify)
+mq_pca %>% 
+  autoplot(label = TRUE, 
+           label.size = 3,
+           label.repel = TRUE,
+           loadings = TRUE,
+           loadings.label = TRUE, 
+           loadings.label.size = 5)
+```
+
 # Conclusion and Discussion

### 2021-08-23 manuscript/Rmd/draft.tex
Start PCA

@@
 
+Paragraph on rank swapping
+
+Paragraph on PCA
+
 \hypertarget{simulations}{%
@@
 
+(Need a more detailed description)
+
 \begin{tabular}{rrrrr}
@@
 
+Olympic Finalist Bassa Mawem pulled out of the sport climbing finals due
+to an injury in his left arm. Reuters reported that Megos, originally
+9th in the total qualification rankings would not be replacing him. We
+suspect that this might be part of the unclear format of the Olympic
+Sport Climbing format. However, according to the climber's rankings,
+Megos technically should not have been next in line to fill in the
+missing finalist spot. Upon removing B. Mawem from the list and
+reranking the remaining competitors, we find that Chon would have moved
+into 8th place with a 640 combined score and Megos would have remained
+in 9th with 648. B. Mawem originally ranked first in the speed climbing,
+and Megos did comparatively poorly with 19th place. Both performed
+better than Mawem in bouldering and lead, leaving those rankings
+unchanged.
+
+Prior, we considered a change in finalists if a non-finalist was
+removed. It is clear now though that the more interesting scenario is
+would would replace an original finalist who is removed from the
+rankings. Should it be the original 9th place competitor, or does it
+make more sense with such a scoring system to rerank? Surely Chon should
+be rewarded for his superior performance in speed compared to Megos with
+the chance to win a medal?
+
+The next set of simulations will evaluate the probability that a removed
+finalist is replaced with someone other than the original 9th place
+competitor.
+
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-14-1} \end{center}
+
 \hypertarget{probability-that-medalists-are-reordered-or-changed}{%
@@
 
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-16-1}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-17-1}
 
@@
 
-\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-16}
+\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-17}
 \end{figure}
@@
 
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-19-1}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-20-1}
 
@@
 For each gender competition, each panel represents the rank of the drop-out athlete, with 0 being the original final results.
-Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-19}
+Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-20}
+\end{figure}
+
+\hypertarget{rank-swapping-analysis}{%
+\subsection{Rank Swapping Analysis}\label{rank-swapping-analysis}}
+
+\hypertarget{principal-component-analysis}{%
+\subsection{Principal Component
+Analysis}\label{principal-component-analysis}}
+
+Olympic results
+
+\begin{figure}
+
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-23-1}
+
+}
+
+\caption{Women's Qualification PCA}\label{fig:unnamed-chunk-23}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-24-1}
+
+}
+
+\caption{Men's Qualification PCA}\label{fig:unnamed-chunk-24}
 \end{figure}
@@
 
-\caption{\label{tab:unnamed-chunk-21}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
+\caption{\label{tab:unnamed-chunk-26}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-22}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
+\caption{\label{tab:unnamed-chunk-27}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-23}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
+\caption{\label{tab:unnamed-chunk-28}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-24}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
+\caption{\label{tab:unnamed-chunk-29}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
 \centering

### 2021-08-31 manuscript/Rmd/draft.Rmd

@@
 
+Women: lead and bouldering belong to the same component, the other is speed.
+
+"overall ability"
+
+"power"
+
 ```{r}
@@
 
+How each variable contributes to each principal component
+
+```{r}
+tidy(wq_pca, matrix = "loadings") %>%
+  ggplot(aes(value, column)) +
+  facet_wrap(~ PC) +
+  geom_col()
+```
+
 ```{r, fig.cap= "Men's Qualification PCA", fig.width = 6, fig.height = 5}
@@
            loadings.label.size = 5)
+```
+
+```{r}
+tidy(mq_pca, matrix = "loadings") %>%
+  ggplot(aes(value, column)) +
+  facet_wrap(~ PC) +
+  geom_col()
 ```

### 2021-08-31 manuscript/Rmd/draft.tex

@@
 
+Women: lead and bouldering belong to the same component, the other is
+speed.
+
+``overall ability''
+
+``power''
+
 \begin{figure}
@@
 
+How each variable contributes to each principal component
+
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-24-1} \end{center}
+
 \begin{figure}
 
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-24-1}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-25-1}
 
@@
 
-\caption{Men's Qualification PCA}\label{fig:unnamed-chunk-24}
+\caption{Men's Qualification PCA}\label{fig:unnamed-chunk-25}
 \end{figure}
+
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-26-1} \end{center}
 
@@
 
-\caption{\label{tab:unnamed-chunk-26}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
+\caption{\label{tab:unnamed-chunk-28}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-27}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
+\caption{\label{tab:unnamed-chunk-29}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-28}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
+\caption{\label{tab:unnamed-chunk-30}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
 \centering
@@
 
-\caption{\label{tab:unnamed-chunk-29}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
+\caption{\label{tab:unnamed-chunk-31}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
 \centering

### 2021-09-05 manuscript/Rmd/draft.Rmd
New abstract

@@
 abstract: |
-  The purpose of this paper is to investigate the controversial competition format and scoring system of sport climbing, which is one of the sports making its debut on the Olympics stage at Tokyo 2020.
-Climbing Olympians are required to participate in three disciplines: speed climbing, bouldering, and lead climbing for only a single set of medals, and a multiplied score of climbers' rankings in all three disciplines will be used to determine the overall standings.
-This work finds great evidence from historical data and simulations that the current competition format is putting speed climbers at a disadvantage.
-Furthermore, it is shown that the product of ranks scoring system violates the independence of irrelevant alternatives property.
+  Sport climbing, which made its Olympic debut at the 2020 Summer Games, generally consists of three separate disciplines: speed climbing, bouldering, and lead climbing.
+However, the International Olympic Committee (IOC) only allowed one set of medals per gender for sport climbing at Tokyo 2020.
+As a result, the governing body of sport climbing, rather than choosing only one of the three disciplines to include in the Olympics, decided to create a competition combining all three disciplines.
+In order to determine a winner, a combined scoring system was created using the product of the ranks across the three disciplines to determine an overall score for each climber.
+In this work, the sport climbing scoring system is evaluated through simulations to investigate its general features such as the advancement probabilities given certain placements and the sensitivity of the scoring system to small changes.
+In addition, analyses of historical climbing contest results are presented and real examples of violations of the independence of irrelevant alternatives are illustrated.
+Finally, this work finds evidence that the current competition format is putting speed climbers at a disadvantage.
 
@@
 ```
+
+gold medal
+
+order of medalists
+
+3 medalists
+
+any rankings
+
+Structure:
+
+sensitivity to:
+
+results
+
+tournament format

### 2021-09-05 manuscript/Rmd/draft.tex
New abstract

@@
 \begin{abstract}
-The purpose of this paper is to investigate the controversial
-competition format and scoring system of sport climbing, which is one of
-the sports making its debut on the Olympics stage at Tokyo 2020.
-Climbing Olympians are required to participate in three disciplines:
-speed climbing, bouldering, and lead climbing for only a single set of
-medals, and a multiplied score of climbers' rankings in all three
-disciplines will be used to determine the overall standings. This work
-finds great evidence from historical data and simulations that the
-current competition format is putting speed climbers at a disadvantage.
-Furthermore, it is shown that the product of ranks scoring system
-violates the independence of irrelevant alternatives property.
+Sport climbing, which made its Olympic debut at the 2020 Summer Games,
+generally consists of three separate disciplines: speed climbing,
+bouldering, and lead climbing. However, the International Olympic
+Committee (IOC) only allowed one set of medals per gender for sport
+climbing at Tokyo 2020. As a result, the governing body of sport
+climbing, rather than choosing only one of the three disciplines to
+include in the Olympics, decided to create a competition combining all
+three disciplines. In order to determine a winner, a combined scoring
+system was created using the product of the ranks across the three
+disciplines to determine an overall score for each climber. In this
+work, the sport climbing scoring system is evaluated through simulations
+to investigate its general features such as the advancement
+probabilities given certain placements and the sensitivity of the
+scoring system to small changes. In addition, analyses of historical
+climbing contest results are presented and real examples of violations
+of the independence of irrelevant alternatives are illustrated. Finally,
+this work finds evidence that the current competition format is putting
+speed climbers at a disadvantage.
 \end{abstract}
@@
 
+gold medal
+
+order of medalists
+
+3 medalists
+
+any rankings
+
+Structure:
+
+sensitivity to:
+
+results
+
+tournament format
+
 \bibliographystyle{agsm}

### 2021-09-07 manuscript/Rmd/draft.tex
Restructure manuscript

@@
 
-The decision to combine the three climbing events and only award one set
-of medals for both men's and women's events in the Olympics has received
-a large amount of criticism from climbing athletes all over the world.
-In a series of interviews conducted by Climbing Magazine in 2016, a
-number of climbers shared their thoughts and concerns about the new
-Olympics climbing format. Climber Lynn Hill compared the idea of
-combining speed climbing, bouldering, and lead climbing to ``asking a
-middle distance runner to compete in the sprint''. She then added
-``Speed climbing is a sport within our sport''. Other climbers also hold
-the same opinion as Hill regarding speed climbing, using words and
-phrases like ``bogus'', ``a bummer'', ``less than ideal'', ``not in
-support'', and ``cheesy and unfair'' to describe the new combined
-competition format. Courtney Woods stated ``Speed climbers will have the
-biggest disadvantage because their realm isn't based on difficult
-movements''. Mike Doyle believed ``Honestly, the people that will suffer
-the most are the ones that focus only on speed climbing. Those
-skills/abilities don't transfer as well to the other disciplines''. Most
-climbers also expressed their hope for a change in the format in future
-tournaments, with some calling for each discipline to have its own set
-of medals. See \citet{blanchard2016} for more information on these
-interviews.
-
 At the 2020 Summer Olympics, both sport climbing competitions for male
@@
 their scores across all the events \citep{crossfit2021}.
+
+The decision to combine the three climbing events and only award one set
+of medals for both men's and women's events in the Olympics has received
+a large amount of criticism from climbing athletes all over the world.
+In a series of interviews conducted by Climbing Magazine in 2016, a
+number of climbers shared their thoughts and concerns about the new
+Olympics climbing format. Climber Lynn Hill compared the idea of
+combining speed climbing, bouldering, and lead climbing to ``asking a
+middle distance runner to compete in the sprint''. She then added
+``Speed climbing is a sport within our sport''. Other climbers also hold
+the same opinion as Hill regarding speed climbing, using words and
+phrases like ``bogus'', ``a bummer'', ``less than ideal'', ``not in
+support'', and ``cheesy and unfair'' to describe the new combined
+competition format. Courtney Woods stated ``Speed climbers will have the
+biggest disadvantage because their realm isn't based on difficult
+movements''. Mike Doyle believed ``Honestly, the people that will suffer
+the most are the ones that focus only on speed climbing. Those
+skills/abilities don't transfer as well to the other disciplines''. Most
+climbers also expressed their hope for a change in the format in future
+tournaments, with some calling for each discipline to have its own set
+of medals. See \citet{blanchard2016} for more information on these
+interviews.
 
@@
 
+\hypertarget{correlation-analysis}{%
+\subsection{Correlation Analysis}\label{correlation-analysis}}
+
 \hypertarget{rank-correlation}{%
-\subsection{Rank Correlation}\label{rank-correlation}}
+\subsubsection{Rank Correlation}\label{rank-correlation}}
 
@@
 
+\hypertarget{principal-component-analysis}{%
+\subsubsection{Principal Component
+Analysis}\label{principal-component-analysis}}
+
+Olympic results
+
+Women: lead and bouldering belong to the same component, the other is
+speed.
+
+``overall ability''
+
+``power''
+
+\begin{figure}
+
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-20-1}
+
+}
+
+\caption{Women's Qualification PCA}\label{fig:unnamed-chunk-20}
+\end{figure}
+
+How each variable contributes to each principal component
+
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-21-1} \end{center}
+
+\begin{figure}
+
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-22-1}
+
+}
+
+\caption{Men's Qualification PCA}\label{fig:unnamed-chunk-22}
+\end{figure}
+
+\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-23-1} \end{center}
+
+\hypertarget{sensitivity-analysis}{%
+\subsection{Sensitivity Analysis}\label{sensitivity-analysis}}
+
 \hypertarget{leave-one-climber-out-analysis}{%
-\subsection{Leave-one-climber-out
+\subsubsection{Leave-one-climber-out
 Analysis}\label{leave-one-climber-out-analysis}}
@@
 
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-20-1}
+{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-26-1}
 
@@
 For each gender competition, each panel represents the rank of the drop-out athlete, with 0 being the original final results.
-Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-20}
+Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-26}
 \end{figure}
@@
 \hypertarget{rank-swapping-analysis}{%
-\subsection{Rank Swapping Analysis}\label{rank-swapping-analysis}}
-
-\hypertarget{principal-component-analysis}{%
-\subsection{Principal Component
-Analysis}\label{principal-component-analysis}}
-
-Olympic results
-
-Women: lead and bouldering belong to the same component, the other is
-speed.
-
-``overall ability''
-
-``power''
-
-\begin{figure}
-
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-23-1}
-
-}
-
-\caption{Women's Qualification PCA}\label{fig:unnamed-chunk-23}
-\end{figure}
-
-How each variable contributes to each principal component
-
-\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-24-1} \end{center}
-
-\begin{figure}
-
-{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-25-1}
-
-}
-
-\caption{Men's Qualification PCA}\label{fig:unnamed-chunk-25}
-\end{figure}
-
-\begin{center}\includegraphics{draft_files/figure-latex/unnamed-chunk-26-1} \end{center}
+\subsubsection{Rank Swapping Analysis}\label{rank-swapping-analysis}}
 

### 2021-10-04 R/displace_medalist.R
updated finalist displacement simulations

@@ -129,7 +129,7 @@ set.seed(80085)
 pct_dispace_medalists <- lapply(4:8, function(x) replicate(100, displace_climber(drop_rank = x)))
 names(pct_dispace_medalists) <- c("drop_r4", "drop_r5", "drop_r6", "drop_r7", "drop_r8")
 
-pct_displace_finalists <- lapply(9:20, function(x) replicate(100, displace_finalist(drop_rank = x)))
+pct_displace_finalists <- lapply(9:20, function(x) replicate(10000, displace_finalist(drop_rank = x)))
 names(pct_displace_finalists) <- paste0("drop", 9:20)
 save(pct_displace_finalists, file = "R/drop_qualifier_sims.rdata")
 ####################################

### 2021-10-04 cmsac2021/cmsac_rrc.Rmd
cmsac folder

---
title: |
  | An Examination of Sport Climbing's 
  | Competition Format and Scoring System
author: 
date: 2021 CMSAC Reproducible Research Competition
output: 
  pdf_document: 
    number_sections: true
    df_print: kable
abstract: "Sport climbing, which made its Olympic debut at the 2020 Summer Games, generally consists of three separate disciplines: speed climbing, bouldering, and lead climbing. However, the International Olympic Committee (IOC) only allowed one set of medals per gender for sport climbing. As a result, the governing body of sport climbing, rather than choosing only one of the three disciplines to include in the Olympics, decided to create a competition combining all three disciplines. In order to determine a winner, a combined scoring system was created using the product of the ranks across the three disciplines to determine an overall score for each climber. In this work, the sport climbing scoring system is evaluated through simulation to investigate general features of this system such as the advancement probabilities given certain placements and the sensitivity of the scoring system to small changes. Additionally, analyses of historical climbing contest results are presented and real examples of violations of the independence of irrelevant alternatives are illustrated. Finally, this work finds evidence that the current competition format is putting speed climbers at a disadvantage."
header-includes:
 \usepackage{setspace}
 \setlength{\parskip}{2mm}
 \setstretch{1.1}
 \usepackage{float}
 \floatplacement{figure}{H}
 \usepackage[skip=3pt]{caption}
bibliography: ref.bib
csl: ref_style.csl
indent: true
geometry: margin=1in
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      echo = FALSE,
                      warning = FALSE,
                      fig.pos = "H")
```

# Introduction

## Combined Competition Format

The 2020 Summer Olympics in Tokyo, Japan marked the first appearance of sport climbing in the Olympic stage. This sport is broken down into three disciplines: speed climbing, bouldering, and lead climbing. However, since only one set of Olympic medals is awarded to each gender event of sport climbing at Tokyo 2020, all three disciplines were included together, forming one single combined event. Under this "triathlon" format, every climber must compete in all three concentrations, and their individual score is determined as the product of the ranks across the three disciplines, with the lowest rank product declared the winner.

The decision to combine the three climbing events and only award one set of medals for both men's and women's events in the Olympics has received a large amount of criticism from climbing athletes all over the world. In a series of interviews conducted by Climbing Magazine in 2016 [@blanchard2016], a number of climbers shared their thoughts and concerns about the new Olympics climbing format. Legendary climber Lynn Hill compared the idea of combining speed climbing, bouldering, and lead climbing to "asking a middle distance runner to compete in the sprint". She then added "Speed climbing is a sport within our sport". Other climbers also hold the same opinion as Hill regarding speed climbing, using words and phrases like "bogus", "a bummer", "less than ideal", "not in support", and "cheesy and unfair" to describe the new combined competition format. Courtney Woods stated "Speed climbers will have the biggest disadvantage because their realm isn’t based on difficult movements". Mike Doyle believed "Honestly, the people that will suffer the most are the ones that focus only on speed climbing. Those skills/abilities don’t transfer as well to the other disciplines". The climbers also expressed their hope for a change in the competition format in future climbing tournaments.

## Rank-product Scoring System

At the 2020 Summer Olympics, both sport climbing competitions for male and female begin with 20 climbers who has previously qualified for the Olympics from qualifying events held in 2019 and 2020. All 20 athletes compete in each of the three disciplines in the qualification round, and their performances in each concentration are ranked from 1 to 20. A competitor's combined score is computed as the product of their ranks in the three events; specifically,
\begin{equation}
Score_i = R^S_i\times R^B_i\times R^L_i,
\end{equation} 
where $R^S_i$, $R^B_i$, and $R^L_i$ are the ranks of the $i$-th competitor in speed climbing, bouldering, and lead climbing, respectively.

The 8 qualifiers with the lowest score in terms of product of ranks across the three disciplines advance to the finals, where they once again compete in all three events. The overall score for each contestant in the final stage is determined by multiplying the rank of speed, bouldering, and lead disciplines, similar to qualification. The climbers with the lowest, second lowest, and third lowest product of ranks in the final wins the gold, silver, and bronze medal, respectively. This type of scoring system heavily rewards high finishes and relatively ignores poor finishing results. For instance, if climber A finished 1st, 20th, and 20th and climber B finished 10th, 10th, and 10th, climber B would have a score of 1000 whereas climber A would have a much better score of 400, despite finishing last in 2 out of 3 of the events.

To the best of our knowledge, we know of no sporting event, team or individual, that uses the product of ranks to determine an overall rankings.  There are examples of team sports that use the sum of ranks to determine the winning team such as cross country, where the team with the lowest sum of ranks of the top five runners is determined to be the winner [@hammond2007; @boudreau2018]. In addition, some individual sports such as the decathlon and heptathlon rely on a sum of scores from the ten or seven events, however, these scores are not determined based on the ranks of the competitors.  That is, a decathlete's score is entirely based upon their times, distances, and heights, and their overall score will be exactly the same if the times, distances, and heights remain the same regardless of the performance of other individuals [@westera2006]. Furthermore, there are other individual competitions consisted of several events combined that do base their scoring on ranks, such as crossfit competitions.  In each event, points are earned based on the competitor's rank in the event based on a scoring table, and a competitors final score is based on the sum of their scores across all the events [@crossfit2021].

In this paper, we attempt to perform statistical analysis to investigate the limitations of sport climbing's combined competition format and ranking system. We will find out whether the concerns of the professional climbers were valid, and whether there is a better format than the current combined competition structure. The manuscript is outlined as follows. We first begin with some descriptions of the data and methods in Section 2. Our analyses and results are then presented in Sections 3 and 4. Finally, in Section 5, we provide a summary of our main findings as well as some discussion to close out the paper.

# Data and Methodology

Our first set of data come from a simulation study that we conducted, with the purpose of examining the rankings and scoring for climbers in both qualification and final rounds. For each round, we performed 10000 simulations, and this was accomplished by randomly assigning the ranks of each event to every participant, with the assumption that the ranks are uniformly distributed. After the completion of the simulations, we calculated the total scores for every simulated round, as well as the final standings for the climbing athletes. The simulation results allow us to examine different properties of sport climbing's rank-product scoring system, including the distributions of total score for qualifying and final rounds, and the probabilities of advancing to the finals and winning a medal, given certain conditions.

We also rely on historical results from major climbing competitions in recent years. In addition to data from Tokyo 2020, we collected data on climbing contests that took place between 2018 and 2020 prior to the Olympics, where the combined format was used to determine the scores and ranks of climbers. The events include the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Youth Olympics. Data were obtained from various sources, including the event websites, Wikipedia, and the International Federation of Sport Climbing (IFSC). The main attributes of our datasets are the name and nationality of the climbers; the finishing place of climbers in speed climbing, bouldering, and lead climbing; the total score (which equals the product of the discipline ranks); the overall placement; and the performance statistics associated with speed (race time), bouldering (tops-zones-attempts), and lead climbing (highest hold reached). We utilize this data to analyze the correlations between the event ranks and final table position, as well as to look at how often the final orderings change if one athlete is dropped and the ranks for each discipline as well as the total score are re-calculated.

Since we are interested in applying a "leave-one-climber-out" type of analysis to sport climbing, there is a connection between this situation and the idea of independence of irrelevant alternatives (IIA). The IIA criterion is a property of a voting system which states that after a winner is determined, if one of the losing candidates drops out and the votes are recounted, there should not be a change in the winner. First mentioned by @arrow1951, the IIA condition is also known as Luce's choice axiom [@luce1959] in probability theory, and it has had a number of applications in the fields of decision theory, economics, and psychology over the years. We notice a link between the concept of IIA and the topic of ranking system in sports. As an illustration, suppose we have 3 players A, B, and C participating in a competition. If A finishes in the first place and C is later disqualified and removed, A should still win. If the original winner (A) loses the modified competition (with C removed), then the Independence of Irrelevant Alternatives has been violated. For our particular case, this sort of analysis can be helpful in examining the overall outcome for a climbing contest, specifically how a disqualification can affect the standings of medalists in the final round.

# Simulation Study

In this section, we discuss the results of our simulations described in Section 2. Our first topic of interest here are following questions:

\vspace{-4mm}

*   For a qualifier, what is the probability that they advance to the final round (i.e. finish in the top 8 of qualification), given that they win any discipline?

*   For a finalist, what is the probability that they win a medal (i.e. finish in the top 3 of the finals), given that they win any discipline?

\vspace{-4mm}

Our simulation study shows that a climber is almost guaranteed to finish in the top 8 of qualification and advance to the final round if they win at least one of the three climbing concentrations (99.48% chance). Regarding the finals, a climber is also very likely to claim a top 3 finish and bring home a medal if they win any event (85.01% chance). Moreover, we notice that if a climber wins any discipline, they are also more likely to finish first overall than any other positions in the eventual qualification and final rankings. This shows how significant winning a discipline is to the overall competition outcome for any given climber.

```{r}
library(tidyverse)
theme_set(theme_minimal())

climbing_sim <- function(nsim = 10000, nplay) {
  sims <- list()
  for (i in 1:nsim) {
    sims[[i]] <-
      bind_cols(
        player = 1:nplay,
        e1 = sample(1:nplay, replace = FALSE),
        e2 = sample(1:nplay, replace = FALSE),
        e3 = sample(1:nplay, replace = FALSE)
      ) %>%
      mutate(sim = i)
  }
  results <- bind_rows(sims) %>% 
    mutate(score = e1 * e2 * e3) %>% 
    group_by(sim) %>% 
    mutate(rank = rank(score, ties.method = "random")) %>%
    ungroup()

  return(results)
}

set.seed(1)
qual <- climbing_sim(nsim = 10000, nplay = 20)
final <- climbing_sim(nsim = 10000, nplay = 8)
```

```{r, fig.cap="The distributions for the probability of finishing at each rank of both qualification and final rounds, given that a climber wins any discipline, obtained from simulations. The probability of finishing exactly at each given rank is coded blue, whereas the probability of finishing at or below each given rank is coded maroon.", fig.height=3.5, fig.width=7}
final_dist <- final %>%
  filter(e1 == 1 | e2 == 1 | e3 == 1) %>%
  count(rank) %>%
  mutate(Probability = n / sum(n),
         Cumulative = cumsum(Probability)) %>%
  select(-n) %>%
  pivot_longer(!rank, names_to = "type") %>%
  mutate(round = "Final")

qual_dist <- qual %>%
  filter(e1 == 1 | e2 == 1 | e3 == 1) %>%
  count(rank) %>%
  mutate(Probability = n / sum(n),
         Cumulative = cumsum(Probability)) %>%
  select(-n) %>%
  pivot_longer(!rank, names_to = "type") %>%
  mutate(round = "Qualification")

final_dist %>%
  bind_rows(qual_dist) %>% 
  mutate(round = fct_relevel(round, "Qualification")) %>% 
  ggplot(aes(rank, value, fill = type)) +
  geom_col(position = "dodge") +
  facet_wrap(~ round, scales = "free") +
  scale_x_reverse(breaks = 1:11) +
  coord_flip() +
  labs(x = "Rank",
       y = "Probability Density",
       fill = "Distribution") +
  scale_fill_manual(values = c("maroon", "midnightblue")) +
  theme(panel.grid.major.y = element_blank()) +
  theme(legend.position = "bottom",
        legend.margin = margin(-5),
        legend.key.size = unit(0.4, "cm"))
```

\vspace{-5mm}

In addition, we are interested in examining the distribution of the total score for both qualification and final rounds.
Figure 2 is a summary of the expected score for each qualification and final placement. According to our simulations, on average, the qualification score that a contestant should aim for in order to move on to the final round is 435 (for 8th rank). Furthermore, we observe that in order to obtain a climbing medal, the average scores that put a finalist in position to stand on the tri-level podium at the end of the competition are 10, 21, and 34 for gold, silver, and bronze medals, respectively.

```{r, fig.cap="The average scores for the top 10 qualification ranks and all 8 final ranks, obtained from simulations.", fig.height=3.3, fig.width=6.6}
final %>% 
  group_by(rank) %>% 
  summarize(avg_score = mean(score)) %>%
  mutate(round = "Final",
         color = ifelse(rank == 1, "gold",
                        ifelse(rank == 2, "#C0C0C0",
                               ifelse(rank == 3, "#A77044", "lightblue")))) %>% 
  bind_rows(
    qual %>% 
      group_by(rank) %>% 
      summarize(avg_score = mean(score)) %>%
      filter(rank <= 10) %>% 
      mutate(round = "Qualification",
             color = ifelse(rank < 9, "maroon", "lightblue"))
  ) %>% 
  mutate(round = fct_relevel(round, "Qualification")) %>% 
  ggplot(aes(rank, avg_score, fill = color)) +
  geom_col() +
  geom_text(aes(label = ceiling(avg_score)), color = "black", size = 3, vjust = -0.3) +
  facet_wrap(~ round, scales = "free") +
  scale_x_continuous(breaks = 1:10) +
  scale_fill_identity() +
  labs(x = "Rank",
       y = "Average Score") +
  theme(panel.grid.major.x = element_blank())
```

# Analysis

## Correlation

Throughout this section, we will be using data from the Women's Qualification at the 2020 Summer Olympics [@datawiki] as a case study for examining the relationship between the climber rankings in each individual discipline and the overall standings. Figure 3 is a multi-panel matrix of scatterplots of the ranks of the individual events and the final women's qualification standings, and we use Kendall’s Tau [@kendall1938] as our measure of ordinal association between the ranked variables. Table 1 shows the Kendall rank correlation coefficients between the overall rank and the ranks of speed, bouldering, and lead disciplines; along with corresponding $95\%$ confidence intervals obtained from bootstrapping [@efron1986].

It is evidently clear that there exists a fairly strong and positive association between the final rank and the ranks of both bouldering ($\tau = 0.432$, $p$-value $=0.00735$, Bootstrapped $95\%$ CI: (0.107, 0.691)) and lead climbing ($\tau = 0.463$, $p$-value $=0.00378$, Bootstrapped $95\%$ CI: (0.112, 0.753)). This implies that climbers that finish high in both bouldering and lead also tend to finish at a higher ranking spot overall. Hence this combined competition format greatly benefits those specializing in bouldering and lead, the two events that require strength and power.

On the other hand, the correlation with the final rank is not as strong for speed climbing as the other two events ($\tau = 0.147$), and there is insufficient evidence for an association between the rank of speed climbing and the overall rank ($p$-value $=0.386$, Bootstrapped $95\%$ CI: (-0.191, 0.469)). Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those with expertise in the other two concentrations. This validates the concerns of climbers from the interviews mentioned in Section 1.1.

```{r, fig.height=3.5, fig.width=8, fig.cap= "Scatterplots of overall rank and speed, bouldering, and lead ranks for Women's Qualification at Tokyo 2020."}
wq <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2020_olympics/wq.csv")

wq %>% 
  select(overall, speed, bouldering, lead) %>% 
  rename_with(str_to_title) %>% 
  pivot_longer(!Overall, 
               names_to = "Discipline",
               values_to = "Rank") %>%
  mutate(Discipline = factor(Discipline, levels = c("Speed", "Bouldering", "Lead"))) %>% 
  ggplot(aes(Rank, Overall)) +
  geom_point() +
  geom_smooth(span = 1, color = "darkblue") +
  facet_wrap(~ Discipline) +
  theme(panel.grid.major.x = element_blank())
```

```{r, eval = FALSE}
wq %>% 
  select(overall, speed, bouldering, lead) %>% 
  cor(method = "kendall")

cor.test(wq$overall, wq$bouldering, method = "kendall")
cor.test(wq$overall, wq$lead, method = "kendall")
cor.test(wq$overall, wq$speed, method = "kendall")

library(NSM3)
set.seed(21)
kendall.ci(wq$overall, wq$bouldering, bootstrap=TRUE, B=1000)
kendall.ci(wq$overall, wq$lead, bootstrap=TRUE, B=1000)
kendall.ci(wq$overall, wq$speed, bootstrap=TRUE, B=1000)
```

```{r}
tribble(
  ~Discipline, ~`Kendall's Tau`, ~`Test Statistic`, ~`p-value`, ~ `Bootstrapped 95% CI`,
  "Speed", 0.147, 109, "0.386", "(-0.191, 0.469)",
  "Bouldering", 0.432, 136, "0.00735", "(0.107, 0.691)",
  "Lead", 0.463, 139, "0.00378", "(0.112, 0.753)"
) %>% 
  knitr::kable(caption = "Kendall's tau values, alongside correlation test statistics, p-values, and bootstrapped 95% confidence intervals, for the overall rank and the rank of speed, bouldering, and lead disciplines of Women's Qualification at Tokyo 2020.")
```

In addition to computing the rank correlation coefficients, we also performed principal component analysis (PCA) to summarize the correlations among a set of observed performance variables associated with the three climbing disciplines. In particular, using the data from the Women's Qualification round at Tokyo 2020, we look at the racing time (in seconds) for speed; the number of successfully completed boulders ("tops"); and the number of holds reached for lead.

Figure 4 is a PCA biplot showing the PC scores of the climbers and loadings of the skill variables. We notice that the lead and bouldering performances strongly influence PC1, while speed time is the only variable contributing to PC2, separated from the other two skills. Moreover, since these two vectors are close and form a small angle, the two variables they represent (bouldering tops and lead holds reached) are positively correlated. This implies that a climber that does well in bouldering is also very likely to deliver a good performance in lead climbing.

```{r, out.width = "72%", fig.cap= "PCA Biplot for Women's Qualification at Tokyo 2020", fig.align="center"}
knitr::include_graphics("pca.png")
```

## Leave-one-climber-out Analysis

Another interesting question that we are interested in investigating is "What would happen to the rankings if a single climber is removed?". This is related to the concept of independence of irrelevant alternatives (IIA), as mentioned in Section 2.

In the analysis that follows, we use data from the 2018 Youth Olympics Women's Final [@2018youth]. This event also implemented the combined format and rank-product scoring system, but only consisted of 6 climbers competing in the final, rather than 8 as Tokyo 2020. To obtain the desired data, after an athlete is dropped (by their rank), the new ranks for each discipline of the remaining players are re-calculated. The new final scores can then be obtained by multiplying the three event ranks, which then determines the new overall ranks.

Figure 5 shows the modified versions of the rankings after each ranked climber is excluded for the Women's Final at the 2018 Youth Olympics. We have clear evidence from this plot that removing a single climber changes the rankings drastically, especially in terms of the order of medalists. One particular interesting case is where an athlete's position change when someone who originally finished behind them drops out. This situation is illustrated by panel 5 of the women's competition, where the fifth-place climber, Krasovskaia, was excluded; and Meul, whose actual final rank was fourth, moved up to the second spot and claim the silver medal. Furthermore, this implies that the ranking system of Olympics sport climbing violates the IIA criterion. This reveals another issue of the sport climbing scoring system, which is that there is a dependency on irrelevant parties.

```{r, fig.height=3.5, fig.width=6.5, fig.cap="This figure illustrates the changes to the final rankings of the 2018 Youth Olympics Women's Final when we leave out one climber. Each panel represents the rank of the drop-out athlete, with 0 being the original final results. Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar"}

wf <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2018_youth_olympics/women_final.csv")

drop_rerank <- function(df) {
  rerank <- list()
  for (i in 1:nrow(df)) {
    rerank[[i]] <- df[-i, ] %>%
      mutate(rank_drop = i)
  }

  rerank_df <- df %>%
    mutate(rank_drop = 0) %>%
    bind_rows(rerank) %>%
    group_by(rank_drop) %>%
    mutate(
      speed = rank(speed),
      bould = rank(bould),
      lead = rank(lead),
      total = speed * bould * lead
    ) %>%
    arrange(total, .by_group = TRUE) %>%
    ungroup() %>%
    group_by(rank_drop, total) %>%
    mutate(
      speed_tb = ifelse(speed < lag(speed), 1, 0),
      bould_tb = ifelse(bould < lag(bould), 1, 0),
      lead_tb = ifelse(lead < lag(lead), 1, 0),
      tb = speed_tb + bould_tb + lead_tb,
      tb = ifelse(is.na(tb), 1, tb)
    ) %>%
    ungroup() %>%
    group_by(rank_drop) %>%
    arrange(total,-tb, .by_group = TRUE) %>%
    mutate(rank = row_number(),
           last = str_to_title(last))

  return(rerank_df)
}

drop_rerank(wf) %>% 
  mutate(last = fct_reorder(last, -rank),
         rank = as.factor(rank),
         rank_change = ifelse(
           rank_drop %in% c(0, 1, 4, 6) |
             rank_drop == 2 & rank %in% 1:3 |
             rank_drop == 3 & rank %in% c(1, 4, 5) |
             rank_drop == 5 & rank %in% c(1, 5),
           "no", "yes")) %>%
  ggplot(aes(x = last, y = total, fill = rank_change)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = rank), hjust = -0.2, size = 3, color = "black") +
  coord_flip() +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
            data = ~ filter(., rank_drop %in% c(2, 3, 5)), 
            color = "black", size = 1.5, fill = NA, inherit.aes = FALSE) +
  facet_wrap(~ rank_drop, nrow = 2, ncol = 4) +
  expand_limits(y = 62, x = 0:7) +
  scale_fill_manual(values = c("grey", "red")) +
  labs(y = "Score", x = NULL) +
  theme(axis.ticks = element_blank(),
        panel.grid.minor = element_blank())
```

# Conclusion and Discussion

In this paper, we examined the general features of the rank-product scoring system of sport climbing, in particular, the advancement probabilities and scores for the climbers. Most importantly, through analyses of past competition data, we pointed out several problems of the combined competition format and rank-product scoring system of sport climbing. First, combining the three disciplines speed, bouldering, and lead together into a "triathlon" competition format is putting speed climbers in an unfavorable position. Second, the sport climbing scoring system violates the independence of irrelevant alternatives, as the orderings of medalists can be affected with a drop-out of a lower-ranked climber.

We suggest that speed climbing should have its own set of medals in future climbing tournaments, whereas bouldering and lead climbing can be combined into one event. In fact, it was confirmed that in the next Summer Olympics held in Paris in 2024, there will be two competitions and two sets of climbing medals for each gender event: combined lead and bouldering, and speed-only [@goh2020]. This is consistent with what we have shown, as bouldering and lead climbing performances are highly correlation with each other and with the overall result, whereas speed climbing should be separated from the combined format.

In closing, there are so many sports, similar to sport climbing, that have generally been untouched in terms of analytics. With the rise of statistics and data science applications in sports, we hope that more people will be interested in working on data problems related to sport climbing in order to make this sport better. One possible future work we could look into is turning the data we collected for this project into an accessible package or database of historical climbing competition results, which would be vital in promoting more engagements in the analytics of sport climbing.

# References
